\subsection*{Psalm 73}
\addcontentsline{toc}{subsection}{Psalm 73}

A Psalm of Asaph.

1 Yet God is good to Israel,\\
to each pure-hearted one.\\
2 But as for me, my steps near slipped,\\
my feet were almost gone.

3 For I envious was, and grudged\\
the foolish folk to see,\\
When I perceived the wicked sort\\
enjoy prosperity.

4 For still their strength continueth firm;\\
their death of bands is free.\\
5 They are not toiled like other men,\\
nor plagued, as others be.

6 Therefore their pride, like to a chain,\\
them compasseth about;\\
And, as a garment, violence\\
doth cover them throughout.

7 Their eyes stand out with fat; they have\\
more than their hearts could wish.\\
8 They are corrupt; their talk of wrong\\
both lewd and lofty is.

9 They set their mouth against the heav’ns\\
in their blasphemous talk;\\
And their reproaching tongue throughout\\
the earth at large doth walk.

10 His people oftentimes for this\\
look back, and turn about;\\
Sith waters of so full a cup\\
to these are poured out.

11 And thus they say, How can it be\\
that God these things doth know?\\
Or, Can there in the Highest be\\
knowledge of things below?

12 Behold, these are the wicked ones,\\
yet prosper at their will\\
In worldly things; they do increase\\
in wealth and riches still.

13 I verily have done in vain\\
my heart to purify;\\
To no effect in innocence\\
washed my hands have I.

14 For daily, and all day throughout,\\
great plagues I suffered have;\\
Yea, ev’ry morning I of new\\
did chastisement receive.

15 If in this manner foolishly\\
to speak I would intend,\\
Thy children’s generation,\\
behold, I should offend.

16 When I this thought to know, it was\\
too hard a thing for me;\\
17 Till to God’s sanctuary I went,\\
then I their end did see.

18 Assuredly thou didst them set\\
a slipp’ry place upon;\\
Them suddenly thou castedst down\\
into destruction.

19 How in a moment suddenly\\
to ruin brought are they!\\
With fearful terrors utterly\\
they are consumed away.

20 Ev’n like unto a dream, when one\\
from sleeping doth arise;\\
So thou, O Lord, when thou awak’st,\\
their image shalt despise.

21 Thus grieved was my heart in me,\\
and me my reins oppressed:\\
22 So rude was I, and ignorant,\\
and in thy sight a beast.

23 Nevertheless continually,\\
O Lord, I am with thee:\\
Thou dost me hold by my right hand,\\
and still upholdest me.

24 Thou, with thy counsel, while I live,\\
wilt me conduct and guide;\\
And to thy glory afterward\\
receive me to abide.

25 Whom have I in the heavens high\\
but thee, O Lord, alone?\\
And in the earth whom I desire\\
besides thee there is none.

26 My flesh and heart doth faint and fail,\\
but God doth fail me never:\\
For of my heart God is the strength\\
and portion for ever.

27 For, lo, they that are far from thee\\
for ever perish shall;\\
Them that a whoring from thee go\\
thou hast destroyed all.

28 But surely it is good for me\\
that I draw near to God:
In God I trust, that all thy works\\
I may declare abroad.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 74}
\addcontentsline{toc}{subsection}{Psalm 74}

Maschil of Asaph.

1 O God, why hast thou cast us off?\\
is it for evermore?\\
Against thy pasture-sheep why doth\\
thine anger smoke so sore?

2 O call to thy rememberance\\
thy congregation,\\
Which thou hast purchased of old;\\
still think the same upon:

The rod of thine inheritance,\\
which thou redeemed hast,\\
This Zion hill, wherein thou hadst\\
thy dwelling in times past.

3 To these long desolations\\
thy feet lift, do not tarry;\\
For all the ills thy foes have done\\
within thy sanctuary.

4 Amidst thy congregations\\
thine enemies do roar:\\
Their ensigns they set up for signs\\
of triumph thee before.

5 A man was famous, and was had\\
in estimation,\\
According as he lifted up\\
his axe thick trees upon.

6 But all at once with axes now\\
and hammers they go to,\\
And down the carved work thereof\\
they break, and quite undo.

7 They fired have thy sanctuary,\\
and have defiled the same,\\
By casting down unto the ground\\
the place where dwelt thy name.

8 Thus said they in their hearts, Let us\\
destroy them out of hand:\\
They burnt up all the synagogues\\
of God within the land.

9 Our signs we do not now behold;\\
there is not us among\\
A prophet more, nor any one\\
that knows the time how long.

10 How long, Lord, shall the enemy\\
thus in reproach exclaim?\\
And shall the adversary thus\\
always blaspheme thy name?

11 Thy hand, ev’n thy right hand of might,\\
why dost thou thus draw back?\\
O from thy bosom pluck it out\\
for our deliv’rance sake.

12 For certainly God is my King,\\
ev’n from the times of old,\\
Working in midst of all the earth\\
salvation manifold.

13 The sea, by thy great pow’r, to part\\
asunder thou didst make;\\
And thou the dragons’ heads, O Lord,\\
within the waters brake.

14 The leviathan’s head thou brak’st\\
in pieces, and didst give\\
Him to be meat unto the folk\\
in wilderness that live.

15 Thou clav’st the fountain and the flood,\\
which did with streams abound:\\
Thou dry’dst the mighty waters up\\
unto the very ground.

16 Thine only is the day, O Lord,\\
thine also is the night;\\
And thou alone prepared hast\\
the sun and shining light.

17 By thee the borders of the earth\\
were settled ev’ry where:\\
The summer and the winter both\\
by thee created were.

18 That th’ enemy reproached hath,\\
O keep it in record;\\
And that the foolish people have\\
blasphemed thy name, O Lord.

19 Unto the multitude do not\\
thy turtle’s soul deliver:\\
The congregation of thy poor\\
do not forget for ever.

20 Unto thy cov’nant have respect;\\
for earth’s dark places be\\
Full of the habitations\\
of horrid cruelty.

21 O let not those that be oppressed\\
return again with shame:\\
Let those that poor and needy are\\
give praise unto thy name.

22 Do thou, O God, arise and plead\\
the cause that is thine own:\\
Remember how thou art reproached\\
still by the foolish one.

23 Do not forget the voice of those \\
that are thine enemies:\\
Of those the tumult ever grows\\
that do against thee rise.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 75}
\addcontentsline{toc}{subsection}{Psalm 75}

To the chief Musician, Al-taschith,
A Psalm \emph{or} Song of Asaph.

1 To thee, O God, do we give thanks,\\
we do give thanks to thee;\\
Because thy wondrous works declare\\
thy great name near to be.

2 I purpose, when I shall receive\\
the congregation,\\
That I shall judgment uprightly\\
render to ev’ry one.

3 Dissolved is the land, with all\\
that in the same do dwell;\\
But I the pillars thereof do\\
bear up, and stablish well.

4 I to the foolish people said,\\
Do not deal foolishly;\\
And unto those that wicked are,\\
Lift not your horn on high.

5 Lift not your horn on high, nor speak\\
6 with stubborn neck. But know,\\
That not from east, nor west, nor south,\\
promotion doth flow.

7 But God is judge; he puts down one,\\
and sets another up.\\
8 For in the hand of God most high\\
of red wine is a cup:

‘Tis full of mixture, he pours forth,\\
and makes the wicked all\\
Wring out the bitter dregs thereof;\\
yea, and they drink them shall.

9 But I for ever will declare,\\
I Jacob’s God will praise.\\
10 All horns of lewd men I’ll cut off;\\
but just men’s horns will raise.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 76}
\addcontentsline{toc}{subsection}{Psalm 76}

To the chief Musician on Neginoth,
A Psalm \emph{or} Song of Asaph.

1 In Judah’s land God is well known,\\
his name’s in Isr’el great:\\
2 In Salem is his tabernacle,\\
in Zion is his seat.

3 There arrows of the bow he brake,\\
the shield, the sword, the war.\\
4 More glorious thou than hills of prey,\\
more excellent art far.

5 Those that were stout of heart are spoiled,\\
they slept their sleep outright;\\
And none of those their hands did find,\\
that were the men of might.

6 When thy rebuke, O Jacob’s God,\\
had forth against them past,\\
Their horses and their chariots both\\
were in a dead sleep cast.

7 Thou, Lord, ev’n thou art he that should\\
be feared; and who is he\\
That may stand up before thy sight,\\
if once thou angry be?

8 From heav’n thou judgment caused be heard;\\
the earth was still with fear,\\
9 When God to judgment rose, to save\\
all meek on earth that were.

10 Surely the very wrath of man\\
unto thy praise redounds:\\
Thou to the remnant of his wrath\\
wilt set restraining bounds.

11 Vow to the Lord your God, and pay:\\
all ye that near him be,\\
Bring gifts and presents unto him;\\
for to be feared is he.

12 By him the sp’rits shall be cut off\\
of those that princes are:\\
Unto the kings that are on earth\\
he fearful doth appear.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 77}
\addcontentsline{toc}{subsection}{Psalm 77}

To the chief Musician, to Jeduthun,
A Psalm of Asaph.

1 Unto the Lord I with my voice,\\
I unto God did cry;\\
Ev’n with my voice, and unto me\\
his ear he did apply.

2 I in my trouble sought the Lord,\\
my sore by night did run,\\
And ceased not; my grieved soul\\
did consolation shun.

3 I to remembrance God did call,\\
yet trouble did remain;\\
And overwhelm’d my spirit was,\\
whilst I did sore complain.

4 Mine eyes, debarred from rest and sleep,\\
thou makest still to wake;\\
My trouble is so great that I\\
unable am to speak.

5 The days of old to mind I called,\\
and oft did think upon\\
The times and ages that are past\\
full many years agone.

6 By night my song I call to mind,\\
and commune with my heart;\\
My sp’rit did carefully enquire\\
how I might ease my smart.

7 For ever will the Lord cast off,\\
and gracious be no more?\\
8 For ever is his mercy gone?\\
fails his word evermore?

9 Is’t true that to be gracious\\
the Lord forgotten hath?\\
And that his tender mercies he\\
hath shut up in his wrath?

10 Then did I say, That surely this\\
is mine infirmity:\\
I’ll mind the years of the right hand\\
of him that is most High.

11 Yea, I remember will the works\\
performed by the Lord:\\
The wonders done of old by thee\\
I surely will record.

12 I also will of all thy works\\
my meditation make;\\
And of thy doings to discourse\\
great pleasure I will take.

13 O God, thy way most holy is\\
within thy sanctuary;\\
And what god is so great in pow’r\\
as is our God most high?

14 Thou art the God that wonders do’st\\
by thy right hand most strong:\\
Thy mighty pow’r thou hast declared\\
the nations among.

15 To thine own people with thine arm\\
thou didst redemption bring;\\
To Jacob’s sons, and to the tribes\\
of Joseph that do spring.

16 The waters, Lord, perceived thee,\\
the waters saw thee well;\\
And they for fear aside did flee;\\
the depths on trembling fell.

17 The clouds in water forth were poured,\\
sound loudly did the sky;\\
And swiftly through the world abroad\\
thine arrows fierce did fly.

18 Thy thunder’s voice alongst the heav’n\\
a mighty noise did make;\\
By lightnings lightened was the world,\\
th’ earth tremble did and shake.

19 Thy way is in the sea, and in\\
the waters great thy path;\\
Yet are thy footsteps hid, O Lord;\\
none knowledge thereof hath.

20 Thy people thou didst safely lead,\\
like to a flock of sheep;\\
By Moses’ hand and Aaron’s thou\\
didst them conduct and keep.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 78}
\addcontentsline{toc}{subsection}{Psalm 78}

Maschil of Asaph.


1 Attend, my people, to my law;\\
thereto give thou an ear;\\
The words that from my mouth proceed\\
attentively do hear.

2 My mouth shall speak a parable,\\
and sayings dark of old;\\
3 The same which we have heard and known,\\
and us our fathers told.

4 We also will them not conceal\\
from their posterity;\\
Them to the generation\\
to come declare will we:

The praises of the Lord our God,\\
and his almighty strength,\\
The wondrous works that he hath done,\\
we will show forth at length.

5 His testimony and his law\\
in Isr’el he did place,\\
And charged our fathers it to show\\
to their succeeding race;

6 That so the race which was to come\\
might well them learn and know;\\
And sons unborn, who should arise,\\
might to their sons them show:

7 That they might set their hope in God,\\
and suffer not to fall\\
His mighty works out of their mind,\\
but keep his precepts all:

8 And might not, like their fathers, be\\
a stiff rebellious race;\\
A race not right in heart; with God\\
whose sp’rit not steadfast was.

9 The sons of Ephraim, who nor bows\\
nor other arms did lack,\\
When as the day of battle was,\\
they faintly turned back.

10 They brake God’s cov’nant, and refused\\
in his commands to go;\\
11 His works and wonders they forgot,\\
which he to them did show.

12 Things marvelous he brought to pass;\\
their fathers them beheld\\
Within the land of Egypt done,\\
yea, ev’n in Zoan’s field.

13 By him divided was the sea,\\
he caused them through to pass;\\
And made the waters so to stand,\\
as like an heap it was.

14 With cloud by day, with light of fire\\
all night, he did them guide.\\
15 In desert rocks he clave, and drink,\\
as from great depths, supplied.

16 He from the rock brought streams, like floods\\
made waters to run down.\\
17 Yet sinning more, in desert they\\
provoked the Highest One.

18 For in their heart they tempted God,\\
and, speaking with mistrust,\\
They greedily did meat require\\
to satisfy their lust.

19 Against the Lord himself they spake,\\
and, murmuring, said thus,\\
A table in the wilderness\\
can God prepare for us?

20 Behold, he smote the rock, and thence\\
came streams and waters great;\\
But can he give his people bread?\\
and send them flesh to eat?

21 The Lord did hear, and waxed wroth;\\
so kindled was a flame\\
‘Gainst Jacob, and ‘gainst Israel\\
up indignation came.

22 For they believed not God, nor trust\\
in his salvation had;\\
23 Though clouds above he did command,\\
and heav’n’s doors open made,

24 And manna rained on them, and gave\\
them corn of heav’n to eat.\\
25 Man angels’ food did eat; to them\\
he to the full sent meat.

26 And in the heaven he did cause\\
an eastern wind to blow;\\
And by his power he let out\\
the southern wind to go.

27 Then flesh as thick as dust he made\\
to rain down them among;\\
And feathered fowls, like as the sand\\
which li’th the shore along.

28 At his command amidst their camp\\
these show’rs of flesh down fell,\\
All round about the tabernacles\\
and tents where they did dwell.

29 So they did eat abundantly,\\
and had of meat their fill;\\
For he did give to them what was\\
their own desire and will.

30 They from their lust had not estranged\\
their heart and their desire;\\
But while the meat was in their mouths,\\
which they did so require,

31 God’s wrath upon them came, and slew\\
the fattest of them all;\\
So that the choice of Israel,\\
o’erthrown by death, did fall.

32 Yet, notwithstanding of all this,\\
they sinned still the more;\\
And though he had great wonders wrought,\\
believed him not therefore:

33 Wherefore their days in vanity\\
he did consume and waste;\\
And by his wrath their wretched years\\
away in trouble past.

34 But when he slew them, then they did\\
to seek him show desire;\\
Yea, they returned, and after God\\
right early did enquire.

35 And that the Lord had been their Rock,\\
they did remember then;\\
Ev’n that the high almighty God\\
had their Redeemer been.

36 Yet with their mouth they flattered him,\\
and spake but feignedly;\\
And they unto the God of truth\\
with their false tongues did lie.

37 For though their words were good, their heart\\
with him was not sincere;\\
Unsteadfast and perfidious\\
they in his cov’nant were.

38 But, full of pity, he forgave\\
their sin, them did not slay;\\
Nor stirred up all his wrath, but oft\\
his anger turned away.

39 For that they were but fading flesh\\
to mind he did recall;\\
A wind that passeth soon away,\\
and not returns at all.

40 How often did they him provoke\\
within the wilderness!\\
And in the desert did him grieve\\
with their rebelliousness!

41 Yea, turning back, they tempted God,\\
and limits set upon\\
Him, who in midst of Isr’el is\\
the only Holy One.

42 They did not call to mind his pow’r,\\
nor yet the day when he\\
Delivered them out of the hand\\
of their fierce enemy;

43 Nor how great signs in Egypt land\\
he openly had wrought;\\
What miracles in Zoan’s field\\
his hand to pass had brought.

44 How lakes and rivers ev’ry where\\
he turned into blood;\\
So that nor man nor beast could drink\\
of standing lake or flood.

45 He brought among them swarms of flies,\\
which did them sore annoy;\\
And divers kinds of filthy frogs\\
he sent them to destroy.

46 He to the caterpillar gave\\
the fruits of all their soil;\\
Their labors he delivered up\\
unto the locusts’ spoil.

47 Their vines with hail, their sycamores\\
he with the frost did blast:\\
48 Their beasts to hail he gave; their flocks\\
hot thunderbolts did waste.

49 Fierce burning wrath he on them cast,\\
and indignation strong,\\
And troubles sore, by sending forth\\
ill angels them among.

50 He to his wrath made way; their soul\\
from death he did not save;\\
But over to the pestilence\\
the lives of them he gave.

51 In Egypt land the first-born all\\
he smote down ev’ry where;\\
Among the tents of Ham, ev’n these\\
chief of their strength that were.

52 But his own people, like to sheep,\\
thence to go forth he made;\\
And he, amidst the wilderness,\\
them, as a flock, did lead.

53 And he them safely on did lead,\\
so that they did not fear;\\
Whereas their en’mies by the sea\\
quite overwhelmed were.

54 To borders of his sanctuary\\
the Lord his people led,\\
Ev’n to the mount which his right hand\\
for them had purchased.

55 The nations of Canaan,\\
by his almighty hand,\\
Before their face he did expel\\
out of their native land;

Which for inheritance to them\\
by line he did divide,\\
And made the tribes of Israel\\
within their tents abide.

56 Yet God most high they did provoke,\\
and tempted ever still;\\
And to observe his testimonies\\
did not incline their will:

57 But, like their fathers, turned back,\\
and dealt unfaithfully:\\
Aside they turned, like a bow\\
that shoots deceitfully.

58 For they to anger did provoke\\
him with their places high;\\
And with their graven images\\
moved him to jealousy.

59 When God heard this, he waxed wroth,\\
and much loathed Isr’el then:\\
60 So Shiloh’s tent he left, the tent\\
which he had placed with men.

61 And he his strength delivered\\
into captivity;\\
He left his glory in the hand\\
of his proud enemy.

62 His people also he gave o’er\\
unto the sword’s fierce rage:\\
So sore his wrath inflamed was\\
against his heritage.

63 The fire consumed their choice young men;\\
their maids no marriage had;\\
64 And when their priests fell by the sword,\\
their wives no mourning made.

65 But then the Lord arose, as one\\
that doth from sleep awake;\\
And like a giant that, by wine\\
refreshed, a shout doth make:

66 Upon his en’mies’ hinder parts\\
he made his stroke to fall;\\
And so upon them he did put\\
a shame perpetual.

67 Moreover, he the tabernacle\\
of Joseph did refuse;\\
The mighty tribe of Ephraim\\
he would in no wise choose:

68 But he did choose Jehudah’s tribe\\
to be the rest above;\\
And of mount Zion he made choice,\\
which he so much did love.

69 And he his sanctuary built\\
like to a palace high,\\
Like to the earth which he did found\\
to perpetuity.

70 Of David, that his servant was,\\
he also choice did make,\\
And even from the folds of sheep\\
was pleased him to take:

71 From waiting on the ewes with young,\\
he brought him forth to feed\\
Israel, his inheritance,\\
his people, Jacob’s seed.

72 So after the integrity\\
he of his heart them fed;\\
And by the good skill of his hands\\
them wisely governed.


% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 79}
\addcontentsline{toc}{subsection}{Psalm 79}

A Psalm of Asaph.

1 O God, the heathen entered have\\
thine heritage; by them\\
Defiled is thy house: on heaps\\
they laid Jerusalem.

2 The bodies of thy servants they\\
have cast forth to be meat\\
To rav’nous fowls; thy dear saints’ flesh\\
they gave to beasts to eat.

3 Their blood about Jerusalem\\
like water they have shed;\\
And there was none to bury them\\
when they were slain and dead.

4 Unto our neighbors a reproach\\
most base become are we;\\
A scorn and laughingstock to them\\
that round about us be.

5 How long, Lord, shall thine anger last?\\
wilt thou still keep the same?\\
And shall thy fervent jealousy\\
burn like unto a flame?

6 On heathen pour thy fury forth,\\
that have thee never known,\\
And on those kingdoms which thy name\\
have never called upon.

7 For these are they who Jacob have\\
devoured cruelly;\\
And they his habitation\\
have caused waste to lie.

8 Against us mind not former sins;\\
thy tender mercies show;\\
Let them prevent us speedily,\\
for we’re brought very low.

9 For thy name’s glory help us, Lord,\\
who hast our Savior been:\\
Deliver us; for thy name’s sake,\\
O purge away our sin.

10 Why say the heathen, where’s their God?\\
let him to them be known;\\
When those who shed thy servants’ blood\\
are in our sight o’erthrown.

11 O let the pris’ner’s sighs ascend\\
before thy sight on high;\\
Preserve those in thy mighty pow’r\\
that are designed to die.

12 And to our neighbors’ bosom cause\\
it sev’n-fold rendered be,\\
Ev’n the reproach wherewith they have,\\
O Lord, reproached thee.

13 So we thy folk, and pasture-sheep,\\
shall give thee thanks always;\\
And unto generations all\\
we will show forth thy praise.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 80}
\addcontentsline{toc}{subsection}{Psalm 80}

To the chief musician upon Shoshannim, Eduth,
A Psalm of Asaph.

1 Hear, Isr’el’s Shepherd! like a flock\\
thou that dost Joseph guide;\\
Shine forth, O thou that dost between\\
the cherubims abide.

2 In Ephraim’s, and Benjamin’s\\
and in Manasseh’s sight,\\
O come for our salvation;\\
stir up thy strength and might.

3 Turn us again, O Lord our God,\\
and upon us vouchsafe\\
To make thy countenance to shine,\\
and so we shall be safe.

4 O Lord of hosts, Almighty God,\\
how long shall kindled be\\
Thy wrath against the prayer made\\
by thine own folk to thee?

5 Thou tears of sorrow giv’st to them\\
instead of bread to eat;\\
Yea, tears instead of drink thou giv’st\\
to them in measure great.

6 Thou makest us a strife unto\\
our neighbours round about;\\
Our enemies among themselves\\
at us do laugh and flout.

7 Turn us again, O God of hosts,\\
and upon us vouchsafe\\
To make thy countenance to shine,\\
and so we shall be safe.

8 A vine from Egypt brought thou hast,\\
by thine outstretched hand;\\
And thou the heathen out didst cast,\\
to plant it in their land.

9 Before it thou a room didst make,\\
where it might grow and stand;\\
Thou causedst it deep root to take,\\
and it did fill the land.

10 The mountains vail’d were with its shade,\\
as with a covering;\\
Like goodly cedars were the boughs\\
which out from it did spring.

11 Upon the one hand to the sea\\
her boughs she did out send;\\
On th’ other side unto the flood\\
her branches did extend.

12 Why hast thou then thus broken down,\\
and ta’en her hedge away?\\
So that all passengers do pluck,\\
and make of her a prey.

13 The boar who from the forest comes\\
doth waste it at his pleasure;\\
The wild beast of the field also\\
devours it out of measure.

14 O God of hosts, we thee beseech,\\
return now unto thine;\\
Look down from heav’n in love, behold,\\
and visit this thy vine:

15 This vineyard, which thine own right hand\\
hath planted us among;\\
And that same branch, which for thyself\\
thou hast made to be strong.

16 Burnt up it is with flaming fire,\\
it also is cut down:\\
They utterly are perished,\\
when as thy face doth frown.

17 O let thy hand be still upon\\
the Man of thy right hand,\\
The Son of man, whom for thyself\\
thou madest strong to stand.

18 So henceforth we will not go back,\\
nor turn from thee at all:\\
O do thou quicken us, and we\\
upon thy name will call.

19 Turn us again, Lord God of hosts,\\
and upon us vouchsafe\\
To make thy countenance to shine,\\
and so we shall be safe.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 81}
\addcontentsline{toc}{subsection}{Psalm 81}

To the chief Musician upon Gittith,
\emph{A Psalm} of Asaph.

1 Sing loud to God our strength; with joy\\
to Jacob’s God do sing.\\
2 Take up a psalm, the pleasant harp,\\
timbrel and psalt’ry bring.

3 Blow trumpets at new-moon, what day\\
our feast appointed is:\\
4 For charge to Isr’el, and a law\\
of Jacob’s God was this.

5 To Joseph this a testimony\\
he made, when Egypt land\\
He traveled through, where speech I heard\\
I did not understand.

6 His shoulder I from burdens took,\\
his hands from pots did free.\\
7 Thou didst in trouble on me call,\\
and I delivered thee:

In secret place of thundering\\
I did thee answer make;\\
And at the streams of Meribah\\
of thee a proof did take.

8 O thou, my people, give an ear,\\
I’ll testify to thee;\\
To thee, O Isr’el, if thou wilt\\
but hearken unto me.

9 In midst of thee there shall not be\\
any strange god at all;\\
Nor unto any god unknown\\
thou bowing down shalt fall.

10 I am the Lord thy God, which did\\
from Egypt land thee guide;\\
I’ll fill thy mouth abundantly,\\
do thou it open wide.

11 But yet my people to my voice\\
would not attentive be;\\
And ev’n my chosen Israel\\
he would have none of me.

12 So to the lust of their own hearts\\
I them delivered;\\
And then in counsels of their own\\
they vainly wandered.

13 O that my people had me heard,\\
Isr’el my ways had chose!\\
14 I had their en’mies soon subdued,\\
my hand turned on their foes.

15 The haters of the Lord to him\\
submission should have feigned;\\
But as for them, their time should have\\
for evermore remained.

16 He should have also fed them with\\
the finest of the wheat;\\
Of honey from the rock thy fill\\
I should have made thee eat.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 82}
\addcontentsline{toc}{subsection}{Psalm 82}

A Psalm of Asaph.

1 In gods’ assembly God doth stand;\\
he judgeth gods among.\\
2 How long, accepting persons vile,\\
will ye give judgment wrong?

3 Defend the poor and fatherless;\\
to poor oppressed do right.\\
4 The poor and needy ones set free;\\
rid them from ill men’s might.

5 They know not, nor will understand;\\
in darkness they walk on:\\
All the foundations of the earth\\
out of their course are gone.

6 I said that ye are gods, and are\\
sons of the Highest all:\\
7 But ye shall die like men, and as\\
one of the princes fall.

8 O God, do thou raise up thyself,\\
the earth to judgment call:\\
For thou, as thine inheritance,\\
shalt take the nations all.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 83}
\addcontentsline{toc}{subsection}{Psalm 83}

A Song \emph{or} Psalm of Asaph.

1 Keep not, O God, we thee entreat,\\
O keep not silence now:\\
Do thou not hold thy peace, O God,\\
and still no more be thou.

2 For, lo, thine enemies a noise\\
tumultuously have made;\\
And they that haters are of thee\\
have lifted up the head.

3 Against thy chosen people they\\
do crafty counsel take;\\
And they against thy hidden ones\\
do consultations make.

4 Come, let us cut them off, said they,\\
from being a nation,\\
That of the name of Isr’el may\\
no more be mention.

5 For with joint heart they plot, in league\\
against thee they combine.\\
6 The tents of Edom, Ishm’elites,\\
Moab’s and Hagar’s line;

7 Gebal, and Ammon, Amalek,\\
Philistines, those of Tyre;\\
8 And Assur join’d with them, to help\\
Lot’s children they conspire.

9 Do to them as to Midian,\\
Jabin at Kison strand;\\
10 And Sis’ra, which at En-dor fell,\\
as dung to fat the land.

11 Like Oreb and like Zeeb make\\
their noble men to fall;\\
Like Zeba and Zalmunna like,\\
make thou their princes all;

12 Who said, For our possession\\
let us God’s houses take.\\
13 My God, them like a wheel, as chaff\\
before the wind, them make.

14 As fire consumes the wood, as flame\\
doth mountains set on fire,\\
15 Chase and affright them with the storm\\
and tempest of thine ire.

16 Their faces fill with shame, O Lord,\\
that they may seek thy name.\\
17 Let them confounded be, and vexed,\\
and perish in their shame:

18 That men may know that thou, to whom\\
alone doth appertain\\
The name Jehovah, dost most high\\
o’er all the earth remain.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 84}
\addcontentsline{toc}{subsection}{Psalm 84}

To the chief Musician upon Gittith,
A Psalm for the sons of Korah.

1 How lovely is thy dwelling-place,\\
O Lord of hosts, to me!\\
The tabernacles of thy grace\\
how pleasant, Lord, they be!

2 My thirsty soul longs veh’mently,\\
yea faints, thy courts to see:\\
My very heart and flesh cry out,\\
O living God, for thee.

3 Behold, the sparrow findeth out\\
an house wherein to rest;\\
The swallow also for herself\\
hath purchased a nest;

Ev’n thine own altars,* where she safe\\
her young ones forth may bring,\\
O thou almighty Lord of hosts,\\
who art my God and King.

4 Blessed are they in thy house that dwell,\\
they ever give thee praise.\\
5 Blessed is the man whose strength thou art,\\
in whose heart are thy ways:

6 Who passing thorough Baca’s vale,\\
therein do dig up wells;\\
Also the rain that falleth down\\
the pools with water fills.

7 So they from strength unwearied go\\
still forward unto strength,\\
Until in Zion they appear\\
before the Lord at length.

8 Lord God of hosts, my prayer hear;\\
O Jacob’s God, give ear.\\
9 See God our shield, look on the face\\
of thine anointed dear.

10 For in thy courts one day excels\\
a thousand; rather in\\
My God’s house will I keep a door,\\
than dwell in tents of sin.

11 For God the Lord’s a sun and shield:\\
he’ll grace and glory give;\\
And will withhold no good from them\\
that uprightly do live.

12 O thou that art the Lord of hosts,\\
that man is truly blest,\\
Who by assured confidence\\
on thee alone doth rest.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 85}
\addcontentsline{toc}{subsection}{Psalm 85}

To the chief Musician,
A Psalm for the sons of Korah.

1 O Lord, thou hast been favorable\\
to thy beloved land:\\
Jacob’s captivity thou hast\\
recalled with mighty hand.

2 Thou pardoned thy people hast\\
all their iniquities;\\
Thou all their trespasses and sins\\
hast covered from thine eyes.

3 Thou took’st off all thine ire, and turn’dst\\
from thy wrath’s furiousness.\\
4 Turn us, God of our health, and cause\\
thy wrath ‘gainst us to cease.

5 Shall thy displeasure thus endure\\
against us without end?\\
Wilt thou to generations all\\
thine anger forth extend?

6 That in thee may thy people joy,\\
wilt thou not us revive?\\
7 Show us thy mercy, Lord, to us\\
do thy salvation give.

8 I’ll hear what God the Lord will speak:\\
to his folk he’ll speak peace,\\
And to his saints; but let them not\\
return to foolishness.

9 To them that fear him surely near\\
is his salvation;\\
That glory in our land may have\\
her habitation.

10 Truth met with mercy, righteousness\\
and peace kissed mutually:\\
11 Truth springs from earth, and righteousness\\
looks down from heaven high.

12 Yea, what is good the Lord shall give;\\
our land shall yield increase:\\
13 Justice, to set us in his steps,\\
shall go before his face.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 86}
\addcontentsline{toc}{subsection}{Psalm 86}

A Prayer of David.

1 O Lord, do thou bow down thine ear,\\
and hear me graciously;\\
Because I sore afflicted am,\\
and am in poverty.

2 Because I’m holy, let my soul\\
by thee preserved be:\\
O thou my God, thy servant save,\\
that puts his trust in thee.

3 Sith unto thee I daily cry,\\
be merciful to me.\\
4 Rejoice thy servant’s soul; for, Lord,\\
I lift my soul to thee.

5 For thou art gracious, O Lord,\\
and ready to forgive;\\
And rich in mercy, all that call\\
upon thee to relieve.

6 Hear, Lord, my pray’r; unto the voice\\
of my request attend:\\
7 In troublous times I’ll call on thee;\\
for thou wilt answer send.

8 Lord, there is none among the gods\\
that may with thee compare;\\
And like the works which thou hast done,\\
not any work is there.

9 All nations whom thou mad’st shall come\\
and worship rev’rently\\
Before thy face; and they, O Lord,\\
thy name shall glorify.

10 Because thou art exceeding great,\\
and works by thee are done\\
Which are to be admired; and thou\\
art God thyself alone.

11 Teach me thy way, and in thy truth,\\
O Lord, then walk will I;\\
Unite my heart, that I thy name\\
may fear continually.

12 O Lord my God, with all my heart\\
to thee I will give praise;\\
And I the glory will ascribe\\
unto thy name always:

13 Because thy mercy toward me\\
in greatness doth excel;\\
And thou delivered hast my soul\\
out from the lowest hell.

14 O God, the proud against me rise,\\
and vi’lent men have met,\\
That for my soul have sought; and thee\\
before them have not set.

15 But thou art full of pity, Lord,\\
a God most gracious,\\
Long-suffering, and in thy truth\\
and mercy plenteous.

16 O turn to me thy countenance,\\
and mercy on me have;\\
Thy servant strengthen, and the son\\
of thine own handmaid save.

17 Show me a sign for good, that they\\
which do me hate may see,\\
And be ashamed; because thou, Lord,\\
didst help and comfort me.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 87}
\addcontentsline{toc}{subsection}{Psalm 87}

A Psalm \emph{or} Song for the sons of Korah.

1 Upon the hills of holiness\\
he his foundation sets.\\
2 God, more than Jacob’s dwellings all,\\
delights in Zion’s gates.

3 Things glorious are said of thee,\\
thou city of the Lord.\\
4 Rahab and Babel I, to those\\
that know me, will record:

Behold ev’n Tyrus, and with it\\
the land of Palestine,\\
And likewise Ethiopia;\\
this man was born therein.

5 And it of Zion shall be said,\\
This man and that man there\\
Was born; and he that is most High\\
himself shall stablish her.

6 When God the people writes, he’ll count\\
that this man born was there.\\
7 There be that sing and play; and all\\
my well-springs in thee are.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 88}
\addcontentsline{toc}{subsection}{Psalm 88}

A Song or Psalm for the sons of Korah, to the chief Musician upon Mahalath Leannoth, Maschil of Heman the Ezrahite.

1 Lord God, my Savior, day and night\\
before thee cried have I.\\
2 Before thee let my prayer come;\\
give ear unto my cry.

3 For troubles great do fill my soul;\\
my life draws nigh the grave.\\
4 I’m counted with those that go down\\
to pit, and no strength have.

5 Ev’n free among the dead, like them\\
that slain in grave do lie;\\
Cut off from thy hand, whom no more\\
thou hast in memory.

6 Thou hast me laid in lowest pit,\\
in deeps and darksome caves.\\
7 Thy wrath lies hard on me, thou hast\\
me pressed with all thy waves.

8 Thou hast put far from me my friends,\\
thou mad’st them to abhor me;\\
And I am so shut up, that I\\
find no evasion for me.

9 By reason of affliction\\
mine eye mourns dolefully:\\
To thee, Lord, do I call, and stretch\\
my hands continually.

10 Wilt thou show wonders to the dead?\\
shall they rise, and thee bless?\\
11 Shall in the grave thy love be told?\\
in death thy faithfulness?

12 Shall thy great wonders in the dark,\\
or shall thy righteousness\\
Be known to any in the land\\
of deep forgetfulness?

13 But, Lord, to thee I cried; my pray’r\\
at morn prevent shall thee.\\
14 Why, Lord, dost thou cast off my soul,\\
and hid’st thy face from me?

15 Distressed am I, and from my youth\\
I ready am to die;\\
Thy terrors I have borne, and am\\
distracted fearfully.

16 The dreadful fierceness of thy wrath\\
quite over me doth go:\\
Thy terrors great have cut me off,\\
they did pursue me so.

17 For round about me ev’ry day,\\
like water, they did roll;\\
And, gathering together, they\\
have compassed my soul.

18 My friends thou hast put far from me,\\
and him that did me love;\\
And those that mine acquaintance were\\
to darkness didst remove.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 89}
\addcontentsline{toc}{subsection}{Psalm 89}

Maschil of Ethan the Ezrahite.

1 God’s mercies I will ever sing;\\
and with my mouth I shall\\
Thy faithfulness make to be known\\
to generations all.

2 For mercy shall be built, said I,\\
for ever to endure;\\
Thy faithfulness, ev’n in the heav’ns,\\
thou wilt establish sure.

3 I with my chosen One have made\\
a cov’nant graciously;\\
And to my servant, whom I loved,\\
to David sworn have I;

4 That I thy seed establish shall\\
for ever to remain,\\
And will to generations all\\
thy throne build and maintain.

5 The praises of thy wonders, Lord,\\
the heavens shall express;\\
And in the congregation\\
of saints thy faithfulness.

6 For who in heaven with the Lord\\
may once himself compare?\\
Who is like God among the sons\\
of those that mighty are?

7 Great fear in meeting of the saints\\
is due unto the Lord;\\
And he of all about him should\\
with rev’rence be adored.

8 O thou that art the Lord of hosts,\\
what Lord in mightiness\\
Is like to thee? who compassed round\\
art with thy faithfulness.

9 Ev’n in the raging of the sea\\
thou over it dost reign;\\
And when the waves thereof do swell,\\
thou stillest them again.

10 Rahab in pieces thou didst break,\\
like one that slaughtered is;\\
And with thy mighty arm thou hast\\
dispersed thine enemies.

11 The heav’ns are thine, thou for thine own\\
the earth dost also take;\\
The world, and fulness of the same,\\
thy pow’r did found and make.

12 The north and south from thee alone\\
their first beginning had;\\
Both Tabor mount and Hermon hill\\
shall in thy name be glad.

13 Thou hast an arm that’s full of pow’r,\\
thy hand is great in might;\\
And thy right hand exceedingly\\
exalted is in height.

14 Justice and judgment of thy throne\\
are made the dwelling-place;\\
Mercy, accompanied with truth,\\
shall go before thy face.

15 O greatly blessed the people are\\
the joyful sound that know;\\
In brightness of thy face, O Lord,\\
they ever on shall go.

16 They in thy name shall all the day\\
rejoice exceedingly;\\
And in thy righteousness shall they\\
exalted be on high.

17 Because the glory of their strength\\
doth only stand in thee;\\
And in thy favor shall our horn\\
and pow’r exalted be.

18 For God is our defense; and he\\
to us doth safety bring:\\
The Holy One of Israel\\
is our almighty King.

19 In vision to thy Holy One\\
thou saidst, I help upon\\
A strong one laid; out of the folk\\
I raised a chosen one;

20 Ev’n David, I have found him out\\
a servant unto me;\\
And with my holy oil my King\\
anointed him to be.

21 With whom my hand shall stablished be;\\
mine arm shall make him strong.\\
22 On him the foe shall not exact,\\
nor son of mischief wrong.

23 I will beat down before his face\\
all his malicious foes;\\
I will them greatly plague who do\\
with hatred him oppose.

24 My mercy and my faithfulness\\
with him yet still shall be;\\
And in my name his horn and pow’r\\
men shall exalted see.

25 His hand and pow’r shall reach afar;\\
I’ll set it in the sea;\\
And his right hand established\\
shall in the rivers be.

26 Thou art my Father, he shall cry,\\
thou art my God alone;\\
And he shall say, Thou art the Rock\\
of my salvation.

27 I’ll make him my first-born, more high\\
than kings of any land.\\
28 My love I’ll ever keep for him,\\
my cov’nant fast shall stand.

29 His seed I by my pow’r will make\\
for ever to endure;\\
And, as the days of heav’n, his throne\\
shall stable be, and sure.

30 But if his children shall forsake\\
my laws, and go astray,\\
And in my judgments shall not walk,\\
but wander from my way:

31 If they my laws break, and do not\\
keep my commandements;\\
32 I’ll visit then their faults with rods,\\
their sins with chastisements.

33 Yet I’ll not take my love from him,\\
nor false my promise make.\\
34 My cov’nant I’ll not break, nor change\\
what with my mouth I spake.

35 Once by my holiness I sware,\\
to David I’ll not lie;\\
36 His seed and throne shall, as the sun,\\
before me last for aye.

37 It, like the moon, shall ever be\\
established stedfastly;\\
And like to that which in the heav’n\\
doth witness faithfully.

38 But thou, displeased, hast cast off,\\
thou didst abhor and loathe;\\
With him that thine anointed is\\
thou hast been very wroth.

39 Thou hast thy servant’s covenant\\
made void, and quite cast by;\\
Thou hast profaned his crown, while it\\
cast on the ground doth lie.

40 Thou all his hedges hast broke down,\\
his strong holds down hast torn.\\
41 He to all passers-by a spoil,\\
to neighbors is a scorn.

42 Thou hast set up his foes’ right hand;\\
mad’st all his en’mies glad:\\
43 Turned his sword’s edge, and him to stand\\
in battle hast not made.

44 His glory thou hast made to cease,\\
his throne to ground down cast;\\
45 Shortened his days of youth, and him\\
with shame thou covered hast.

46 How long, Lord, wilt thou hide thyself?\\
for ever, in thine ire?\\
And shall thine indignation\\
burn like unto a fire?

47 Remember, Lord, how short a time\\
I shall on earth remain:\\
O wherefore is it so that thou\\
has made all men in vain?

48 What man is he that liveth here,\\
and death shall never see?\\
Or from the power of the grave\\
what man his soul shall free?

49 Thy former loving-kindnesses,\\
O Lord, where be they now?\\
Those which in truth and faithfulness\\
to David sworn hast thou?

50 Mind, Lord, thy servant’s sad reproach;\\
how I in bosom bear\\
The scornings of the people all,\\
who strong and mighty are.

51 Wherewith thy raging enemies\\
reproached, O Lord, think on;\\
Wherewith they have reproached the steps\\
of thine anointed one.

52 All blessing to the Lord our God\\
let be ascribed then:\\
For evermore so let it be.\\
Amen, yea, and amen.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

