\subsection*{Psalm 90}
\addcontentsline{toc}{subsection}{Psalm 90}

A Prayer of Moses the man of God.

1 Lord, thou hast been our dwelling-place\\
in generations all.\\
2 Before thou ever hadst brought forth\\
the mountains great or small;

Ere ever thou hadst formed the earth,\\
and all the world abroad;\\
Ev’n thou from everlasting art\\
to everlasting God.

3 Thou dost unto destruction\\
man that is mortal turn;\\
And unto them thou say’st, Again,\\
ye sons of men, return.

4 Because a thousand years appear\\
no more before thy sight\\
Than yesterday, when it is past,\\
or than a watch by night.

5 As with an overflowing flood\\
thou carry’st them away:\\
They like a sleep are, like the grass\\
that grows at morn are they.

6 At morn it flourishes and grows,\\
cut down at ev’n doth fade.\\
7 For by thine anger we’re consumed,\\
thy wrath makes us afraid.

8 Our sins thou and iniquities\\
dost in thy presence place,\\
And sett’st our secret faults before\\
the brightness of thy face.

9 For in thine anger all our days\\
do pass on to an end;\\
And as a tale that hath been told,\\
so we our years do spend.

10 Threescore and ten years do sum up\\
our days and years, we see;\\
Or, if, by reason of more strength,\\
in some fourscore they be:

Yet doth the strength of such old men\\
but grief and labor prove;\\
For it is soon cut off, and we\\
fly hence, and soon remove.

11 Who knows the power of thy wrath?\\
according to thy fear\\
12 So is thy wrath: Lord, teach thou us\\
our end in mind to bear;

And so to count our days, that we\\
our hearts may still apply\\
To learn thy wisdom and thy truth,\\
that we may live thereby.

13 Turn yet again to us, O Lord,\\
how long thus shall it be?\\
Let it repent thee now for those\\
that servants are to thee.

14 O with thy tender mercies, Lord,\\
us early satisfy;\\
So we rejoice shall all our days,\\
and still be glad in thee.

15 According as the days have been,\\
wherein we grief have had,\\
And years wherein we ill have seen,\\
so do thou make us glad.

16 O let thy work and pow’r appear\\
thy servants’ face before;\\
And show unto their children dear\\
thy glory evermore:

17 And let the beauty of the Lord\\
our God be us upon:\\
Our handy-works establish thou,\\
establish them each one.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 91}
\addcontentsline{toc}{subsection}{Psalm 91}

1 He that doth in the secret place\\
of the most High reside,\\
Under the shade of him that is\\
th’ Almighty shall abide.

2 I of the Lord my God will say,\\
He is my refuge still,\\
He is my fortress, and my God,\\
and in him trust I will.

3 Assuredly he shall thee save,\\
and give deliverance\\
From subtle fowler’s snare, and from\\
the noisome pestilence.

4 His feathers shall thee hide; thy trust\\
under his wings shall be:\\
His faithfulness shall be a shield\\
and buckler unto thee.

5 Thou shalt not need to be afraid\\
for terrors of the night;\\
Nor for the arrow that doth fly\\
by day, while it is light;

6 Nor for the pestilence, that walks\\
in darkness secretly;\\
Nor for destruction, that doth waste\\
at noon-day openly.

7 A thousand at thy side shall fall,\\
on thy right hand shall lie\\
Ten thousand dead; yet unto thee\\
it shall not once come nigh.

8 Only thou with thine eyes shalt look,\\
and a beholder be;\\
And thou therein the just reward\\
of wicked men shalt see.

9 Because the Lord, who constantly\\
my refuge is alone,\\
Ev’n the most High, is made by thee\\
thy habitation;

10 No plague shall near thy dwelling come;\\
no ill shall thee befall:\\
11 For thee to keep in all thy ways\\
his angels charge he shall.

12 They in their hands shall bear thee up,\\
still waiting thee upon;\\
Lest thou at any time should’st dash\\
thy foot against a stone.

13 Upon the adder thou shalt tread,\\
and on the lion strong;\\
Thy feet on dragons trample shall,\\
and on the lions young.

14 Because on me he set his love,\\
I’ll save and set him free;\\
Because my great name he hath known,\\
I will him set on high.

15 He’ll call on me, I’ll answer him;\\
I will be with him still\\
In trouble, to deliver him,\\
and honor him I will.

16 With length of days unto his mind\\
I will him satisfy;\\
I also my salvation\\
will cause his eyes to see.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 92}
\addcontentsline{toc}{subsection}{Psalm 92}

A Psalm \emph{or} Song for the sabbath day.

1 To render thanks unto the Lord\\
it is a comely thing,\\
And to thy name, O thou most High,\\
due praise aloud to sing.

2 Thy loving-kindness to show forth\\
when shines the morning light;\\
And to declare thy faithfulness\\
with pleasure ev’ry night.

3 On a ten-stringed instrument,\\
upon the psaltery,\\
And on the harp with solemn sound,\\
and grave sweet melody.

4 For thou, Lord, by thy mighty works\\
hast made my heart right glad;\\
And I will triumph in the works\\
which by thine hands were made.

5 How great, Lord, are thy works! each thought\\
of thine a deep it is:\\
6 A brutish man it knoweth not;\\
fools understand not this.

7 When those that lewd and wicked are\\
spring quickly up like grass,\\
And workers of iniquity\\
do flourish all apace;

It is that they for ever may\\
destroyed be and slain;\\
8 But thou, O Lord, art the most High,\\
for ever to remain.

9 For, lo, thine enemies, O Lord,\\
thine en’mies perish shall;\\
The workers of iniquity\\
shall be dispersed all.

10 But thou shalt, like unto the horn\\
of th’ unicorn, exalt\\
My horn on high: thou with fresh oil\\
anoint me also shalt.

11 Mine eyes shall also my desire\\
see on mine enemies;\\
Mine ears shall of the wicked hear\\
that do against me rise.

12 But like the palm-tree flourishing\\
shall be the righteous one;\\
He shall like to the cedar grow\\
that is in Lebanon.

13 Those that within the house of God\\
are planted by his grace,\\
They shall grow up, and flourish all\\
in our God’s holy place.

14 And in old age, when others fade,\\
they fruit still forth shall bring;\\
They shall be fat, and full of sap,\\
and aye be flourishing;

15 To show that upright is the Lord:\\
he is a rock to me;\\
And he from all unrighteousness\\
is altogether free.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 93}
\addcontentsline{toc}{subsection}{Psalm 93}

1 The Lord doth reign, and clothed is he
with majesty most bright;
His works do show him clothed to be,\\
and girt about with might.

The world is also stablished,\\
that it cannot depart.\\
2 Thy throne is fixed of old, and thou\\
from everlasting art.

3 The floods, O Lord, have lifted up,\\
they lifted up their voice;\\
The floods have lifted up their waves,\\
and made a mighty noise.

4 But yet the Lord, that is on high,\\
is more of might by far\\
Than noise of many waters is,\\
or great sea-billows are.

5 Thy testimonies ev’ry one\\
in faithfulness excel;\\
And holiness for ever, Lord,\\
thine house becometh well.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 94}
\addcontentsline{toc}{subsection}{Psalm 94}

1 O Lord God, unto whom alone\\
all vengeance doth belong;\\
O mighty God, who vengeance own’st,\\
shine forth, avenging wrong.

2 Lift up thyself, thou of the earth\\
the sov’reign Judge that art;\\
And unto those that are so proud\\
a due reward impart.

3 How long, O mighty God, shall they\\
who lewd and wicked be,\\
How long shall they who wicked are\\
thus triumph haughtily?

4 How long shall things most hard by them\\
be uttered and told?\\
And all that work iniquity\\
to boast themselves be bold?

5 Thy folk they break in pieces, Lord,\\
thine heritage oppress:\\
6 The widow they and stranger slay,\\
and kill the fatherless.

7 Yet say they, God it shall not see,\\
nor God of Jacob know.\\
8 Ye brutish people! understand;\\
fools! when wise will ye grow?

9 The Lord did plant the ear of man,\\
and hear then shall not he?\\
He only formed the eye, and then\\
shall he not clearly see?

10 He that the nations doth correct,\\
shall he not chastise you?\\
He knowledge unto man doth teach,\\
and shall himself not know?

11 Man’s thoughts to be but vanity\\
the Lord doth well discern.\\
12 Blessed is the man thou chast’nest, Lord,\\
and mak’st thy law to learn:

13 That thou may’st give him rest from days\\
of sad adversity,\\
Until the pit be digged for those\\
that work iniquity.

14 For sure the Lord will not cast off\\
those that his people be,\\
Neither his own inheritance\\
quit and forsake will he:

15 But judgment unto righteousness\\
shall yet return again;\\
And all shall follow after it\\
that are right-hearted men.

16 Who will rise up for me against\\
those that do wickedly?\\
Who will stand up for me ‘gainst those\\
that work iniquity?

17 Unless the Lord had been my help\\
when I was sore oppressed,\\
Almost my soul had in the house\\
of silence been at rest.

18 When I had uttered this word,\\
(my foot doth slip away,)\\
Thy mercy held me up, O Lord,\\
thy goodness did me stay.

19 Amidst the multitude of thoughts\\
which in my heart do fight,\\
My soul, lest it be overcharged,\\
thy comforts do delight.

20 Shall of iniquity the throne\\
have fellowship with thee,\\
Which mischief, cunningly contrived,\\
doth by a law decree?

21 Against the righteous souls they join,\\
they guiltless blood condemn.\\
22 But of my refuge God’s the rock,\\
and my defense from them.

23 On them their own iniquity\\
the Lord shall bring and lay,\\
And cut them off in their own sin;\\
our Lord God shall them slay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 95}
\addcontentsline{toc}{subsection}{Psalm 95}

1 O come, let us sing to the Lord:\\
come, let us ev’ry one\\
A joyful noise make to the Rock\\
of our salvation.

2 Let us before his presence come\\
with praise and thankful voice;\\
Let us sing psalms to him with grace,\\
and make a joyful noise.

3 For God, a great God, and great King,\\
above all gods he is.\\
4 Depths of the earth are in his hand,\\
the strength of hills is his.

5 To him the spacious sea belongs,\\
for he the same did make;\\
The dry land also from his hands\\
its form at first did take.

6 O come, and let us worship him,\\
let us bow down withal,\\
And on our knees before the Lord\\
our Maker let us fall.

7 For he’s our God, the people we\\
of his own pasture are,\\
And of his hand the sheep; to-day,\\
if ye his voice will hear,

8 Then harden not your hearts, as in\\
the provocation,\\
As in the desert, on the day\\
of the tentation:

9 When me your fathers tempted and proved,\\
and did my working see;\\
10 Ev’n for the space of forty years\\
this race hath grieved me.

I said, This people errs in heart,\\
my ways they do not know:\\
11 To whom I swore in wrath, that to\\
my rest they should not go.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 96}
\addcontentsline{toc}{subsection}{Psalm 96}

1 O sing a new song to the Lord:\\
sing all the earth to God.\\
2 To God sing, bless his name, show still\\
his saving health abroad.

3 Among the heathen nations\\
his glory do declare;\\
And unto all the people show\\
his works that wondrous are.

4 For great’s the Lord, and greatly he\\
is to be magnified;\\
Yea, worthy to be feared is he\\
above all gods beside.

5 For all the gods are idols dumb,\\
which blinded nations fear;\\
But our God is the Lord, by whom\\
the heav’ns created were.

6 Great honor is before his face,\\
and majesty divine;\\
Strength is within his holy place,\\
and there doth beauty shine.

7 Do ye ascribe unto the Lord,\\
of people ev’ry tribe,\\
Glory do ye unto the Lord,\\
and mighty pow’r ascribe.

8 Give ye the glory to the Lord\\
that to his name is due;\\
Come ye into his courts, and bring\\
an offering with you.

9 In beauty of his holiness,\\
O do the Lord adore;\\
Likewise let all the earth throughout\\
tremble his face before.

10 Among the heathen say, God reigns;\\
the world shall steadfastly\\
Be fixed from moving; he shall judge\\
the people righteously.

11 Let heav’ns be glad before the Lord,\\
and let the earth rejoice;\\
Let seas, and all that is therein,\\
cry out, and make a noise.

12 Let fields rejoice, and ev’ry thing\\
that springeth of the earth:\\
Then woods and ev’ry tree shall sing\\
with gladness and with mirth

13 Before the Lord; because he comes,\\
to judge the earth comes he:\\
He’ll judge the world with righteousness,\\
the people faithfully.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 97}
\addcontentsline{toc}{subsection}{Psalm 97}

1 God reigneth, let the earth be glad,\\
and isles rejoice each one.\\
2 Dark clouds him compass; and in right\\
with judgment dwells his throne.

3 Fire goes before him, and his foes\\
it burns up round about:\\
4 His lightnings lighten did the world;\\
earth saw, and shook throughout.

5 Hills at the presence of the Lord,\\
like wax, did melt away;\\
Ev’n at the presence of the Lord\\
of all the earth, I say.

6 The heav’ns declare his righteousness,\\
all men his glory see.\\
7 All who serve graven images,\\
confounded let them be.

Who do of idols boast themselves,\\
let shame upon them fall:\\
Ye that are called gods, see that\\
ye do him worship all.

8 Zion did hear, and joyful was,\\
glad Judah’s daughters were;\\
They much rejoiced, O Lord, because\\
thy judgments did appear.

9 For thou, O Lord, art high above\\
all things on earth that are;\\
Above all other gods thou art\\
exalted very far.

10 Hate ill, all ye that love the Lord:\\
his saints’ souls keepeth he;\\
And from the hands of wicked men\\
he sets them safe and free.

11 For all those that be righteous\\
sown is a joyful light,\\
And gladness sown is for all those\\
that are in heart upright.

12 Ye righteous, in the Lord rejoice;\\
express your thankfulness,\\
When ye into your memory\\
do call his holiness.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 98}
\addcontentsline{toc}{subsection}{Psalm 98}

1 O sing a new song to the Lord,\\
for wonders he hath done:\\
His right hand and his holy arm\\
him victory hath won.

2 The Lord God his salvation\\
hath caused to be known;\\
His justice in the heathen’s sight\\
he openly hath shown.

3 He mindful of his grace and truth\\
to Isr’el’s house hath been;\\
And the salvation of our God\\
all ends of th’ earth have seen.

4 Let all the earth unto the Lord\\
send forth a joyful noise;\\
Lift up your voice aloud to him,\\
sing praises, and rejoice.

5 With harp, with harp, and voice of psalms,\\
unto Jehovah sing:\\
6 With trumpets, cornets, gladly sound\\
before the Lord the King.

7 Let seas and all their fullness roar;\\
the world, and dwellers there;\\
8 Let floods clap hands, and let the hills\\
together joy declare

9 Before the Lord; because he comes,\\
to judge the earth comes he:\\
He’ll judge the world with righteousness,\\
his folk with equity.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 99}
\addcontentsline{toc}{subsection}{Psalm 99}

1 Th’ eternal Lord doth reign as king,\\
let all the people quake;\\
He sits between the cherubims,\\
let th’ earth be moved and shake.

2 The Lord in Zion great and high\\
above all people is;\\
3 Thy great and dreadful name (for it\\
is holy) let them bless.

4 The king’s strength also judgment loves;\\
thou settlest equity:\\
Just judgment thou dost execute\\
in Jacob righteously.

5 The Lord our God exalt on high,\\
and rev’rently do ye\\
Before his footstool worship him:\\
the Holy One is he.

6 Moses and Aaron ‘mong his priests,\\
Samuel, with them that call\\
Upon his name: these called on God,\\
and he them answered all.

7 Within the pillar of the cloud\\
he unto them did speak:\\
The testimonies he them taught,\\
and laws, they did not break.

8 Thou answer’dst them, O Lord our God;\\
thou wast a God that gave\\
Pardon to them, though on their deeds\\
thou wouldest vengeance have.

9 Do ye exalt the Lord our God,\\
and at his holy hill\\
Do ye him worship: for the Lord\\
our God is holy still.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 100}
\addcontentsline{toc}{subsection}{Psalm 100}

\emph{First Version---Long Meter (L.M., 8.8.8.8)}

A Psalm of praise.

1 All people that on earth do dwell,\\
Sing to the Lord with cheerful voice.\\
2 Him serve with mirth, his praise forth tell,\\
Come ye before him and rejoice.

3 Know that the Lord is God indeed;\\
Without our aid he did us make:\\
We are his flock, he doth us feed,\\
And for his sheep he doth us take.

4 O enter then his gates with praise,\\
Approach with joy his courts unto:\\
Praise, laud, and bless his name always,\\
For it is seemly so to do.

5 For why? the Lord our God is good,\\
His mercy is for ever sure;\\
His truth at all times firmly stood,\\
And shall from age to age endure.

\emph{Second Version---Common Meter (C.M., 8.6.8.6)}

A Psalm of praise.

1 O all ye lands, unto the Lord\\
make ye a joyful noise.\\
2 Serve God with gladness, him before\\
come with a singing voice.

3 Know ye the Lord that he is God;\\
not we, but he us made:\\
We are his people, and the sheep\\
within his pasture fed.

4 Enter his gates and courts with praise,\\
to thank him go ye thither:\\
To him express your thankfulness,\\
and bless his name together.

5 Because the Lord our God is good,\\
his mercy faileth never;\\
And to all generations\\
his truth endureth ever.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 101}
\addcontentsline{toc}{subsection}{Psalm 101}

1 I mercy will and judgment sing,\\
Lord, I will sing to thee.\\
2 With wisdom in a perfect way\\
shall my behavior be.

O when, in kindness unto me,\\
wilt thou be pleased to come?\\
I with a perfect heart will walk\\
within my house at home.

3 I will endure no wicked thing\\
before mine eyes to be:\\
I hate their work that turn aside,\\
it shall not cleave to me.

4 A stubborn and a froward heart\\
depart quite from me shall;\\
A person giv’n to wickedness\\
I will not know at all.

5 I’ll cut him off that slandereth\\
his neighbor privily:\\
The haughty heart I will not bear,\\
nor him that looketh high.

6 Upon the faithful of the land\\
mine eyes shall be, that they\\
May dwell with me: he shall me serve\\
that walks in perfect way.

7 Who of deceit a worker is\\
in my house shall not dwell;\\
And in my presence shall he not\\
remain that lies doth tell.

8 Yea, all the wicked of the land\\
early destroy will I;\\
All from God’s city to cut off\\
that work iniquity.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 102}
\addcontentsline{toc}{subsection}{Psalm 102}

\emph{First Version---Common Meter (C.M. 8.6.8.6)}

A Prayer of the afflicted, when he is overwhelmed,
and poureth out his complaint before the Lord.

1 O Lord, unto my pray’r give ear,\\
my cry let come to thee;\\
2 And in the day of my distress\\
hide not thy face from me.

Give ear to me; what time I call,\\
to answer me make haste:\\
3 For, as an hearth, my bones are burnt,\\
my days, like smoke, do waste.

4 My heart within me smitten is,\\
and it is withered\\
Like very grass; so that I do\\
forget to eat my bread.

5 By reason of my groaning voice\\
my bones cleave to my skin.\\
6 Like pelican in wilderness\\
forsaken I have been:

I like an owl in desert am,\\
that nightly there doth moan;\\
7 I watch, and like a sparrow am\\
on the house-top alone.

8 My bitter en’mies all the day\\
reproaches cast on me;\\
And, being mad at me, with rage\\
against me sworn they be.

9 For why? I ashes eaten have\\
like bread, in sorrows deep;\\
My drink I also mingled have\\
with tears that I did weep.

10 Thy wrath and indignation\\
did cause this grief and pain;\\
For thou hast lift me up on high,\\
and cast me down again.

11 My days are like unto a shade,\\
which doth declining pass;\\
And I am dried and withered,\\
ev’n like unto the grass.

12 But thou, Lord, everlasting art,\\
and thy remembrance shall\\
Continually endure, and be\\
to generations all.

13 Thou shalt arise, and mercy have\\
upon thy Zion yet;\\
The time to favor her is come,\\
the time that thou hast set.

14 For in her rubbish and her stones\\
thy servants pleasure take;\\
Yea, they the very dust thereof\\
do favor for her sake.

15 So shall the heathen people fear\\
the Lord’s most holy name;\\
And all the kings on earth shall dread\\
thy glory and thy fame.

16 When Zion by the mighty Lord\\
built up again shall be,\\
In glory then and majesty\\
to men appear shall he.

17 The prayer of the destitute\\
he surely will regard;\\
Their prayer will he not despise,\\
by him it shall be heard.

18 For generations yet to come\\
this shall be on record:\\
So shall the people that shall be\\
created praise the Lord.

19 He from his sanctuary’s height\\
hath downward cast his eye;\\
And from his glorious throne in heav’n\\
the Lord the earth did spy;

20 That of the mournful prisoner\\
the groanings he might hear,\\
To set them free that unto death\\
by men appointed are:

21 That they in Zion may declare\\
the Lord’s most holy name,\\
And publish in Jerusalem\\
the praises of the same;

22 When as the people gather shall\\
in troops with one accord,\\
When kingdoms shall assembled be\\
to serve the highest Lord.

23 My wonted strength and force he hath\\
abated in the way,\\
And he my days hath shortened:\\
24 Thus therefore did I say,

My God, in mid-time of my days\\
take thou me not away:\\
From age to age eternally\\
thy years endure and stay.

25 The firm foundation of the earth\\
of old time thou hast laid;\\
The heavens also are the work\\
which thine own hands have made.

26 Thou shalt for evermore endure,\\
but they shall perish all;\\
Yea, ev’ry one of them wax old,\\
like to a garment, shall:

Thou, as a vesture, shalt them change,\\
and they shall changed be:\\
27 But thou the same art, and thy years\\
are to eternity.

28 The children of thy servants shall\\
continually endure;\\
And in thy sight, O Lord, their seed\\
shall be established sure.

\emph{Second Version---Long Meter (L.M. 8.8.8.8)}

A Prayer of the afflicted, when he is overwhelmed,
and poureth out his complaint before the Lord.

1 Lord, hear my pray’r, and let my cry\\
Have speedy access unto thee;\\
2 In day of my calamity\\
O hide not thou thy face from me.

Hear when I call to thee; that day\\
An answer speedily return:\\
3 My days, like smoke, consume away,\\
And, as an hearth, my bones do burn.

4 My heart is wounded very sore,\\
And withered, like grass doth fade:\\
I am forgetful grown therefore\\
To take and eat my daily bread.

5 By reason of my smart within,\\
And voice of my most grievous groans,\\
My flesh consumed is, my skin,\\
All parched, doth cleave unto my bones.

6 The pelican of wilderness,\\
The owl in desert, I do match;\\
7 And, sparrow-like, companionless,\\
Upon the house’s top, I watch.

8 I all day long am made a scorn,\\
Reproached by my malicious foes:\\
The madmen are against me sworn,\\
The men against me that arose.

9 For I have ashes eaten up,\\
To me as if they had been bread;\\
And with my drink I in my cup\\
Of bitter tears a mixture made.

10 Because thy wrath was not appeased,\\
And dreadful indignation:\\
Therefore it was that thou me raised,\\
And thou again didst cast me down.

11 My days are like a shade alway,\\
Which doth declining swiftly pass;\\
And I am withered away,\\
Much like unto the fading grass.

12 But thou, O Lord, shalt still endure,\\
From change and all mutation free,\\
And to all generations sure\\
Shall thy remembrance ever be.

13 Thou shalt arise, and mercy yet\\
Thou to mount Zion shalt extend:\\
Her time for favor which was set,\\
Behold, is now come to an end.

14 Thy saints take pleasure in her stones,\\
Her very dust to them is dear.\\
15 All heathen lands and kingly thrones\\
On earth thy glorious name shall fear.

16 God in his glory shall appear,\\
When Zion he builds and repairs.\\
17 He shall regard and lend his ear\\
Unto the needy’s humble pray’rs:

Th’ afflicted’s pray’r he will not scorn.\\
18 All times this shall be on record:\\
And generations yet unborn\\
Shall praise and magnify the Lord.

19 He from his holy place looked down,\\
The earth he viewed from heav’n on high;\\
20 To hear the pris’ner’s mourning groan,\\
And free them that are doomed to die;

21 That Zion, and Jerus’lem too,\\
His name and praise may well record,\\
22 When people and the kingdoms do\\
Assemble all to praise the Lord.

23 My strength he weakened in the way,\\
My days of life he shortened.\\
24 My God, O take me not away\\
In mid-time of my days, I said:

Thy years throughout all ages last.\\
25 Of old thou hast established\\
The earth’s foundation firm and fast:\\
Thy mighty hands the heav’ns have made.

26 They perish shall, as garments do,\\
But thou shalt evermore endure;\\
As vestures, thou shalt change them so;\\
And they shall all be changed sure:

27 But from all changes thou art free;\\
Thy endless years do last for aye.\\
28 Thy servants, and their seed who be,\\
Established shall before thee stay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 103}
\addcontentsline{toc}{subsection}{Psalm 103}

\emph{A Psalm} of David.


1 O thou my soul, bless God the Lord;\\
and all that in me is\\
Be stirred up his holy name\\
to magnify and bless.

2 Bless, O my soul, the Lord thy God,\\
and not forgetful be\\
Of all his gracious benefits\\
he hath bestowed on thee.

3 All thine iniquities who doth\\
most graciously forgive:\\
Who thy diseases all and pains\\
doth heal, and thee relieve.

4 Who doth redeem thy life, that thou\\
to death may’st not go down;\\
Who thee with loving-kindness doth\\
and tender mercies crown:

5 Who with abundance of good things\\
doth satisfy thy mouth;\\
So that, ev’n as the eagle’s age,\\
renewed is thy youth.

6 God righteous judgment executes\\
for all oppressed ones.\\
7 His ways to Moses, he his acts\\
made known to Isr’el’s sons.

8 The Lord our God is merciful,\\
and he is gracious,\\
Long-suffering, and slow to wrath,\\
in mercy plenteous.

9 He will not chide continually,\\
nor keep his anger still.\\
10 With us he dealt not as we sinned,\\
nor did requite our ill.

11 For as the heaven in its height\\
the earth surmounteth far;\\
So great to those that do him fear\\
his tender mercies are:

12 As far as east is distant from\\
the west, so far hath he\\
From us removed, in his love,\\
all our iniquity.

13 Such pity as a father hath\\
unto his children dear;\\
Like pity shews the Lord to such\\
as worship him in fear.

14 For he remembers we are dust,\\
and he our frame well knows.\\
15 Frail man, his days are like the grass,\\
as flow’r in field he grows:

16 For over it the wind doth pass,\\
and it away is gone;\\
And of the place where once it was\\
it shall no more be known.

17 But unto them that do him fear\\
God’s mercy never ends;\\
And to their children’s children still\\
his righteousness extends:

18 To such as keep his covenant,\\
and mindful are alway\\
Of his most just commandements,\\
that they may them obey.

19 The Lord prepared hath his throne\\
in heavens firm to stand;\\
And ev’ry thing that being hath\\
his kingdom doth command.

20 O ye his angels, that excel\\
in strength, bless ye the Lord;\\
Ye who obey what he commands,\\
and hearken to his word.

21 O bless and magnify the Lord,\\
ye glorious hosts of his;\\
Ye ministers, that do fulfil\\
whate’er his pleasure is.

22 O bless the Lord, all ye his works,\\
wherewith the world is stored\\
In his dominions ev’ry where.\\
My soul, bless thou the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 104}
\addcontentsline{toc}{subsection}{Psalm 104}

1 Bless God, my soul. O Lord my God,\\
thou art exceeding great;\\
With honour and with majesty\\
thou clothed art in state.

2 With light, as with a robe, thyself\\
thou coverest about;\\
And, like unto a curtain, thou\\
the heavens stretchest out.

3 Who of his chambers doth the beams\\
within the waters lay;\\
Who doth the clouds his chariot make,\\
on wings of wind make way.

4 Who flaming fire his ministers,\\
his angels sp’rits, doth make:\\
5 Who earth’s foundations did lay,\\
that it should never shake.

6 Thou didst it cover with the deep,\\
as with a garment spread:\\
The waters stood above the hills,\\
when thou the word but said.

7 But at the voice of thy rebuke\\
they fled, and would not stay;\\
They at thy thunder’s dreadful voice\\
did haste them fast away.

8 They by the mountains do ascend,\\
and by the valley-ground\\
Descend, unto that very place\\
which thou for them didst found.

9 Thou hast a bound unto them set,\\
that they may not pass over,\\
That they do not return again\\
the face of earth to cover.

10 He to the valleys sends the springs,\\
which run among the hills:\\
11 They to all beasts of field give drink,\\
wild asses drink their fills.

12 By them the fowls of heav’n shall have\\
their habitation,\\
Which do among the branches sing\\
with delectation.

13 He from his chambers watereth\\
the hills, when they are dry’d:\\
With fruit and increase of thy works\\
the earth is satisfied.

14 For cattle he makes grass to grow,\\
he makes the herb to spring\\
For th’ use of man, that food to him\\
he from the earth may bring;

15 And wine, that to the heart of man\\
doth cheerfulness impart,\\
Oil that his face makes shine, and bread\\
that strengtheneth his heart.

16 The trees of God are full of sap;\\
the cedars that do stand\\
In Lebanon, which planted were\\
by his almighty hand.

17 Birds of the air upon their boughs\\
do choose their nests to make;\\
As for the stork, the fir-tree she\\
doth for her dwelling take.

18 The lofty mountains for wild goats\\
a place of refuge be;\\
The conies also to the rocks\\
do for their safety flee.

19 He sets the moon in heav’n, thereby\\
the seasons to discern:\\
From him the sun his certain time\\
of going down doth learn.

20 Thou darkness mak’st, ’tis night, then beasts\\
of forests creep abroad.\\
21 The lions young roar for their prey,\\
and seek their meat from God.

22 The sun doth rise, and home they flock,\\
down in their dens they lie.\\
23 Man goes to work, his labour he\\
doth to the ev’ning ply.

24 How manifold, Lord, are thy works!\\
in wisdom wonderful\\
Thou ev’ry one of them hast made;\\
earth’s of thy riches full:

25 So is this great and spacious sea,\\
wherein things creeping are,\\
Which numbered cannot be; and beasts\\
both great and small are there.

26 There ships go; there thou mak’st to play\\
that leviathan great.\\
27 These all wait on thee, that thou may’st\\
in due time give them meat.

28 That which thou givest unto them\\
they gather for their food;\\
Thine hand thou open’st lib’rally,\\
they fill-ed are with good.

29 Thou hid’st thy face; they troubled are,\\
their breath thou tak’st away;\\
Then do they die, and to their dust\\
return again do they.

30 Thy quick’ning spirit thou send’st forth,\\
then they created be;\\
And then the earth’s decayed face\\
renewed is by thee.

31 The glory of the mighty Lord\\
continue shall for ever:\\
The Lord Jehovah shall rejoice\\
in all his works together.

32 Earth, as affrighted, trembleth all,\\
if he on it but look;\\
And if the mountains he but touch,\\
they presently do smoke.

33 I will sing to the Lord most high,\\
so long as I shall live;\\
And while I being have I shall\\
to my God praises give.

34 Of him my meditation shall\\
sweet thoughts to me afford;\\
And as for me, I will rejoice\\
in God, my only Lord.

35 From earth let sinners be consumed,\\
let ill men no more be.\\
O thou my soul, bless thou the Lord.\\
Praise to the Lord give ye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 105}
\addcontentsline{toc}{subsection}{Psalm 105}

1 Give thanks to God, call on his name;\\
to men his deeds make known.\\
2 Sing ye to him, sing psalms; proclaim\\
his wondrous works each one.

3 See that ye in his holy name\\
to glory do accord;\\
And let the heart of ev’ry one\\
rejoice that seeks the Lord.

4 The Lord Almighty, and his strength,\\
with steadfast hearts seek ye:\\
His blessed and his gracious face\\
seek ye continually.

5 Think on the works that he hath done,\\
which admiration breed;\\
His wonders, and the judgments all\\
which from his mouth proceed;

6 O ye that are of Abr’ham’s race,\\
his servant well approv’n;\\
And ye that Jacob’s children are,\\
whom he chose for his own.

7 Because he, and he only, is\\
the mighty Lord our God;\\
And his most righteous judgments are\\
in all the earth abroad.

8 His cov’nant he remembered hath,\\
that it may ever stand:\\
To thousand generations\\
the word he did command.

9 Which covenant he firmly made\\
with faithful Abraham,\\
And unto Isaac, by his oath,\\
he did renew the same:

10 And unto Jacob, for a law,\\
he made it firm and sure,\\
A covenant to Israel,\\
which ever should endure.

11 He said, I’ll give Canaan’s land\\
for heritage to you;\\
12 While they were strangers there, and few,\\
in number very few:

13 While yet they went from land to land\\
without a sure abode;\\
And while through sundry kingdoms they\\
did wander far abroad;

14 Yet, notwithstanding suffered he\\
no man to do them wrong:\\
Yea, for their sakes, he did reprove\\
kings, who were great and strong.

15 Thus did he say, Touch ye not those\\
that mine anointed be,\\
Nor do the prophets any harm\\
that do pertain to me.

16 He called for famine on the land,\\
he brake the staff of bread:\\
17 But yet he sent a man before,\\
by whom they should be fed;

Ev’n Joseph, whom unnat’rally\\
sell for a slave did they;\\
18 Whose feet with fetters they did hurt,\\
and he in irons lay;

19 Until the time that his word came\\
to give him liberty;\\
The word and purpose of the Lord\\
did him in prison try.

20 Then sent the king, and did command\\
that he enlarged should be:\\
He that the people’s ruler was\\
did send to set him free.

21 A lord to rule his family\\
he raised him, as most fit;\\
To him of all that he possessed\\
he did the charge commit:

22 That he might at his pleasure bind\\
the princes of the land;\\
And he might teach his senators\\
wisdom to understand.

23 The people then of Israel\\
down into Egypt came;\\
And Jacob also sojourned\\
within the land of Ham.

24 And he did greatly by his pow’r\\
increase his people there;\\
And stronger than their enemies\\
they by his blessing were.

25 Their heart he turned to envy\\
his folk maliciously,\\
With those that his own servants were\\
to deal in subtlety.

26 His servant Moses he did send,\\
Aaron his chosen one.\\
27 By these his signs and wonders great\\
in Ham’s land were made known.

28 Darkness he sent, and made it dark;\\
his word they did obey.\\
29 He turned their waters into blood,\\
and he their fish did slay.

30 The land in plenty brought forth frogs\\
in chambers of their kings.\\
31 His word all sorts of flies and lice\\
in all their borders brings.

32 He hail for rain, and flaming fire\\
into their land he sent:\\
33 And he their vines and fig-trees smote:\\
trees of their coasts he rent.

34 He spake, and caterpillars came,\\
locusts did much abound;\\
35 Which in their land all herbs consumed,\\
and all fruits of their ground.

36 He smote all first-born in their land,\\
chief of their strength each one.\\
37 With gold and silver brought them forth,\\
weak in their tribes were none.

38 Egypt was glad when forth they went,\\
their fear on them did light.\\
39 He spread a cloud for covering,\\
and fire to shine by night.

40 They asked, and he brought quails: with bread\\
of heav’n he fill-ed them.\\
41 He opened rocks, floods gushed, and ran\\
in deserts like a stream.

42 For on his holy promise he,\\
and servant Abr’ham, thought.\\
43 With joy his people, his elect\\
with gladness, forth he brought.

44 And unto them the pleasant lands\\
he of the heathen gave;\\
That of the people’s labor they\\
inheritance might have.

45 That they his statutes might observe\\
according to his word;\\
And that they might his laws obey.\\
Give praise unto the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 106}
\addcontentsline{toc}{subsection}{Psalm 106}

1 Give praise and thanks unto the Lord,\\
for bountiful is he;\\
His tender mercy doth endure\\
unto eternity.

2 God’s mighty works who can express?\\
or shew forth all his praise?\\
3 Blessed are they that judgment keep,\\
and justly do always.

4 Remember me, Lord, with that love\\
which thou to thine dost bear;\\
With thy salvation, O my God,\\
to visit me draw near:

5 That I thy chosen’s good may see,\\
and in their joy rejoice;\\
And may with thine inheritance\\
triumph with cheerful voice.

6 We with our fathers sinned have,\\
and of iniquity\\
Too long we have the workers been;\\
we have done wickedly.

7 The wonders great, which thou, O Lord,\\
didst work in Egypt land,\\
Our fathers, though they saw, yet them\\
they did not understand:

And they thy mercies’ multitude\\
kept not in memory;\\
But at the sea, ev’n the Red sea,\\
provoked him grievously.

8 Nevertheless he saved them,\\
ev’n for his own name’s sake;\\
That so he might to be well known\\
his mighty power make.

9 When he the Red sea did rebuke,\\
then dried up it was:\\
Through depths, as through the wilderness,\\
he safely made them pass.

10 From hands of those that hated them\\
he did his people save;\\
And from the en’my’s cruel hand\\
to them redemption gave.

11 The waters overwhelmed their foes;\\
not one was left alive.\\
12 Then they believed his word, and praise\\
to him in songs did give.

13 But soon did they his mighty works\\
forget unthankfully,\\
And on his counsel and his will\\
did not wait patiently;

14 But much did lust in wilderness,\\
and God in desert tempt.\\
15 He gave them what they sought, but to\\
their soul he leanness sent.

16 And against Moses in the camp\\
their envy did appear;\\
At Aaron they, the saint of God,\\
envious also were.

17 Therefore the earth did open wide,\\
and Dathan did devour,\\
And all Abiram’s company\\
did cover in that hour.

18 Likewise among their company\\
a fire was kindled then;\\
And so the hot consuming flame\\
burnt up these wicked men.

19 Upon the hill of Horeb they\\
an idol-calf did frame,\\
A molten image they did make,\\
and worshipped the same.

20 And thus their glory, and their God,\\
most vainly changed they\\
Into the likeness of an ox\\
that eateth grass or hay.

21 They did forget the mighty God,\\
that had their savior been,\\
By whom such great things brought to pass\\
they had in Egypt seen.

22 In Ham’s land he did wondrous works,\\
things terrible did he,\\
When he his mighty hand and arm\\
stretched out at the Red sea.

23 Then said he, He would them destroy,\\
had not, his wrath to stay,\\
His chosen Moses stood in breach,\\
that them he should not slay.

24 Yea, they despised the pleasant land,\\
believed not his word:\\
25 But in their tents they murmured,\\
not heark’ning to the Lord.

26 Therefore in desert them to slay\\
he lifted up his hand:\\
27 ‘Mong nations to o’erthrow their seed,\\
and scatter in each land.

28 They unto Baal-peor did\\
themselves associate;\\
The sacrifices of the dead\\
they did profanely eat.

29 Thus, by their lewd inventions,\\
they did provoke his ire;\\
And then upon them suddenly\\
the plague brake in as fire.

30 Then Phin’has rose, and justice did,\\
and so the plague did cease;\\
31 That to all ages counted was\\
to him for righteousness.

32 And at the waters, where they strove,\\
they did him angry make,\\
In such sort, that it fared ill\\
with Moses for their sake:

33 Because they there his spirit meek\\
provoked bitterly,\\
So that he uttered with his lips\\
words unadvisedly.

34 Nor, as the Lord commanded them,\\
did they the nations slay:\\
35 But with the heathen mingled were,\\
and learned of them their way.

36 And they their idols served, which did\\
a snare unto them turn.\\
37 Their sons and daughters they to dev’ls\\
in sacrifice did burn.

38 In their own children’s guiltless blood\\
their hands they did imbrue,\\
Whom to Canaan’s idols they\\
for sacrifices slew:

So was the land defiled with blood.\\
39 They stained with their own way,\\
And with their own inventions\\
a whoring they did stray.

40 Against his people kindled was\\
the wrath of God therefore,\\
Insomuch that he did his own\\
inheritance abhor.

41 He gave them to the heathen’s hand;\\
their foes did them command.\\
42 Their en’mies them oppressed, they were\\
made subject to their hand.

43 He many times delivered them;\\
but with their counsel so\\
They him provoked, that for their sin\\
they were brought very low.

44 Yet their affliction he beheld,\\
when he did hear their cry:\\
45 And he for them his covenant\\
did call to memory;

After his mercies’ multitude\\
46 he did repent: And made\\
Them to be pitied of all those\\
who did them captive lead.

47 O Lord our God, us save, and gather\\
the heathen from among,\\
That we thy holy name may praise\\
in a triumphant song.

48 Blessed be Jehovah, Isr’el’s God,\\
to all eternity:\\
Let all the people say, Amen.\\
Praise to the Lord give ye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}



