\subsection*{Psalm 1 \blfootnote{\emph{Psalm 1 Tunes:} \sc{Azmon / Denfield}}}

\addcontentsline{toc}{subsection}{Psalm 1}

1 That man hath perfect blessedness,\\
who walketh not astray\\
In counsel of ungodly men,\\
nor stands in sinners’ way,

Nor sitteth in the scorner’s chair:\\
2 But placeth his delight\\
Upon God’s law, and meditates\\
on his law day and night.

3 He shall be like a tree that grows\\
near planted by a river,\\
Which in his season yields his fruit,\\
and his leaf fadeth never:

And all he doth shall prosper well\\
4 The wicked are not so;\\
But like they are unto the chaff,\\
which wind drives to and fro.

5 In judgment therefore shall not stand \\
such as ungodly are; \\
Nor in th’ assembly of the just \\
shall wicked men appear.

6 For why? the way of godly men \\
unto the Lord is known:\\
Whereas the way of wicked men \\
shall quite be overthrown.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 2 \blfootnote{\emph{Psalm 2 Tune:} \sc{York}}}
\addcontentsline{toc}{subsection}{Psalm 2}

1 Why rage the heathen? and vain things\\
why do the people mind?\\
2 Kings of the earth do set themselves,\\
and princes are combin’d,

To plot against the Lord, and his\\
Anointed, saying thus,\\
3 Let us asunder break their bands,\\
and cast their cords from us.

4 He that in heaven sits shall laugh;\\
the Lord shall scorn them all.\\
5 Then shall he speak to them in wrath,\\
in rage he vex them shall.

6 Yet, notwithstanding, I have him\\
to be my King appointed;\\
And over Sion, my holy hill,\\
I have him King anointed.

7 The sure decree I will declare:\\
The Lord hath said to me,\\
Thou art mine only Son; this day\\
I have begotten thee.

8 Ask of me, and for heritage\\
the heathen I’ll make thine;\\
And, for possession, I to thee\\
will give earth’s utmost line.

9 Thou shalt, as with a weighty rod\\
of iron, break them all;\\
And, as a potter’s sherd, thou shalt\\
them dash in pieces small.

10 Now therefore, kings, be wise; be taught, \\
ye judges of the earth:\\
11 Serve God in fear, and see that ye\\
join trembling with your mirth.

12 Kiss ye the Son, lest in his ire \\
ye perish from the way, \\
If once his wrath begin to burn: \\
bless’d all that on him stay.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 3}
\addcontentsline{toc}{subsection}{Psalm 3}

A Psalm of David, when he fled from Absalom his son.

1 O Lord, how are my foes increas’d?\\
against me many rise.\\
2 Many say of my soul, For him\\
in God no succour lies.

3 Yet thou my shield and glory art,\\
th’ uplifter of mine head.\\
4 I cry’d, and, from his holy hill,\\
the Lord me answer made.

5 I laid me down and slept; I wak’d;\\
for God sustained me.\\
6 I will not fear though thousands ten\\
set round against me be.

7 Arise, O Lord; save me, my God;\\
for thou my foes hast stroke\\
All on the cheek-bone, and the teeth\\
of wicked men hast broke.

8 Salvation doth appertain\\
unto the Lord alone:\\
Thy blessing, Lord, for evermore\\
thy people is upon.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 4 \blfootnote{\emph{Psalm 4 Tune:} \sc{Crimond}}}
\addcontentsline{toc}{subsection}{Psalm 4}


To the chief Musician on Neginoth\footnote{Stringed Instruments?},\\
A Psalm of David.

1 Give ear unto me when I call,\\
God of my righteousness:\\
Have mercy, hear my pray’r; thou hast\\
enlarg’d me in distress.

2 O ye the sons of men! how long\\
will ye love vanities?\\
How long my glory turn to shame,\\
and will ye follow lies?

3 But know, that for himself the Lord\\
the godly man doth chuse:\\
The Lord, when I on him do call,\\
to hear will not refuse.

4 Fear, and sin not; talk with your heart\\
on bed, and silent be.\\
5 Off ‘rings present of righteousness,\\
and in the Lord trust ye.

6 O who will shew us any good?\\
is that which many say:\\
But of thy countenance the light,\\
Lord, lift on us alway.

7 Upon my heart, bestow’d by thee,\\
more gladness I have found\\
Than they, ev’n then, when corn and wine\\
did most with them abound.

8 I will both lay me down in peace,\\
and quiet sleep will take;\\
Because thou only me to dwell\\
in safety, Lord, dost make.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 5}
\addcontentsline{toc}{subsection}{Psalm 5}

To the chief Musician, upon Nehiloth\footnote{Wind Instruments?},
A \emph{Psalm} of David.

1 Give ear unto my words, O Lord,\\
my meditation weigh.\\
2 Hear my loud cry, my King, my God;\\
for I to thee will pray.

3 Lord, thou shalt early hear my voice:\\
I early will direct\\
My pray’r to thee; and, looking up,\\
an answer will expect.

4 For thou art not a God that doth\\
in wickedness delight;\\
Neither shall evil dwell with thee,\\
5 Nor fools stand in thy sight.

All that ill-doers are thou hat’st;\\
6 Cutt’st off that liars be:\\
The bloody and deceitful man\\
abhorred is by thee.

7 But I into thy house will come\\
in thine abundant grace;\\
And I will worship in thy fear\\
toward thy holy place.

8 Because of those mine enemies,\\
Lord, in thy righteousness\\
Do thou me lead; do thou thy way\\
make straight before my face.

9 For in their mouth there is no truth,\\
their inward part is ill;\\
Their throat’s an open sepulchre,\\
their tongue doth flatter still.

10 O God, destroy them; let them be\\
by their own counsel quell’d:\\
Them for their many sins cast out,\\
for they ‘gainst thee rebell’d.

11 But let all joy that trust in thee,\\
and still make shouting noise;\\
For them thou sav’st; let all that love\\
thy name in thee rejoice.

12 For, Lord, unto the righteous man\\
thou wilt thy blessing yield:\\
With favour thou wilt compass him \\
about, as with a shield.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 6}
\addcontentsline{toc}{subsection}{Psalm 6}

\emph{First Version---Long Meter (L.M., 8.8.8.8)} \blfootnote{\emph{Psalm 6 L.M. Tune:} \sc{Finnart}}

To the chief Musician on Neginoth\footnote{String Instruments?} upon Sheminith\footnote{Eighth?},
A Psalm of David.

1 Lord, in thy wrath rebuke me not;\\
Nor in thy hot rage chasten me.\\
2 Lord, pity me, for I am weak:\\
Heal me, for my bones vexed be.

3 My soul is also vexed sore;\\
But, Lord, how long stay wilt thou make?\\
4 Return, O Lord, my soul set free;\\
O save me, for thy mercies’ sake.

5 Because those that deceased are\\
Of thee shall no remembrance have;\\
And who is he that will to thee\\
Give praises lying in the grave?

6 I with my groaning weary am,\\
I also all the night my bed\\
Have caused for to swim; and I\\
With tears my couch have watered.

7 Mine eye, consum’d with grief, grows old,\\
Because of all mine enemies.\\
8 Hence from me, wicked workers all;\\
For God hath heard my weeping cries.

9 God hath my supplication heard,\\
My pray’r received graciously\\
10 Sham’d and sore vex’d be all my foes,\\
Sham’d and back turned suddenly.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (C.M., 8.6.8.6)}

To the chief Musician on Neginoth\footnote{String Instruments?} upon Sheminith\footnote{Eighth?},
A Psalm of David.

1 In thy great indignation,\\
O Lord, rebuke me not;\\
Nor on me lay thy chast ‘ning hand,\\
in thy displeasure hot.

2 Lord, I am weak, therefore on me\\
have mercy, and me spare:\\
Heal me, O Lord, because thou know’st\\
my bones much vexed are.

3 My soul is vexed sore: but, Lord,\\
how long stay wilt thou make?\\
4 Return, Lord, free my soul; and save\\
me, for thy mercies’ sake.

5 Because of thee in death there shall\\
no more remembrance be:\\
Of those that in the grave do lie,\\
who shall give thanks to thee?

6 I with my groaning weary am,\\
and all the night my bed\\
I caused for to swim; with tears\\
my couch I watered.

7 By reason of my vexing grief,\\
mine eye consumed is;\\
It waxeth old, because of all\\
that be mine enemies.

8 But now, depart from me all ye\\
that work iniquity:\\
For why? the Lord hath heard my voice,\\
when I did mourn and cry.

9 Unto my supplication\\
the Lord did hearing give:\\
When I to him my prayer make,\\
the Lord will it receive.

10 Let all be sham’d and troubled sore,\\
That en’mies are to me;\\
Let them turn back, and suddenly\\
ashamed let them be.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 7}
\addcontentsline{toc}{subsection}{Psalm 7}


Shiggaion of David, which he sang unto the Lord,
concerning the words of Cush the Benjamite.

1 O Lord my God, in thee do I\\
my confidence repose:\\
Save and deliver me from all\\
my persecuting foes;

2 Lest that the enemy my soul\\
should, like a lion, tear,\\
In pieces rending it, while there\\
is no deliverer.

3 O Lord my God, if it be so\\
that I committed this;\\
If it be so that in my hands\\
iniquity there is:

4 If I rewarded ill to him\\
that was at peace with me;\\
(Yea, ev’n the man that without cause\\
my foe was I did free;)

5 Then let the foe pursue and take\\
my soul, and my life thrust\\
Down to the earth, and let him lay\\
mine honour in the dust.

6 Rise in thy wrath, Lord, raise thyself,\\
for my foes raging be;\\
And, to the judgment which thou hast\\
commanded, wake for me.

7 So shall th’ assembly of thy folk\\
about encompass thee:\\
Thou, therefore, for their sakes, return\\
unto thy place on high.

8 The Lord he shall the people judge:\\
my judge, Jehovah, be,\\
After my righteousness, and mine\\
integrity in me.

9 O let the wicked’s malice end;\\
but stablish stedfastly\\
The righteous: for the righteous God\\
the hearts and reins doth try.

10 In God, who saves th’ upright in heart,\\
is my defence and stay.\\
11 God just men judgeth, God is wroth\\
with ill men ev’ry day.

12 If he do not return again,\\
then he his sword will whet;\\
His bow he hath already bent,\\
and hath it ready set:

13 He also hath for him prepar’d\\
the instruments of death;\\
Against the persecutors he\\
his shafts ordained hath.

14 Behold, he with iniquity\\
doth travail, as in birth;\\
A mischief he conceived hath,\\
and falsehood shall bring forth.

15 He made a pit and digg’d it deep,\\
another there to take;\\
But he is fall’n into the ditch\\
which he himself did make.

16 Upon his own head his mischief\\
shall be returned home;\\
His vi’lent dealing also down\\
on his own pate shall come.

17 According to his righteousness\\
the Lord I’ll magnify;\\
And will sing praise unto the name\\
of God that is most high.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 8 \blfootnote{\emph{Psalm 8 Tune:} \sc{Richmond}}}
\addcontentsline{toc}{subsection}{Psalm 8}


To the chief Musician upon Gittith,\\
A Psalm of David.

1 How excellent in all the earth,\\
Lord, our Lord, is thy name!\\
Who hast thy glory far advanc’d\\
above the starry frame.

2 From infants’ and from sucklings’ mouth\\
thou didest strength ordain,\\
For thy foes’ cause, that so thou might’st\\
th’ avenging foe restrain.

3 When I look up unto the heav’ns,\\
which thine own fingers fram’d,\\
Unto the moon, and to the stars,\\
which were by thee ordain’d;

4 Then say I, What is man, that he\\
remember’d is by thee?\\
Or what the son of man, that thou\\
so kind to him should’st be?

5 For thou a little lower hast\\
him than the angels made;\\
With glory and with dignity\\
thou crowned hast his head.

6 Of thy hands’ works thou mad’st him lord,\\
all under’s feet didst lay;\\
7 All sheep and oxen, yea, and beasts\\
that in the field do stray;

8 Fowls of the air, fish of the sea,\\
all that pass through the same.\\
9 How excellent in all the earth,\\
Lord, our Lord, is thy name!

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 9 \blfootnote{\emph{Psalm 9 Tune:} \sc{Stroudwater}}}
\addcontentsline{toc}{subsection}{Psalm 9}


To the chief Musician upon Muth-labben,\\
A Psalm of David.

1 Lord, thee I’ll praise with all my heart,\\
thy wonders all proclaim.\\
2 In thee, most High, I’ll greatly joy,\\
and sing unto thy name.

3 When back my foes were turn’d, they fell,\\
and perish’d at thy sight:\\
4 For thou maintain’dst my right and cause;\\
on throne sat’st judging right.

5 The heathen thou rebuked hast,\\
the wicked overthrown;\\
Thou hast put out their names, that they\\
may never more be known.

6 O en’my! now destructions have\\
an end perpetual:\\
Thou cities raz’d; perish’d with them\\
is their memorial.

7 God shall endure for aye; he doth\\
for judgment set his throne;\\
8 In righteousness to judge the world,\\
justice to give each one.

9 God also will a refuge be\\
for those that are oppress’d;\\
A refuge will he be in times\\
of trouble to distress’d.

10 And they that know thy name, in thee\\
their confidence will place:\\
For thou hast not forsaken them\\
that truly seek thy face.

11 O sing ye praises to the Lord,\\
that dwells in Sion hill;\\
And all the nations among\\
his deeds record ye still.

12 When he enquireth after blood,\\
he then rememb’reth them:\\
The humble folk he not forgets\\
that call upon his name.

13 Lord, pity me; behold the grief\\
which I from foes sustain;\\
Ev’n thou, who from the gates of death\\
dost raise me up again;

14 That I, in Sion’s daughters’ gates,\\
may all thy praise advance;\\
And that I may rejoice always\\
in thy deliverance.

15 The heathen are sunk in the pit\\
which they themselves prepar’d;\\
And in the net which they have hid\\
their own feet fast are snar’d.

16 The Lord is by the judgment known\\
which he himself hath wrought:\\
The sinners’ hands do make the snares\\
wherewith themselves are caught.

17 They who are wicked into hell\\
each one shall turned be;\\
And all the nations that forget\\
to seek the Lord most high.

18 For they that needy are shall not\\
forgotten be alway;\\
The expectation of the poor\\
shall not be lost for aye.

19 Arise, Lord, let not man prevail;\\
judge heathen in thy sight:\\
20 That they may know themselves but men,\\
the nations, Lord, affright.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 10 \blfootnote{\emph{Psalm 10 Tune:} \sc {Glenluce}}}
\addcontentsline{toc}{subsection}{Psalm 10}

1 Wherefore is it that thou, O Lord,\\
dost stand from us afar?\\
And wherefore hidest thou thyself,\\
when times so troublous are?

2 The wicked in his loftiness\\
doth persecute the poor:\\
In these devices they have fram’d\\
let them be taken sure.

3 The wicked of his heart’s desire\\
doth talk with boasting great;\\
He blesseth him that’s covetous,\\
whom yet the Lord doth hate.

4 The wicked, through his pride of face,\\
on God he doth not call;\\
And in the counsels of his heart\\
the Lord is not at all.

5 His ways they always grievous are;\\
thy judgments from his sight\\
Removed are: at all his foes\\
he puffeth with despight.

6 Within his heart he thus hath said,\\
I shall not moved be;\\
And no adversity at all\\
shall ever come to me.

7 His mouth with cursing, fraud, deceit,\\
is fill’d abundantly;\\
And underneath his tongue there is\\
mischief and vanity.

8 He closely sits in villages;\\
he slays the innocent:\\
Against the poor that pass him by\\
his cruel eyes are bent.

9 He, lion-like, lurks in his den;\\
he waits the poor to take;\\
And when he draws him in his net,\\
his prey he doth him make.

10 Himself he humbleth very low,\\
he croucheth down withal,\\
That so a multitude of poor\\
may by his strong ones fall.

11 He thus hath said within his heart,\\
The Lord hath quite forgot;\\
He hides his countenance, and he\\
for ever sees it not.

12 O Lord, do thou arise; O God,\\
lift up thine hand on high:\\
Put not the meek afflicted ones\\
out of thy memory.

13 Why is it that the wicked man\\
thus doth the Lord despise?\\
Because that God will it require\\
he in his heart denies.

14 Thou hast it seen; for their mischief\\
and spite thou wilt repay:\\
The poor commits himself to thee;\\
thou art the orphan’s stay.

15 The arm break of the wicked man,\\
and of the evil one;\\
Do thou seek out his wickedness,\\
until thou findest none.

16 The Lord is King through ages all,\\
ev’n to eternity;\\
The heathen people from his land\\
are perish’d utterly.

17 O Lord, of those that humble are\\
thou the desire didst hear;\\
Thou wilt prepare their heart, and thou\\
to hear wilt bend thine ear;

18 To judge the fatherless, and those\\
that are oppressed sore;\\
That man, that is but sprung of earth,\\
may them oppress no more.

 % Separation Symbol

\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 11 \blfootnote{\emph{Psalm 11 Tune:} \sc {Salisbury}}}
\addcontentsline{toc}{subsection}{Psalm 11}

To the chief Musician,\\
\emph{A Psalm} of David.

1 I in the Lord do put my trust:\\
how is it then that ye\\
Say to my soul, Flee, as a bird,\\
unto your mountain high?

2 For, lo, the wicked bend their bow,\\
their shafts on string they fit,\\
That those who upright are in heart\\
they privily may hit.

3 If the foundations be destroy’d,\\
what hath the righteous done?\\
4 God in his holy temple is,\\
in heaven is his throne:

His eyes do see, his eye-lids try\\
5 men’s sons. The just he proves:\\
But his soul hates the wicked man,\\
and him that vi’lence loves.

6 Snares, fire and brimstone, furious storms,\\
on sinners he shall rain:\\
This, as the portion of their cup,\\
doth unto them pertain.

7 Because the Lord most righteous doth\\
in righteousness delight;\\
And with a pleasant countenance\\
beholdeth the upright.

%  % Separation Symbol
% \begin{center}
% \adforn{21}\quad\adforn{11}\quad\adforn{49}
% \end{center}

\subsection*{Psalm 12 \blfootnote{\emph{Psalm 12 Tune:} \sc {Farrant}}}
\addcontentsline{toc}{subsection}{Psalm 12}

To the chief Musician upon Sheminith,\\
A Psalm of David.

1 Help, Lord, because the godly man\\
doth daily fade away;\\
And from among the sons of men\\
the faithful do decay.

2 Unto his neighbour ev’ry one\\
doth utter vanity:\\
They with a double heart do speak,\\
and lips of flattery.

3 God shall cut off all flatt’ring lips,\\
tongues that speak proudly thus,\\
4 We’ll with our tongue prevail, our lips\\
are ours: who’s lord o’er us?

5 For poor oppress’d, and for the sighs\\
of needy, rise will I,\\
Saith God, and him in safety set\\
from such as him defy.

6 The words of God are words most pure;\\
they be like silver try’d\\
In earthen furnace, seven times\\
that hath been purify’d.

7 Lord, thou shalt them preserve and keep\\
for ever from this race.\\
8 On each side walk the wicked, when\\
vile men are high in place.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 13 \blfootnote{\emph{Psalm 13 Tune:} \sc {Stornoway}}}
\addcontentsline{toc}{subsection}{Psalm 13}

To the chief Musician,\\
A Psalm of David.

1 How long wilt thou forget me, Lord?\\
shall it for ever be?\\
O how long shall it be that thou\\
wilt hide thy face from me?

2 How long take counsel in my soul,\\
still sad in heart, shall I?\\
How long exalted over me\\
shall be mine enemy?

3 O Lord my God, consider well,\\
and answer to me make:\\
Mine eyes enlighten, lest the sleep\\
of death me overtake:

4 Lest that mine enemy should say,\\
Against him I prevail’d;\\
And those that trouble me rejoice,\\
when I am mov’d and fail’d.

5 But I have all my confidence\\
thy mercy set upon;\\
My heart within me shall rejoice\\
in thy salvation.

6 I will unto the Lord my God\\
sing praises cheerfully,\\
Because he hath his bounty shown\\
to me abundantly.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 14  \blfootnote{\emph{Psalm 14 Tune:} \sc {St. Thomas}}}
\addcontentsline{toc}{subsection}{Psalm 14}

To the chief Musician,\\
\emph{A Psalm} of David.

1 That there is not a God, the fool\\
doth in his heart conclude:\\
They are corrupt, their works are vile;\\
not one of them doth good.

2 Upon men’s sons the Lord from heav’n\\
did cast his eyes abroad,\\
To see if any understood,\\
and did seek after God.

3 They altogether filthy are,\\
they all aside are gone;\\
And there is none that doeth good,\\
yea, sure there is not one.

4 These workers of iniquity\\
do they not know at all,\\
That they my people eat as bread,\\
and on God do not call?

5 There fear’d they much; for God is with\\
the whole race of the just.\\
6 You shame the counsel of the poor,\\
because God is his trust.

7 Let Isr’el’s help from Sion come:\\
when back the Lord shall bring\\
His captives, Jacob shall rejoice,\\
and Israel shall sing.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 15 \blfootnote{\emph{Psalm 15 Tune:} \sc {Grafenberg}}}
\addcontentsline{toc}{subsection}{Psalm 15}

A Psalm of David.

1 Within thy tabernacle, Lord,\\
who shall abide with thee?\\
And in thy high and holy hill\\
who shall a dweller be?

2 The man that walketh uprightly,\\
and worketh righteousness,\\
And as he thinketh in his heart,\\
so doth he truth express.

3 Who doth not slander with his tongue,\\
nor to his friend doth hurt;\\
Nor yet against his neighbour doth\\
take up an ill report.

4 In whose eyes vile men are despis’d;\\
but those that God do fear\\
He honoureth; and changeth not,\\
though to his hurt he swear.

5 His coin puts not to usury,\\
nor take reward will he\\
Against the guiltless. Who doth thus\\
shall never moved be.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 16}
\addcontentsline{toc}{subsection}{Psalm 16}

Michtam of David.

1 Lord, keep me; for I trust in thee.\\
2 To God thus was my speech,\\
Thou art my Lord; and unto thee\\
my goodness doth not reach:

3 To saints on earth, to th’ excellent,\\
where my delight’s all plac’d.\\
4 Their sorrows shall be multiply’d\\
to other gods that haste:

Of their drink-offerings of blood\\
I will no off ‘ring make;\\
Yea, neither I their very names\\
up in my lips will take.

5 God is of mine inheritance\\
and cup the portion;\\
The lot that fallen is to me\\
thou dost maintain alone.

6 Unto me happily the lines\\
in pleasant places fell;\\
Yea, the inheritance I got\\
in beauty doth excel.

7 I bless the Lord, because he doth\\
by counsel me conduct;\\
And in the seasons of the night\\
my reins do me instruct.

8 Before me still the Lord I set:\\
sith it is so that he\\
Doth ever stand at my right hand,\\
I shall not moved be.

9 Because of this my heart is glad,\\
and joy shall be exprest\\
Ev’n by my glory; and my flesh\\
in confidence shall rest.

10 Because my soul in grave to dwell\\
shall not be left by thee;\\
Nor wilt thou give thine Holy One\\
corruption to see.

11 Thou wilt me shew the path of life:\\
of joys there is full store\\
Before thy face; at thy right hand\\
are pleasures evermore.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 17 \blfootnote{\emph{Psalm 17 Tune:} \sc{St. Bernard}}}
\addcontentsline{toc}{subsection}{Psalm 17}

A Prayer of David.

1 Lord, hear the right, attend my cry,\\
unto my pray’r give heed,\\
That doth not in hypocrisy\\
from feigned lips proceed.

2 And from before thy presence forth\\
my sentence do thou send:\\
Toward these things that equal are\\
do thou thine eyes intend.

3 Thou prov’dst mine heart, thou visit’dst me\\
by night, thou didst me try,\\
Yet nothing found’st; for that my mouth\\
shall not sin, purpos’d I.

4 As for men’s works, I, by the word\\
that from thy lips doth flow,\\
Did me preserve out of the paths\\
wherein destroyers go.

5 Hold up my goings, Lord, me guide\\
in those thy paths divine,\\
So that my footsteps may not slide\\
out of those ways of thine.

6 I called have on thee, O God,\\
because thou wilt me hear:\\
That thou may’st hearken to my speech,\\
to me incline thine ear.

7 Thy wondrous loving-kindness show,\\
thou that, by thy right hand,\\
Sav’st them that trust in thee from those\\
that up against them stand.

8 As th’ apple of the eye me keep;\\
in thy wings shade me close\\
9 From lewd oppressors, compassing\\
me round, as deadly foes.

10 In their own fat they are inclos’d;\\
their mouth speaks loftily.\\
11 Our steps they compass’d; and to ground\\
down bowing set their eye.

12 He like unto a lion is\\
that’s greedy of his prey,\\
Or lion young, which lurking doth\\
in secret places stay.

13 Arise, and disappoint my foe,\\
and cast him down, O Lord:\\
My soul save from the wicked man,\\
the man which is thy sword.

14 From men, which are thy hand, O Lord,\\
from worldly men me save,\\
Which only in this present life\\
their part and portion have.

Whose belly with thy treasure hid\\
thou fill’st: they children have\\
In plenty; of their goods the rest\\
they to their children leave.

15 But as for me, I thine own face\\
in righteousness will see;\\
And with thy likeness, when I wake,\\
I satisfy’d shall be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 18 \blfootnote{\emph{Psalm 18 Tune:} \sc{Warwick}}}
\addcontentsline{toc}{subsection}{Psalm 18}


To the chief Musician, \emph{A Psalm} of David, the servant of the Lord, who spake unto the Lord the words of this song in the day that the Lord delivered him from the hand of all his enemies, and from the hand of Saul: And he said,

1 Thee will I love, O Lord, my strength.\\
2 My fortress is the Lord,\\
My rock, and he that doth to me\\
deliverance afford:

My God, my strength, whom I will trust,\\
a buckler unto me,\\
The horn of my salvation,\\
and my high tow’r, is he.

3 Upon the Lord, who worthy is\\
of praises, will I cry;\\
And then shall I preserved be\\
safe from mine enemy.

4 Floods of ill men affrighted me,\\
death’s pangs about me went;\\
5 Hell’s sorrows me environed;\\
death’s snares did me prevent.

6 In my distress I call’d on God,\\
cry to my God did I;\\
He from his temple heard my voice,\\
to his ears came my cry.

7 Th’ earth, as affrighted, then did shake,\\
trembling upon it seiz’d:\\
The hills’ foundations moved were,\\
because he was displeas’d.

8 Up from his nostrils came a smoke,\\
and from his mouth there came\\
Devouring fire, and coals by it\\
were turned into flame.

9 He also bowed down the heav’ns,\\
and thence he did descend;\\
And thickest clouds of darkness did\\
under his feet attend.

10 And he upon a cherub rode,\\
and thereon he did fly;\\
Yea, on the swift wings of the wind\\
his flight was from on high.

11 He darkness made his secret place:\\
about him, for his tent,\\
Dark waters were, and thickest clouds\\
of th’ airy firmament.

12 And at the brightness of that light,\\
which was before his eye,\\
His thick clouds pass’d away, hailstones\\
and coals of fire did fly.

13 The Lord God also in the heav’ns\\
did thunder in his ire;\\
And there the Highest gave his voice,\\
hailstones and coals of fire.

14 Yea, he his arrows sent abroad,\\
and them he scattered;\\
His lightnings also he shot out,\\
and them discomfited.

15 The waters’ channels then were seen,\\
the world’s foundations vast\\
At thy rebuke discover’d were,\\
and at thy nostrils’ blast.

16 And from above the Lord sent down,\\
and took me from below;\\
From many waters he me drew,\\
which would me overflow.

17 He me reliev’d from my strong foes,\\
and such as did me hate;\\
Because he saw that they for me\\
too strong were, and too great.

18 They me prevented in the day\\
of my calamity;\\
But even then the Lord himself\\
a stay was unto me.

19 He to a place where liberty\\
and room was hath me brought;\\
Because he took delight in me,\\
he my deliv’rance wrought.

20 According to my righteousness\\
he did me recompense,\\
He me repaid according to\\
my hands’ pure innocence.

21 For I God’s ways kept, from my God\\
did not turn wickedly.\\
His judgments were before me, I\\
his laws put not from me.

23 Sincere before him was my heart;\\
with him upright was I;\\
And watchfully I kept myself\\
from mine iniquity.

24 After my righteousness the Lord\\
hath recompensed me,\\
After the cleanness of my hands\\
appearing in his eye.

25 Thou gracious to the gracious art,\\
to upright men upright:\\
26 Pure to the pure, froward thou kyth’st\\
unto the froward wight.

27 For thou wilt the afflicted save\\
in grief that low do lie:\\
But wilt bring down the countenance\\
of them whose looks are high.

28 The Lord will light my candle so,\\
that it shall shine full bright:\\
The Lord my God will also make\\
my darkness to be light.

29 By thee through troops of men I break,\\
and them discomfit all;\\
And, by my God assisting me,\\
I overleap a wall.

30 As for God, perfect is his way:\\
the Lord his word is try’d;\\
He is a buckler to all those\\
who do in him confide.

31 Who but the Lord is God? but he\\
who is a rock and stay?\\
32 ‘Tis God that girdeth me with strength,\\
and perfect makes my way.

33 He made my feet swift as the hinds,\\
set me on my high places.\\
34 Mine hands to war he taught, mine arms\\
brake bows of steel in pieces.

35 The shield of thy salvation\\
thou didst on me bestow:\\
Thy right hand held me up, and great\\
thy kindness made me grow.

36 And in my way my steps thou hast\\
enlarged under me,\\
That I go safely, and my feet\\
are kept from sliding free.

37 Mine en’mies I pursued have,\\
and did them overtake;\\
Nor did I turn again till I\\
an end of them did make.

38 I wounded them, they could not rise;\\
they at my feet did fall.\\
39 Thou girdedst me with strength for war;\\
my foes thou brought’st down all:

40 And thou hast giv’n to me the necks\\
of all mine enemies;\\
That I might them destroy and slay,\\
who did against me rise.

41 They cried out, but there was none\\
that would or could them save;\\
Yea, they did cry unto the Lord,\\
but he no answer gave.

42 Then did I beat them small as dust\\
before the wind that flies;\\
And I did cast them out like dirt\\
upon the street that lies.

43 Thou mad’st me free from people’s strife,\\
and heathen’s head to be:\\
A people whom I have not known\\
shall service do to me.

44 At hearing they shall me obey,\\
to me they shall submit.\\
45 Strangers for fear shall fade away,\\
who in close places sit.

46 God lives, bless’d be my Rock; the God\\
of my health praised be.\\
47 God doth avenge me, and subdues\\
the people under me.

48 He saves me from mine enemies;\\
yea, thou hast lifted me\\
Above my foes; and from the man\\
of vi’lence set me free.

49 Therefore to thee will I give thanks\\
the heathen folk among;\\
And to thy name, O Lord, I will\\
sing praises in a song.

50 He great deliv’rance gives his king:\\
he mercy doth extend\\
To David, his anointed one,\\
and his seed without end.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 19 \blfootnote{\emph{Psalm 19 Tune:} \sc{Crediton / Moravia / St. Andrew}}}
\addcontentsline{toc}{subsection}{Psalm 19}

To the chief Musician,\\
A Psalm of David.

1 The heav’ns God’s glory do declare,\\
the skies his hand-works preach:\\
2 Day utters speech to day, and night\\
to night doth knowledge teach.

3 There is no speech nor tongue to which\\
their voice doth not extend:\\
4 Their line is gone through all the earth,\\
their words to the world’s end.

In them he set the sun a tent;\\
5 Who, bridegroom-like, forth goes\\
From’s chamber, as a strong man doth\\
to run his race rejoice.

6 From heav’n’s end is his going forth,\\
circling to th’ end again;\\
And there is nothing from his heat\\
that hidden doth remain.

7 God’s law is perfect, and converts\\
the soul in sin that lies:\\
God’s testimony is most sure,\\
and makes the simple wise.

8 The statutes of the Lord are right,\\
and do rejoice the heart:\\
The Lord’s command is pure, and doth\\
light to the eyes impart.

9 Unspotted is the fear of God,\\
and doth endure for ever:\\
The judgments of the Lord are true\\
and righteous altogether.

10 They more than gold, yea, much fine gold,\\
to be desired are:\\
Than honey, honey from the comb\\
that droppeth, sweeter far.

11 Moreover, they thy servant warn\\
how he his life should frame:\\
A great reward provided is\\
for them that keep the same.

12 Who can his errors understand?\\
O cleanse thou me within\\
13 From secret faults. Thy servant keep\\
from all presumptuous sin:

And do not suffer them to have\\
dominion over me:\\
Then, righteous and innocent,\\
I from much sin shall be.

14 The words which from my mouth proceed,\\
the thoughts sent from my heart,\\
Accept, O Lord, for thou my strength\\
and my Redeemer art.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 20 \blfootnote{\emph{Psalm 20 Tune:} \sc{Orton}}}
\addcontentsline{toc}{subsection}{Psalm 20}

To the chief Musician,\\
A Psalm of David.

1 Jehovah hear thee in the day\\
when trouble he doth send:\\
And let the name of Jacob’s God\\
thee from all ill defend.

2 O let him help send from above,\\
out of his sanctuary:\\
From Sion, his own holy hill,\\
let him give strength to thee.

3 Let him remember all thy gifts,\\
accept thy sacrifice:\\
4 Grant thee thine heart’s wish, and fulfil\\
thy thoughts and counsel wise.

5 In thy salvation we will joy;\\
in our God’s name we will\\
Display our banners: and the Lord\\
thy prayers all fulfil.

6 Now know I God his king doth save:\\
he from his holy heav’n\\
Will hear him, with the saving strength\\
by his own right hand giv’n.

7 In chariots some put confidence,\\
some horses trust upon:\\
But we remember will the name\\
of our Lord God alone.

8 We rise, and upright stand, when they\\
are bowed down, and fall.\\
9 Deliver, Lord; and let the King\\
us hear, when we do call.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 21 \blfootnote{\emph{Psalm 21 Tune:} \sc{Arnold}}}
\addcontentsline{toc}{subsection}{Psalm 21}

To the chief Musician,\\
A \emph{Psalm} of David.

1 The king in thy great strength, O Lord,\\
shall very joyful be:\\
In thy salvation rejoice\\
how veh’mently shall he!

2 Thou hast bestowed upon him\\
all that his heart would have;\\
And thou from him didst not withhold\\
whate’er his lips did crave.

3 For thou with blessings him prevent’st\\
of goodness manifold;\\
And thou hast set upon his head\\
a crown of purest gold.

4 When he desired life of thee,\\
thou life to him didst give;\\
Ev’n such a length of days, that he\\
for evermore should live.

5 In that salvation wrought by thee\\
his glory is made great;\\
Honour and comely majesty\\
thou hast upon him set.

6 Because that thou for evermore\\
most blessed hast him made;\\
And thou hast with thy countenance\\
made him exceeding glad.

7 Because the king upon the Lord\\
his confidence doth lay;\\
And through the grace of the most High\\
shall not be mov’d away.

8 Thine hand shall all those men find out\\
that en’mies are to thee;\\
Ev’n thy right hand shall find out those\\
of thee that haters be.

9 Like fiery ov’n thou shalt them make,\\
when kindled is thine ire;\\
God shall them swallow in his wrath,\\
devour them shall the fire.

10 Their fruit from earth thou shalt destroy,\\
their seed men from among:\\
11 For they beyond their might ‘gainst thee\\
did plot mischief and wrong.

12 Thou therefore shalt make them turn back,\\
when thou thy shafts shalt place\\
Upon thy strings, made ready all\\
to fly against their face.

13 In thy great pow’r and strength, O Lord,\\
be thou exalted high;\\
So shall we sing with joyful hearts,\\
thy power praise shall we.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 22 \blfootnote{\emph{Psalm 22 Tune:} \sc{Bangor / Old Twenty-Second}}}
\addcontentsline{toc}{subsection}{Psalm 22}

To the chief Musician upon Aijeleth Shahar\footnote{The hind of the morning.},\\
A Psalm of David.

1 My God, my God, why hast thou me\\
forsaken? why so far\\
Art thou from helping me, and from\\
my words that roaring are?

2 All day, my God, to thee I cry,\\
yet am not heard by thee;\\
And in the season of the night\\
I cannot silent be.

3 But thou art holy, thou that dost\\
inhabit Isr’el’s praise.\\
4 Our fathers hop’d in thee, they hop’d\\
and thou didst them release.

5 When unto thee they sent their cry,\\
to them deliv’rance came:\\
Because they put their trust in thee,\\
they were not put to shame.

6 But as for me, a worm I am,\\
and as no man am priz’d:\\
Reproach of men I am, and by\\
the people am despis’d.

7 All that me see laugh me to scorn;\\
shoot out the lip do they;\\
They nod and shake their heads at me,\\
and, mocking, thus do say,

8 This man did trust in God, that he\\
would free him by his might:\\
Let him deliver him, sith he\\
had in him such delight.

9 But thou art he out of the womb\\
that didst me safely take;\\
When I was on my mother’s breasts\\
thou me to hope didst make.

10 And I was cast upon thy care,\\
ev’n from the womb till now;\\
And from my mother’s belly, Lord,\\
my God and guide art thou.

11 Be not far off, for grief is near,\\
and none to help is found.\\
12 Bulls many compass me, strong bulls\\
of Bashan me surround.

13 Their mouths they open’d wide on me,\\
upon me gape did they,\\
Like to a lion ravening\\
and roaring for his prey.

14 Like water I’m pour’d out, my bones\\
all out of joint do part:\\
Amidst my bowels, as the wax,\\
so melted is my heart.

15 My strength is like a potsherd dry’d;\\
my tongue it cleaveth fast\\
Unto my jaws; and to the dust\\
of death thou brought me hast.

16 For dogs have compass’d me about:\\
the wicked, that did meet\\
In their assembly, me inclos’d;\\
they pierc’d my hands and feet.

17 I all my bones may tell; they do\\
upon me look and stare.\\
18 Upon my vesture lots they cast,\\
and clothes among them share.

19 But be not far, O Lord, my strength;\\
haste to give help to me.\\
20 From sword my soul, from pow’r of dogs\\
my darling set thou free.

21 Out of the roaring lion’s mouth\\
do thou me shield and save:\\
For from the horns of unicorns\\
an ear to me thou gave.

22 I will shew forth thy name unto\\
those that my brethren are;\\
Amidst the congregation\\
thy praise I will declare.

23 Praise ye the Lord, who do him fear;\\
him glorify all ye\\
The seed of Jacob: fear him all\\
that Isr’el’s children be.

24 For he despis’d not nor abhorr’d\\
th’ afflicted’s misery;\\
Nor from him hid his face, but heard\\
when he to him did cry.

25 Within the congregation great\\
my praise shall be of thee;\\
My vows before them that him fear\\
shall be perform’d by me.

26 The meek shall eat, and shall be fill’d;\\
they also praise shall give\\
Unto the Lord that do him seek:\\
your heart shall ever live.

27 All ends of th’ earth remember shall,\\
and turn the Lord unto;\\
All kindreds of the nations\\
to him shall homage do:

28 Because the kingdom to the Lord\\
doth appertain as his;\\
Likewise among the nations\\
the Governor he is.

29 Earth’s fat ones eat, and worship shall:\\
all who to dust descend\\
Shall bow to him; none of them can\\
his soul from death defend.

30 A seed shall service do to him;\\
unto the Lord it shall\\
Be for a generation\\
reckon’d in ages all.

31 They shall come, and they shall declare\\
his truth and righteousness\\
Unto a people yet unborn,\\
and that he hath done this.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 23 \blfootnote{\emph{Psalm 23 Tune:} \sc{Bays of Harris / Crimond}}}
\addcontentsline{toc}{subsection}{Psalm 23}

A Psalm of David.

1 The Lord’s my shepherd, I’ll not want.\\
2 He makes me down to lie\\
In pastures green: he leadeth me\\
the quiet waters by.

3 My soul he doth restore again;\\
and me to walk doth make\\
Within the paths of righteousness,\\
ev’n for his own name’s sake.

4 Yea, though I walk in death’s dark vale,\\
yet will I fear none ill:\\
For thou art with me; and thy rod\\
and staff me comfort still.

5 My table thou hast furnished\\
in presence of my foes;\\
My head thou dost with oil anoint,\\
and my cup overflows.

6 Goodness and mercy all my life\\
shall surely follow me:\\
And in God’s house for evermore\\
my dwelling-place shall be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 24 \blfootnote{\emph{Psalm 24 Tune:} \sc{Rest / St. Georges Edinburgh}}}
\addcontentsline{toc}{subsection}{Psalm 24}

A Psalm of David.\footnote{This title was not included in the 1650 Psalter}

1 The earth belongs unto the Lord,\\
and all that it contains;\\
The world that is inhabited,\\
and all that there remains.

2 For the foundations thereof\\
he on the seas did lay,\\
And he hath it established\\
upon the floods to stay.

3 Who is the man that shall ascend\\
into the hill of God?\\
Or who within his holy place\\
shall have a firm abode?

4 Whose hands are clean, whose heart is pure,\\
and unto vanity\\
Who hath not lifted up his soul,\\
nor sworn deceitfully.

5 He from th’ Eternal shall receive\\
the blessing him upon,\\
And righteousness, ev’n from the God\\
of his salvation.

6 This is the generation\\
that after him enquire,\\
O Jacob, who do seek thy face\\
with their whole heart’s desire.

7 Ye gates, lift up your heads on high;\\
ye doors that last for aye,\\
Be lifted up, that so the King\\
of glory enter may.

8 But who of glory is the King?\\
The mighty Lord is this;\\
Ev’n that same Lord, that great in might\\
and strong in battle is.

9 Ye gates, lift up your heads; ye doors,\\
doors that do last for aye,\\
Be lifted up, that so the King\\
of glory enter may.

10 But who is he that is the King\\
of glory? who is this?\\
The Lord of hosts, and none but he,\\
the King of glory is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 25 \blfootnote{\emph{Psalm 25 S.M. Tune:} \sc{Dennis}}}
\addcontentsline{toc}{subsection}{Psalm 25}

\emph{First Version---Short Meter (S.M., 6.6.8.6)}

\emph{A Psalm} of David.

1 To thee I lift my soul:\\
2 O Lord, I trust in thee:\\
My God, let me not be asham’d,\\
nor foes triumph o’er me.

3 Let none that wait on thee\\
be put to shame at all;\\
But those that without cause transgress,\\
let shame upon them fall.

4 Shew me thy ways, O Lord;\\
thy paths, O teach thou me:\\
5 And do thou lead me in thy truth,\\
therein my teacher be:

For thou art God that dost\\
to me salvation send,\\
And I upon thee all the day\\
expecting do attend.

6 Thy tender mercies, Lord,\\
I pray thee to remember,\\
And loving-kindnesses; for they\\
have been of old for ever.

7 My sins and faults of youth\\
do thou, O Lord, forget:\\
After thy mercy think on me,\\
and for thy goodness great.

8 God good and upright is:\\
the way he’ll sinners show.\\
9 The meek in judgment he will guide,\\
and make his path to know.

10 The whole paths of the Lord\\
are truth and mercy sure,\\
To those that do his cov’nant keep,\\
and testimonies pure.

11 Now, for thine own name’s sake,\\
O Lord, I thee entreat\\
To pardon mine iniquity;\\
for it is very great.

12 What man is he that fears\\
the Lord, and doth him serve?\\
Him shall he teach the way that he\\
shall choose, and still observe.

13 His soul shall dwell at ease;\\
and his posterity\\
Shall flourish still, and of the earth\\
inheritors shall be.

14 With those that fear him is\\
the secret of the Lord;\\
The knowledge of his covenant\\
he will to them afford.

15 Mine eyes upon the Lord\\
continually are set:\\
For he it is that shall bring forth\\
my feet out of the net.

16 Turn unto me thy face,\\
and to me mercy show;\\
Because that I am desolate,\\
and am brought very low.

17 My heart’s griefs are increas’d:\\
me from distress relieve.\\
18 See mine affliction and my pain,\\
and all my sins forgive.

19 Consider thou my foes,\\
because they many are;\\
And it a cruel hatred is\\
which they against me bear.

20 O do thou keep my soul,\\
do thou deliver me:\\
And let me never be asham’d,\\
because I trust in thee.

21 Let uprightness and truth\\
keep me, who thee attend.\\
22 Redemption, Lord, to Israel\\
from all his troubles send.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (C.M., 8.6.8.6)}

\emph{A Psalm} of David.

1 To thee I lift my soul, O Lord:\\
2 My God, I trust in thee:\\
Let me not be asham’d; let not\\
my foes triumph o’er me.

3 Yea, let thou none ashamed be\\
that do on thee attend:\\
Ashamed let them be, O Lord,\\
who without cause offend.

4 Thy ways, Lord, shew; teach me thy paths:\\
5 Lead me in truth, teach me:\\
For of my safety thou art God;\\
all day I wait on thee.

6 Thy mercies, that most tender are,\\
do thou, O Lord, remember,\\
And loving-kindnesses; for they\\
have been of old for ever.

7 Let not the errors of my youth,\\
nor sins, remember’d be:\\
In mercy, for thy goodness’ sake,\\
O Lord, remember me.

8 The Lord is good and gracious,\\
he upright is also:\\
He therefore sinners will instruct\\
in ways that they should go.

9 The meek and lowly he will guide\\
in judgment just alway:\\
To meek and poor afflicted ones\\
he’ll clearly teach his way.

10 The whole paths of the Lord our God\\
are truth and mercy sure,\\
To such as keep his covenant,\\
and testimonies pure.

11 Now, for thine own name’s sake, O Lord,\\
I humbly thee entreat\\
To pardon mine iniquity;\\
for it is very great.

12 What man fears God? him shall he teach\\
the way that he shall chuse.\\
13 His soul shall dwell at ease; his seed\\
the earth, as heirs, shall use.

14 The secret of the Lord is with\\
such as do fear his name;\\
And he his holy covenant\\
will manifest to them.

15 Towards the Lord my waiting eyes\\
continually are set;\\
For he it is that shall bring forth\\
my feet out of the net.

16 O turn thee unto me, O God,\\
have mercy me upon;\\
Because I solitary am,\\
and in affliction.

17 Enlarg’d the griefs are of mine heart;\\
me from distress relieve.\\
18 See mine affliction and my pain,\\
and all my sins forgive.

19 Consider thou mine enemies,\\
because they many are;\\
And it a cruel hatred is\\
which they against me bear.

20 O do thou keep my soul; O God,\\
do thou deliver me:\\
Let me not be asham’d; for I\\
do put my trust in thee.

21 O let integrity and truth\\
keep me, who thee attend.\\
22 Redemption, Lord, to Israel\\
from all his troubles send.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 26 \blfootnote{\emph{Psalm 26 Tune:} \sc{St. Thomas}}}
\addcontentsline{toc}{subsection}{Psalm 26}

\emph{A Psalm} of David

1 Judge me, O Lord, for I have walk’d\\
in mine integrity:\\
I trusted also in the Lord;\\
slide therefore shall not I.

2 Examine me, and do me prove;\\
try heart and reins, O God:\\
3 For thy love is before mine eyes,\\
thy truth’s paths I have trode.

4 With persons vain I have not sat,\\
nor with dissemblers gone:\\
5 Th’ assembly of ill men I hate;\\
to sit with such I shun.

6 Mine hands in innocence, O Lord,\\
I’ll wash and purify;\\
So to thine holy altar go,\\
and compass it will I:

7 That I, with voice of thanksgiving,\\
may publish and declare,\\
And tell of all thy mighty works,\\
that great and wondrous are.

8 The habitation of thy house,\\
Lord, I have loved well;\\
Yea, in that place I do delight\\
where doth thine honour dwell.

9 With sinners gather not my soul,\\
and such as blood would spill:\\
10 Whose hands mischievous plots, right hand\\
corrupting bribes do fill.

11 But as for me, I will walk on\\
in mine integrity:\\
Do thou redeem me, and, O Lord,\\
be merciful to me.

12 My foot upon an even place\\
doth stand with stedfastness:\\
Within the congregations\\
th’ Eternal I will bless.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 27 \blfootnote{\emph{Psalm 27 Tune:} \sc{St. Botolph}}}
\addcontentsline{toc}{subsection}{Psalm 27}

\emph{A Psalm} of David.

1 The Lord’s my light and saving health,\\
who shall make me dismayed?\\
My life’s strength is the Lord, of whom\\
then shall I be afraid?

2 When as mine enemies and foes,\\
most wicked persons all,\\
To eat my flesh against me rose,\\
they stumbled and did fall.

3 Against me though an host encamp,\\
my heart yet fearless is:\\
Though war against me rise, I will\\
be confident in this.

4 One thing I of the Lord desired,\\
and will seek to obtain,\\
That all days of my life I may\\
within God’s house remain;

That I the beauty of the Lord\\
behold may and admire,\\
And that I in his holy place\\
may rev’rently enquire.

5 For he in his pavilion shall\\
me hide in evil days;\\
In secret of his tent me hide,\\
and on a rock me raise.

6 And now, ev’n at this present time,\\
mine head shall lifted be\\
Above all those that are my foes,\\
and round encompass me:

Therefore unto his tabernacle\\
I’ll sacrifices bring\\
Of joyfulness; I’ll sing, yea, I\\
to God will praises sing.

7 O Lord, give ear unto my voice,\\
when I do cry to thee;\\
Upon me also mercy have,\\
and do thou answer me.

8 When thou didst say, seek ye my face,\\
then unto thee reply\\
Thus did my heart, above all things\\
thy face, Lord, seek will I.

9 Far from me hide not thou thy face;\\
put not away from thee\\
Thy servant in thy wrath: thou hast\\
an helper been to me.

O God of my salvation,\\
leave me not, nor forsake:\\
10 Though me my parents both should leave,\\
the Lord will me up take.

11 O Lord, instruct me in thy way,\\
to me a leader be\\
In a plain path, because of those\\
that hatred bear to me.

12 Give me not to mine en’mies’ will;\\
for witnesses that lie\\
Against me risen are, and such\\
as breathe out cruelty.

13 I fainted had, unless that I\\
believed had to see\\
The Lord’s own goodness in the land\\
of them that living be.

14 Wait on the Lord, and be thou strong,\\
and he shall strength afford\\
Unto thine heart; yea, do thou wait,\\
I say, upon the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 28 \blfootnote{\emph{Psalm 28 Tune:} \sc{Caithness}}}
\addcontentsline{toc}{subsection}{Psalm 28}

\emph{A Psalm} of David.

1 To thee I’ll cry, O Lord, my rock;\\
hold not thy peace to me;\\
Lest like those that to pit descend\\
I by thy silence be.

2 The voice hear of my humble pray’rs,\\
when unto thee I cry;\\
When to thine holy oracle\\
I lift mine hands on high.

3 With ill men draw me not away\\
that work iniquity;\\
That speak peace to their friends, while in\\
their hearts doth mischief lie.

4 Give them according to their deeds\\
and ills endeavored:\\
And as their handy-works deserve,\\
to them be rendered.

5 God shall not build, but them destroy,\\
who would not understand\\
The Lord’s own works, nor did regard\\
the doing of his hand.

6 For ever blessed be the Lord,\\
for graciously he heard\\
The voice of my petitions,\\
and prayers did regard.

7 The Lord’s my strength and shield; my heart\\
upon him did rely;\\
And I am helped: hence my heart\\
doth joy exceedingly,

And with my song I will him praise.\\
8 My strength is God alone:\\
He also is the saving strength\\
of his anointed one.\\

9 O thine own people do thou save,\\
bless thine inheritance;\\
Them also do thou feed, and them\\
for evermore advance.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 29}
\addcontentsline{toc}{subsection}{Psalm 29}

A Psalm of David.

1 Give ye unto the Lord, ye sons\\
that of the mighty be,\\
All strength and glory to the Lord\\
with cheerfulness give ye.

2 Unto the Lord the glory give\\
that to his name is due;\\
And in the beauty of holiness\\
unto Jehovah bow.

3 The Lord’s voice on the waters is;\\
the God of majesty\\
Doth thunder, and on multitudes\\
of waters sitteth he.

4 A pow’rful voice it is that comes\\
out from the Lord most high;\\
The voice of that great Lord is full\\
of glorious majesty.

5 The voice of the Eternal doth\\
asunder cedars tear;\\
Yea, God the Lord doth cedars break\\
that Lebanon doth bear.

6 He makes them like a calf to skip,\\
ev’n that great Lebanon,\\
And, like to a young unicorn,\\
the mountain Sirion.

7 God’s voice divides the flames of fire;\\
8 The desert it doth shake:\\
The Lord doth make the wilderness\\
of Kadesh all to quake.

9 God’s voice doth make the hinds to calve,\\
it makes the forest bare:\\
And in his temple ev’ry one\\
his glory doth declare.

10 The Lord sits on the floods; the Lord\\
sits King, and ever shall.\\
11 The Lord will give his people strength,\\
and with peace bless them all.\\

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 30}
\addcontentsline{toc}{subsection}{Psalm 30}

A Psalm; A Song at the Dedication of the House.\footnote{The preceding title information is no the 1650 Psalter.}
\emph{A Psalm} of David.

1 Lord, I will thee extol, for thou\\
hast lifted me on high,\\
And over me thou to rejoice\\
mad’st not mine enemy.

2 O thou who art the Lord my God,\\
I in distress to thee,\\
With loud cries lifted up my voice,\\
and thou hast healed me.

3 O Lord, my soul thou hast brought up,\\
and rescued from the grave;\\
That I to pit should not go down,\\
alive thou didst me save.

4 O ye that are his holy ones,\\
sing praise unto the Lord;\\
And give unto him thanks, when ye\\
his holiness record.

5 For but a moment lasts his wrath;\\
life in his favor lies:\\
Weeping may for a night endure,\\
at morn doth joy arise.

6 In my prosperity I said,\\
that nothing shall me move.\\
7 O Lord, thou hast my mountain made\\
to stand strong by thy love:

But when that thou, O gracious God,\\
didst hide thy face from me,\\
Then quickly was my prosp’rous state\\
turned into misery.

8 Wherefore unto the Lord my cry\\
I caused to ascend:\\
My humble supplication\\
I to the Lord did send.

9 What profit is there in my blood,\\
when I go down to pit?\\
Shall unto thee the dust give praise?\\
thy truth declare shall it?

10 Hear, Lord, have mercy; help me, Lord:\\
11 Thou turned hast my sadness\\
To dancing; yea, my sackcloth loosed,\\
and girded me with gladness;

12 That sing thy praise my glory may,\\
and never silent be.\\
O Lord my God, for evermore\\
I will give thanks to thee.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 31 \blfootnote{\emph{Psalm 31 Tune:} \sc{Consolation / St. Lawrence}}}
\addcontentsline{toc}{subsection}{Psalm 31}

To the chief Musician, A Psalm of David.

1 In thee, O Lord, I put my trust,\\
shamed let me never be;\\
According to thy righteousness\\
do thou deliver me.

2 Bow down thine ear to me, with speed\\
send me deliverance:\\
To save me, my strong rock be thou,\\
and my house of defense.

3 Because thou art my rock, and thee\\
I for my fortress take;\\
Therefore do thou me lead and guide,\\
ev’n for thine own name’s sake.

4 And since thou art my strength, therefore\\
pull me out of the net,\\
Which they in subtlety for me\\
so privily have set.

5 Into thine hands I do commit\\
my sp’rit: for thou art he,\\
O thou, Jehovah, God of truth,\\
that hast redeemed me.

6 Those that do lying vanities\\
regard, I have abhorred:\\
But as for me, my confidence\\
is fixed on the Lord.

7 I’ll in thy mercy gladly joy:\\
for thou my miseries\\
Considered hast; thou hast my soul\\
known in adversities:

8 And thou hast not enclosed me\\
within the en’my’s hand;\\
And by thee have my feet been made\\
in a large room to stand.

9 O Lord, upon me mercy have,\\
for trouble is on me:\\
Mine eye, my belly, and my soul,\\
with grief consumed be.

10 Because my life with grief is spent,\\
my years with sighs and groans:\\
My strength doth fail; and for my sin\\
consumed are my bones.

11 I was a scorn to all my foes,\\
and to my friends a fear;\\
And specially reproached of those\\
that were my neighbors near:

When they me saw they from me fled.\\
12 Ev’n so I am forgot,\\
As men are out of mind when dead:\\
I’m like a broken pot.

13 For slanders I of many heard;\\
fear compassed me, while they\\
Against me did consult, and plot\\
to take my life away.

14 But as for me, O Lord, my trust\\
upon thee I did lay;\\
And I to thee, Thou art my God,\\
did confidently say.

15 My times are wholly in thine hand:\\
do thou deliver me\\
From their hands that mine enemies\\
and persecutors be.

16 Thy countenance to shine do thou\\
upon thy servant make:\\
Unto me give salvation,\\
for thy great mercies’ sake.

17 Let me not be ashamed, O Lord,\\
for on thee called I have:\\
Let wicked men be shamed, let them\\
be silent in the grave.

18 To silence put the lying lips,\\
that grievous things do say,\\
And hard reports, in pride and scorn,\\
on righteous men do lay.

19 How great’s the goodness thou for them\\
that fear thee keep’st in store,\\
And wrought’st for them that trust in thee\\
the sons of men before!

20 In secret of thy presence thou\\
shalt hide them from man’s pride:\\
From strife of tongues thou closely shalt,\\
as in a tent, them hide.

21 All praise and thanks be to the Lord;\\
for he hath magnified\\
His wondrous love to me within\\
a city fortified.

22 For from thine eyes cut off I am,\\
I in my haste had said;\\
My voice yet heard’st thou, when to thee\\
with cries my moan I made.

23 O love the Lord, all ye his saints;\\
because the Lord doth guard\\
The faithful, and he plenteously\\
proud doers doth reward.

24 Be of good courage, and He strength\\
unto your heart shall send,\\
All ye whose hope and confidence\\
doth on the Lord depend.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 32 \blfootnote{\emph{Psalm 32 Tune:} \sc{Salzburg}}}
\addcontentsline{toc}{subsection}{Psalm 32}

\emph{A Psalm} of David, Maschil.

1 O blessed is the man to whom\\
is freely pardoned\\
All the transgression he hath done,\\
whose sin is covered.

2 Blessed is the man to whom the Lord\\
imputeth not his sin,\\
And in whose sp’rit there is no guile,\\
nor fraud is found therein.

3 When as I did refrain my speech,\\
and silent was my tongue,\\
My bones then waxed old, because\\
I roared all day long.

4 For upon me both day and night\\
thine hand did heavy lie,\\
So that my moisture turned is\\
in summer’s drought thereby.

5 I thereupon have unto thee\\
my sin acknowledged,\\
And likewise mine iniquity\\
I have not covered:

I will confess unto the Lord\\
my trespasses, said I;\\
And of my sin thou freely didst\\
forgive th’ iniquity.

6 For this shall ev’ry godly one\\
his prayer make to thee;\\
In such a time he shall thee seek,\\
as found thou mayest be.

Surely, when floods of waters great\\
do swell up to the brim,\\
They shall not overwhelm his soul,\\
nor once come near to him.

7 Thou art my hiding-place, thou shalt\\
from trouble keep me free:\\
Thou with songs of deliverance\\
about shalt compass me.

8 I will instruct thee, and thee teach\\
the way that thou shalt go;\\
And, with mine eye upon thee set,\\
I will direction show.

9 Then be not like the horse or mule,\\
which do not understand;\\
Whose mouth, lest they come near to thee,\\
a bridle must command.

10 Unto the man that wicked is\\
his sorrows shall abound;\\
But him that trusteth in the Lord\\
mercy shall compass round.

11 Ye righteous, in the Lord be glad,\\
in him do ye rejoice:\\
All ye that upright are in heart,\\
for joy lift up your voice.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 33 \blfootnote{\emph{Psalm 33 Tune:} \sc{Irish}}}
\addcontentsline{toc}{subsection}{Psalm 33}

1 Ye righteous, in the Lord rejoice;\\
it comely is and right,\\
That upright men, with thankful voice,\\
should praise the Lord of might.

2 Praise God with harp, and unto him\\
sing with the psaltery;\\
Upon a ten-stringed instrument\\
make ye sweet melody.

3 A new song to him sing, and play\\
with loud noise skilfully;\\
4 For right is God’s word, all his works\\
are done in verity.

5 To judgment and to righteousness\\
a love he beareth still;\\
The loving-kindness of the Lord\\
the earth throughout doth fill.

6 The heavens by the word of God\\
did their beginning take;\\
And by the breathing of his mouth\\
he all their hosts did make.

7 The waters of the seas he brings\\
together as an heap;\\
And in storehouses, as it were,\\
he layeth up the deep.

8 Let earth, and all that live therein,\\
with rev’rence fear the Lord;\\
Let all the world’s inhabitants\\
dread him with one accord.

9 For he did speak the word, and done\\
it was without delay;\\
Established it firmly stood,\\
whatever he did say.

10 God doth the counsel bring to naught\\
which heathen folk do take;\\
And what the people do devise\\
of none effect doth make.

11 O but the counsel of the Lord\\
doth stand for ever sure;\\
And of his heart the purposes\\
from age to age endure.

12 That nation blessed is, whose God\\
Jehovah is, and those\\
A blessed people are, whom for\\
his heritage he chose.

13 The Lord from heav’n sees and beholds\\
all sons of men full well:\\
14 He views all from his dwelling-place\\
that in the earth do dwell.

15 He forms their hearts alike, and all\\
their doings he observes.\\
16 Great hosts save not a king, much strength\\
no mighty man preserves.

17 An horse for preservation\\
is a deceitful thing;\\
And by the greatness of his strength\\
can no deliv’rance bring.

18 Behold on those that do him fear\\
the Lord doth set his eye;\\
Ev’n those who on his mercy do\\
with confidence rely.

19 From death to free their soul, in dearth\\
life unto them to yield.\\
20 Our soul doth wait upon the Lord;\\
he is our help and shield.

21 Since in his holy name we trust,\\
our heart shall joyful be.\\
22 Lord, let thy mercy be on us,\\
as we do hope in thee.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 34 \blfootnote{\emph{Psalm 34 Tune:} \sc{St. Matthew \ Wiltshire}}}
\addcontentsline{toc}{subsection}{Psalm 34}

\emph{A Psalm} of David, when he changed his behaviour before Abimelech; who drove him away, and he departed.

1 God will I bless all times; his praise\\
my mouth shall still express.\\
2 My soul shall boast in God: the meek\\
shall hear with joyfulness.

3 Extol the Lord with me, let us\\
exalt his name together.\\
4 I sought the Lord, he heard, and did\\
me from all fears deliver.

5 They looked to him, and lightened were:\\
not shamed were their faces.\\
6 This poor man cried, God heard, and saved\\
him from all his distresses.

7 The angel of the Lord encamps,\\
and round encompasseth\\
All those about that do him fear,\\
and them delivereth.

8 O taste and see that God is good:\\
who trusts in him is blessed.\\
9 Fear God his saints: none that him fear\\
shall be with want oppressed.

10 The lions young may hungry be,\\
and they may lack their food:\\
But they that truly seek the Lord\\
shall not lack any good.

11 O children, hither do ye come,\\
and unto me give ear;\\
I shall you teach to understand\\
how ye the Lord should fear.

12 What man is he that life desires,\\
to see good would live long?\\
13 Thy lips refrain from speaking guile,\\
and from ill words thy tongue.

14 Depart from ill, do good, seek peace,\\
pursue it earnestly.\\
15 God’s eyes are on the just; his ears\\
are open to their cry.

16 The face of God is set against\\
those that do wickedly,\\
That he may quite out from the earth\\
cut off their memory.

17 The righteous cry unto the Lord,\\
he unto them gives ear;\\
And they out of their troubles all\\
by him delivered are.

18 The Lord is ever nigh to them\\
that be of broken sp’rit;\\
To them he safety doth afford\\
that are in heart contrite.

19 The troubles that afflict the just\\
in number many be;\\
But yet at length out of them all\\
the Lord doth set him free.

20 He carefully his bones doth keep,\\
whatever can befall;\\
That not so much as one of them\\
can broken be at all.

21 Ill shall the wicked slay; laid waste\\
shall be who hate the just.\\
22 The Lord redeems his servants’ souls;\\
none perish that him trust.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 35}
\addcontentsline{toc}{subsection}{Psalm 35}

\emph{A Psalm} of David.

1 Plead, Lord, with those that plead; and fight\\
with those that fight with me.\\
2 Of shield and buckler take thou hold,\\
stand up mine help to be.

3 Draw also out the spear, and do\\
against them stop the way\\
That me pursue: unto my soul,\\
I’m thy salvation, say.

4 Let them confounded be and shamed\\
that for my soul have sought:\\
Who plot my hurt turned back be they,\\
and to confusion brought.

5 Let them be like unto the chaff\\
that flies before the wind;\\
And let the angel of the Lord\\
pursue them hard behind.

6 With darkness cover thou their way,\\
and let it slipp’ry prove;\\
And let the angel of the Lord\\
pursue them from above.

7 For without cause have they for me\\
their net hid in a pit,\\
They also have without a cause\\
for my soul digged it.

8 Let ruin seize him unawares;\\
his net he hid withal\\
Himself let catch; and in the same\\
destruction let him fall.

9 My soul in God shall joy; and glad\\
in his salvation be:\\
10 And all my bones shall say, O Lord,\\
who is like unto thee,

Which dost the poor set free from him\\
that is for him too strong;\\
The poor and needy from the man\\
that spoils and does him wrong?

11 False witnesses rose; to my charge\\
things I not knew they laid.\\
12 They, to the spoiling of my soul,\\
me ill for good repaid.

13 But as for me, when they were sick,\\
in sackcloth sad I mourned:\\
My humbled soul did fast, my pray’r\\
into my bosom turned.

14 Myself I did behave as he\\
had been my friend or brother;\\
I heavily bowed down, as one\\
that mourneth for his mother.

15 But in my trouble they rejoiced,\\
gath’ring themselves together;\\
Yea, objects vile together did\\
themselves against me gather:

I knew it not; they did me tear,\\
and quiet would not be.\\
16 With mocking hypocrites, at feasts\\
they gnashed their teeth at me.

17 How long, Lord, look’st thou on? from those\\
destructions they intend\\
Rescue my soul, from lions young\\
my darling do defend.

18 I will give thanks to thee, O Lord,\\
within th’ assembly great;\\
And where much people gathered are\\
thy praises forth will set.

19 Let not my wrongful enemies\\
proudly rejoice o’er me;\\
Nor who me hate without a cause,\\
let them wink with the eye.

20 For peace they do not speak at all;\\
but crafty plots prepare\\
Against all those within the land\\
that meek and quiet are.

21 With mouths set wide, they ‘gainst me said,\\
Ha, ha! our eye doth see.\\
22 Lord, thou hast seen, hold not thy peace;\\
Lord, be not far from me.

23 Stir up thyself; wake, that thou may’st\\
judgment to me afford,\\
Ev’n to my cause, O thou that art\\
my only God and Lord.

24 O Lord my God, do thou me judge\\
after thy righteousness;\\
And let them not their joy ‘gainst me\\
triumphantly express:

25 Nor let them say within their hearts,\\
Ah, we would have it thus;\\
Nor suffer them to say, that he\\
is swallowed up by us.

26 Shamed and confounded be they all\\
that at my hurt are glad;\\
Let those against me that do boast\\
with shame and scorn be clad.

27 Let them that love my righteous cause\\
be glad, shout, and not cease\\
To say, The Lord be magnified,\\
who loves his servant’s peace.

28 Thy righteousness shall also be\\
declared by my tongue;\\
The praises that belong to thee\\
speak shall it all day long.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 36 \blfootnote{\emph{Psalm 36 Tune:} \sc{Elijah}}}
\addcontentsline{toc}{subsection}{Psalm 36}

To the chief Musician,
\emph{A Psalm} of David, the servant of the Lord.

1 The wicked man’s transgression\\
within my heart thus says,\\
Undoubtedly the fear of God\\
is not before his eyes.

2 Because himself he flattereth\\
in his own blinded eye,\\
Until the hatefulness be found\\
of his iniquity.

3 Words from his mouth proceeding are,\\
fraud and iniquity:\\
He to be wise, and to do good,\\
hath left off utterly.

4 He mischief, lying on his bed,\\
most cunningly doth plot:\\
He sets himself in ways not good,\\
ill he abhorreth not.

5 Thy mercy, Lord, is in the heav’ns;\\
thy truth doth reach the clouds:\\
6 Thy justice is like mountains great;\\
thy judgments deep as floods:

Lord, thou preservest man and beast.\\
7 How precious is thy grace!\\
Therefore in shadow of thy wings\\
men’s sons their trust shall place.

8 They with the fatness of thy house\\
shall be well satisfied;\\
From rivers of thy pleasures thou\\
wilt drink to them provide.

9 Because of life the fountain pure\\
remains alone with thee;\\
And in that purest light of thine\\
we clearly light shall see.

10 Thy loving-kindness unto them\\
continue that thee know;\\
And still on men upright in heart\\
thy righteousness bestow.

11 Let not the foot of cruel pride\\
come, and against me stand;\\
And let me not removed be,\\
Lord, by the wicked’s hand.

12 There fallen are they, and ruined,\\
that work iniquities:\\
Cast down they are, and never shall\\
be able to arise.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 37 \blfootnote{\emph{Psalm 37 Tunes:} \sc{Azmon / Denfield}}}
\addcontentsline{toc}{subsection}{Psalm 37}

\emph{A Psalm} of David.

1 For evil-doers fret thou not\\
thyself unquietly;\\
Nor do thou envy bear to those\\
that work iniquity.

2 For, even like unto the grass,\\
soon be cut down shall they;\\
And, like the green and tender herb,\\
they wither shall away.

3 Set thou thy trust upon the Lord,\\
and be thou doing good;\\
And so thou in the land shalt dwell,\\
and verily have food.

4 Delight thyself in God; he’ll give\\
thine heart’s desire to thee.\\
5 Thy way to God commit, him trust,\\
it bring to pass shall he.

6 And, like unto the light, he shall\\
thy righteousness display;\\
And he thy judgment shall bring forth\\
like noon-tide of the day.

7 Rest in the Lord, and patiently\\
wait for him: do not fret\\
For him who, prosp’ring in his way,\\
success in sin doth get.

8 Do thou from anger cease, and wrath\\
see thou forsake also:\\
Fret not thyself in any wise,\\
that evil thou should’st do.

9 For those that evil doers are\\
shall be cut off and fall:\\
But those that wait upon the Lord\\
the earth inherit shall.

10 For yet a little while, and then\\
the wicked shall not be;\\
His place thou shalt consider well,\\
but it thou shalt not see.

11 But by inheritance the earth\\
the meek ones shall possess:\\
They also shall delight themselves\\
in an abundant peace.

12 The wicked plots against the just,\\
and at him whets his teeth:\\
13 The Lord shall laugh at him, because\\
his day he coming seeth.

14 The wicked have drawn out the sword,\\
and bent their bow, to slay\\
The poor and needy, and to kill\\
men of an upright way.

15 But their own sword, which they have drawn,\\
shall enter their own heart:\\
Their bows which they have bent shall break,\\
and into pieces part.

16 A little that a just man hath\\
is more and better far\\
Than is the wealth of many such\\
as lewd and wicked are.

17 For sinners’ arms shall broken be;\\
but God the just sustains.\\
18 God knows the just man’s days, and still\\
their heritage remains.

19 They shall not be ashamed when they\\
the evil time do see;\\
And when the days of famine are,\\
they satisfied shall be.

20 But wicked men, and foes of God,\\
as fat of lambs, decay;\\
They shall consume, yea, into smoke\\
they shall consume away.

21 The wicked borrows, but the same\\
again he doth not pay;\\
Whereas the righteous mercy shows,\\
and gives his own away.

22 For such as blessed be of him\\
the earth inherit shall;\\
And they that cursed are of him\\
shall be destroyed all.

23 A good man’s footsteps by the Lord\\
are ordered aright;\\
And in the way wherein he walks\\
he greatly doth delight.

24 Although he fall, yet shall he not\\
be cast down utterly;\\
Because the Lord with his own hand\\
upholds him mightily.

25 I have been young, and now am old,\\
yet have I never seen\\
The just man left, nor that his seed\\
for bread have beggars been.

26 He’s ever merciful, and lends:\\
his seed is blessed therefore.\\
27 Depart from evil, and do good,\\
and dwell for evermore.

28 For God loves judgment, and his saints\\
leaves not in any case;\\
They are kept ever: but cut off\\
shall be the sinner’s race.

29 The just inherit shall the land,\\
and ever in it dwell:\\
30 The just man’s mouth doth wisdom speak;\\
his tongue doth judgment tell.

31 In’s heart the law is of his God,\\
his steps slide not away.\\
32 The wicked man doth watch the just,\\
and seeketh him to slay.

33 Yet him the Lord will not forsake,\\
nor leave him in his hands:\\
The righteous will he not condemn,\\
when he in judgment stands.

34 Wait on the Lord, and keep his way,\\
and thee exalt shall he\\
Th’ earth to inherit; when cut off\\
the wicked thou shalt see.

35 I saw the wicked great in pow’r,\\
spread like a green bay-tree:\\
36 He passed, yea, was not; him I sought,\\
but found he could not be.

37 Mark thou the perfect, and behold\\
the man of uprightness;\\
Because that surely of this man\\
the latter end is peace.

38 But those men that transgressors are\\
shall be destroyed together;\\
The latter end of wicked men\\
shall be cut off for ever.

39 But the salvation of the just\\
is from the Lord above;\\
He in the time of their distress\\
their stay and strength doth prove.

40 The Lord shall help, and them deliver:\\
he shall them free and save\\
From wicked men; because in him\\
their confidence they have.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 38}
\addcontentsline{toc}{subsection}{Psalm 38}

A Psalm of David, to bring to remembrance.

1 In thy great indignation,\\
O Lord, rebuke me not;\\
Nor on me lay thy chast’ning hand,\\
in thy displeasure hot.

2 For in me fast thine arrows stick,\\
thine hand doth press me sore:\\
3 And in my flesh there is no health,\\
nor soundness any more.

This grief I have, because thy wrath\\
is forth against me gone;\\
And in my bones there is no rest,\\
for sin that I have done.

4 Because gone up above mine head\\
my great transgressions be;\\
And, as a weighty burden, they\\
too heavy are for me.

5 My wounds do stink, and are corrupt;\\
my folly makes it so.\\
6 I troubled am, and much bowed down;\\
all day I mourning go.

7 For a disease that loathsome is\\
so fills my loins with pain,\\
That in my weak and weary flesh\\
no soundness doth remain.

8 So feeble and infirm am I,\\
and broken am so sore,\\
That, through disquiet of my heart,\\
I have been made to roar.

9 O Lord, all that I do desire\\
is still before thine eye;\\
And of my heart the secret groans\\
not hidden are from thee.

10 My heart doth pant incessantly,\\
my strength doth quite decay;\\
As for mine eyes, their wonted light\\
is from me gone away.

11 My lovers and my friends do stand\\
at distance from my sore;\\
And those do stand aloof that were\\
kinsmen and kind before.

12 Yea, they that seek my life lay snares:\\
who seek to do me wrong\\
Speak things mischievous, and deceits\\
imagine all day long.

13 But, as one deaf, that heareth not,\\
I suffered all to pass;\\
I as a dumb man did become,\\
whose mouth not opened was:

14 As one that hears not, in whose mouth\\
are no reproofs at all.\\
15 For, Lord, I hope in thee; my God,\\
thou’lt hear me when I call.

16 For I said, Hear me, lest they should\\
rejoice o’er me with pride;\\
And o’er me magnify themselves,\\
when as my foot doth slide.

17 For I am near to halt, my grief\\
is still before mine eye:\\
18 For I’ll declare my sin, and grieve\\
for mine iniquity.

19 But yet mine en’mies lively are,\\
and strong are they beside;\\
And they that hate me wrongfully\\
are greatly multiplied.

20 And they for good that render ill,\\
as en’mies me withstood;\\
Yea, ev’n for this, because that I\\
do follow what is good.

21 Forsake me not, O Lord; my God,\\
far from me never be.\\
22 O Lord, thou my salvation art,\\
haste to give help to me.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 39}
\addcontentsline{toc}{subsection}{Psalm 39}

To the chief Musician, even to Jeduthan,\\
A Psalm of David.

1 I said, I will look to my ways,\\
lest with my tongue I sin:\\
In sight of wicked men my mouth\\
with bridle I’ll keep in.

2 With silence I as dumb became,\\
I did myself restrain\\
From speaking good; but then the more\\
increased was my pain.

3 My heart within me waxed hot;\\
and, while I musing was,\\
The fire did burn; and from my tongue\\
these words I did let pass:

4 Mine end, and measure of my days,\\
O Lord, unto me show\\
What is the same; that I thereby\\
my frailty well may know.

5 Lo, thou my days an handbreadth mad’st;\\
mine age is in thine eye\\
As nothing: sure each man at best\\
is wholly vanity.

6 Sure each man walks in a vain show;\\
they vex themselves in vain:\\
He heaps up wealth, and doth not know\\
to whom it shall pertain.

7 And now, O Lord, what wait I for?\\
my hope is fix’d on thee.\\
8 Free me from all my trespasses,\\
the fool’s scorn make not me.

9 Dumb was I, op’ning not my mouth,\\
because this work was thine.\\
10 Thy stroke take from me; by the blow\\
of thine hand I do pine.

11 When with rebukes thou dost correct\\
man for iniquity,\\
Thou wastes his beauty like a moth:\\
sure each man’s vanity.

12 Attend my cry, Lord, at my tears\\
and pray’rs not silent be:\\
I sojourn as my fathers all,\\
and stranger am with thee.

13 O spare thou me, that I my strength\\
recover may again,\\
Before from hence I do depart,\\
and here no more remain.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 40}
\addcontentsline{toc}{subsection}{Psalm 40}

To the chief Musician,
A Psalm of David.

1 I waited for the Lord my God,\\
and patiently did bear;\\
At length to me he did incline\\
my voice and cry to hear.

2 He took me from a fearful pit,\\
and from the miry clay,\\
And on a rock he set my feet,\\
establishing my way.

3 He put a new song in my mouth,\\
our God to magnify:\\
Many shall see it, and shall fear,\\
and on the Lord rely.

4 O blessed is the man whose trust\\
upon the Lord relies;\\
Respecting not the proud, nor such\\
as turn aside to lies.

5 O Lord my God, full many are\\
the wonders thou hast done;\\
Thy gracious thoughts to us-ward far\\
above all thoughts are gone:

In order none can reckon them\\
to thee: if them declare,\\
And speak of them I would, they more\\
than can be numbered are.

6 No sacrifice nor offering\\
didst thou at all desire;\\
Mine ears thou bored: sin-off ‘ring thou\\
and burnt didst not require:

7 Then to the Lord these were my words,\\
I come, behold and see;\\
Within the volume of the book\\
it written is of me:

8 To do thy will I take delight,\\
O thou my God that art;\\
Yea, that most holy law of thine\\
I have within my heart.

9 Within the congregation great\\
I righteousness did preach:\\
Lo, thou dost know, O Lord, that I\\
refrained not my speech.

10 I never did within my heart\\
conceal thy righteousness;\\
I thy salvation have declared,\\
and shown thy faithfulness:

Thy kindness, which most loving is,\\
concealed have not I,\\
Nor from the congregation great\\
have hid thy verity.

11 Thy tender mercies, Lord, from me\\
O do thou not restrain;\\
Thy loving-kindness, and thy truth,\\
let them me still maintain.

12 For ills past reck’ning compass me,\\
and mine iniquities\\
Such hold upon me taken have,\\
I cannot lift mine eyes:

They more than hairs are on mine head,\\
thence is my heart dismayed.\\
13 Be pleased, Lord, to rescue me;\\
Lord, hasten to mine aid.

14 Shamed and confounded be they all\\
that seek my soul to kill;\\
Yea, let them backward driven be,\\
and shamed, that wish me ill.

15 For a reward of this their shame\\
confounded let them be.\\
That in this manner scoffing say,\\
Aha, aha! to me.

16 In thee let all be glad, and joy,\\
who seeking thee abide;\\
Who thy salvation love, say still,\\
The Lord be magnified.

17 I’m poor and needy, yet the Lord\\
of me a care doth take:\\
Thou art my help and savior,\\
my God, no tarrying make.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 41}
\addcontentsline{toc}{subsection}{Psalm 41}

To the chief Musician,
A Psalm of David.

1 Blessed is he that wisely doth\\
the poor man’s case consider;\\
For when the time of trouble is,\\
the Lord will him deliver.

2 God will him keep, yea, save alive;\\
on earth he blessed shall live;\\
And to his enemies’ desire\\
thou wilt him not up give.

3 God will give strength when he on bed\\
of languishing doth mourn;\\
And in his sickness sore, O Lord,\\
thou all his bed wilt turn.

4 I said, O Lord, do thou extend\\
thy mercy unto me;\\
O do thou heal my soul; for why?\\
I have offended thee.

5 Those that to me are enemies,\\
of me do evil say,\\
When shall he die, that so his name\\
may perish quite away?

6 To see me if he comes, he speaks\\
vain words: but then his heart\\
Heaps mischief to it, which he tells,\\
when forth he doth depart.

7 My haters jointly whispering,\\
‘gainst me my hurt devise.\\
8 Mischief, say they, cleaves fast to him;\\
he li’th, and shall not rise.

9 Yea, ev’n mine own familiar friend,\\
on whom I did rely,\\
Who ate my bread, ev’n he his heel\\
against me lifted high.

10 But, Lord, be merciful to me,\\
and up again me raise,\\
That I may justly them requite\\
according to their ways.

11 By this I know that certainly\\
I favored am by thee;\\
Because my hateful enemy\\
triumphs not over me.

12 But as for me, thou me uphold’st\\
in mine integrity;\\
And me before thy countenance\\
thou sett’st continually.

13 The Lord, the God of Israel,\\
be blessed for ever then,\\
From age to age eternally.\\
Amen, yea, and amen.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}




