\subsection*{Psalm 42}
\addcontentsline{toc}{subsection}{Psalm 42}

To the chief Musician, Maschil, for the sons of Korah.

1 Like as the hart for water-brooks\\
in thirst doth pant and bray;\\
So pants my longing soul, O God,\\
that come to thee I may.

2 My soul for God, the living God,\\
doth thirst: when shall I near\\
Unto thy countenance approach,\\
and in God’s sight appear?

3 My tears have unto me been meat,\\
both in the night and day,\\
While unto me continually,\\
Where is thy God? they say.

4 My soul is poured out in me,\\
when this I think upon;\\
Because that with the multitude\\
I heretofore had gone:

With them into God’s house I went,\\
with voice of joy and praise;\\
Yea, with the multitude that kept\\
the solemn holy days.

5 O why art thou cast down, my soul?\\
why in me so dismayed?\\
Trust God, for I shall praise him yet,\\
his count’nance is mine aid.

6 My God, my soul’s cast down in me;\\
thee therefore mind I will\\
From Jordan’s land, the Hermonites,\\
and ev’n from Mizar hill.

7 At the noise of thy water-spouts\\
deep unto deep doth call;\\
Thy breaking waves pass over me,\\
yea, and thy billows all.

8 His loving-kindness yet the Lord\\
command will in the day,\\
His song’s with me by night; to God,\\
by whom I live, I’ll pray:

9 And I will say to God my rock,\\
Why me forgett’st thou so?\\
Why, for my foes’ oppression,\\
thus mourning do I go?

10 ‘Tis as a sword within my bones,\\
when my foes me upbraid;\\
Ev’n when by them, Where is thy God?\\
’tis daily to me said.

11 O why art thou cast down, my soul?\\
why, thus with grief oppresed,\\
Art thou disquieted in me?\\
in God still hope and rest:

For yet I know I shall him praise,\\
who graciously to me\\
The health is of my countenance,\\
yea, mine own God is 

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 43}
\addcontentsline{toc}{subsection}{Psalm 43}

1 Judge me, O God, and plead my cause\\
against th’ ungodly nation;\\
From the unjust and crafty man,\\
O be thou my salvation.

2 For thou the God art of my strength;\\
why thrusts thou me thee fro’?\\
For th’ enemy’s oppression\\
why do I mourning go?

3 O send thy light forth and thy truth;\\
let them be guides to me,\\
And bring me to thine holy hill,\\
ev’n where thy dwellings be.

4 Then will I to God’s altar go,\\
to God my chiefest joy:\\
Yea, God, my God, thy name to praise\\
my harp I will employ.

5 Why art thou then cast down, my soul?\\
what should discourage thee?\\
And why with vexing thoughts art thou\\
disquieted in me?

Still trust in God; for him to praise\\
good cause I yet shall have:\\
He of my count’nance is the health,\\
my God that doth me save.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 44}
\addcontentsline{toc}{subsection}{Psalm 44}

To the chief Musician
for the sons of Korah, Maschil.

1 O God, we with our ears have heard,\\
our fathers have us told,\\
What works thou in their days hadst done,\\
ev’n in the days of old.

2 Thy hand did drive the heathen out,\\
and plant them in their place;\\
Thou didst afflict the nations,\\
but them thou didst increase.

3 For neither got their sword the land,\\
nor did their arm them save;\\
But thy right hand, arm, countenance;\\
for thou them favor gave.

4 Thou art my King: for Jacob, Lord,\\
deliv’rances command.\\
5 Through thee we shall push down our foes,\\
that do against us stand:

We, through thy name, shall tread down those\\
that ris’n against us have.\\
6 For in my bow I shall not trust,\\
nor shall my sword me save.

7 But from our foes thou hast us saved,\\
our haters put to shame.\\
8 In God we all the day do boast,\\
and ever praise thy name.

9 But now we are cast off by thee,\\
and us thou putt’st to shame;\\
And when our armies do go forth,\\
thou go’st not with the same.

10 Thou mak’st us from the enemy,\\
faint-hearted, to turn back;\\
And they who hate us for themselves\\
our spoils away do take.

11 Like sheep for meat thou gavest us;\\
‘mong heathen cast we be.\\
12 Thou didst for nought thy people sell;\\
their price enrich’d not thee.

13 Thou mak’st us a reproach to be\\
unto our neighbors near;\\
Derision and a scorn to them\\
that round about us are.

14 A by-word also thou dost us\\
among the heathen make;\\
The people, in contempt and spite,\\
at us their heads do shake.

15 Before me my confusion\\
continually abides;\\
And of my bashful countenance\\
the shame me ever hides:

16 For voice of him that doth reproach,\\
and speaketh blasphemy;\\
By reason of th’ avenging foe,\\
and cruel enemy.

17 All this is come on us, yet we\\
have not forgotten thee;\\
Nor falsely in thy covenant\\
behaved ourselves have we.

18 Back from thy way our heart not turned;\\
our steps no straying made;\\
19 Though us thou brak’st in dragons’ place,\\
and cover’dst with death’s shade.

20 If we God’s name forgot, or stretched\\
to a strange god our hands,\\
21 Shall not God search this out? for he\\
heart’s secrets understands.

22 Yea, for thy sake we’re killed all day,\\
counted as slaughter-sheep.\\
23 Rise, Lord, cast us not ever off;\\
awake, why dost thou sleep?

24 O wherefore hidest thou thy face?\\
forgett’st our cause distressed,\\
25 And our oppression? For our soul\\
is to the dust down pressed:

Our belly also on the earth\\
fast cleaving, hold doth take.\\
26 Rise for our help, and us redeem,\\
ev’n for thy mercies’ sake.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{First Version (C.M., 8.6.8.6)}

\subsection*{Psalm 45}
\addcontentsline{toc}{subsection}{Psalm 45}

To the chief Musician upon Shoshannim,
for the sons of Korah, Maschil, A Song of loves.

1 My heart brings forth a goodly thing;\\
my words that I indite\\
Concern the King: my tongue’s a pen\\
of one that swift doth write.

2 Thou fairer art than sons of men:\\
into thy lips is store\\
Of grace infused; God therefore thee\\
hath blessed for evermore.

3 O thou that art the mighty One,\\
thy sword gird on thy thigh;\\
Ev’n with thy glory excellent,\\
and with thy majesty.

4 For meekness, truth, and righteousness,\\
in state ride prosp’rously;\\
And thy right hand shall thee instruct\\
in things that fearful be.

5 Thine arrows sharply pierce the heart\\
of th’ en’mies of the King;\\
And under thy subjection\\
the people down do bring.

6 For ever and for ever is,\\
O God, thy throne of might;\\
The scepter of thy kingdom is\\
a scepter that is right.

7 Thou lovest right, and hatest ill;\\
for God, thy God, most high,\\
Above thy fellows hath with th’ oil\\
of joy anointed thee.

8 Of aloes, myrrh, and cassia,\\
a smell thy garments had,\\
Out of the iv’ry palaces,\\
whereby they made thee glad.

9 Among thy women honorable\\
kings’ daughters were at hand:\\
Upon thy right hand did the queen\\
in gold of Ophir stand.

10 O daughter, hearken and regard,\\
and do thine ear incline;\\
Likewise forget thy father’s house,\\
and people that are thine.

11 Then of the King desired shall be\\
thy beauty veh’mently:\\
Because he is thy Lord, do thou\\
him worship rev’rently.

12 The daughter there of Tyre shall be\\
with gifts and off ‘rings great:\\
Those of the people that are rich\\
thy favor shall entreat.

13 Behold, the daughter of the King\\
all glorious is within;\\
And with embroideries of gold\\
her garments wrought have been.

14 She shall be brought unto the King\\
in robes with needle wrought;\\
Her fellow-virgins following\\
shall unto thee be brought.

15 They shall be brought with gladness great,\\
and mirth on ev’ry side,\\
Into the palace of the King,\\
and there they shall abide.

16 Instead of those thy fathers dear,\\
thy children thou may’st take,\\
And in all places of the earth\\
them noble princes make.

17 Thy name remembered I will make\\
through ages all to be:\\
The people therefore evermore\\
shall praises give to thee.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (S.M, 6.6.8.6)}

To the chief Musician upon Shoshannim,
for the sons of Korah, Maschil, A Song of loves.

1 My heart inditing is\\
good matter in a song:\\
I speak the things that I have made,\\
which to the King belong:

My tongue shall be as quick,\\
his honour to indite,\\
As is the pen of any scribe\\
that useth fast to write.

2 Thou’rt fairest of all men;\\
grace in thy lips doth flow:\\
And therefore blessings evermore\\
on thee doth God bestow.

3 Thy sword gird on thy thigh,\\
thou that art most of might:\\
Appear in dreadful majesty,\\
and in thy glory bright.

4 For meekness, truth, and right,\\
ride prosp’rously in state;\\
And thy right hand shall teach to thee\\
things terrible and great.

5 Thy shafts shall pierce their hearts\\
that foes are to the King;\\
Whereby into subjection\\
the people thou shalt bring.

6 Thy royal seat, O Lord,\\
for ever shall remain:\\
The scepter of thy kingdom doth\\
all righteousness maintain.

7 Thou lov’st right, and hat’st ill;\\
for God, thy God, most high,\\
Above thy fellows hath with th’ oil\\
of joy anointed thee.

8 Of myrrh and spices sweet\\
a smell thy garments had,\\
Out of the iv’ry palaces,\\
whereby they made thee glad.

9 And in thy glorious train\\
kings’ daughters waiting stand;\\
And thy fair queen, in Ophir gold,\\
doth stand at thy right hand.

10 O daughter, take good heed,\\
incline, and give good ear;\\
Thou must forget thy kindred all,\\
and father’s house most dear.

11 Thy beauty to the King\\
shall then delightful be:\\
And do thou humbly worship him,\\
because thy Lord is he.

12 The daughter then of Tyre\\
there with a gift shall be,\\
And all the wealthy of the land\\
shall make their suit to thee.

13 The daughter of the King\\
all glorious is within;\\
And with embroideries of gold\\
her garments wrought have been.

14 She cometh to the King\\
in robes with needle wrought;\\
The virgins that do follow her\\
shall unto thee be brought.

15 They shall be brought with joy,\\
and mirth on ev’ry side,\\
Into the palace of the King,\\
and there they shall abide.

16 And in thy fathers’ stead,\\
thy children thou may’st take,\\
And in all places of the earth\\
them noble princes make.

17 I will show forth thy name\\
to generations all:\\
Therefore the people evermore\\
to thee give praises shal

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 46}
\addcontentsline{toc}{subsection}{Psalm 46}

To the chief Musician, for the sons of Korah,
A Song upon Alomoth.

1 God is our refuge and our strength,\\
in straits a present aid;\\
2 Therefore, although the earth remove,\\
we will not be afraid:

Though hills amidst the seas be cast;\\
3 Though waters roaring make,\\
And troubled be; yea, though the hills,\\
by swelling seas do shake.

4 A river is, whose streams do glad\\
the city of our God;\\
The holy place, wherein the Lord\\
most high hath his abode.

5 God in the midst of her doth dwell;\\
nothing shall her remove:\\
The Lord to her an helper will,\\
and that right early, prove.

6 The heathen raged tumultuously,\\
the kingdoms moved were:\\
The Lord God uttered his voice,\\
the earth did melt for fear.

7 The Lord of hosts upon our side\\
doth constantly remain:\\
The God of Jacob’s our refuge,\\
us safely to maintain.

8 Come, and behold what wondrous works\\
have by the Lord been wrought;\\
Come, see what desolations\\
he on the earth hath brought.

9 Unto the ends of all the earth\\
wars into peace he turns:\\
The bow he breaks, the spear he cuts,\\
in fire the chariot burns.

10 Be still, and know that I am God;\\
among the heathen I\\
Will be exalted; I on earth\\
will be exalted high.

11 Our God, who is the Lord of hosts,\\
is still upon our side;\\
The God of Jacob our refuge\\
for ever will abide.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 47}
\addcontentsline{toc}{subsection}{Psalm 47}

To the chief Musician,
A Psalm for the sons of Korah.

1 All people, clap your hands; to God\\
with voice of triumph shout:\\
2 For dreadful is the Lord most high,\\
great King the earth throughout.

3 The heathen people under us\\
he surely shall subdue;\\
And he shall make the nations\\
under our feet to bow.

4 The lot of our inheritance\\
choose out for us shall he,\\
Of Jacob, whom he loved well,\\
ev’n the excellency.

5 God is with shouts gone up, the Lord\\
with trumpets sounding high.\\
6 Sing praise to God, sing praise, sing praise,\\
praise to our King sing ye.

7 For God is King of all the earth;\\
with knowledge praise express.\\
8 God rules the nations: God sits on\\
his throne of holiness.

9 The princes of the people are\\
assembled willingly;\\
Ev’n of the God of Abraham\\
they who the people be.

For why? the shields that do defend\\
the earth are only his:\\
They to the Lord belong; yea, he\\
exalted greatly is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 48}
\addcontentsline{toc}{subsection}{Psalm 48}

A Song \emph{and} Psalm for the sons of Korah.

1 Great is the Lord, and greatly he\\
is to be praised still,\\
Within the city of our God,\\
upon his holy hill.

2 Mount Zion stands most beautiful,\\
the joy of all the land;\\
The city of the mighty King\\
on her north side doth stand.

3 The Lord within her palaces\\
is for a refuge known.\\
4 For, lo, the kings that gathered were\\
together, by have gone.

5 But when they did behold the same,\\
they, wond’ring, would not stay;\\
But, being troubled at the sight,\\
they thence did haste away.

6 Great terror there took hold on them;\\
they were possessed with fear;\\
Their grief came like a woman’s pain,\\
when she a child doth bear.

7 Thou Tarshish ships with east wind break’st:\\
8 As we have heard it told,\\
So, in the city of the Lord,\\
our eyes did it behold;

In our God’s city, which his hand\\
for ever stablish will.\\
9 We of thy loving-kindness thought,\\
Lord, in thy temple still.

10 O Lord, according to thy name,\\
through all the earth’s thy praise;\\
And thy right hand, O Lord, is full\\
of righteousness always.

11 Because thy judgments are made known,\\
let Zion mount rejoice;\\
Of Judah let the daughters all\\
send forth a cheerful voice.

12 Walk about Zion, and go round;\\
the high tow’rs thereof tell:\\
13 Consider ye her palaces,\\
and mark her bulwarks well;

That ye may tell posterity.\\
14 For this God doth abide\\
Our God for evermore; he will\\
ev’n unto death us guide.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 49}
\addcontentsline{toc}{subsection}{Psalm 49}

To the chief Musician,
A Psalm for the sons of Korah.

1 Hear this, all people, and give ear,\\
all in the world that dwell;\\
2 Both low and high, both rich and poor.\\
3 My mouth shall wisdom tell:

My heart shall knowledge meditate.\\
4 I will incline mine ear\\
To parables, and on the harp\\
my sayings dark declare.

5 Amidst those days that evil be,\\
why should I, fearing, doubt?\\
When of my heels th’ iniquity\\
shall compass me about.

6 Whoe’er they be that in their wealth\\
their confidence do pitch,\\
And boast themselves, because they are\\
become exceeding rich:

7 Yet none of these his brother can\\
redeem by any way;\\
Nor can he unto God for him\\
sufficient ransom pay,

8 (Their soul’s redemption precious is,\\
and it can never be,)\\
9 That still he should for ever live,\\
and not corruption see.

10 For why? he seeth that wise men die,\\
and brutish fools also\\
Do perish; and their wealth, when dead,\\
to others they let go.

11 Their inward thought is, that their house\\
and dwelling-places shall\\
Stand through all ages; they their lands\\
by their own names do call.

12 But yet in honor shall not man\\
abide continually;\\
But passing hence, may be compared\\
unto the beasts that die.

13 Thus brutish folly plainly is\\
their wisdom and their way;\\
Yet their posterity approve\\
what they do fondly say.

14 Like sheep they in the grave are laid,\\
and death shall them devour;\\
And in the morning upright men\\
shall over them have pow’r:

Their beauty from their dwelling shall\\
consume within the grave.\\
15 But from hell’s hand God will me free,\\
for he shall me receive.

16 Be thou not then afraid when one\\
enriched thou dost see,\\
Nor when the glory of his house\\
advanced is on high:

17 For he shall carry nothing hence\\
when death his days doth end;\\
Nor shall his glory after him\\
into the grave descend.

18 Although he his own soul did bless\\
whilst he on earth did live;\\
(And when thou to thyself dost well,\\
men will thee praises give;)

19 He to his fathers’ race shall go,\\
they never shall see light.\\
20 Man honored wanting knowledge is\\
like beasts that perish quite.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 50}
\addcontentsline{toc}{subsection}{Psalm 50}

\emph{First Version (S.M., 6.6.8.6)}

A Psalm of Asaph.

1 The mighty God, the Lord,\\
hath spoken, and did call\\
The earth, from rising of the sun,\\
to where he hath his fall.

2 From out of Zion hill,\\
which of excellency\\
And beauty the perfection is,\\
God shined gloriously.

3 Our God shall surely come,\\
keep silence shall not he:\\
Before him fire shall waste, great storms\\
shall round about him be.

4 Unto the heavens clear\\
he from above shall call,\\
And to the earth likewise, that he\\
may judge his people all.

5 Together let my saints\\
unto me gathered be,\\
Those that by sacrifice have made\\
a covenant with me.

6 And then the heavens shall\\
his righteousness declare:\\
Because the Lord himself is he\\
by whom men judged are.

7 My people Isr’el hear,\\
speak will I from on high,\\
Against thee I will testify;\\
God, ev’n thy God, am I.

8 I for thy sacrifice\\
no blame will on thee lay,\\
Nor for burnt-off’rings, which to me\\
thou offer’dst ev’ry day.

9 I’ll take no calf nor goats\\
from house or fold of thine:\\
10 For beasts of forests, cattle all\\
on thousand hills, are mine.

11 The fowls on mountains high\\
are all to me well known;\\
Wild beasts which in the fields do lie,\\
ev’n they are all mine own.

12 Then, if I hungry were,\\
I would not tell it thee;\\
Because the world, and fullness all\\
thereof, belongs to me.

13 Will I eat flesh of bulls?\\
or goats’ blood drink will I?\\
14 Thanks offer thou to God, and pay\\
thy vows to the most High.

15 And call upon me when\\
in trouble thou shalt be;\\
I will deliver thee, and thou\\
my name shalt glorify.

16 But to the wicked man\\
God saith, My laws and truth\\
Should’st thou declare? how dar’st thou take\\
my cov’nant in thy mouth?

17 Since thou instruction hat’st,\\
which should thy ways direct;\\
And since my words behind thy back\\
thou cast’st, and dost reject.

18 When thou a thief didst see,\\
with him thou didst consent;\\
And with the vile adulterers\\
partaker on thou went.

19 Thou giv’st thy mouth to ill,\\
thy tongue deceit doth frame;\\
20 Thou sitt’st, and ‘gainst thy brother speak’st,\\
thy mother’s son dost shame.

21 Because I silence kept,\\
while thou these things hast wrought;\\
That I was altogether like\\
thyself, hath been thy thought;

Yet I will thee reprove,\\
and set before thine eyes,\\
In order ranked, thy misdeeds\\
and thine iniquities.

22 Now, ye that God forget,\\
this carefully consider;\\
Lest I in pieces tear you all,\\
and none can you deliver.

23 Whoso doth offer praise\\
me glorifies; and I\\
Will show him God’s salvation,\\
that orders right his way.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (C.M., 8.6.8.6)}

A Psalm of Asaph.

1 The mighty God, the Lord, hath spoke,\\
and called the earth upon,\\
Ev’n from the rising of the sun\\
unto his going down.

2 From out of Zion, his own hill,\\
where the perfection high\\
Of beauty is, from thence the Lord\\
hath shined gloriously.

3 Our God shall come, and shall no more\\
be silent, but speak out:\\
Before him fire shall waste, great storms\\
shall compass him about.

4 He to the heavens from above,\\
and to the earth below,\\
Shall call, that he his judgments may\\
before his people show.

5 Let all my saints together be\\
unto me gathered;\\
Those that by sacrifice with me\\
a covenant have made.

6 And then the heavens shall declare\\
his righteousness abroad:\\
Because the Lord himself doth come;\\
none else is judge but God.

7 Hear, O my people, and I’ll speak;\\
O Israel by name,\\
Against thee I will testify;\\
God, ev’n thy God, I am.

8 I for thy sacrifices few\\
reprove thee never will,\\
Nor for burnt-off’rings to have been\\
before me offered still.

9 I’ll take no bullock nor he-goats\\
from house nor folds of thine:\\
10 For beasts of forests, cattle all\\
on thousand hills, are mine.

11 The fowls are all to me well known\\
that mountains high do yield;\\
And I do challenge as mine own\\
the wild beasts of the field.

12 If I were hungry, I would not\\
to thee for need complain;\\
For earth, and all its fullness, doth\\
to me of right pertain.

13 That I to eat the flesh of bulls\\
take pleasure dost thou think?\\
Or that I need, to quench my thirst,\\
the blood of goats to drink?

14 Nay, rather unto me, thy God,\\
thanksgiving offer thou:\\
To the most High perform thy word,\\
and fully pay thy vow:

15 And in the day of trouble great\\
see that thou call on me;\\
I will deliver thee, and thou\\
my name shalt glorify.

16 But God unto the wicked saith,\\
Why should’st thou mention make\\
Of my commands? how dar’st thou in\\
thy mouth my cov’nant take?

17 Since it is so that thou dost hate\\
all good instruction;\\
And since thou cast’st behind thy back,\\
and slight’st my words each one.

18 When thou a thief didst see, then straight\\
thou join’dst with him in sin,\\
And with the vile adulterers\\
thou hast partaker been.

19 Thy mouth to evil thou dost give,\\
thy tongue deceit doth frame.\\
20 Thou sitt’st, and ‘gainst thy brother speak’st,\\
thy mother’s son to shame.

21 These things thou wickedly hast done,\\
and I have silent been:\\
Thou thought’st that I was like thyself,\\
and did approve thy sin:

But I will sharply thee reprove,\\
and I will order right\\
Thy sins and thy transgressions\\
in presence of thy sight.

22 Consider this, and be afraid,\\
ye that forget the Lord,\\
Lest I in pieces tear you all,\\
when none can help afford.

23 Who off’reth praise me glorifies:\\
I will shew God’s salvation\\
To him that ordereth aright\\
his life and conversation.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 51}
\addcontentsline{toc}{subsection}{Psalm 51}

To the chief Musician, A Psalm of David,
when Nathan the prophet came unto him, after he had gone into Bathsheba.

1 After thy loving-kindness, Lord,\\
have mercy upon me:\\
For thy compassions great, blot out\\
all mine iniquity.

2 Me cleanse from sin, and thoroughly wash\\
from mine iniquity:\\
3 For my transgressions I confess;\\
my sin I ever see.

4 ‘Gainst thee, thee only, have I sinned,\\
in thy sight done this ill;\\
That when thou speak’st thou may’st be just,\\
and clear in judging still.

5 Behold, I in iniquity\\
was formed the womb within;\\
My mother also me conceived\\
in guiltiness and sin.

6 Behold, thou in the inward parts\\
with truth delighted art;\\
And wisdom thou shalt make me know\\
within the hidden part.

7 Do thou with hyssop sprinkle me,\\
I shall be cleansed so;\\
Yea, wash thou me, and then I shall\\
be whiter than the snow.

8 Of gladness and of joyfulness\\
make me to hear the voice;\\
That so these very bones which thou\\
hast broken may rejoice.

9 All mine iniquities blot out,\\
thy face hide from my sin.\\
10 Create a clean heart, Lord, renew\\
a right sp’rit me within.

11 Cast me not from thy sight, nor take\\
thy Holy Sp’rit away.\\
12 Restore me thy salvation’s joy;\\
with thy free Sp’rit me stay.

13 Then will I teach thy ways unto\\
those that transgressors be;\\
And those that sinners are shall then\\
be turned unto thee.

14 O God, of my salvation God,\\
me from blood-guiltiness\\
Set free; then shall my tongue aloud\\
sing of thy righteousness.

15 My closed lips, O Lord, by thee\\
let them be opened;\\
Then shall thy praises by my mouth\\
abroad be published.

16 For thou desir’st not sacrifice,\\
else would I give it thee;\\
Nor wilt thou with burnt-offering\\
at all delighted be.

17 A broken spirit is to God\\
a pleasing sacrifice:\\
A broken and a contrite heart,\\
Lord, thou wilt not despise.

18 Show kindness, and do good, O Lord,\\
to Zion, thine own hill:\\
The walls of thy Jerusalem\\
build up of thy good will.

19 Then righteous off’rings shall thee please,\\
and off’rings burnt, which they\\
With whole burnt-off’rings, and with calves,\\
shall on thine altar lay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 52}
\addcontentsline{toc}{subsection}{Psalm 52}

To the chief Musician, Maschil, \emph{A Psalm} of David,
when Doeg the Edomite came and told Saul, and said unto him, David is come to the house of Ahimelech.

1 Why dost thou boast, O mighty man,\\
of mischief and of ill?\\
The goodness of Almighty God\\
endureth ever still.

2 Thy tongue mischievous calumnies\\
deviseth subtilely,\\
Like to a razor sharp to cut,\\
working deceitfully.

3 Ill more than good, and more than truth\\
thou lovest to speak wrong:\hfill \emph{Selah}\\
4 Thou lovest all-devouring words,\\
O thou deceitful tongue.

5 So God shall thee destroy for aye,\\
remove thee, pluck thee out\\
Quite from thy house, out of the land\\
of life he shall thee root.\hfill\emph{Selah}

6 The righteous shall it see, and fear,\\
and laugh at him they shall:\\
7 Lo, this the man is that did not\\
make God his strength at all:

But he in his abundant wealth\\
his confidence did place;\\
And he took strength unto himself\\
from his own wickedness.

8 But I am in the house of God\\
like to an olive green:\\
My confidence for ever hath\\
upon God’s mercy been.

9 And I for ever will thee praise,\\
because thou hast done this:\\
I on thy name will wait; for good\\
before thy saints it is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 53}
\addcontentsline{toc}{subsection}{Psalm 53}

To the chief Musician upon Mahalath, Maschil,
\emph{A Psalm} of David.

1 That there is not a God, the fool\\
doth in his heart conclude:\\
They are corrupt, their works are vile,\\
not one of them doth good.

2 The Lord upon the sons of men\\
from heav’n did cast his eyes,\\
To see if any one there was\\
that sought God, and was wise.

3 They altogether filthy are,\\
they all are backward gone;\\
And there is none that doeth good,\\
no, not so much as one.

4 These workers of iniquity,\\
do they not know at all,\\
That they my people eat as bread,\\
and on God do not call?

5 Ev’n there they were afraid, and stood\\
with trembling, all dismayed,\\
Whereas there was no cause at all\\
why they should be afraid:

For God his bones that thee besieged\\
hath scattered all abroad;\\
Thou hast confounded them, for they\\
despised are of God.

6 Let Isr’el’s help from Zion come:\\
when back the Lord shall bring\\
His captives, Jacob shall rejoice,\\
and Israel shall sing.


% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 54}
\addcontentsline{toc}{subsection}{Psalm 54}

To the chief Musician on Neginoth, Maschil, \emph{A Psalm} of David, when the Ziphims came and said to Saul, Doth not David hide himself with us?

1 Save me, O God, by thy great name,\\
and judge me by thy strength:\\
2 My prayer hear, O God; give ear\\
unto my words at length.

3 For they that strangers are to me\\
do up against me rise;\\
Oppressors seek my soul, and God\\
set not before their eyes.

4 The Lord my God my helper is,\\
lo, therefore I am bold:\\
He taketh part with ev’ry one\\
that doth my soul uphold.

5 Unto mine enemies he shall\\
mischief and ill repay:\\
O for thy truth’s sake cut them off,\\
and sweep them clean away.

6 I will a sacrifice to thee\\
give with free willingness;\\
Thy name, O Lord, because ’tis good,\\
with praise I will confess.

7 For he hath me delivered\\
from all adversities;\\
And his desire mine eye hath seen\\
upon mine enemies.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 55}
\addcontentsline{toc}{subsection}{Psalm 55}

To the chief Musician on Neginoth, Maschil,
\emph{A Psalm} of David.

1 Lord, hear my pray’r, hide not thyself\\
from my entreating voice:\\
2 Attend and hear me; in my plaint\\
I mourn and make a noise.

3 Because of th’ en’my’s voice, and for\\
lewd men’s oppression great:\\
On me they cast iniquity,\\
and they in wrath me hate.

4 Sore pained within me is my heart:\\
death’s terrors on me fall.\\
5 On me comes trembling, fear and dread\\
o’erwhelmed me withal.

6 O that I, like a dove, had wings,\\
said I, then would I flee\\
Far hence, that I might find a place\\
where I in rest might be.

7 Lo, then far off I wander would,\\
and in the desert stay;\\
8 From windy storm and tempest I\\
would haste to ‘scape away.

9 O Lord, on them destruction bring,\\
and do their tongues divide;\\
For in the city violence\\
and strife I have espied.

10 They day and night upon the walls\\
do go about it round:\\
There mischief is, and sorrow there\\
in midst of it is found.

11 Abundant wickedness there is\\
within her inward part;\\
And from her streets deceitfulness\\
and guile do not depart.

12 He was no foe that me reproached,\\
then that endure I could;\\
Nor hater that did ‘gainst me boast,\\
from him me hide I would.

13 But thou, man, who mine equal, guide,\\
and mine acquaintance wast:\\
14 We joined sweet counsels, to God’s house\\
in company we past.

15 Let death upon them seize, and down\\
let them go quick to hell;\\
For wickedness doth much abound\\
among them where they dwell.

16 I’ll call on God: God will me save.\\
17 I’ll pray, and make a noise\\
At ev’ning, morning, and at noon;\\
and he shall hear my voice.

18 He hath my soul delivered,\\
that it in peace might be\\
From battle that against me was;\\
for many were with me.

19 The Lord shall hear, and them afflict,\\
of old who hath abode:\\
Because they never changes have,\\
therefore they fear not God.

20 ‘Gainst those that were at peace with him\\
he hath put forth his hand:\\
The covenant that he had made,\\
by breaking he profaned.

21 More smooth than butter were his words,\\
while in his heart was war;\\
His speeches were more soft than oil,\\
and yet drawn swords they are.

22 Cast thou thy burden on the Lord,\\
and he shall thee sustain;\\
Yea, he shall cause the righteous man\\
unmoved to remain.

23 But thou, O Lord my God, those men\\
in justice shalt o’erthrow,\\
And in destruction’s dungeon dark\\
at last shalt lay them low:

The bloody and deceitful men\\
shall not live half their days:\\
But upon thee with confidence\\
I will depend always.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 56}
\addcontentsline{toc}{subsection}{Psalm 56}


To the chief Musician upon Jonath-elem-rechokim, Michtam of David, when the Philistines took him in Gath.

1 Show mercy, Lord, to me, for man\\
would swallow me outright;\\
He me oppresseth, while he doth\\
against me daily fight.

2 They daily would me swallow up\\
that hate me spitefully;\\
For they be many that do fight\\
against me, O most High.

3 When I’m afraid I’ll trust in thee:\\
4 In God I’ll praise his word;\\
I will not fear what flesh can do,\\
my trust is in the Lord.

5 Each day they wrest my words; their thoughts\\
‘gainst me are all for ill.\\
6 They meet, they lurk, they mark my steps,\\
waiting my soul to kill.

7 But shall they by iniquity\\
escape thy judgments so?\\
O God, with indignation down\\
do thou the people throw.

8 My wand’rings all what they have been\\
thou know’st, their number took;\\
Into thy bottle put my tears:\\
are they not in thy book?

9 My foes shall, when I cry, turn back;\\
I know’t, God is for me.\\
10 In God his word I’ll praise; his word\\
in God shall praised be.

11 In God I trust; I will not fear\\
what man can do to me.\\
12 Thy vows upon me are, O God:\\
I’ll render praise to thee.

13 Wilt thou not, who from death me saved,\\
my feet from falls keep free,\\
To walk before God in the light\\
of those that living be?

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 57}
\addcontentsline{toc}{subsection}{Psalm 57}

To the chief Musician, Al-taschith, Michtam
of David, when he fled from Saul in the cave.

1 Be merciful to me, O God;\\
thy mercy unto me\\
Do thou extend; because my soul\\
doth put her trust in thee:

Yea, in the shadow of thy wings\\
my refuge I will place,\\
Until these sad calamities\\
do wholly overpass.

2 My cry I will cause to ascend\\
unto the Lord most high;\\
To God, who doth all things for me\\
perform most perfectly.

3 From heav’n he shall send down, and me\\
from his reproach defend\\
That would devour me: God his truth\\
and mercy forth shall send.

4 My soul among fierce lions is,\\
I firebrands live among,\\
Men’s sons, whose teeth are spears and darts,\\
a sharp sword is their tongue.

5 Be thou exalted very high\\
above the heav’ns, O God;\\
Let thou thy glory be advanced\\
o’er all the earth abroad.

6 My soul’s bowed down; for they a net\\
have laid, my steps to snare:\\
Into the pit which they have digged\\
for me, they fallen are.

7 My heart is fixed, my heart is fixed,\\
O God; I’ll sing and praise.\\
8 My glory wake; wake psalt’ry, harp;\\
myself I’ll early raise.

9 I’ll praise thee ‘mong the people, Lord;\\
‘mong nations sing will I:\\
10 For great to heav’n thy mercy is,\\
thy truth is to the sky.

11 O Lord, exalted be thy name\\
above the heav’ns to stand:\\
Do thou thy glory far advance\\
above both sea and land.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 58}
\addcontentsline{toc}{subsection}{Psalm 58}

To the chief Musician, Al-taschith, Michtam of David.

1 Do ye, O congregation,\\
indeed speak righteousness?\\
O ye that are the sons of men,\\
judge ye with uprightness?

2 Yea, ev’n within your very hearts\\
ye wickedness have done;\\
And ye the vi’lence of your hands\\
do weigh the earth upon.

3 The wicked men estranged are,\\
ev’n from the very womb;\\
They, speaking lies, do stray as soon\\
as to the world they come.

4 Unto a serpent’s poison like\\
their poison doth appear;\\
Yea, they are like the adder deaf,\\
that closely stops her ear;

5 That so she may not hear the voice\\
of one that charm her would,\\
No, not though he most cunning were,\\
and charm most wisely could.

6 Their teeth, O God, within their mouth\\
break thou in pieces small;\\
The great teeth break thou out, O Lord,\\
of these young lions all.

7 Let them like waters melt away,\\
which downward still do flow:\\
In pieces cut his arrows all,\\
when he shall bend his bow.

8 Like to a snail that melts away,\\
let each of them be gone;\\
Like woman’s birth untimely, that\\
they never see the sun.

9 He shall them take away before\\
your pots the thorns can find,\\
Both living, and in fury great,\\
as with a stormy wind.

10 The righteous, when he vengeance sees,\\
he shall be joyful then;\\
The righteous one shall wash his feet\\
in blood of wicked men.

11 So men shall say, The righteous man\\
reward shall never miss:\\
And verily upon the earth\\
a God to judge there is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 59}
\addcontentsline{toc}{subsection}{Psalm 59}

To the chief Musician, Al-taschith, Michtam of David;
when Saul sent, and they watched the house to kill him.

1 My God, deliver me from those\\
that are mine enemies;\\
And do thou me defend from those\\
that up against me rise.

2 Do thou deliver me from them\\
that work iniquity;\\
And give me safety from the men\\
of bloody cruelty.

3 For, lo, they for my soul lay wait:\\
the mighty do combine\\
Against me, Lord; not for my fault,\\
nor any sin of mine.

4 They run, and, without fault in me,\\
themselves do ready make:\\
Awake to meet me with thy help;\\
and do thou notice take.

5 Awake therefore, Lord God of hosts,\\
thou God of Israel,\\
To visit heathen all: spare none\\
that wickedly rebel.

6 At ev’ning they go to and fro;\\
they make great noise and sound,\\
Like to a dog, and often walk\\
about the city round.

7 Behold, they belch out with their mouth,\\
and in their lips are swords:\\
For they do say thus, Who is he\\
that now doth hear our words?

8 But thou, O Lord, shalt laugh at them,\\
and all the heathen mock.\\
9 While he’s in pow’r I’ll wait on thee;\\
for God is my high rock.

10 He of my mercy that is God\\
betimes shall me prevent;\\
Upon mine en’mies God shall let\\
me see mine heart’s content.

11 Them slay not, lest my folk forget;\\
but scatter them abroad\\
By thy strong pow’r; and bring them down,\\
O thou our shield and God.

12 For their mouth’s sin, and for the words\\
that from their lips do fly,\\
Let them be taken in their pride;\\
because they curse and lie.

13 In wrath consume them, them consume,\\
that so they may not be:\\
And that in Jacob God doth rule\\
to th’ earth’s ends let them see.

14 At ev’ning let thou them return,\\
making great noise and sound,\\
Like to a dog, and often walk\\
about the city round.

15 And let them wander up and down,\\
in seeking food to eat;\\
And let them grudge when they shall not\\
be satisfied with meat.

16 But of thy pow’r I’ll sing aloud;\\
at morn thy mercy praise:\\
For thou to me my refuge wast,\\
and tow’r, in troublous days.

17 O God, thou art my strength, I will\\
sing praises unto thee;\\
For God is my defense, a God\\
of mercy unto me.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 60}
\addcontentsline{toc}{subsection}{Psalm 60}
To the chief Musician upon Shushan-eduth, Michtam of David, to teach; when he strove with Aram-naharaim, and with Aram-zobah, when Joab returned and smote of Edom, in the valley of Salt, twelve thousand.

1 O Lord, thou hast rejected us,\\
and scattered us abroad;\\
Thou justly hast displeased been;\\
return to us, O God.

2 The earth to tremble thou hast made;\\
therein didst breaches make:\\
Do thou thereof the breaches heal,\\
because the land doth shake.

3 Unto thy people thou hard things\\
hast showed, and on them sent;\\
And thou hast caused us to drink\\
wine of astonishment.

4 And yet a banner thou hast giv’n\\
to them who thee do fear;\\
That it by them, because of truth,\\
displayed may appear.

5 That thy beloved people may\\
delivered be from thrall,\\
Save with the pow’r of thy right hand,\\
and hear me when I call.

6 God in his holiness hath spoke;\\
herein I will take pleasure:\\
Shechem I will divide, and forth\\
will Succoth’s valley measure.

7 Gilead I claim as mine by right;\\
Manasseh mine shall be;\\
Ephraim is of mine head the strength;\\
Judah gives laws for me;

8 Moab’s my washing-pot; my shoe\\
I’ll over Edom throw;\\
And over Palestina’s land\\
I will in triumph go.

9 O who is he will bring me to\\
the city fortified?\\
O who is he that to the land\\
of Edom will me guide?

10 O God, which hadest us cast off,\\
this thing wilt thou not do?\\
Ev’n thou, O God, which didest not\\
forth with our armies go?

11 Help us from trouble; for the help\\
is vain which man supplies.\\
12 Through God we’ll do great acts; he shall\\
tread down our enemies.



% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 61}
\addcontentsline{toc}{subsection}{Psalm 61}

To the chief Musician upon Neginoth,
\emph{A Psalm} of David.

1 O God, give ear unto my cry;\\
unto my pray’r attend.\\
2 From th’ utmost corner of the land\\
my cry to thee I’ll send.

What time my heart is overwhelmed,\\
and in perplexity,\\
Do thou me lead unto the Rock\\
that higher is than I.

3 For thou hast for my refuge been\\
a shelter by thy pow’r;\\
And for defense against my foes\\
thou hast been a strong tow’r.

4 Within thy tabernacle I\\
for ever will abide;\\
And under covert of thy wings\\
with confidence me hide.

5 For thou the vows that I did make,\\
O Lord my God, didst hear:\\
Thou hast giv’n me the heritage\\
of those thy name that fear.

6 A life prolonged for many days\\
thou to the king shalt give;\\
Like many generations be\\
the years which he shall live.

7 He in God’s presence his abode\\
for evermore shall have:\\
O do thou truth and mercy both\\
prepare, that may him save.

8 And so will I perpetually\\
sing praise unto thy name;\\
That having made my vows, I may\\
each day perform the same.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 62}
\addcontentsline{toc}{subsection}{Psalm 62}

To the chief Musician, to Jeduthun, A Psalm of David.

1 My soul with expectation\\
depends on God indeed;\\
My strength and my salvation doth\\
from him alone proceed.

2 He only my salvation is,\\
and my strong rock is he:\\
He only is my sure defense;\\
much moved I shall not be.

3 How long will ye against a man\\
plot mischief? ye shall all\\
Be slain; ye as a tott’ring fence\\
shall be, and bowing wall.

4 They only plot to cast him down\\
from his excellency:\\
They joy in lies; with mouth they bless,\\
but they curse inwardly.

5 My soul, wait thou with patience\\
upon thy God alone;\\
On him dependeth all my hope\\
and expectation.

6 He only my salvation is,\\
and my strong rock is he;\\
He only is my sure defense:\\
I shall not moved be.

7 In God my glory placed is,\\
and my salvation sure;\\
In God the rock is of my strength,\\
my refuge most secure.

8 Ye people, place your confidence\\
in him continually;\\
Before him pour ye out your heart:\\
God is our refuge high.

9 Surely mean men are vanity,\\
and great men are a lie;\\
In balance laid, they wholly are\\
more light than vanity.

10 Trust ye not in oppression,\\
in robb’ry be not vain;\\
On wealth set not your hearts, when as\\
increased is your gain.

11 God hath it spoken once to me,\\
yea, this I heard again,\\
That power to Almighty God\\
alone doth appertain.

12 Yea, mercy also unto thee\\
belongs, O Lord, alone:\\
For thou according to his work\\
rewardest ev’ry one.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 63}
\addcontentsline{toc}{subsection}{Psalm 63}

A Psalm of David, when he was in the wilderness of Judah.

1 Lord, thee my God, I’ll early seek:\\
my soul doth thirst for thee;\\
My flesh longs in a dry parched land,\\
wherein no waters be:

2 That I thy power may behold,\\
and brightness of thy face,\\
As I have seen thee heretofore\\
within thy holy place.

3 Since better is thy love than life,\\
my lips thee praise shall give.\\
4 I in thy name will lift my hands,\\
and bless thee while I live.

5 Ev’n as with marrow and with fat\\
my soul shall filled be;\\
Then shall my mouth with joyful lips\\
sing praises unto thee:

6 When I do thee upon my bed\\
remember with delight,\\
And when on thee I meditate\\
in watches of the night.

7 In shadow of thy wings I’ll joy;\\
for thou mine help hast been.\\
8 My soul thee follows hard; and me\\
thy right hand doth sustain.

9 Who seek my soul to spill shall sink\\
down to earth’s lowest room.\\
10 They by the sword shall be cut off,\\
and foxes’ prey become.

11 Yet shall the king in God rejoice,\\
and each one glory shall\\
That swear by him: but stopped shall be\\
the mouth of liars all.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 64}
\addcontentsline{toc}{subsection}{Psalm 64}

To the chief Musician,
A Psalm of David.

1 When I to thee my prayer make,\\
Lord, to my voice give ear;\\
My life save from the enemy,\\
of whom I stand in fear.

2 Me from their secret counsel hide\\
who do live wickedly;\\
From insurrection of those men\\
that work iniquity:

3 Who do their tongues with malice whet,\\
and make them cut like swords;\\
In whose bent bows are arrows set,\\
ev’n sharp and bitter words:

4 That they may at the perfect man\\
in secret aim their shot;\\
Yea, suddenly they dare at him\\
to shoot, and fear it not.

5 In ill encourage they themselves,\\
and their snares close do lay:\\
Together conference they have;\\
Who shall them see? they say.

6 They have searched out iniquities,\\
a perfect search they keep:\\
Of each of them the inward thought,\\
and very heart, is deep.

7 God shall an arrow shoot at them,\\
and wound them suddenly:\\
8 So their own tongue shall them confound;\\
all who them see shall fly.

9 And on all men a fear shall fall,\\
God’s works they shall declare;\\
For they shall wisely notice take\\
what these his doings are.

10 In God the righteous shall rejoice,\\
and trust upon his might;\\
Yea, they shall greatly glory all\\
in heart that are upright.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 65}
\addcontentsline{toc}{subsection}{Psalm 65}

1 Praise waits for thee in Zion, Lord:\\
to thee vows paid shall be.\\
2 O thou that hearer art of pray’r,\\
all flesh shall come to thee.

3 Iniquities, I must confess,\\
prevail against me do:\\
But as for our transgressions,\\
them purge away shalt thou.

4 Blessed is the man whom thou dost choose,\\
and mak’st approach to thee,\\
That he within thy courts, O Lord,\\
may still a dweller be:

We surely shall be satisfied\\
with thy abundant grace,\\
And with the goodness of thy house,\\
ev’n of thy holy place.

5 O God of our salvation,\\
thou, in thy righteousness,\\
By fearful works unto our pray’rs\\
thine answer dost express:

Therefore the ends of all the earth,\\
and those afar that be\\
Upon the sea, their confidence,\\
O Lord, will place in thee.

6 Who, being girt with pow’r, sets fast\\
by his great strength the hills.\\
7 Who noise of seas, noise of their waves,\\
and people’s tumult, stills.

8 Those in the utmost parts that dwell\\
are at thy signs afraid:\\
Th’ outgoings of the morn and ev’n\\
by thee are joyful made.

9 The earth thou visit’st, wat’ring it;\\
thou mak’st it rich to grow\\
With God’s full flood; thou corn prepar’st,\\
when thou provid’st it so.

10 Her rigs thou wat’rest plenteously,\\
her furrows settelest:\\
With show’rs thou dost her mollify,\\
her spring by thee is blest.

11 So thou the year most lib’rally\\
dost with thy goodness crown;\\
And all thy paths abundantly\\
on us drop fatness down.

12 They drop upon the pastures wide,\\
that do in deserts lie;\\
The little hills on ev’ry side\\
rejoice right pleasantly.

13 With flocks the pastures clothed be,\\
the vales with corn are clad;\\
And now they shout and sing to thee,\\
for thou hast made them glad.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 66}
\addcontentsline{toc}{subsection}{Psalm 66}

To the chief Musician,
A Song \emph{or} Psalm.

1 All lands to God in joyful sounds,\\
aloft your voices raise.\\
2 Sing forth the honor of his name,\\
and glorious make his praise.

3 Say unto God, How terrible\\
in all thy works art thou!\\
Through thy great pow’r thy foes to thee\\
shall be constrained to bow.

4 All on the earth shall worship thee,\\
they shall thy praise proclaim\\
In songs: they shall sing cheerfully\\
unto thy holy name.

5 Come, and the works that God hath wrought\\
with admiration see:\\
In’s working to the sons of men\\
most terrible is he.

6 Into dry land the sea he turned,\\
and they a passage had;\\
Ev’n marching through the flood on foot,\\
there we in him were glad.

7 He ruleth ever by his pow’r;\\
his eyes the nations see:\\
O let not the rebellious ones\\
lift up themselves on high.

8 Ye people, bless our God; aloud\\
the voice speak of his praise:\\
9 Our soul in life who safe preserves,\\
our foot from sliding stays.

10 For thou didst prove and try us, Lord,\\
as men do silver try;\\
11 Brought’st us into the net, and mad’st\\
bands on our loins to lie.

12 Thou hast caused men ride o’er our heads;\\
and though that we did pass\\
Through fire and water, yet thou brought’st\\
us to a wealthy place.

13 I’ll bring burnt off ‘rings to thy house;\\
to thee my vows I’ll pay,\\
14 Which my lips uttered, my mouth spake,\\
when trouble on me lay.

15 Burnt-sacrifices of fat rams\\
with incense I will bring;\\
Of bullocks and of goats I will\\
present an offering.

16 All that fear God, come, hear, I’ll tell\\
what he did for my soul.\\
17 I with my mouth unto him cried,\\
my tongue did him extol.

18 If in my heart I sin regard,\\
the Lord me will not hear:\\
19 But surely God me heard, and to\\
my prayer’s voice gave ear.

20 O let the Lord, our gracious God,\\
for ever blessed be,\\
Who turned not my pray’r from him,\\
nor yet his grace from me.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 67}
\addcontentsline{toc}{subsection}{Psalm 67}

To the chief Musician on Neginoth,
A Psalm \emph{or} Song.

\emph{First Version---Short Meter (S.M. or 6.6.8.6)}

1 Lord, bless and pity us,\\
shine on us with thy face:\\
2 That th’ earth thy way, and nations all\\
may know thy saving grace.

3 Let people praise thee, Lord;\\
let people all thee praise.\\
4 O let the nations be glad,\\
in songs their voices raise:

Thou’lt justly people judge,\\
on earth rule nations all.\\
5 Let people praise thee, Lord; let them\\
praise thee, both great and small.

6 The earth her fruit shall yield,\\
our God shall blessing send.\\
7 God shall us bless; men shall him fear\\
unto earth’s utmost end.

\emph{Second Version---Common Meter (C.M. 8.6.8.6}

1 Lord, unto us be merciful,\\
do thou us also bless;\\
And graciously cause shine on us\\
the brightness of thy face:

2 That so thy way upon the earth\\
to all men may be known;\\
Also among the nations all\\
thy saving health be shown.

3 O let the people praise thee, Lord;\\
let people all thee praise.\\
4 O let the nations be glad,\\
and sing for joy always:

For rightly thou shalt people judge,\\
and nations rule on earth.\\
5 Let people praise thee, Lord; let all\\
the folk praise thee with mirth.

6 Then shall the earth yield her increase;\\
God, our God, bless us shall.\\
7 God shall us bless; and of the earth\\
the ends shall fear him all.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 68}
\addcontentsline{toc}{subsection}{Psalm 68}

To the chief Musician,
A Psalm \emph{or} Song of David.

1 Let God arise, and scattered\\
let all his en’mies be;\\
And let all those that do him hate\\
before his presence flee.

2 As smoke is driv’n, so drive thou them;\\
as fire melts wax away,\\
Before God’s face let wicked men\\
so perish and decay.

3 But let the righteous be glad:\\
let them before God’s sight\\
Be very joyful; yea, let them\\
rejoice with all their might.

4 To God sing, to his name sing praise;\\
extol him with your voice,\\
That rides on heav’n, by his name Jah,\\
before his face rejoice.

5 Because the Lord a father is\\
unto the fatherless;\\
God is the widow’s judge, within\\
his place of holiness.

6 God doth the solitary set\\
in fam’lies: and from bands\\
The chain’d doth free; but rebels do\\
inhabit parched lands.

7 O God, what time thou didst go forth\\
before thy people’s face;\\
And when through the great wilderness\\
thy glorious marching was;

8 Then at God’s presence shook the earth,\\
then drops from heaven fell;\\
This Sinai shook before the Lord,\\
the God of Israel.

9 O God, thou to thine heritage\\
didst send a plenteous rain,\\
Whereby thou, when it weary was,\\
didst it refresh again.

10 Thy congregation then did make\\
their habitation there:\\
Of thine own goodness for the poor,\\
O God, thou didst prepare.

11 The Lord himself did give the word,\\
the word abroad did spread;\\
Great was the company of them\\
the same who published.

12 Kings of great armies foiled were,\\
and forced to flee away;\\
And women, who remained at home,\\
did distribute the prey.

13 Though ye have lien among the pots,\\
like doves ye shall appear,\\
Whose wings with silver, and with gold\\
whose feathers covered are.

14 When there th’ Almighty scatter’d kings,\\
like Salmon’s snow ’twas white.\\
15 God’s hill is like to Bashan hill,\\
like Bashan hill for height.

16 Why do ye leap, ye mountains high?\\
this is the hill where God\\
Desires to dwell; yea, God in it\\
for aye will make abode.

17 God’s chariots twenty thousand are,\\
thousands of angels strong;\\
In’s holy place God is, as in\\
mount Sinai, them among.

18 Thou hast, O Lord, most glorious,\\
ascended up on high;\\
And in triumph victorious led\\
captive captivity:

Thou hast received gifts for men,\\
for such as did rebel;\\
Yea, ev’n for them, that God the Lord\\
in midst of them might dwell.

19 Blessed be the Lord, who is to us\\
of our salvation God;\\
Who daily with his benefits\\
us plenteously doth load.

20 He of salvation is the God,\\
who is our God most strong;\\
And unto God the Lord from death\\
the issues do belong.

21 But surely God shall wound the head\\
of those that are his foes;\\
The hairy scalp of him that still\\
on in his trespass goes.

22 God said, My people I will bring\\
again from Bashan hill;\\
Yea, from the sea’s devouring depths\\
them bring again I will;

23 That in the blood of enemies\\
thy foot imbrued may be,\\
And of thy dogs dipped in the same\\
the tongues thou mayest see.

24 Thy goings they have seen, O God;\\
the steps of majesty\\
Of my God, and my mighty King,\\
within the sanctuary.

25 Before went singers, players next\\
on instruments took way;\\
And them among the damsels were\\
that did on timbrels play.

26 Within the congregations\\
bless God with one accord:\\
From Isr’el’s fountain do ye bless\\
and praise the mighty Lord.

27 With their prince, little Benjamin,\\
princes and council there\\
Of Judah were, there Zabulon’s\\
and Napht’li’s princes were.

28 Thy God commands thy strength; make strong\\
what thou wrought’st for us, Lord.\\
29 For thy house at Jerusalem\\
kings shall thee gifts afford.

30 The spearmen’s host, the multitude\\
of bulls, which fiercely look,\\
Those calves which people have forth sent,\\
O Lord our God, rebuke,

Till ev’ry one submit himself,\\
and silver pieces bring:\\
The people that delight in war\\
disperse, O God and King.

31 Those that be princes great shall then\\
come out of Egypt lands;\\
And Ethiopia to God\\
shall soon stretch out her hands.

32 O all ye kingdoms of the earth,\\
sing praises to this King;\\
For he is Lord that ruleth all,\\
unto him praises sing.

33 To him that rides on heav’ns of heav’ns,\\
which he of old did found;\\
Lo, he sends out his voice, a voice\\
in might that doth abound.

34 Strength unto God do ye ascribe;\\
for his excellency\\
Is over Israel, his strength\\
is in the clouds most high.

35 Thou’rt from thy temple dreadful, Lord;\\
Isr’el’s own God is he,\\
Who gives his people strength and pow’r:\\
O let God blessed be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}
\subsection*{Psalm 69}
\addcontentsline{toc}{subsection}{Psalm 69}

To the chief Musician upon Shoshannim,
\emph{A Psalm} of David.

1 Save me, O God, because the floods\\
do so environ me,\\
That ev’n unto my very soul\\
come in the waters be.

2 I downward in deep mire do sink,\\
where standing there is none:\\
I am into deep waters come,\\
where floods have o’er me gone.

3 I weary with my crying am,\\
my throat is also dried;\\
Mine eyes do fail, while for my God\\
I waiting do abide.

4 Those men that do without a cause\\
bear hatred unto me,\\
Than are the hairs upon my head\\
in number more they be:

They that would me destroy, and are\\
mine en’mies wrongfully,\\
Are mighty: so what I took not,\\
to render forced was I.

5 Lord, thou my folly know’st, my sins\\
not covered are from thee.\\
6 Let none that wait on thee be shamed,\\
Lord God of hosts, for me.

O Lord, the God of Israel,\\
let none, who search do make,\\
And seek thee, be at any time\\
confounded for my sake.

7 For I have borne reproach for thee,\\
my face is hid with shame.\\
8 To brethren strange, to mother’s sons\\
an alien I became.

9 Because the zeal did eat me up,\\
which to thine house I bear;\\
And the reproaches cast at thee,\\
upon me fallen are.

10 My tears and fasts, t’ afflict my soul,\\
were turned to my shame.\\
11 When sackcloth I did wear, to them\\
a proverb I became.

12 The men that in the gate do sit\\
against me evil spake;\\
They also that vile drunkards were\\
of me their song did make.

13 But, in an acceptable time,\\
my pray’r, Lord, is to thee:\\
In truth of thy salvation, Lord,\\
and mercy great, hear me.

14 Deliver me out of the mire,\\
from sinking do me keep;\\
Free me from those that do me hate,\\
and from the waters deep.

15 Let not the flood on me prevail,\\
whose water overflows;\\
Nor deep me swallow, nor the pit\\
her mouth upon me close.

16 Hear me, O Lord, because thy love\\
and kindness is most good;\\
Turn unto me, according to\\
thy mercies’ multitude.

17 Nor from thy servant hide thy face:\\
I’m troubled, soon attend.\\
18 Draw near my soul, and it redeem;\\
me from my foes defend.

19 To thee is my reproach well known,\\
my shame, and my disgrace:\\
Those that mine adversaries be\\
are all before thy face.

20 Reproach hath broke my heart; I’m full\\
of grief: I looked for one\\
To pity me, but none I found;\\
comforters found I none.

21 They also bitter gall did give\\
unto me for my meat:\\
They gave me vinegar to drink,\\
when as my thirst was great.

22 Before them let their table prove\\
a snare; and do thou make\\
Their welfare and prosperity\\
a trap themselves to take.

23 Let thou their eyes so darkened be,\\
that sight may them forsake;\\
And let their loins be made by thee\\
continually to shake.

24 Thy fury pour thou out on them,\\
and indignation;\\
And let thy wrathful anger, Lord,\\
fast hold take them upon.

25 All waste and desolate let be\\
their habitation;\\
And in their tabernacles all\\
inhabitants be none.

26 Because him they do persecute,\\
whom thou didst smite before;\\
They talk unto the grief of those\\
whom thou hast wounded sore.

27 Add thou iniquity unto\\
their former wickedness;\\
And do not let them come at all\\
into thy righteousness.

28 Out of the book of life let them\\
be razed and blotted quite;\\
Among the just and righteous\\
let not their names be writ.

29 But now become exceeding poor\\
and sorrowful am I:\\
By thy salvation, O my God,\\
let me be set on high.

30 The name of God I with a song\\
most cheerfully will praise;\\
And I, in giving thanks to him,\\
his name shall highly raise.

31 This to the Lord a sacrifice\\
more gracious shall prove\\
Than bullock, ox, or any beast\\
that hath both horn and hoof.

32 When this the humble men shall see,\\
it joy to them shall give:\\
O all ye that do seek the Lord,\\
your hearts shall ever live.

33 For God the poor hears, and will not\\
his prisoners contemn.\\
34 Let heav’n, and earth, and seas, him praise,\\
and all that move in them.

35 For God will Judah’s cities build,\\
and he will Zion save,\\
That they may dwell therein, and it\\
in sure possession have.

36 And they that are his servants’ seed\\
inherit shall the same;\\
So shall they have their dwelling there\\
that love his blessed name.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 70}
\addcontentsline{toc}{subsection}{Psalm 70}

To the chief Musician,
\emph{A Psalm} of David, to bring to remembrance.

\emph{First Version---Short Meter (S.M. 6.6.8.6)}

1 Lord, haste me to deliver;\\
with speed, Lord, succor me.\\
2 Let them that for my soul do seek\\
shamed and confounded be:

Turned back be they, and shamed,\\
that in my hurt delight.\\
3 Turned back be they, Ha, ha! that say,\\
their shaming to requite.

4 In thee let all be glad,\\
and joy that seek for thee:\\
Let them who thy salvation love\\
say still, God praised be.

5 I poor and needy am;\\
come, Lord, and make no stay:\\
My help thou and deliv’rer art;\\
O Lord, make no delay.

\emph{Second Version---Common Meter (C.M. 8.6.8.6.)}

To the chief Musician,
\emph{A Psalm} of David, to bring to remembrance.

1 Make haste, O God, me to preserve;\\
with speed, Lord, succor me.\\
2 Let them that for my soul do seek\\
shamed and confounded be:

Let them be turned back, and shamed,\\
that in my hurt delight.\\
3 Turned back be they, Ha, ha! that say,\\
their shaming to requite.

4 O Lord, in thee let all be glad,\\
and joy that seek for thee:\\
Let them who thy salvation love\\
say still, God praised be.

5 But I both poor and needy am;\\
come, Lord, and make no stay:\\
My help thou and deliv’rer art;\\
O Lord, make no delay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}
\subsection*{Psalm 71}
\addcontentsline{toc}{subsection}{Psalm 71}

1 O Lord, my hope and confidence\\
is placed in thee alone;\\
Then let thy servant never be\\
put to confusion.

2 And let me, in thy righteousness,\\
from thee deliv’rance have;\\
Cause me escape, incline thine ear\\
unto me, and me save.

3 Be thou my dwelling-rock, to which\\
I ever may resort:\\
Thou gav’st commandment me to save,\\
for thou’rt my rock and fort.

4 Free me, my God, from wicked hands,\\
hands cruel and unjust:\\
5 For thou, O Lord God, art my hope,\\
and from my youth my trust.

6 Thou from the womb didst hold me up;\\
thou art the same that me\\
Out of my mother’s bowels took;\\
I ever will praise thee.

7 To many I a wonder am;\\
but thou’rt my refuge strong.\\
8 Filled let my mouth be with thy praise\\
and honor all day long.

9 O do not cast me off, when as\\
old age doth overtake me;\\
And when my strength decayed is,\\
then do not thou forsake me.

10 For those that are mine enemies\\
against me speak with hate;\\
And they together counsel take\\
that for my soul lay wait.

11 They said, God leaves him; him pursue\\
and take: none will him save.\\
12 Be thou not far from me, my God:\\
thy speedy help I crave.

13 Confound, consume them, that unto\\
my soul are enemies:\\
Clothed be they with reproach and shame\\
that do my hurt devise.

14 But I with expectation\\
will hope continually;\\
And yet with praises more and more\\
I will thee magnify.

15 Thy justice and salvation\\
my mouth abroad shall show,\\
Ev’n all the day; for I thereof\\
the numbers do not know.

16 And I will constantly go on\\
in strength of God the Lord;\\
And thine own righteousness, ev’n thine\\
alone, I will record.

17 For even from my youth, O God,\\
by thee I have been taught;\\
And hitherto I have declared\\
the wonders thou hast wrought.

18 And now, Lord, leave me not, when I\\
old and gray-headed grow:\\
Till to this age thy strength and pow’r\\
to all to come I show.

19 And thy most perfect righteousness\\
O Lord, is very high,\\
Who hast so great things done: O God,\\
who is like unto thee?

20 Thou, Lord, who great adversities,\\
and sore, to me didst show,\\
Shalt quicken, and bring me again\\
from depths of earth below.

21 My greatness and my pow’r thou wilt\\
increase, and far extend:\\
On ev’ry side against all grief\\
thou wilt me comfort send.

22 Thee, ev’n thy truth, I’ll also praise,\\
my God, with psaltery:\\
Thou Holy One of Israel,\\
with harp I’ll sing to thee.

23 My lips shall much rejoice in thee,\\
when I thy praises sound;\\
My soul, which thou redeemed hast,\\
in joy shall much abound.

24 My tongue thy justice shall proclaim,\\
continuing all day long;\\
For they confounded are, and shamed,\\
that seek to do me wrong.


% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 72}
\addcontentsline{toc}{subsection}{Psalm 72}


\emph{A Psalm} for Solomon.

1 O Lord, thy judgments give the king,\\
his son thy righteousness.\\
2 With right he shall thy people judge,\\
thy poor with uprightness.

3 The lofty mountains shall bring forth\\
unto the people peace;\\
Likewise the little hills the same\\
shall do by righteousness.

4 The people’s poor ones he shall judge,\\
the needy’s children save;\\
And those shall he in pieces break\\
who them oppressed have.

5 They shall thee fear, while sun and moon\\
do last, through ages all.\\
6 Like rain on mown grass he shall drop,\\
or show’rs on earth that fall.

7 The just shall flourish in his days,\\
and prosper in his reign:\\
He shall, while doth the moon endure,\\
abundant peace maintain.

8 His large and great dominion shall\\
from sea to sea extend:\\
It from the river shall reach forth\\
unto earth’s utmost end.

9 They in the wilderness that dwell\\
bow down before him must;\\
And they that are his enemies\\
shall lick the very dust.

10 The kings of Tarshish, and the isles,\\
to him shall presents bring;\\
And unto him shall offer gifts\\
Sheba’s and Seba’s king.

11 Yea, all the mighty kings on earth\\
before him down shall fall;\\
And all the nations of the world\\
do service to him shall.

12 For he the needy shall preserve,\\
when he to him doth call;\\
The poor also, and him that hath\\
no help of man at all.

13 The poor man and the indigent\\
in mercy he shall spare;\\
He shall preserve alive the souls\\
of those that needy are.

14 Both from deceit and violence\\
their soul he shall set free;\\
And in his sight right precious\\
and dear their blood shall be.

15 Yea, he shall live, and giv’n to him\\
shall be of Sheba’s gold:\\
For him still shall they pray, and he\\
shall daily be extolled.

16 Of corn an handful in the earth\\
on tops of mountains high,\\
With prosp’rous fruit shall shake, like trees\\
on Lebanon that be.

The city shall be flourishing,\\
her citizens abound\\
In number shall, like to the grass\\
that grows upon the ground.

17 His name for ever shall endure;\\
last like the sun it shall:\\
Men shall be blessed in him, and blessed\\
all nations shall him call.

18 Now blessed be the Lord our God,\\
the God of Israel,\\
For he alone doth wondrous works,\\
in glory that excel.

19 And blessed be his glorious name\\
to all eternity:\\
The whole earth let his glory fill.\\
Amen, so let it be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}




