\subsection*{Psalm 107}
\addcontentsline{toc}{subsection}{Psalm 107}

 1 Praise God, for he is good: for still\\
his mercies lasting be.\\
2 Let God’s redeemed say so, whom he\\
from th’ en’my’s hand did free;

3 And gathered them out of the lands,\\
from north, south, east, and west.\\
4 They strayed in desert’s pathless way,\\
no city found to rest.

5 For thirst and hunger in them faints\\
6 their soul. When straits them press,\\
They cry unto the Lord, and he\\
them frees from their distress.

7 Them also in a way to walk\\
that right is he did guide,\\
That they might to a city go,\\
wherein they might abide.

8 O that men to the Lord would give\\
praise for his goodness then,\\
And for his works of wonder done\\
unto the sons of men!

9 For he the soul that longing is\\
doth fully satisfy;\\
With goodness he the hungry soul\\
doth fill abundantly.

10 Such as shut up in darkness deep,\\
and in death’s shade abide,\\
Whom strongly hath affliction bound,\\
and irons fast have tried:

11 Because against the words of God\\
they wrought rebelliously,\\
And they the counsel did contemn\\
of him that is most High:

12 Their heart he did bring down with grief,\\
they fell, no help could have.\\
13 In trouble then they cried to God,\\
he them from straits did save.

14 He out of darkness did them bring,\\
and from death’s shade them take;\\
These bands, wherewith they had been bound,\\
asunder quite he brake.

15 O that men to the Lord would give\\
praise for his goodness then,\\
And for his works of wonder done\\
unto the sons of men!

16 Because the mighty gates of brass\\
in pieces he did tear,\\
By him in sunder also cut\\
the bars of iron were.

17 Fools, for their sin, and their offence,\\
do sore affliction bear;\\
18 All kind of meat their soul abhors;\\
they to death’s gates draw near.

19 In grief they cry to God; he saves\\
them from their miseries.\\
20 He sends his word, them heals, and them\\
from their destructions frees.

21 O that men to the Lord would give\\
praise for his goodness then,\\
And for his works of wonder done\\
unto the sons of men!

22 And let them sacrifice to him\\
off ‘rings of thankfulness;\\
And let them shew abroad his works\\
in songs of joyfulness.

23 Who go to sea in ships, and in\\
great waters trading be,\\
24 Within the deep these men God’s works\\
and his great wonders see.

25 For he commands, and forth in haste\\
the stormy tempest flies,\\
Which makes the sea with rolling waves\\
aloft to swell and rise.

26 They mount to heav’n, then to the depths\\
they do go down again;\\
Their soul doth faint and melt away\\
with trouble and with pain.

27 They reel and stagger like one drunk,\\
at their wit’s end they be:\\
28 Then they to God in trouble cry,\\
who them from straits doth free.

29 The storm is changed into a calm\\
at his command and will;\\
So that the waves, which raged before,\\
now quiet are and still.

30 Then are they glad, because at rest\\
and quiet now they be:\\
So to the haven he them brings,\\
which they desired to see.

31 O that men to the Lord would give\\
praise for his goodness then,\\
And for his works of wonder done\\
unto the sons of men!

32 Among the people gathered\\
let them exalt his name;\\
Among assembled elders spread\\
his most renowned fame.

33 He to dry land turns water-springs,\\
and floods to wilderness;\\
34 For sins of those that dwell therein,\\
fat land to barrenness.

35 The burnt and parched wilderness\\
to water-pools he brings;\\
The ground that was dried up before\\
he turns to water-springs:

36 And there, for dwelling, he a place\\
doth to the hungry give,\\
That they a city may prepare\\
commodiously to live.

37 There sow they fields, and vineyards plant,\\
to yield fruits of increase.\\
38 His blessing makes them multiply,\\
lets not their beasts decrease.

39 Again they are diminished,\\
and very low brought down,\\
Through sorrow and affliction,\\
and great oppression.

40 He upon princes pours contempt,\\
and causeth them to stray,\\
And wander in a wilderness,\\
wherein there is no way.

41 Yet setteth he the poor on high\\
from all his miseries,\\
And he, much like unto a flock,\\
doth make him families.

42 They that are righteous shall rejoice,\\
when they the same shall see;\\
And, as ashamed, stop her mouth\\
shall all iniquity.

43 Whoso is wise, and will these things\\
observe, and them record,\\
Ev’n they shall understand the love\\
and kindness of the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 108}
\addcontentsline{toc}{subsection}{Psalm 108}

A Song \emph{or} Psalm of David.

1 My heart is fixed, Lord; I will sing,\\
and with my glory praise.\\
2 Awake up psaltery and harp;\\
myself I’ll early raise.

3 I’ll praise thee ‘mong the people, Lord;\\
‘mong nations sing will I:\\
4 For above heav’n thy mercy’s great,\\
thy truth doth reach the sky.

5 Be thou above the heavens, Lord,\\
exalted gloriously;\\
Thy glory all the earth above\\
be lifted up on high.

6 That those who thy beloved are\\
delivered may be,\\
O do thou save with thy right hand,\\
and answer give to me.

7 God in his holiness hath said,\\
Herein I will take pleasure;\\
Shechem I will divide, and forth\\
will Succoth’s valley measure.

8 Gilead I claim as mine by right;\\
Manasseh mine shall be;\\
Ephraim is of my head the strength;\\
Judah gives laws for me;

9 Moab’s my washing-pot; my shoe\\
I’ll over Edom throw;\\
Over the land of Palestine\\
I will in triumph go.

10 O who is he will bring me to\\
the city fortified?\\
O who is he that to the land\\
of Edom will me guide?

11 O God, thou who hadst cast us off,\\
this thing wilt thou not do?\\
And wilt not thou, ev’n thou, O God,\\
forth with our armies go?

12 Do thou from trouble give us help,\\
for helpless is man’s aid.\\
13 Through God we shall do valiantly;\\
our foes he shall down tread.


% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 109}
\addcontentsline{toc}{subsection}{Psalm 109}


A Song \emph{or} Psalm of David.

1 O thou the God of all my praise,\\
do thou not hold thy peace;\\
2 For mouths of wicked men to speak\\
against me do not cease:

The mouths of vile deceitful men\\
against me opened be;\\
And with a false and lying tongue\\
they have accused me.

3 They did beset me round about\\
with words of hateful spite:\\
And though to them no cause I gave,\\
against me they did fight.

4 They for my love became my foes,\\
but I me set to pray.\\
5 Evil for good, hatred for love,\\
to me they did repay.

6 Set thou the wicked over him;\\
and upon his right hand\\
Give thou his greatest enemy,\\
ev’n Satan, leave to stand.

7 And when by thee he shall be judged,\\
let him condemned be;\\
And let his pray’r be turn’d to sin,\\
when he shall call on thee.

8 Few be his days, and in his room\\
his charge another take.\\
9 His children let be fatherless,\\
his wife a widow make.

10 His children let be vagabonds,\\
and beg continually;\\
And from their places desolate\\
seek bread for their supply.

11 Let covetous extortioners\\
catch all he hath away:\\
Of all for which he laboured hath\\
let strangers make a prey.

12 Let there be none to pity him,\\
let there be none at all\\
That on his children fatherless\\
will let his mercy fall.

13 Let his posterity from earth\\
cut off for ever be,\\
And in the foll’wing age their name\\
be blotted out by thee.

14 Let God his father’s wickedness\\
still to remembrance call;\\
And never let his mother’s sin\\
be blotted out at all.

15 But let them all before the Lord\\
appear continually,\\
That he may wholly from the earth\\
cut off their memory.

16 Because he mercy minded not,\\
but persecuted still\\
The poor and needy, that he might\\
the broken-hearted kill.

17 As he in cursing pleasure took,\\
so let it to him fall;\\
As he delighted not to bless,\\
so bless him not at all.

18 As cursing he like clothes put on,\\
into his bowels so,\\
Like water, and into his bones,\\
like oil, down let it go.

19 Like to the garment let it be\\
which doth himself array,\\
And for a girdle, wherewith he\\
is girt about alway.

20 From God let this be their reward\\
that en’mies are to me,\\
And their reward that speak against\\
my soul maliciously.

21 But do thou, for thine own name’s sake,\\
O God the Lord, for me:\\
Sith good and sweet thy mercy is,\\
from trouble set me free.

22 For I am poor and indigent,\\
afflicted sore am I,\\
My heart within me also is\\
wounded exceedingly.

23 I pass like a declining shade,\\
am like the locust tossed:\\
24 My knees through fasting weakened are,\\
my flesh hath fatness lost.

25 I also am a vile reproach\\
unto them made to be;\\
And they that did upon me look\\
did shake their heads at me.

26 O do thou help and succour me,\\
who art my God and Lord:\\
And, for thy tender mercy’s sake,\\
safety to me afford:

27 That thereby they may know that this\\
is thy almighty hand;\\
And that thou, Lord, hast done the same,\\
they may well understand.

28 Although they curse with spite, yet, Lord,\\
bless thou with loving voice:\\
Let them ashamed be when they rise;\\
thy servant let rejoice.

29 Let thou mine adversaries all\\
with shame be clothed over;\\
And let their own confusion\\
them, as a mantle, cover.

30 But as for me, I with my mouth\\
will greatly praise the Lord;\\
And I among the multitude\\
his praises will record.

31 For he shall stand at his right hand\\
who is in poverty,\\
To save him from all those that would\\
condemn his soul to die.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 110}
\addcontentsline{toc}{subsection}{Psalm 110}

A Psalm of David.

1 The Lord did say unto my Lord,\\
Sit thou at my right hand,\\
Until I make thy foes a stool,\\
whereon thy feet may stand.

2 The Lord shall out of Zion send\\
the rod of thy great pow’r:\\
In midst of all thine enemies\\
be thou the governor.

3 A willing people in thy day\\
of pow’r shall come to thee,\\
In holy beauties from morn’s womb;\\
thy youth like dew shall be.

4 The Lord himself hath made an oath,\\
and will repent him never,\\
Of th’ order of Melchisedec\\
thou art a priest for ever.

5 The glorious and mighty Lord,\\
that sits at thy right hand,\\
Shall, in his day of wrath, strike through\\
kings that do him withstand.

6 He shall among the heathen judge,\\
he shall with bodies dead\\
The places fill: o’er many lands\\
he wound shall ev’ry head.

7 The brook that runneth in the way\\
with drink shall him supply;\\
And, for this cause, in triumph he\\
shall lift his head on high.


% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 111}
\addcontentsline{toc}{subsection}{Psalm 111}

1 Praise ye the Lord: with my whole heart\\
I will God’s praise declare,\\
Where the assemblies of the just\\
and congregations are.

2 The whole works of the Lord our God\\
are great above all measure,\\
Sought out they are of ev’ry one\\
that doth therein take pleasure.

3 His work most honourable is,\\
most glorious and pure,\\
And his untainted righteousness\\
for ever doth endure.

4 His works most wonderful he hath\\
made to be thought upon:\\
The Lord is gracious, and he is\\
full of compassion.

5 He giveth meat unto all those\\
that truly do him fear;\\
And evermore his covenant\\
he in his mind will bear.

6 He did the power of his works\\
unto his people show,\\
When he the heathen’s heritage\\
upon them did bestow.

7 His handy-works are truth and right;\\
all his commands are sure:\\
8 And, done in truth and uprightness,\\
they evermore endure.

9 He sent redemption to his folk;\\
his covenant for aye\\
He did command: holy his name\\
and rev’rend is alway.

10 Wisdom’s beginning is God’s fear:\\
good understanding they\\
Have all that his commands fulfill:\\
his praise endures for aye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 112}
\addcontentsline{toc}{subsection}{Psalm 112}

1 Praise ye the Lord. The man is blessed\\
that fears the Lord aright,\\
He who in his commandments\\
doth greatly take delight.

2 His seed and offspring powerful\\
shall be the earth upon:\\
Of upright men blessed shall be\\
the generation.

3 Riches and wealth shall ever be\\
within his house in store;\\
And his unspotted righteousness\\
endures for evermore.

4 Unto the upright light doth rise,\\
though he in darkness be:\\
Compassionate, and merciful,\\
and righteous, is he.

5 A good man doth his favor show,\\
and doth to others lend:\\
He with discretion his affairs\\
will guide unto the end.

6 Surely there is not any thing\\
that ever shall him move:\\
The righteous man’s memorial\\
shall everlasting prove.

7 When he shall evil tidings hear,\\
he shall not be afraid:
His heart is fixed, his confidence\\
upon the Lord is stayed.

8 His heart is firmly stablished,\\
afraid he shall not be,\\
Until upon his enemies\\
he his desire shall see.

9 He hath dispersed, giv’n to the poor;\\
his righteousness shall be\\
To ages all; with honor shall\\
his horn be raised high.

10 The wicked shall it see, and fret,\\
his teeth gnash, melt away:\\
What wicked men do most desire\\
shall utterly decay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 113}
\addcontentsline{toc}{subsection}{Psalm 113}

1 Praise God: ye servants of the Lord,\\
O praise, the Lord’s name praise.\\
2 Yea, blessed be the name of God\\
from this time forth always.

3 From rising sun to where it sets,\\
God’s name is to be praised.\\
4 Above all nations God is high,\\
‘bove heav’ns his glory raised.

5 Unto the Lord our God that dwells\\
on high, who can compare?\\
6 Himself that humbleth things to see\\
in heav’n and earth that are.

7 He from the dust doth raise the poor,\\
that very low doth lie;\\
And from the dunghill lifts the man\\
oppressed with poverty;

8 That he may highly him advance,\\
and with the princes set;\\
With those that of his people are\\
the chief, ev’n princes great.

9 The barren woman house to keep\\
he maketh, and to be\\
Of sons a mother full of joy.\\
Praise to the Lord give ye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 114}
\addcontentsline{toc}{subsection}{Psalm 114}

1 When Isr’el out of Egypt went,\\
and did his dwelling change,\\
When Jacob’s house went out from those\\
that were of language strange,

2 He Judah did his sanctuary,\\
his kingdom Isr’el make:\\
3 The sea it saw, and quickly fled,\\
Jordan was driven back.

4 Like rams the mountains, and like lambs\\
the hills skipped to and fro.\\
5 O sea, why fledd’st thou? Jordan, back\\
why wast thou driven so?

6 Ye mountains great, wherefore was it\\
that ye did skip like rams?\\
And wherefore was it, little hills,\\
that ye did leap like lambs?

7 O at the presence of the Lord,\\
earth, tremble thou for fear,\\
While as the presence of the God\\
of Jacob doth appear:

8 Who from the hard and stony rock\\
did standing water bring;\\
And by his pow’r did turn the flint\\
into a water-spring.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 115}
\addcontentsline{toc}{subsection}{Psalm 115}


1 Not unto us, Lord, not to us,\\
but do thou glory take\\
Unto thy name, ev’n for thy truth,\\
and for thy mercy’s sake.

2 O wherefore should the heathen say,\\
Where is their God now gone?\\
3 But our God in the heavens is,\\
what pleased him he hath done.

4 Their idols silver are and gold,\\
work of men’s hands they be.\\
5 Mouths have they, but they do not speak;\\
and eyes, but do not see;

6 Ears have they, but they do not hear;\\
noses, but savor not;\\
7 Hands, feet, but handle not, nor walk;\\
nor speak they through their throat.

8 Like them their makers are, and all\\
on them their trust that build.\\
9 O Isr’el, trust thou in the Lord,\\
he is their help and shield.

10 O Aaron’s house, trust in the Lord,\\
their help and shield is he.\\
11 Ye that fear God, trust in the Lord,\\
their help and shield he’ll be.

12 The Lord of us hath mindful been,\\
and he will bless us still:\\
He will the house of Isr’el bless,\\
bless Aaron’s house he will.

13 Both small and great, that fear the Lord,\\
he will them surely bless.\\
14 The Lord will you, you and your seed,\\
aye more and more increase.

15 O blessed are ye of the Lord,\\
who made the earth and heav’n.\\
16 The heav’n, ev’n heav’ns, are God’s, but he\\
earth to men’s sons hath giv’n.

17 The dead, nor who to silence go,\\
God’s praise do not record.\\
18 But henceforth we for ever will\\
bless God. Praise ye the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 116}
\addcontentsline{toc}{subsection}{Psalm 116}

1 I love the Lord, because my voice\\
and prayers he did hear.\\
2 I, while I live, will call on him,\\
who bowed to me his ear.

3 Of death the cords and sorrows did\\
about me compass round;\\
The pains of hell took hold on me,\\
I grief and trouble found.

4 Upon the name of God the Lord\\
then did I call, and say,\\
Deliver thou my soul, O Lord,\\
I do thee humbly pray.

5 God merciful and righteous is,\\
yea, gracious is our Lord.\\
6 God saves the meek: I was brought low,\\
he did me help afford.

7 O thou my soul, do thou return\\
unto thy quiet rest;\\
For largely, lo, the Lord to thee\\
his bounty hath expressed.

8 For my distressed soul from death\\
delivered was by thee:\\
Thou didst my mourning eyes from tears,\\
my feet from falling, free.

9 I in the land of those that live\\
will walk the Lord before.\\
10 I did believe, therefore I spake:\\
I was afflicted sore.

11 I said, when I was in my haste,\\
that all men liars be.\\
12 What shall I render to the Lord\\
for all his gifts to me?

13 I’ll of salvation take the cup,\\
on God’s name will I call:\\
14 I’ll pay my vows now to the Lord\\
before his people all.

15 Dear in God’s sight is his saints’ death.\\
16 Thy servant, Lord, am I;\\
Thy servant sure, thine handmaid’s son:\\
my bands thou didst untie.

17 Thank off’rings I to thee will give,\\
and on God’s name will call.\\
18 I’ll pay my vows now to the Lord\\
before his people all;

19 Within the courts of God’s own house,\\
within the midst of thee,\\
O city of Jerusalem.\\
Praise to the Lord give ye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 117}
\addcontentsline{toc}{subsection}{Psalm 117}

1 O give ye praise unto the Lord,\\
all nations that be;\\
Likewise, ye people all, accord\\
his name to magnify.

2 For great to us-ward ever are\\
his loving-kindnesses:\\
His truth endures for evermore.\\
The Lord O do ye bless.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 118}
\addcontentsline{toc}{subsection}{Psalm 118}

1 O praise the Lord, for he is good;\\
his mercy lasteth ever.\\
2 Let those of Israel now say,\\
His mercy faileth never.

3 Now let the house of Aaron say,\\
His mercy lasteth ever.\\
4 Let those that fear the Lord now say,\\
His mercy faileth never.

5 I in distress call’d on the Lord;\\
the Lord did answer me:\\
He in a large place did me set,\\
from trouble made me free.

6 The mighty Lord is on my side,\\
I will not be afraid;\\
For any thing that man can do\\
I shall not be dismay’d.

7 The Lord doth take my part with them\\
that help to succour me:\\
Therefore on those that do me hate\\
I my desire shall see.

8 Better it is to trust in God\\
than trust in man’s defence;\\
9 Better to trust in God than make\\
princes our confidence.

10 The nations, joining all in one,\\
did compass me about:\\
But in the Lord’s most holy name\\
I shall them all root out.

11 They compass’d me about; I say,\\
they compass’d me about:\\
But in the Lord’s most holy name\\
I shall them all root out.

12 Like bees they compass’d me about;\\
like unto thorns that flame\\
They quenched are: for them shall I\\
destroy in God’s own name.

13 Thou sore hast thrust, that I might fall,\\
but my Lord helped me.\\
14 God my salvation is become,\\
my strength and song is he.

15 In dwellings of the righteous\\
is heard the melody\\
Of joy and health: the Lord’s right hand\\
doth ever valiantly.

16 The right hand of the mighty Lord\\
exalted is on high;\\
The right hand of the mighty Lord\\
doth ever valiantly.

17 I shall not die, but live, and shall\\
the works of God discover.\\
18 The Lord hath me chastised sore,\\
but not to death giv’n over.

19 O set ye open unto me\\
the gates of righteousness;\\
Then will I enter into them,\\
and I the Lord will bless.

20 This is the gate of God, by it\\
the just shall enter in.\\
21 Thee will I praise, for thou me heard’st\\
and hast my safety been.

22 That stone is made head corner-stone,
which builders did despise:
23 This is the doing of the Lord,\\
and wondrous in our eyes.

24 This is the day God made, in it\\
we’ll joy triumphantly.\\
25 Save now, I pray thee, Lord; I pray,\\
send now prosperity.

26 Blessed is he in God’s great name\\
that cometh us to save:\\
We, from the house which to the Lord\\
pertains, you blessed have.

27 God is the Lord, who unto us\\
hath made light to arise:\\
Bind ye unto the altar’s horns\\
with cords the sacrifice.

28 Thou art my God, I’ll thee exalt;\\
my God, I will thee praise.\\
29 Give thanks to God, for he is good:\\
his mercy lasts always.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 119}
\addcontentsline{toc}{subsection}{Psalm 119}

\subsubsection*{Aleph}

1 Blessed are they that undefiled,\\
and straight are in the way;\\
Who in the Lord’s most holy law\\
do walk, and do not stray.

2 Blessed are they who to observe\\
his statutes are inclined;\\
And who do seek the living God\\
with their whole heart and mind.

3 Such in his ways do walk, and they\\
do no iniquity.\\
4 Thou hast commanded us to keep\\
thy precepts carefully.

5 O that thy statutes to observe\\
thou would’st my ways direct!\\
6 Then shall I not be shamed, when I\\
thy precepts all respect.

7 Then with integrity of heart\\
thee will I praise and bless,\\
When I the judgments all have learned\\
of thy pure righteousness.

8 That I will keep thy statutes all\\
firmly resolved have I:\\
O do not then, most gracious God,\\
forsake me utterly.

\subsubsection*{Beth}

9 By what means shall a young man learn\\
his way to purify?\\
If he according to thy word\\
thereto attentive be.

10 Unfeignedly thee have I sought\\
with all my soul and heart:\\
O let me not from the right path\\
of thy commands depart.

11 Thy word I in my heart have hid,\\
that I offend not thee.\\
12 O Lord, thou ever blessed art,\\
thy statutes teach thou me.

13 The judgments of thy mouth each one\\
my lips declared have:\\
14 More joy thy testimonies' way\\
than riches all me gave.

15 I will thy holy precepts make\\
my meditation;\\
And carefully I’ll have respect\\
unto thy ways each one.

16 Upon thy statutes my delight\\
shall constantly be set:\\
And, by thy grace, I never will\\
thy holy word forget.

\subsubsection*{Gimel}

17 With me thy servant, in thy grace,\\
deal bountifully, Lord;\\
That by thy favor I may live,\\
and duly keep thy word.

18 Open mine eyes, that of thy law\\
the wonders I may see.\\
19 I am a stranger on this earth,\\
hide not thy laws from me.

20 My soul within me breaks, and doth\\
much fainting still endure,\\
Through longing that it hath all times\\
unto thy judgments pure.

21 Thou hast rebuked the cursed proud,\\
who from thy precepts swerve.\\
22 Reproach and shame remove from me,\\
for I thy laws observe.

23 Against me princes spake with spite,\\
while they in council sat:\\
But I thy servant did upon\\
thy statutes meditate.

24 My comfort, and my heart’s delight,\\
thy testimonies be;\\
And they, in all my doubts and fears,\\
are counselors to me

\subsubsection*{Daleth}

25 My soul to dust cleaves: quicken me,\\
according to thy word.\\
26 My ways I showed, and me thou heard’st:\\
teach me thy statutes, Lord.

27 The way of thy commandements\\
make me aright to know;\\
So all thy works that wondrous are\\
I shall to others show.

28 My soul doth melt, and drop away,\\
for heaviness and grief:\\
To me, according to thy word,\\
give strength, and send relief.

29 From me the wicked way of lies\\
let far removed be;\\
And graciously thy holy law\\
do thou grant unto me.

30 I chosen have the perfect way\\
of truth and verity:\\
Thy judgments that most righteous are\\
before me laid have I.

31 I to thy testimonies cleave;\\
shame do not on me cast.\\
32 I’ll run thy precepts’ way, when thou\\
my heart enlarged hast.

\subsubsection*{He}

33 Teach me, O Lord, the perfect way\\
of thy precepts divine,\\
And to observe it to the end\\
I shall my heart incline.

34 Give understanding unto me,\\
so keep thy law shall I;\\
Yea, ev’n with my whole heart I shall\\
observe it carefully.

35 In thy law’s path make me to go;\\
for I delight therein.\\
36 My heart unto thy testimonies,\\
and not to greed, incline.

37 Turn thou away my sight and eyes\\
from viewing vanity;\\
And in thy good and holy way\\
be pleased to quicken me.

38 Confirm to me thy gracious word,\\
which I did gladly hear,\\
Ev’n to thy servant, Lord, who is\\
devoted to thy fear.

39 Turn thou away my fear’d reproach;\\
for good thy judgments be.\\
40 Lo, for thy precepts I have long’d;\\
in thy truth quicken me.

\subsubsection*{Vau}

41 Let thy sweet mercies also come\\
and visit me, O Lord;\\
Ev’n thy benign salvation,\\
according to thy word.

42 So shall I have wherewith I may\\
give him an answer just,\\
Who spitefully reproacheth me;\\
for in thy word I trust.

43 The word of truth out of my mouth\\
take thou not utterly;\\
For on thy judgments righteous\\
my hope doth still rely.

44 So shall I keep for evermore\\
thy law continually.\\
45 And, sith that I thy precepts seek,\\
I’ll walk at liberty.

46 I’ll speak thy word to kings, and I\\
with shame shall not be moved;\\
47 And will delight myself always\\
in thy laws, which I loved.

48 To thy commandements, which I loved,\\
my hands lift up I will;\\
And I will also meditate\\
upon thy statutes still

\subsubsection*{Zain}

49 Remember, Lord, thy gracious word\\
thou to thy servant spake,\\
Which, for a ground of my sure hope,\\
thou causedst me to take.

50 This word of thine my comfort is\\
in mine affliction:\\
For in my straits I am revived\\
by this thy word alone.

51 The men whose hearts with pride are stuffed\\
did greatly me deride;\\
Yet from thy straight commandements\\
I have not turned aside.

52 Thy judgments righteous, O Lord,\\
which thou of old forth gave,\\
I did remember, and myself\\
by them comforted have.

53 Horror took hold on me, because\\
ill men thy law forsake.\\
54 I in my house of pilgrimage\\
thy laws my songs do make.

55 Thy name by night, Lord, I did mind,\\
and I have kept thy law.\\
56 And this I had, because thy word\\
I kept, and stood in awe.

\subsubsection*{Cheth}

57 Thou my sure portion art alone,\\
which I did choose, O Lord:\\
I have resolved, and said, that I\\
would keep thy holy word.

58 With my whole heart I did entreat\\
thy face and favor free:\\
According to thy gracious word\\
be merciful to me.

59 I thought upon my former ways,\\
and did my life well try;\\
And to thy testimonies pure\\
my feet then turned I.

60 I did not stay, nor linger long,\\
as those that slothful are;\\
But hastily thy laws to keep\\
myself I did prepare.

61 Bands of ill men me robbed; yet I\\
thy precepts did not slight.\\
62 I’ll rise at midnight thee to praise,\\
ev’n for thy judgments right.

63 I am companion to all those\\
who fear, and thee obey.\\
64 O Lord, thy mercy fills the earth:\\
teach me thy laws, I pray.

\subsubsection*{Teth}

65 Well hast thou with thy servant dealt,\\
as thou didst promise give.\\
66 Good judgment me, and knowledge teach,\\
for I thy word believe.

67 Ere I afflicted was I strayed;\\
but now I keep thy word.\\
68 Both good thou art, and good thou do’st:\\
teach me thy statutes, Lord.

69 The men that are puffed up with pride\\
against me forged a lie;\\
Yet thy commandements observe\\
with my whole heart will I.

70 Their hearts, through worldly ease and wealth,\\
as fat as grease they be:\\
But in thy holy law I take\\
delight continually.

71 It hath been very good for me\\
that I afflicted was,\\
That I might well instructed be,\\
and learn thy holy laws.

72 The word that cometh from thy mouth\\
is better unto me\\
Than many thousands and great sums\\
of gold and silver be.

\subsubsection*{Jod}

73 Thou mad’st and fashion’dst me: thy laws\\
to know give wisdom, Lord.\\
74 So who thee fear shall joy to see\\
me trusting in thy word.

75 That very right thy judgments are\\
I know, and do confess;\\
And that thou hast afflicted me\\
in truth and faithfulness.

76 O let thy kindness merciful,\\
I pray thee, comfort me,\\
As to thy servant faithfully\\
was promised by thee.

77 And let thy tender mercies come\\
to me, that I may live;\\
Because thy holy laws to me\\
sweet delectation give.

78 Lord, let the proud ashamed be;\\
for they, without a cause,\\
With me perversely dealt: but I\\
will muse upon thy laws.

79 Let such as fear thee, and have known\\
thy statutes, turn to me.\\
80 My heart let in thy laws be sound,\\
that shamed I never be.


Caph

81 My soul for thy salvation faints;\\
yet I thy word believe.\\
82 Mine eyes fail for thy word: I say,\\
When wilt thou comfort give?

83 For like a bottle I’m become,\\
that in the smoke is set:\\
I’m black, and parched with grief; yet I\\
thy statutes not forget.

84 How many are thy servant’s days?\\
when wilt thou execute\\
Just judgment on these wicked men\\
that do me persecute?

85 The proud have digged pits for me,\\
which is against thy laws.\\
86 Thy words all faithful are: help me,\\
pursued without a cause.

87 They so consumed me, that on earth\\
my life they scarce did leave:\\
Thy precepts yet forsook I not,\\
but close to them did cleave.

88 After thy loving-kindness, Lord,\\
me quicken, and preserve:\\
The testimony of thy mouth\\
so shall I still observe.

\subsubsection*{Lamed}

89 Thy word for ever is, O Lord,\\
in heaven settled fast;\\
90 Unto all generations\\
thy faithfulness doth last:

The earth thou hast established,\\
and it abides by thee.\\
91 This day they stand as thou ordain’dst;\\
for all thy servants be.

92 Unless in thy most perfect law\\
my soul delights had found,\\
I should have perished, when as\\
my troubles did abound.

93 Thy precepts I will ne’er forget;\\
they quick’ning to me brought.\\
94 Lord, I am thine; O save thou me:\\
thy precepts I have sought.

95 For me the wicked have laid wait,\\
me seeking to destroy:\\
But I thy testimonies true\\
consider will with joy.

96 An end of all perfection\\
here have I seen, O God:\\
But as for thy commandment,\\
it is exceeding broad.

\subsubsection*{Mem}

97 O how love I thy law! it is\\
my study all the day:\\
98 It makes me wiser than my foes;\\
for it doth with me stay.

99 Than all my teachers now I have\\
more understanding far;\\
Because my meditation\\
thy testimonies are.

100 In understanding I excel\\
those that are ancients;\\
For I endeavored to keep\\
all thy commandements.

101 My feet from each ill way I stayed,\\
that I may keep thy word.\\
102 I from thy judgments have not swerved;\\
for thou hast taught me, Lord.

103 How sweet unto my taste, O Lord,\\
are all thy words of truth!\\
Yea, I do find them sweeter far\\
than honey to my mouth.

104 I through thy precepts, that are pure,\\
do understanding get;\\
I therefore ev’ry way that’s false\\
with all my heart do hate

\subsubsection*{Nun}

105 Thy word is to my feet a lamp,\\
and to my path a light.\\
106 I sworn have, and I will perform,\\
to keep thy judgments right.

107 I am with sore affliction\\
ev’n overwhelmed, O Lord:\\
In mercy raise and quicken me,\\
according to thy word.

108 The free-will-off ‘rings of my mouth\\
accept, I thee beseech:\\
And unto me thy servant, Lord,\\
thy judgments clearly teach.

109 Though still my soul be in my hand,\\
thy laws I’ll not forget.\\
110 I erred not from them, though for me\\
the wicked snares did set.

111 I of thy testimonies have\\
above all things made choice,\\
To be my heritage for aye;\\
for they my heart rejoice.

112 I carefully inclined have\\
my heart still to attend;\\
That I thy statutes may perform\\
alway unto the end.

\subsubsection*{Samech}

113 I hate the thoughts of vanity,\\
but love thy law do I.\\
114 My shield and hiding-place thou art:\\
I on thy word rely.

115 All ye that evil-doers are\\
from me depart away;\\
For the commandements of my God\\
I purpose to obey.

116 According to thy faithful word\\
uphold and stablish me,\\
That I may live, and of my hope\\
ashamed never be.

117 Hold thou me up, so shall I be\\
in peace and safety still;\\
And to thy statutes have respect\\
continually I will.

118 Thou tread’st down all that love to stray;\\
false their deceit doth prove.\\
119 Lewd men, like dross, away thou putt’st;\\
therefore thy law I love.

120 For fear of thee my very flesh\\
doth tremble, all dismayed;\\
And of thy righteous judgments, Lord,\\
my soul is much afraid.

\subsubsection*{Ain}

121 To all men I have judgment done,\\
performing justice right;\\
Then let me not be left unto\\
my fierce oppressors’ might.

122 For good unto thy servant, Lord,\\
thy servant’s surety be:\\
From the oppression of the proud\\
do thou deliver me.

123 Mine eyes do fail with looking long\\
for thy salvation,\\
The word of thy pure righteousness\\
while I do wait upon.

124 In mercy with thy servant deal,\\
thy laws me teach and show.\\
125 I am thy servant, wisdom give,\\
that I thy laws may know.

126 ‘Tis time thou work, Lord; for they have\\
made void thy law divine.\\
127 Therefore thy precepts more I love\\
than gold, yea, gold most fine.

128 Concerning all things thy commands\\
all right I judge therefore;\\
And ev’ry false and wicked way\\
I perfectly abhor.


\subsubsection*{Pe}

129 Thy statutes, Lord, are wonderful,\\
my soul them keeps with care.\\
130 The entrance of thy words gives light,\\
makes wise who simple are.

131 My mouth I have wide opened,\\
and panted earnestly,\\
While after thy commandements\\
I longed exceedingly.

132 Look on me, Lord, and merciful\\
do thou unto me prove,\\
As thou art wont to do to those\\
thy name who truly love.

133 O let my footsteps in thy word\\
aright still ordered be:\\
Let no iniquity obtain\\
dominion over me.

134 From man’s oppression save thou me;\\
so keep thy laws I will.\\
135 Thy face make on thy servant shine;\\
teach me thy statutes still.

136 Rivers of waters from mine eyes\\
did run down, when I saw\\
How wicked men run on in sin,\\
and do not keep thy law.

\subsubsection*{Tzaddi}

137 O Lord, thou art most righteous;\\
thy judgments are upright.\\
138 Thy testimonies thou command’st\\
most faithful are and right.

139 My zeal hath ev’n consumed me,\\
because mine enemies\\
Thy holy words forgotten have,\\
and do thy laws despise.

140 Thy word’s most pure, therefore on it\\
thy servant’s love is set.\\
141 Small, and despised I am, yet I\\
thy precepts not forget.

142 Thy righteousness is righteousness\\
which ever doth endure:\\
Thy holy law, Lord, also is\\
the very truth most pure.

143 Trouble and anguish have me found,\\
and taken hold on me:\\
Yet in my trouble my delight\\
thy just commandements be.

144 Eternal righteousness is in\\
thy testimonies all:\\
Lord, to me understanding give,\\
and ever live I shall.

\subsubsection*{Koph}

145 With my whole heart I cried, Lord, hear;\\
I will thy word obey.\\
146 I cried to thee; save me, and I\\
will keep thy laws alway.

147 I of the morning did prevent\\
the dawning, and did cry:\\
For all mine expectation\\
did on thy word rely.

148 Mine eyes did timeously prevent\\
the watches of the night,\\
That in thy word with careful mind\\
then meditate I might.

149 After thy loving-kindness hear\\
my voice, that calls on thee:\\
According to thy judgment, Lord,\\
revive and quicken me.

150 Who follow mischief they draw nigh;\\
they from thy law are far:\\
151 But thou art near, Lord; most firm truth\\
all thy commandements are.

152 As for thy testimonies all,\\
of old this have I tried,\\
That thou hast surely founded them\\
for ever to abide.

\subsubsection*{Resh}

153 Consider mine affliction,\\
in safety do me set:\\
Deliver me, O Lord, for I\\
thy law do not forget.

154 After thy word revive thou me:\\
save me, and plead my cause.\\
155 Salvation is from sinners far;\\
for they seek not thy laws.

156 O Lord, both great and manifold\\
thy tender mercies be:\\
According to thy judgments just,\\
revive and quicken me.

157 My persecutors many are,\\
and foes that do combine;\\
Yet from thy testimonies pure\\
my heart doth not decline.

158 I saw transgressors, and was grieved;\\
for they keep not thy word.\\
159 See how I love thy law! as thou\\
art kind, me quicken, Lord.

160 From the beginning all thy word\\
hath been most true and sure:\\
Thy righteous judgments ev’ry one\\
for evermore endure.

\subsubsection*{Schin}

161 Princes have persecuted me,\\
although no cause they saw:\\
But still of thy most holy word\\
my heart doth stand in awe.

162 I at thy word rejoice, as one\\
of spoil that finds great store.\\
163 Thy law I love; but lying all\\
I hate and do abhor.

164 Sev’n times a-day it is my care\\
to give due praise to thee;\\
Because of all thy judgments, Lord,\\
which righteous ever be.

165 Great peace have they who love thy law;\\
offence they shall have none.\\
166 I hoped for thy salvation, Lord,\\
and thy commands have done.

167 My soul thy testimonies pure\\
observed carefully;\\
On them my heart is set, and them\\
I love exceedingly.

168 Thy testimonies and thy laws\\
I kept with special care;\\
For all my works and ways each one\\
before thee open are.

\subsubsection*{Tau}

169 O let my earnest pray’r and cry\\
come near before thee, Lord:\\
Give understanding unto me,\\
according to thy word.

170 Let my request before thee come:\\
after thy word me free.\\
171 My lips shall utter praise, when thou\\
hast taught thy laws to me.

172 My tongue of thy most blessed word\\
shall speak, and it confess;\\
Because all thy commandements\\
are perfect righteousness.

173 Let thy strong hand make help to me:\\
thy precepts are my choice.\\
174 I longed for thy salvation, Lord,\\
and in thy law rejoice.

175 O let my soul live, and it shall\\
give praises unto thee;\\
And let thy judgments gracious\\
be helpful unto me.

176 I, like a lost sheep, went astray;\\
thy servant seek, and find:\\
For thy commands I suffered not\\
to slip out of my mind.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 120}
\addcontentsline{toc}{subsection}{Psalm 120}

A Song of degrees.

1 In my distress to God I cried,\\
and he gave ear to me.\\
2 From lying lips, and guileful tongue,\\
O Lord, my soul set free.

3 What shall be giv’n thee? or what shall\\
be done to thee, false tongue?\\
4 Ev’n burning coals of juniper,\\
sharp arrows of the strong.

5 Woe’s me that I in Mesech am\\
a sojourner so long;\\
That I in tabernacles dwell\\
to Kedar that belong.

6 My soul with him that hateth peace\\
hath long a dweller been.\\
7 I am for peace; but when I speak,\\
for battle they are keen.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 121}
\addcontentsline{toc}{subsection}{Psalm 121}

A Song of degrees.

1 I to the hills will lift mine eyes,\\
from whence doth come mine aid.\\
2 My safety cometh from the Lord,\\
who heav’n and earth hath made.

3 Thy foot he’ll not let slide, nor will\\
he slumber that thee keeps.\\
4 Behold, he that keeps Israel,\\
he slumbers not, nor sleeps.

5 The Lord thee keeps, the Lord thy shade\\
on thy right hand doth stay:\\
6 The moon by night thee shall not smite,\\
nor yet the sun by day.

7 The Lord shall keep thy soul; he shall\\
preserve thee from all ill.\\
8 Henceforth thy going out and in\\
God keep for ever will.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 122}
\addcontentsline{toc}{subsection}{Psalm 122}

A Song of degrees of David.

1 I joyed when to the house of God,\\
Go up, they said to me.\\
2 Jerusalem, within thy gates\\
our feet shall standing be.

3 Jerus’lem, as a city, is\\
compactly built together:\\
4 Unto that place the tribes go up,\\
the tribes of God go thither:

To Isr’el’s testimony, there\\
to God’s name thanks to pay.\\
5 For thrones of judgment, ev’n the thrones\\
of David’s house, there stay.

6 Pray that Jerusalem may have\\
peace and felicity:\\
Let them that love thee and thy peace\\
have still prosperity.

7 Therefore I wish that peace may still\\
within thy walls remain,\\
And ever may thy palaces\\
prosperity retain.

8 Now, for my friends’ and brethren’s sakes,\\
Peace be in thee, I’ll say.\\
9 And for the house of God our Lord,\\
I’ll seek thy good alway.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 123}
\addcontentsline{toc}{subsection}{Psalm 123}

A Song of degrees.

1 O thou that dwellest in the heav’ns,\\
I lift mine eyes to thee.\\
2 Behold, as servants’ eyes do look\\
their masters’ hand to see,

As handmaid’s eyes her mistress’ hand;\\
so do our eyes attend\\
Upon the Lord our God, until\\
to us he mercy send.

3 O Lord, be gracious unto us,\\
unto us gracious be;\\
Because replenished with contempt\\
exceedingly are we.

4 Our soul is filled with scorn of those\\
that at their ease abide,\\
And with the insolent contempt\\
of those that swell in pride.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 124}
\addcontentsline{toc}{subsection}{Psalm 124}

A Song of degrees of David.

1 Had not the Lord been on our side,\\
may Israel now say;\\
2 Had not the Lord been on our side,\\
when men rose us to slay;

3 They had us swallow’d quick, when as\\
their wrath ’gainst us did flame:\\
4 Waters had cover’d us, our soul\\
had sunk beneath the stream.

5 Then had the waters, swelling high,\\
over our soul made way.\\
6 Bless’d be the Lord, who to their teeth\\
us gave not for a prey.

7 Our soul’s escaped, as a bird\\
out of the fowler’s snare;\\
The snare asunder broken is,\\
and we escaped are.

8 Our sure and all-sufficient help\\
is in JEHOVAH’s name;\\
His name who did the heav’n create,\\
and who the earth did frame.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 125}
\addcontentsline{toc}{subsection}{Psalm 125}

A Song of degrees.

1 They in the Lord that firmly trust\\
shall be like Zion hill,\\
Which at no time can be removed,\\
but standeth ever still.

2 As round about Jerusalem\\
the mountains stand alway,\\
The Lord his folk doth compass so,\\
from henceforth and for aye.

3 For ill men’s rod upon the lot\\
of just men shall not lie;\\
Lest righteous men stretch forth their hands\\
unto iniquity.

4 Do thou to all those that be good\\
thy goodness, Lord, impart;\\
And do thou good to those that are\\
upright within their heart.

5 But as for such as turn aside\\
after their crooked way,\\
God shall lead forth with wicked men:\\
on Isr’el peace shall stay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 126}
\addcontentsline{toc}{subsection}{Psalm 126}

A Song of degrees.

1 When Zion’s bondage God turned back,\\
as men that dreamed were we.\\
2 Then filled with laughter was our mouth,\\
our tongue with melody:

They ‘mong the heathen said, The Lord\\
great things for them hath wrought.\\
3 The Lord hath done great things for us,\\
whence joy to us is brought.

4 As streams of water in the south,\\
our bondage, Lord, recall.\\
5 Who sow in tears, a reaping time\\
of joy enjoy they shall.

6 That man who, bearing precious seed,\\
in going forth doth mourn,\\
He doubtless, bringing back his sheaves,\\
rejoicing shall return.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 127}
\addcontentsline{toc}{subsection}{Psalm 127}

A Song of degrees for Solomon.

1 Except the Lord do build the house,\\
the builders lose their pain:\\
Except the Lord the city keep,\\
the watchmen watch in vain.

2 ‘Tis vain for you to rise betimes,\\
or late from rest to keep,\\
To feed on sorrows’ bread; so gives\\
he his beloved sleep.

3 Lo, children are God’s heritage,\\
the womb’s fruit his reward.\\
4 The sons of youth as arrows are,\\
for strong men’s hands prepared.

5 O happy is the man that hath\\
his quiver filled with those;\\
They unashamed in the gate\\
shall speak unto their foes.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 128}
\addcontentsline{toc}{subsection}{Psalm 128}

A Song of degrees.

1 Blessed is each one that fears the Lord,\\
and walketh in his ways;\\
2 For of thy labour thou shalt eat,\\
and happy be always.

3 Thy wife shall as a fruitful vine\\
by thy house’ sides be found:\\
Thy children like to olive-plants\\
about thy table round.

4 Behold, the man that fears the Lord,\\
thus blessed shall he be.\\
5 The Lord shall out of Zion give\\
his blessing unto thee:

Thou shalt Jerus’lem’s good behold\\
whilst thou on earth dost dwell.\\
6 Thou shalt thy children’s children see,\\
and peace on Israel.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 129}
\addcontentsline{toc}{subsection}{Psalm 129}


A Song of degrees.

1 Oft did they vex me from my youth,\\
may Isr’el now declare;\\
2 Oft did they vex me from my youth,\\
yet not victorious were.

3 The plowers plowed upon my back;\\
they long their furrows drew.\\
4 The righteous Lord did cut the cords\\
of the ungodly crew.

5 Let Zion’s haters all be turned\\
back with confusion.\\
6 As grass on houses’ tops be they,\\
which fades ere it be grown:

7 Whereof enough to fill his hand\\
the mower cannot find;\\
Nor can the man his bosom fill,\\
whose work is sheaves to bind.

8 Neither say they who do go by,\\
God’s blessing on you rest:\\
We in the name of God the Lord\\
do wish you to be blest.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 130}
\addcontentsline{toc}{subsection}{Psalm 130}

A Song of degrees.

1 Lord, from the depths to thee I cried.\\
2 My voice, Lord, do thou hear:\\
Unto my supplication’s voice\\
give an attentive ear.

3 Lord, who shall stand, if thou, O Lord,\\
should’st mark iniquity?\\
4 But yet with thee forgiveness is,\\
that feared thou mayest be.

5 I wait for God, my soul doth wait,\\
my hope is in his word.\\
6 More than they that for morning watch,\\
my soul waits for the Lord;

I say, more than they that do watch\\
the morning light to see.\\
7 Let Israel hope in the Lord,\\
for with him mercies be;

And plenteous redemption\\
is ever found with him.\\
8 And from all his iniquities\\
he Isr’el shall redeem.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 131}
\addcontentsline{toc}{subsection}{Psalm 131}

A Song of degrees of David.

1 My heart not haughty is, O Lord,\\
mine eyes not lofty be;\\
Nor do I deal in matters great,\\
or things too high for me.

2 I surely have myself behaved\\
with quiet sp’rit and mild,\\
As child of mother weaned: my soul\\
is like a weaned child.

3 Upon the Lord let all the hope\\
of Israel rely,\\
Ev’n from the time that present is\\
unto eternity.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 132}
\addcontentsline{toc}{subsection}{Psalm 132}

A Song of degrees.

1 David, and his afflictions all,\\
Lord, do thou think upon;\\
2 How unto God he sware, and vowed\\
to Jacob’s mighty One.

3 I will not come within my house,\\
nor rest in bed at all;\\
4 Nor shall mine eyes take any sleep,\\
nor eyelids slumber shall;

5 Till for the Lord a place I find,\\
where he may make abode;\\
A place of habitation\\
for Jacob’s mighty God.

6 Lo, at the place of Ephratah\\
of it we understood;\\
And we did find it in the fields,\\
and city of the wood.

7 We’ll go into his tabernacles,\\
and at his footstool bow.\\
8 Arise, O Lord, into thy rest,\\
th’ ark of thy strength, and thou.

9 O let thy priests be clothed, Lord,\\
with truth and righteousness;\\
And let all those that are thy saints\\
shout loud for joyfulness.

10 For thine own servant David’s sake,\\
do not deny thy grace;\\
Nor of thine own anointed one\\
turn thou away the face.

11 The Lord in truth to David sware,\\
he will not turn from it,\\
I of thy body’s fruit will make\\
upon thy throne to sit.

12 My cov’nant if thy sons will keep,\\
and laws to them made known,\\
Their children then shall also sit\\
for ever on thy throne.

13 For God of Zion hath made choice;\\
there he desires to dwell.\\
14 This is my rest, here still I’ll stay;\\
for I do like it well.

15 Her food I’ll greatly bless; her poor\\
with bread will satisfy.\\
16 Her priests I’ll clothe with health; her saints\\
shall shout forth joyfully.

17 And there will I make David’s horn\\
to bud forth pleasantly:\\
For him that mine anointed is\\
a lamp ordained have I.

18 As with a garment I will clothe\\
with shame his en’mies all:\\
But yet the crown that he doth wear\\
upon him flourish shall.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 133}
\addcontentsline{toc}{subsection}{Psalm 133}

A Song of degrees of David.

1 Behold, how good a thing it is,\\
and how becoming well,\\
Together such as brethren are\\
in unity to dwell!

2 Like precious ointment on the head,\\
that down the beard did flow,\\
Ev’n Aaron’s beard, and to the skirts,\\
did of his garments go.

3 As Hermon’s dew, the dew that doth\\
on Zion’ hills descend:\\
For there the blessing God commands,\\
life that shall never end.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 134}
\addcontentsline{toc}{subsection}{Psalm 134}

A Song of degrees.

1 Behold, bless ye the Lord, all ye\\
that his attendants are,\\
Ev’n you that in God’s temple be,\\
and praise him nightly there.

2 Your hands within God’s holy place\\
lift up, and praise his name.\\
3 From Zion’ hill the Lord thee bless,\\
that heav’n and earth did frame.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 135}
\addcontentsline{toc}{subsection}{Psalm 135}

1 Praise ye the Lord, the Lord’s name praise;\\
his servants, praise ye God.\\
2 Who stand in God’s house, in the courts\\
of our God make abode.

3 Praise ye the Lord, for he is good;\\
unto him praises sing:\\
Sing praises to his name, because\\
it is a pleasant thing.

4 For Jacob to himself the Lord\\
did choose of his good pleasure,\\
And he hath chosen Israel\\
for his peculiar treasure.

5 Because I know assuredly\\
the Lord is very great,\\
And that our Lord above all gods\\
in glory hath his seat.

6 What things soever pleased the Lord,\\
that in the heav’n did he,\\
And in the earth, the seas, and all\\
the places deep that be.

7 He from the ends of earth doth make\\
the vapors to ascend;\\
With rain he lightnings makes, and wind\\
doth from his treasures send.

8 Egypt’s first-born, from man to beast\\
9 who smote. Strange tokens he\\
On Phar’oh and his servants sent,\\
Egypt, in midst of thee.

10 He smote great nations, slew great kings:\\
11 Sihon of Heshbon king,\\
And Og of Bashan, and to nought\\
did Canaan’s kingdoms bring:

12 And for a wealthy heritage\\
their pleasant land he gave,\\
An heritage which Israel,\\
his chosen folk, should have.

13 Thy name, O Lord, shall still endure,\\
and thy memorial\\
With honor shall continued be\\
to generations all.

14 For why? the righteous God will judge\\
his people righteously;\\
Concerning those that do him serve,\\
himself repent will he.

15 The idols of the nations\\
of silver are and gold,\\
And by the hands of men is made\\
their fashion and mould.

16 Mouths have they, but they do not speak;\\
eyes, but they do not see;\\
17 Ears have they, but hear not; and in\\
their mouths no breathing be.

18 Their makers are like them; so are\\
all that on them rely.\\
19 O Isr’el’s house, bless God; bless God,\\
O Aaron’s family.

20 O bless the Lord, of Levi’s house\\
ye who his servants are;\\
And bless the holy name of God,\\
all ye the Lord that fear.

21 And blessed be the Lord our God\\
from Zion’s holy hill,\\
Who dwelleth at Jerusalem.\\
The Lord O praise ye still.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 136}
\addcontentsline{toc}{subsection}{Psalm 136}

\emph{First Version}

1 Give thanks to God, for good is he:\\
for mercy hath he ever.\\
2 Thanks to the God of gods give ye:\\
for his grace faileth never.

3 Thanks give the Lord of lords unto:\\
for mercy hath he ever.\\
4 Who only wonders great can do:\\
for his grace faileth never.

5 Who by his wisdom made heav’ns high:\\
for mercy hath he ever.\\
6 Who stretch’d the earth above the sea:\\
for his grace faileth never.

7 To him that made the great lights shine:\\
for mercy hath he ever.\\
8 The sun to rule till day decline:\\
for his grace faileth never.

9 The moon and stars to rule by night:\\
for mercy hath he ever.\\
10 Who Egypt’s first-born kill’d outright:\\
for his grace faileth never.

11 And Isr’el brought from Egypt land:\\
for mercy hath he ever.\\
12 With stretch’d-out arm, and with strong hand:\\
for his grace faileth never.

13 By whom the Red sea parted was:\\
for mercy hath he ever.\\
14 And through its midst made Isr’el pass:\\
for his grace faileth never.

15 But Phar’oh and his host did drown:\\
for mercy hath he ever.\\
16 Who through the desert led his own:\\
for his grace faileth never.

17 To him great kings who overthrew:\\
for he hath mercy ever.\\
18 Yea, famous kings in battle slew:\\
for his grace faileth never.

19 Ev’n Sihon king of Amorites:\\
for he hath mercy ever.\\
20 And Og the king of Bashanites:\\
for his grace faileth never.

21 Their land in heritage to have:\\
(for mercy hath he ever.)\\
22 His servant Isr’el right he gave:\\
for his grace faileth never.

23 In our low state who on us thought:\\
for he hath mercy ever.\\
24 And from our foes our freedom wrought:\\
for his grace faileth never.

25 Who doth all flesh with food relieve:\\
for he hath mercy ever.\\
26 Thanks to the God of heaven give:\\
for his grace faileth never.

\emph{Second Version}

1 Praise God, for he is kind:\\
His mercy lasts for aye.\\
2 Give thanks with heart and mind\\
To God of gods alway:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

3 The Lord of lords praise ye,\\
Whose mercies still endure.\\
4 Great wonders only he\\
Doth work by his great pow’r\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

5 Which God omnipotent,\\
By might and wisdom high,\\
The heav’n and firmament\\
Did frame, as we may see:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

6 To him who did outstretch\\
This earth so great and wide,\\
Above the waters’ reach\\
Making it to abide:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

7 Great lights he made to be;\\
For his grace lasteth aye:\\
8 Such as the sun we see,\\
To rule the lightsome day:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

9 Also the moon so clear,\\
Which shineth in our sight;\\
The stars that do appear,\\
To guide the darksome night:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

10 To him that Egypt smote,\\
Who did his message scorn;\\
And in his anger hot\\
Did kill all their first-born:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

11 Thence Isr’el out he brought;\\
For his grace lasteth ever.\\
12 With a strong hand he wrought,\\
And stretch’d-out arm deliver:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

13 The sea he cut in two;\\
For his grace lasteth still.\\
14 And through its midst to go\\
Made his own Israel:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

15 But overwhelm’d and lost\\
Was proud king Pharaoh,\\
With all his mighty host,\\
And chariots there also:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

16 To him who pow’rfully\\
His chosen people led,\\
Ev’n through the desert dry,\\
And in that place them fed:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

17 To him great kings who smote;\\
For his grace hath no bound.\\
18 Who slew, and spared not\\
Kings famous and renown’d:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

19 Sihon the Am’rites’ king;\\
For his grace lasteth ever:\\
20 Og also, who did reign\\
The land of Bashan over:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

21 Their land by lot he gave;\\
For his grace faileth never,\\
22 That Isr’el might it have\\
In heritage for ever:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

23 Who hath remembered\\
Us in our low estate;\\
24 And us delivered\\
From foes which did us hate:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

25 Who to all flesh gives food;\\
For his grace faileth never.\\
26 Give thanks to God most good,\\
The God of heav’n, for ever:\\
For certainly\\
His mercies dure\\
Most firm and sure\\
Eternally.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 137}
\addcontentsline{toc}{subsection}{Psalm 137}

1 By Babel’s streams we sat and wept,\\
when Zion we thought on.\\
2 In midst thereof we hanged our harps\\
the willow-trees upon.

3 For there a song required they,\\
who did us captive bring:\\
Our spoilers called for mirth, and said,\\
A song of Zion sing.

4 O how the Lord’s song shall we sing\\
within a foreign land?\\
5 If thee, Jerus’lem, I forget,\\
skill part from my right hand.

6 My tongue to my mouth’s roof let cleave,\\
if I do thee forget,\\
Jerusalem, and thee above\\
my chief joy do not set.

7 Remember Edom’s children, Lord,\\
who in Jerus’lems day,\\
Ev’n unto its foundation,\\
Raze, raze it quite, did say.

8 O daughter thou of Babylon,\\
near to destruction;\\
Blessed shall he be that thee rewards,\\
as thou to us hast done.

9 Yea, happy surely shall he be\\
thy tender little ones\\
Who shall lay hold upon, and them\\
shall dash against the stones.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 138}
\addcontentsline{toc}{subsection}{Psalm 138}

\emph{A Psalm} of David.

1 Thee will I praise with all my heart,\\
I will sing praise to thee\\
2 Before the gods: And worship will\\
toward thy sanctuary.

I’ll praise thy name, ev’n for thy truth,\\
and kindness of thy love;\\
For thou thy word hast magnified\\
all thy great name above.

3 Thou didst me answer in the day\\
when I to thee did cry;\\
And thou my fainting soul with strength\\
didst strengthen inwardly.

4 All kings upon the earth that are\\
shall give thee praise, O Lord,\\
When as they from thy mouth shall hear\\
thy true and faithful word.

5 Yea, in the righteous ways of God\\
with gladness they shall sing:\\
For great’s the glory of the Lord;\\
who doth for ever reign.

6 Though God be high, yet he respects\\
all those that lowly be;\\
Whereas the proud and lofty ones\\
afar off knoweth he.

7 Though I in midst of trouble walk,\\
I life from thee shall have:\\
‘Gainst my foes’ wrath thou’lt stretch thine hand;\\
thy right hand shall me save.

8 Surely that which concerneth me\\
the Lord will perfect make:\\
Lord, still thy mercy lasts; do not\\
thine own hands’ works forsake.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 139}
\addcontentsline{toc}{subsection}{Psalm 139}

To the chief Musician,
A Psalm of David.

1 O Lord, thou hast me searched and known.\\
2 Thou know’st my sitting down,\\
And rising up; yea, all my thoughts\\
afar to thee are known.

3 My footsteps, and my lying down,\\
thou compassest always;\\
Thou also most entirely art\\
acquaint with all my ways.

4 For in my tongue, before I speak,\\
not any word can be,\\
But altogether, lo, O Lord,\\
it is well known to thee.

5 Behind, before, thou hast beset,\\
and laid on me thine hand.\\
6 Such knowledge is too strange for me,\\
too high to understand.

7 From thy Sp’rit whither shall I go?\\
or from thy presence fly?\\
8 Ascend I heav’n, lo, thou art there;\\
there, if in hell I lie.

9 Take I the morning wings, and dwell\\
in utmost parts of sea;\\
10 Ev’n there, Lord, shall thy hand me lead,\\
thy right hand hold shall me.

11 If I do say that darkness shall\\
me cover from thy sight,\\
Then surely shall the very night\\
about me be as light.

12 Yea, darkness hideth not from thee,\\
but night doth shine as day:\\
To thee the darkness and the light\\
are both alike alway.

13 For thou possessed hast my reins,\\
and thou hast covered me,\\
When I within my mother’s womb\\
enclosed was by thee.

14 Thee will I praise; for fearfully\\
and strangely made I am;\\
Thy works are marv’llous, and right well\\
my soul doth know the same.

15 My substance was not hid from thee,\\
when as in secret I\\
Was made; and in earth’s lowest parts\\
was wrought most curiously.

16 Thine eyes my substance did behold,\\
yet being unperfect;\\
And in the volume of thy book\\
my members all were writ;

Which after in continuance\\
were fashioned ev’ry one,\\
When as they yet all shapeless were,\\
and of them there was none.

17 How precious also are thy thoughts,\\
O gracious God, to me!\\
And in their sum how passing great\\
and numberless they be!

18 If I should count them, than the sand\\
they more in number be:\\
What time soever I awake,\\
I ever am with thee.

19 Thou, Lord, wilt sure the wicked slay:\\
hence from me bloody men.\\
20 Thy foes against thee loudly speak,\\
and take thy name in vain.

21 Do not I hate all those, O Lord,\\
that hatred bear to thee?\\
With those that up against thee rise\\
can I but grieved be?

22 With perfect hatred them I hate,\\
my foes I them do hold.\\
23 Search me, O God, and know my heart,\\
try me, my thoughts unfold:

24 And see if any wicked way\\
there be at all in me;\\
And in thine everlasting way\\
to me a leader be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 140}
\addcontentsline{toc}{subsection}{Psalm 140}

To the chief Musician,
A Psalm of David.

1 Lord, from the ill and froward man\\
give me deliverance,\\
And do thou safe preserve me from\\
the man of violence:

2 Who in their heart mischievous things\\
are meditating ever;\\
And they for war assembled are\\
continually together.

3 Much like unto a serpent’s tongue\\
their tongues they sharp do make;\\
And underneath their lips there lies\\
the poison of a snake.

4 Lord, keep me from the wicked’s hands,\\
from vi’lent men me save;\\
Who utterly to overthrow\\
my goings purposed have.

5 The proud for me a snare have hid,\\
and cords; yea, they a net\\
Have by the way-side for me spread;\\
they gins for me have set.

6 I said unto the Lord, Thou art\\
my God: unto the cry\\
Of all my supplications,\\
Lord, do thine ear apply.\\

7 O God the Lord, who art the strength\\
of my salvation:\\
A cov’ring in the day of war\\
my head thou hast put on.

8 Unto the wicked man, O Lord,\\
his wishes do not grant;\\
Nor further thou his ill device,\\
lest they themselves should vaunt.

9 As for the head and chief of those\\
about that compass me,\\
Ev’n by the mischief of their lips\\
let thou them cover’d be.

10 Let burning coals upon them fall,\\
them throw in fiery flame,\\
And in deep pits, that they no more\\
may rise out of the same.

11 Let not an evil speaker be\\
on earth established:\\
Mischief shall hunt the vi’lent man,\\
till he be ruined.

12 I know God will th’ afflicted’s cause\\
maintain, and poor men’s right.\\
13 Surely the just shall praise thy name;\\
th’ upright dwell in thy sight

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 141}
\addcontentsline{toc}{subsection}{Psalm 141}

A Psalm of David.

1 O Lord, I unto thee do cry,\\
do thou make haste to me,\\
And give an ear unto my voice,\\
when I cry unto thee.

2 As incense let my prayer be\\
directed in thine eyes;\\
And the uplifting of my hands\\
as th’ ev ‘ning sacrifice.

3 Set, Lord, a watch before my mouth,\\
keep of my lips the door.\\
4 My heart incline thou not unto\\
the ills I should abhor,

To practice wicked works with men\\
that work iniquity;\\
And with their delicates my taste\\
let me not satisfy.

5 Let him that righteous is me smite,\\
it shall a kindness be;\\
Let him reprove, I shall it count\\
a precious oil to me:

Such smiting shall not break my head;\\
for yet the time shall fall,\\
When I in their calamities\\
to God pray for them shall.

6 When as their judges down shall be\\
in stony places cast,\\
Then shall they hear my words; for they\\
shall sweet be to their taste.

7 About the grave’s devouring mouth\\
our bones are scattered round,\\
As wood which men do cut and cleave\\
lies scattered on the ground.

8 But unto thee, O God the Lord,\\
mine eyes uplifted be:\\
My soul do not leave destitute;\\
my trust is set on thee.

9 Lord, keep me safely from the snares\\
which they for me prepare;\\
And from the subtle gins of them\\
that wicked workers are.

10 Let workers of iniquity\\
into their own nets fall,\\
Whilst I do, by thine help, escape\\
the danger of them all.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 142}
\addcontentsline{toc}{subsection}{Psalm 142}

Maschil of David; A Prayer when he was in the cave.

1 I with my voice cried to the Lord,\\
with it made my request:\\
2 Poured out to him my plaint, to him\\
my trouble I expressed.

3 When in me was o’erwhelmed my sp’rit,\\
then well thou knew’st my way;\\
Where I did walk a snare for me\\
they privily did lay.

4 I looked on my right hand, and viewed,\\
but none to know me were;\\
All refuge failed me, no man\\
did for my soul take care.

5 I cried to thee; I said, Thou art\\
my refuge, Lord, alone;\\
And in the land of those that live\\
thou art my portion.

6 Because I am brought very low,\\
attend unto my cry:\\
Me from my persecutors save,\\
who stronger are than I.

7 From prison bring my soul, that I\\
thy name may glorify:\\
The just shall compass me, when thou\\
with me deal’st bounteously.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 143}
\addcontentsline{toc}{subsection}{Psalm 143}

\emph{First Version---Common Meter (C.M. 8.6.8.6}

A Psalm of David.

1 Lord, hear my pray’r, attend my suits;\\
and in thy faithfulness\\
Give thou an answer unto me,\\
and in thy righteousness.

2 Thy servant also bring thou not\\
in judgment to be tried:\\
Because no living man can be\\
in thy sight justified.

3 For th’ en’my hath pursued my soul,\\
my life to ground down tread:\\
In darkness he hath made me dwell,\\
as who have long been dead.

4 My sp’rit is therefore overwhelmed\\
in me perplexedly;\\
Within me is my very heart\\
amazed wondrously.

5 I call to mind the days of old,\\
to meditate I use\\
On all thy works; upon the deeds\\
I of thy hands do muse.

6 My hands to thee I stretch; my soul\\
thirsts, as dry land, for thee.\\
7 Haste, Lord, to hear, my spirit fails:\\
hide not thy face from me;

Lest like to them I do become\\
that go down to the dust.\\
8 At morn let me thy kindness hear;\\
for in thee do I trust.

Teach me the way that I should walk:\\
I lift my soul to thee.\\
9 Lord, free me from my foes; I flee\\
to thee to cover me.

10 Because thou art my God, to do\\
thy will do me instruct:\\
Thy Sp’rit is good, me to the land\\
of uprightness conduct.

11 Revive and quicken me, O Lord,\\
ev’n for thine own name’s sake;\\
And do thou, for thy righteousness,\\
my soul from trouble take.

12 And of thy mercy slay my foes;\\
let all destroyed be\\
That do afflict my soul: for I\\
a servant am to thee.

\emph{Second Version (6.6.6.6 D)}

1 Oh, hear my prayer, Lord,\\
And unto my desire\\
To bow thine ear accord,\\
I humbly thee require;\\
And, in thy faithfulness,\\
Unto me answer make,\\
And, in thy righteousness,\\
Upon me pity take.

2 In judgment enter not\\
With me thy servant poor;\\
For why, this well I wot,\\
No sinner can endure\\
The sight of thee, O God:\\
If thou his deeds shalt try,\\
He dare make none abode\\
Himself to justify.

3 Behold, the cruel foe\\
Me persecutes with spite,\\
My soul to overthrow:\\
Yea, he my life down quite\\
Unto the ground hath smote,\\
And made me dwell full low\\
In darkness, as forgot,\\
Or men dead long ago.

4 Therefore my sp’rit much vexed,\\
O’erwhelmed is me within;\\
My heart right sore perplexed\\
And desolate hath been.\\
5 Yet I do call to mind\\
What ancient days record,\\
Thy works of ev’ry kind\\
I think upon, O Lord.

6 Lo, I do stretch my hands\\
To thee, my help alone;\\
For thou well understands\\
All my complaint and moan:\\
My thirsting soul desires,\\
And longeth after thee,\\
As thirsty ground requires\\
With rain refreshed to be.

7 Lord, let my pray’r prevail,\\
To answer it make speed;\\
For, lo, my sp’rit doth fail:\\
Hide not thy face in need;\\
Lest I be like to those\\
That do in darkness sit,\\
Or him that downward goes\\
Into the dreadful pit.

8 Because I trust in thee,\\
O Lord, cause me to hear\\
Thy loving-kindness free,\\
When morning doth appear:\\
Cause me to know the way\\
Wherein my path should be;\\
For why, my soul on high\\
I do lift up to thee.

9 From my fierce enemy\\
In safety do me guide,\\
Because I flee to thee,\\
Lord, that thou may’st me hide.\\
10 My God alone art thou,\\
Teach me thy righteousness:\\
Thy Sp’rit’s good, lead me to\\
The land of uprightness.

11 O Lord, for thy name’s sake,\\
Be pleased to quicken me;\\
And, for thy truth, forth take\\
My soul from misery.\\
12 And of thy grace destroy\\
My foes, and put to shame\\
All who my soul annoy;\\
For I thy servant am.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 144}
\addcontentsline{toc}{subsection}{Psalm 144}

\emph{A Psalm} of David.

1 O blessed ever be the Lord,\\
who is my strength and might,\\
Who doth instruct my hands to war,\\
my fingers teach to fight.

2 My goodness, fortress, my high tow’r,\\
deliverer, and shield,\\
In whom I trust: who under me\\
my people makes to yield.

3 Lord, what is man, that thou of him\\
dost so much knowledge take?\\
Or son of man, that thou of him\\
so great account dost make?

4 Man is like vanity; his days,\\
as shadows, pass away.\\
5 Lord, bow thy heav’ns, come down,\\
touch thou the hills, and smoke shall they.

6 Cast forth thy lightning, scatter them;\\
thine arrows shoot, them rout.\\
7 Thine hand send from above, me save;\\
from great depths draw me out;

And from the hand of children strange,\\
8 Whose mouth speaks vanity;\\
And their right hand is a right hand\\
that works deceitfully.

9 A new song I to thee will sing,\\
Lord, on a psaltery;\\
I on a ten-stringed instrument\\
will praises sing to thee.

10 Ev’n he it is that unto kings\\
salvation doth send;\\
Who his own servant David doth\\
from hurtful sword defend.

11 O free me from strange children’s hand,\\
whose mouth speaks vanity;\\
And their right hand a right hand is\\
that works deceitfully.

12 That, as the plants, our sons may be\\
in youth grown up that are;\\
Our daughters like to corner-stones,\\
carved like a palace fair.

13 That to afford all kind of store\\
our garners may be filled;\\
That our sheep thousands, in our streets\\
ten thousands they may yield.

14 That strong our oxen be for work,\\
that no in-breaking be,\\
Nor going out; and that our streets\\
may from complaints be free.

15 Those people blessed are who be\\
in such a case as this;\\
Yea, blessed all those people are,\\
whose God Jehovah is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 145}
\addcontentsline{toc}{subsection}{Psalm 145}

\emph{First Version}

David’s \emph{Psalm} of praise.

1 I’ll thee extol, my God, O King;
I’ll bless thy name always.
2 Thee will I bless each day, and will
thy name for ever praise.

3 Great is the Lord, much to be praised;
his greatness search exceeds.
4 Race unto race shall praise thy works,
and show thy mighty deeds.

5 I of thy glorious majesty
the honor will record;
I’ll speak of all thy mighty works,
which wondrous are, O Lord.

6 Men of thine acts the might shall show,
thine acts that dreadful are;
And I, thy glory to advance,
thy greatness will declare.

7 The mem’ry of thy goodness great
they largely shall express;
With songs of praise they shall extol
thy perfect righteousness.

8 The Lord is very gracious,
in him compassions flow;
In mercy he is very great,
and is to anger slow.

9 The Lord Jehovah unto all
his goodness doth declare;
And over all his other works
his tender mercies are.

10 Thee all thy works shall praise, O Lord,
and thee thy saints shall bless;
11 They shall thy kingdom’s glory show,
thy pow’r by speech express:

12 To make the sons of men to know
his acts done mightily,
And of his kingdom th’ excellent
and glorious majesty.

13 Thy kingdom shall for ever stand,
thy reign through ages all.
14 God raiseth all that are bowed down,
upholdeth all that fall.

15 The eyes of all things wait on thee,
the giver of all good;
And thou, in time convenient,
bestow’st on them their food:

16 Thine hand thou open’st lib’rally,
and of thy bounty gives
Enough to satisfy the need
of ev’ry thing that lives.

17 The Lord is just in all his ways,
holy in his works all.
18 God’s near to all that call on him,
in truth that on him call.

19 He will accomplish the desire
of those that do him fear:
He also will deliver them,
and he their cry will hear.

20 The Lord preserves all who him love,
that naught can them annoy:
But he all those that wicked are
will utterly destroy.

21 My mouth the praises of the Lord
to publish cease shall never:
Let all flesh bless his holy name
for ever and for ever.

\emph{Second Version}

David’s \emph{Psalm} of praise.

1 O Lord, thou art my God and King;
Thee will I magnify and praise:
I will thee bless, and gladly sing
Unto thy holy name always.

2 Each day I rise I will thee bless,
And praise thy name time without end.
3 Much to be praised, and great God is;
His greatness none can comprehend.

4 Race shall thy works praise unto race,
The mighty acts show done by thee.
5 I will speak of the glorious grace,
And honor of thy majesty;

Thy wondrous works I will record.
6 By men the might shall be extolled
Of all thy dreadful acts, O Lord:
And I thy greatness will unfold.

7 They utter shall abundantly
The mem’ry of thy goodness great;
And shall sing praises cheerfully,
Whilst they thy righteousness relate.

8 The Lord our God is gracious,
Compassionate is he also;
In mercy he is plenteous,
But unto wrath and anger slow.

9 Good unto all men is the Lord:
O’er all his works his mercy is.
10 Thy works all praise to thee afford:
Thy saints, O Lord, thy name shall bless.

11 The glory of thy kingdom show
Shall they, and of thy power tell:
12 That so men’s sons his deeds may know,
His kingdom’s grace that doth excel.

13 Thy kingdom hath none end at all,
It doth through ages all remain.
14 The Lord upholdeth all that fall,
The cast-down raiseth up again.

15 The eyes of all things, Lord, attend,
And on thee wait that here do live,
And thou, in season due, dost send
Sufficient food them to relieve.

16 Yea, thou thine hand dost open wide,
And ev’ry thing dost satisfy
That lives, and doth on earth abide,
Of thy great liberality.

17 The Lord is just in his ways all,
And holy in his works each one.
18 He’s near to all that on him call,
Who call in truth on him alone.

19 God will the just desire fulfill
Of such as do him fear and dread:
Their cry regard, and hear he will,
And save them in the time of need.

20 The Lord preserves all, more and less,
That bear to him a loving heart:
But workers all of wickedness
Destroy will he, and clean subvert.

21 Therefore my mouth and lips I’ll frame
To speak the praises of the Lord:
To magnify his holy name
For ever let all flesh accord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 146}
\addcontentsline{toc}{subsection}{Psalm 146}

1 Praise God. The Lord praise, O my soul.
2 I’ll praise God while I live;
While I have being to my God
in songs I’ll praises give.

3 Trust not in princes, nor man’s son,
in whom there is no stay:
4 His breath departs, to’s earth he turns;
that day his thoughts decay.

5 O happy is that man and blest,
whom Jacob’s God doth aid;
Whose hope upon the Lord doth rest,
and on his God is stayed:

6 Who made the earth and heavens high,
who made the swelling deep,
And all that is within the same;
who truth doth ever keep:

7 Who righteous judgment executes
for those oppressed that be,
Who to the hungry giveth food;
God sets the pris’ners free.

8 The Lord doth give the blind their sight,
the bowed down doth raise:
The Lord doth dearly love all those
that walk in upright ways.

9 The stranger’s shield, the widow’s stay,
the orphan’s help, is he:
But yet by him the wicked’s way
turned upside down shall be.

10 The Lord shall reign for evermore:
thy God, O Zion, he
Reigns to all generations.
Praise to the Lord give ye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 147}
\addcontentsline{toc}{subsection}{Psalm 147}

1 Praise ye the Lord; for it is good
praise to our God to sing:
For it is pleasant, and to praise
it is a comely thing.

2 God doth build up Jerusalem;
and he it is alone
That the dispersed of Israel
doth gather into one.

3 Those that are broken in their heart,
and grieved in their minds,
He healeth, and their painful wounds
he tenderly up-binds.

4 He counts the number of the stars;
he names them ev’ry one.
5 Great is our Lord, and of great pow’r;
his wisdom search can none.

6 The Lord lifts up the meek; and casts
the wicked to the ground.
7 Sing to the Lord, and give him thanks;
on harp his praises sound;

8 Who covereth the heav’n with clouds,
who for the earth below
Prepareth rain, who maketh grass
upon the mountains grow.

9 He gives the beast his food, he feeds
the ravens young that cry.
10 His pleasure not in horses’ strength,
nor in man’s legs, doth lie.

11 But in all those that do him fear
the Lord doth pleasure take;
In those that to his mercy do
by hope themselves betake.

12 The Lord praise, O Jerusalem;
Zion, thy God confess:
13 For thy gates’ bars he maketh strong;
thy sons in thee doth bless.

14 He in thy borders maketh peace;
with fine wheat filleth thee.
15 He sends forth his command on earth,
his word runs speedily.

16 Hoar-frost, like ashes, scatt’reth he;
like wool he snow doth give:
17 Like morsels casteth forth his ice;
who in its cold can live?

18 He sendeth forth his mighty word,
and melteth them again;
His wind he makes to blow, and then
the waters flow amain.

19 The doctrine of his holy word
to Jacob he doth show;
His statutes and his judgments he
gives Israel to know.

20 To any nation never he
such favor did afford;
For they his judgments have not known.
O do ye praise the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 148}
\addcontentsline{toc}{subsection}{Psalm 148}

\emph{First Version---Common (C.M. 8.6.8.6}

1 Praise God. From heavens praise the Lord,\\
in heights praise to him be.\\
2 All ye his angels, praise ye him;\\
his hosts all, praise him ye.

3 O praise ye him, both sun and moon,\\
praise him, all stars of light.\\
4 Ye heav’ns of heav’ns him praise, and floods\\
above the heavens’ height.

5 Let all the creatures praise the name\\
of our almighty Lord:\\
For he commanded, and they were\\
created by his word.

6 He also, for all times to come,\\
hath them established sure;\\
He hath appointed them a law,\\
which ever shall endure.

7 Praise ye Jehovah from the earth,\\
dragons, and ev’ry deep:\\
8 Fire, hail, snow, vapor, stormy wind,\\
his word that fully keep.

9 All hills and mountains, fruitful trees,\\
and all ye cedars high:\\
10 Beasts, and all cattle, creeping things,\\
and all ye birds that fly.

11 Kings of the earth, all nations,\\
princes, earth’s judges all:\\
12 Both young men, yea, and maidens too,\\
old men, and children small.

13 Let them God’s name praise; for his name\\
alone is excellent:\\
His glory reacheth far above\\
the earth and firmament.

14 His people’s horn, the praise of all\\
his saints, exalteth he;\\
Ev’n Isr’el’s seed, a people near\\
to him. The Lord praise ye.

\emph{Second Version---Short Meter (S.M. 6.6.8.6)}

1 The Lord of heav’n confess,\\
On high his glory raise.\\
2 Him let all angels bless,\\
Him all his armies praise.\\
3 Him glorify\\
Sun, moon, and stars;\\
4 Ye higher spheres,\\
And cloudy sky.

5 From God your beings are,\\
Him therefore famous make;\\
You all created were,\\
When he the word but spake.\\
6 And from that place,\\
Where fixed you be\\
By his decree,\\
You cannot pass.

7 Praise God from earth below,\\
Ye dragons, and ye deeps:\\
8 Fire, hail, clouds, wind, and snow.\\
Whom in command he keeps.\\
9 Praise ye his name,\\
Hills great and small,\\
Trees low and tall;\\
10 Beasts wild and tame;

All things that creep or fly.\\
11 Ye kings, ye vulgar throng,\\
All princes mean or high;\\
12 Both men and virgins young,\\
Ev’n young and old,\\
13 Exalt his name;\\
For much his fame\\
Should be extolled.

O let God’s name be praised\\
Above both earth and sky;\\
14 For he his saints hath raised,\\
And set their horn on high;\\
Ev’n those that be\\
Of Isr’el’s race,\\
Near to his grace.\\
The Lord praise ye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 149}
\addcontentsline{toc}{subsection}{Psalm 149}

1 Praise ye the Lord: unto him sing\\
a new song, and his praise\\
In the assembly of his saints\\
in sweet psalms do ye raise.

2 Let Isr’el in his Maker joy,\\
and to him praises sing:\\
Let all that Zion’s children are\\
be joyful in their King.

3 O let them unto his great name\\
give praises in the dance;\\
Let them with timbrel and with harp\\
in songs his praise advance.

4 For God doth pleasure take in those\\
that his own people be;\\
And he with his salvation\\
the meek will beautify.

5 And in his glory excellent\\
let all his saints rejoice:\\
Let them to him upon their beds\\
aloud lift up their voice.

6 Let in their mouth aloft be raised\\
the high praise of the Lord,\\
And let them have in their right hand\\
a sharp two-edged sword;

7 To execute the vengeance due\\
upon the heathen all,\\
And make deserved punishment\\
upon the people fall.

8 And ev’n with chains, as pris’ners, bind\\
their kings that them command;\\
Yea, and with iron fetters strong,\\
the nobles of their land.

9 On them the judgment to perform\\
found written in his word:\\
This honor is to all his saints.\\
O do ye praise the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 150}
\addcontentsline{toc}{subsection}{Psalm 150}

1 Praise ye the Lord. God’s praise within\\
his sanctuary raise;\\
And to him in the firmament\\
of his pow’r give ye praise.

2 Because of all his mighty acts,\\
with praise him magnify:\\
O praise him, as he doth excel\\
in glorious majesty.

3 Praise him with trumpet’s sound; his praise\\
with psaltery advance:\\
4 With timbrel, harp, stringed instruments,\\
and organs, in the dance.

5 Praise him on cymbals loud; him praise\\
on cymbals sounding high.\\
6 Let each thing breathing praise the Lord.\\
Praise to the Lord give ye.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

