
\begin{center}
\Huge \sc{The Revised Metrical Scottish Psalter}\vspace{24pt}\\
\Large \sc{Update the 1650 Psalter}
\end{center}


\begin{center}
\Large Revised on 27 July 2023
\end{center}

% \newpage 
% \tableofcontents

\newpage 
\begin{center}
\Huge \sc{The Revised Metrical Scottish Psalter }
\end{center}

\subsection*{Comments}

All the Psalms are set in Common Meter (8.6.8.6); some Psalms have an additional version in another meter.

The texts of the Psalms are from https://thewestminsterstandard.org/1650-scottish-metrical-psalter/.

Examples of tunes in common meter include:

\begin{multicols}{3}
\begin{itemize}
\sc\item Amazing Grace
  \item Aberdeen
	\item Antioch (\textup{Joy to the World!}) 
	\item Arlington
	\item Avon (Alas, and Did My Savior Bleed)
	\item Azmon (\textup{O For a Thousand Tongues to Sing})
	\item Bradford
	\item Bristol
	\item Caddo
	\item Clinton
	\item Christian Love
	\item Coronation \textup{(All Hail the Power of Jesus' Name)}
	\item Detroit (Composer: Bradshaw, 1820)
	\item Dominus Regit Me
	\item Dundee
	\item Ellacombe
	\item Fountain
	\item Gloria Patri
	\item Hamburg (\textup{When I Survey the Wondrous Cross})
  \item I Sing the Almight Power of God
	\item Land of Rest
	\item Lord of the Dance
	\item Maccabaeus
	\item Manoah
	\item Martyrdom (\textup{Alas! and Did My Savior Bleed})
	\item McKee
	\item Naomi
	\item Pisgah
	\item Prayer
	\item Repton
	\item Richmond
	\item Shanti
	\item St. Agnes
	\item St. Anne
	\item St. Columba
	\item St. Etheldreda
	\item St. Peter
	\item St. Stephen
	\item St. Theodulph
	\item Walsall
	\item Winchester Old
	\item Woodlands
	\item \textup{and thousands more at \href{https://hymnary.org/search?page=3&qu=meter\%3A8.6.8.6\%20in\%3Atunes&sort=totalInstances}{hymnary.org}}
\end{itemize}
\end{multicols}

\newpage

\begin{center}
\Large{\sc{Book I}}
\addcontentsline{toc}{section}{Book I}
\end{center}
\begin{multicols}{3}



\subsection*{Psalm 1 \blfootnote{\emph{Psalm 1 Tunes:} \sc{Azmon / Denfield}}}

\addcontentsline{toc}{subsection}{Psalm 1}

1 That man hath perfect blessedness,\\
who walketh not astray\\
In counsel of ungodly men,\\
nor stands in sinners’ way,

Nor sitteth in the scorner’s chair:\\
2 But placeth his delight\\
Upon God’s law, and meditates\\
on his law day and night.

3 He shall be like a tree that grows\\
near planted by a river,\\
Which in his season yields his fruit,\\
and his leaf fadeth never:

And all he doth shall prosper well\\
4 The wicked are not so;\\
But like they are unto the chaff,\\
which wind drives to and fro.

5 In judgment therefore shall not stand \\
such as ungodly are; \\
Nor in th’ assembly of the just \\
shall wicked men appear.

6 For why? the way of godly men \\
unto the Lord is known:\\
Whereas the way of wicked men \\
shall quite be overthrown.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 2 \blfootnote{\emph{Psalm 2 Tune:} \sc{York}}}
\addcontentsline{toc}{subsection}{Psalm 2}

1 Why rage the heathen? and vain things\\
why do the people mind?\\
2 Kings of the earth do set themselves,\\
and princes are combin’d,

To plot against the Lord, and his\\
Anointed, saying thus,\\
3 Let us asunder break their bands,\\
and cast their cords from us.

4 He that in heaven sits shall laugh;\\
the Lord shall scorn them all.\\
5 Then shall he speak to them in wrath,\\
in rage he vex them shall.

6 Yet, notwithstanding, I have him\\
to be my King appointed;\\
And over Sion, my holy hill,\\
I have him King anointed.

7 The sure decree I will declare:\\
The Lord hath said to me,\\
Thou art mine only Son; this day\\
I have begotten thee.

8 Ask of me, and for heritage\\
the heathen I’ll make thine;\\
And, for possession, I to thee\\
will give earth’s utmost line.

9 Thou shalt, as with a weighty rod\\
of iron, break them all;\\
And, as a potter’s sherd, thou shalt\\
them dash in pieces small.

10 Now therefore, kings, be wise; be taught, \\
ye judges of the earth:\\
11 Serve God in fear, and see that ye\\
join trembling with your mirth.

12 Kiss ye the Son, lest in his ire \\
ye perish from the way, \\
If once his wrath begin to burn: \\
bless’d all that on him stay.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 3}
\addcontentsline{toc}{subsection}{Psalm 3}

A \emph{Psalm} of David, when he fled from Absalom his son.

1 O Lord, how are my foes increas’d?\\
against me many rise.\\
2 Many say of my soul, For him\\
in God no succour lies.

3 Yet thou my shield and glory art,\\
th’ uplifter of mine head.\\
4 I cry’d, and, from his holy hill,\\
the Lord me answer made.

5 I laid me down and slept; I wak’d;\\
for God sustained me.\\
6 I will not fear though thousands ten\\
set round against me be.

7 Arise, O Lord; save me, my God;\\
for thou my foes hast stroke\\
All on the cheek-bone, and the teeth\\
of wicked men hast broke.

8 Salvation doth appertain\\
unto the Lord alone:\\
Thy blessing, Lord, for evermore\\
thy people is upon.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 4 \blfootnote{\emph{Psalm 4 Tune:} \sc{Crimond}}}
\addcontentsline{toc}{subsection}{Psalm 4}


To the chief Musician on \emph{Neginoth},\\
A \emph{Psalm} of David.

1 Give ear unto me when I call,\\
God of my righteousness:\\
Have mercy, hear my pray’r; thou hast\\
enlarg’d me in distress.

2 O ye the sons of men! how long\\
will ye love vanities?\\
How long my glory turn to shame,\\
and will ye follow lies?

3 But know, that for himself the Lord\\
the godly man doth chuse:\\
The Lord, when I on him do call,\\
to hear will not refuse.

4 Fear, and sin not; talk with your heart\\
on bed, and silent be.\\
5 Off ‘rings present of righteousness,\\
and in the Lord trust ye.

6 O who will shew us any good?\\
is that which many say:\\
But of thy countenance the light,\\
Lord, lift on us alway.

7 Upon my heart, bestow’d by thee,\\
more gladness I have found\\
Than they, ev’n then, when corn and wine\\
did most with them abound.

8 I will both lay me down in peace,\\
and quiet sleep will take;\\
Because thou only me to dwell\\
in safety, Lord, dost make.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 5}
\addcontentsline{toc}{subsection}{Psalm 5}

To the chief Musician, upon Nehiloth,
A \emph{Psalm} of David.

1 Give ear unto my words, O Lord,\\
my meditation weigh.\\
2 Hear my loud cry, my King, my God;\\
for I to thee will pray.

3 Lord, thou shalt early hear my voice:\\
I early will direct\\
My pray’r to thee; and, looking up,\\
an answer will expect.

4 For thou art not a God that doth\\
in wickedness delight;\\
Neither shall evil dwell with thee,\\
5 Nor fools stand in thy sight.

All that ill-doers are thou hat’st;\\
6 Cutt’st off that liars be:\\
The bloody and deceitful man\\
abhorred is by thee.

7 But I into thy house will come\\
in thine abundant grace;\\
And I will worship in thy fear\\
toward thy holy place.

8 Because of those mine enemies,\\
Lord, in thy righteousness\\
Do thou me lead; do thou thy way\\
make straight before my face.

9 For in their mouth there is no truth,\\
their inward part is ill;\\
Their throat’s an open sepulchre,\\
their tongue doth flatter still.

10 O God, destroy them; let them be\\
by their own counsel quell’d:\\
Them for their many sins cast out,\\
for they ‘gainst thee rebell’d.

11 But let all joy that trust in thee,\\
and still make shouting noise;\\
For them thou sav’st; let all that love\\
thy name in thee rejoice.

12 For, Lord, unto the righteous man\\
thou wilt thy blessing yield:\\
With favour thou wilt compass him \\
about, as with a shield.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 6}
\addcontentsline{toc}{subsection}{Psalm 6}

\emph{First Version---Long Meter (L.M., 8.8.8.8)} \blfootnote{\emph{Psalm 6 L.M. Tune:} \sc{Finnart}}

To the chief Musician on \emph{Neginoth} upon Sheminith,
A \emph{Psalm} of David.

1 Lord, in thy wrath rebuke me not;\\
Nor in thy hot rage chasten me.\\
2 Lord, pity me, for I am weak:\\
Heal me, for my bones vexed be.

3 My soul is also vexed sore;\\
But, Lord, how long stay wilt thou make?\\
4 Return, O Lord, my soul set free;\\
O save me, for thy mercies’ sake.

5 Because those that deceased are\\
Of thee shall no remembrance have;\\
And who is he that will to thee\\
Give praises lying in the grave?

6 I with my groaning weary am,\\
I also all the night my bed\\
Have caused for to swim; and I\\
With tears my couch have watered.

7 Mine eye, consum’d with grief, grows old,\\
Because of all mine enemies.\\
8 Hence from me, wicked workers all;\\
For God hath heard my weeping cries.

9 God hath my supplication heard,\\
My pray’r received graciously\\
10 Sham’d and sore vex’d be all my foes,\\
Sham’d and back turned suddenly.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (C.M., 8.6.8.6)}

To the chief Musician on Neginoth upon Sheminith,\\
A \emph{Psalm} of David.

1 In thy great indignation,\\
O Lord, rebuke me not;\\
Nor on me lay thy chast ‘ning hand,\\
in thy displeasure hot.

2 Lord, I am weak, therefore on me\\
have mercy, and me spare:\\
Heal me, O Lord, because thou know’st\\
my bones much vexed are.

3 My soul is vexed sore: but, Lord,\\
how long stay wilt thou make?\\
4 Return, Lord, free my soul; and save\\
me, for thy mercies’ sake.

5 Because of thee in death there shall\\
no more remembrance be:\\
Of those that in the grave do lie,\\
who shall give thanks to thee?

6 I with my groaning weary am,\\
and all the night my bed\\
I caused for to swim; with tears\\
my couch I watered.

7 By reason of my vexing grief,\\
mine eye consumed is;\\
It waxeth old, because of all\\
that be mine enemies.

8 But now, depart from me all ye\\
that work iniquity:\\
For why? the Lord hath heard my voice,\\
when I did mourn and cry.

9 Unto my supplication\\
the Lord did hearing give:\\
When I to him my prayer make,\\
the Lord will it receive.

10 Let all be sham’d and troubled sore,\\
That en’mies are to me;\\
Let them turn back, and suddenly\\
ashamed let them be.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 7}
\addcontentsline{toc}{subsection}{Psalm 7}


\emph{Shiggaion} of David, which he sang unto the Lord,
concerning the words of Cush the Benjamite.

1 O Lord my God, in thee do I\\
my confidence repose:\\
Save and deliver me from all\\
my persecuting foes;

2 Lest that the enemy my soul\\
should, like a lion, tear,\\
In pieces rending it, while there\\
is no deliverer.

3 O Lord my God, if it be so\\
that I committed this;\\
If it be so that in my hands\\
iniquity there is:

4 If I rewarded ill to him\\
that was at peace with me;\\
(Yea, ev’n the man that without cause\\
my foe was I did free;)

5 Then let the foe pursue and take\\
my soul, and my life thrust\\
Down to the earth, and let him lay\\
mine honour in the dust.

6 Rise in thy wrath, Lord, raise thyself,\\
for my foes raging be;\\
And, to the judgment which thou hast\\
commanded, wake for me.

7 So shall th’ assembly of thy folk\\
about encompass thee:\\
Thou, therefore, for their sakes, return\\
unto thy place on high.

8 The Lord he shall the people judge:\\
my judge, Jehovah, be,\\
After my righteousness, and mine\\
integrity in me.

9 O let the wicked’s malice end;\\
but stablish stedfastly\\
The righteous: for the righteous God\\
the hearts and reins doth try.

10 In God, who saves th’ upright in heart,\\
is my defence and stay.\\
11 God just men judgeth, God is wroth\\
with ill men ev’ry day.

12 If he do not return again,\\
then he his sword will whet;\\
His bow he hath already bent,\\
and hath it ready set:

13 He also hath for him prepar’d\\
the instruments of death;\\
Against the persecutors he\\
his shafts ordained hath.

14 Behold, he with iniquity\\
doth travail, as in birth;\\
A mischief he conceived hath,\\
and falsehood shall bring forth.

15 He made a pit and digg’d it deep,\\
another there to take;\\
But he is fall’n into the ditch\\
which he himself did make.

16 Upon his own head his mischief\\
shall be returned home;\\
His vi’lent dealing also down\\
on his own pate shall come.

17 According to his righteousness\\
the Lord I’ll magnify;\\
And will sing praise unto the name\\
of God that is most high.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 8 \blfootnote{\emph{Psalm 8 Tune:} \sc{Richmond}}}
\addcontentsline{toc}{subsection}{Psalm 8}


To the chief Musician upon \emph{Gittith},\\
A \emph{Psalm} of David.

1 How excellent in all the earth,\\
Lord, our Lord, is thy name!\\
Who hast thy glory far advanc’d\\
above the starry frame.

2 From infants’ and from sucklings’ mouth\\
thou didest strength ordain,\\
For thy foes’ cause, that so thou might’st\\
th’ avenging foe restrain.

3 When I look up unto the heav’ns,\\
which thine own fingers fram’d,\\
Unto the moon, and to the stars,\\
which were by thee ordain’d;

4 Then say I, What is man, that he\\
remember’d is by thee?\\
Or what the son of man, that thou\\
so kind to him should’st be?

5 For thou a little lower hast\\
him than the angels made;\\
With glory and with dignity\\
thou crowned hast his head.

6 Of thy hands’ works thou mad’st him lord,\\
all under’s feet didst lay;\\
7 All sheep and oxen, yea, and beasts\\
that in the field do stray;

8 Fowls of the air, fish of the sea,\\
all that pass through the same.\\
9 How excellent in all the earth,\\
Lord, our Lord, is thy name!

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 9 \blfootnote{\emph{Psalm 9 Tune:} \sc{Stroudwater}}}
\addcontentsline{toc}{subsection}{Psalm 9}


To the chief Musician upon Muth-labben,\\
A \emph{Psalm} of David.

1 Lord, thee I’ll praise with all my heart,\\
thy wonders all proclaim.\\
2 In thee, most High, I’ll greatly joy,\\
and sing unto thy name.

3 When back my foes were turn’d, they fell,\\
and perish’d at thy sight:\\
4 For thou maintain’dst my right and cause;\\
on throne sat’st judging right.

5 The heathen thou rebuked hast,\\
the wicked overthrown;\\
Thou hast put out their names, that they\\
may never more be known.

6 O en’my! now destructions have\\
an end perpetual:\\
Thou cities raz’d; perish’d with them\\
is their memorial.

7 God shall endure for aye; he doth\\
for judgment set his throne;\\
8 In righteousness to judge the world,\\
justice to give each one.

9 God also will a refuge be\\
for those that are oppress’d;\\
A refuge will he be in times\\
of trouble to distress’d.

10 And they that know thy name, in thee\\
their confidence will place:\\
For thou hast not forsaken them\\
that truly seek thy face.

11 O sing ye praises to the Lord,\\
that dwells in Sion hill;\\
And all the nations among\\
his deeds record ye still.

12 When he enquireth after blood,\\
he then rememb’reth them:\\
The humble folk he not forgets\\
that call upon his name.

13 Lord, pity me; behold the grief\\
which I from foes sustain;\\
Ev’n thou, who from the gates of death\\
dost raise me up again;

14 That I, in Sion’s daughters’ gates,\\
may all thy praise advance;\\
And that I may rejoice always\\
in thy deliverance.

15 The heathen are sunk in the pit\\
which they themselves prepar’d;\\
And in the net which they have hid\\
their own feet fast are snar’d.

16 The Lord is by the judgment known\\
which he himself hath wrought:\\
The sinners’ hands do make the snares\\
wherewith themselves are caught.

17 They who are wicked into hell\\
each one shall turned be;\\
And all the nations that forget\\
to seek the Lord most high.

18 For they that needy are shall not\\
forgotten be alway;\\
The expectation of the poor\\
shall not be lost for aye.

19 Arise, Lord, let not man prevail;\\
judge heathen in thy sight:\\
20 That they may know themselves but men,\\
the nations, Lord, affright.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 10 \blfootnote{\emph{Psalm 10 Tune:} \sc {Glenluce}}}
\addcontentsline{toc}{subsection}{Psalm 10}

1 Wherefore is it that thou, O Lord,\\
dost stand from us afar?\\
And wherefore hidest thou thyself,\\
when times so troublous are?

2 The wicked in his loftiness\\
doth persecute the poor:\\
In these devices they have fram’d\\
let them be taken sure.

3 The wicked of his heart’s desire\\
doth talk with boasting great;\\
He blesseth him that’s covetous,\\
whom yet the Lord doth hate.

4 The wicked, through his pride of face,\\
on God he doth not call;\\
And in the counsels of his heart\\
the Lord is not at all.

5 His ways they always grievous are;\\
thy judgments from his sight\\
Removed are: at all his foes\\
he puffeth with despight.

6 Within his heart he thus hath said,\\
I shall not moved be;\\
And no adversity at all\\
shall ever come to me.

7 His mouth with cursing, fraud, deceit,\\
is fill’d abundantly;\\
And underneath his tongue there is\\
mischief and vanity.

8 He closely sits in villages;\\
he slays the innocent:\\
Against the poor that pass him by\\
his cruel eyes are bent.

9 He, lion-like, lurks in his den;\\
he waits the poor to take;\\
And when he draws him in his net,\\
his prey he doth him make.

10 Himself he humbleth very low,\\
he croucheth down withal,\\
That so a multitude of poor\\
may by his strong ones fall.

11 He thus hath said within his heart,\\
The Lord hath quite forgot;\\
He hides his countenance, and he\\
for ever sees it not.

12 O Lord, do thou arise; O God,\\
lift up thine hand on high:\\
Put not the meek afflicted ones\\
out of thy memory.

13 Why is it that the wicked man\\
thus doth the Lord despise?\\
Because that God will it require\\
he in his heart denies.

14 Thou hast it seen; for their mischief\\
and spite thou wilt repay:\\
The poor commits himself to thee;\\
thou art the orphan’s stay.

15 The arm break of the wicked man,\\
and of the evil one;\\
Do thou seek out his wickedness,\\
until thou findest none.

16 The Lord is King through ages all,\\
ev’n to eternity;\\
The heathen people from his land\\
are perish’d utterly.

17 O Lord, of those that humble are\\
thou the desire didst hear;\\
Thou wilt prepare their heart, and thou\\
to hear wilt bend thine ear;

18 To judge the fatherless, and those\\
that are oppressed sore;\\
That man, that is but sprung of earth,\\
may them oppress no more.

 % Separation Symbol

\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 11 \blfootnote{\emph{Psalm 11 Tune:} \sc {Salisbury}}}
\addcontentsline{toc}{subsection}{Psalm 11}

To the chief Musician,\\
A \emph{Psalm} of David.

1 I in the Lord do put my trust:\\
how is it then that ye\\
Say to my soul, Flee, as a bird,\\
unto your mountain high?

2 For, lo, the wicked bend their bow,\\
their shafts on string they fit,\\
That those who upright are in heart\\
they privily may hit.

3 If the foundations be destroy’d,\\
what hath the righteous done?\\
4 God in his holy temple is,\\
in heaven is his throne:

His eyes do see, his eye-lids try\\
5 men’s sons. The just he proves:\\
But his soul hates the wicked man,\\
and him that vi’lence loves.

6 Snares, fire and brimstone, furious storms,\\
on sinners he shall rain:\\
This, as the portion of their cup,\\
doth unto them pertain.

7 Because the Lord most righteous doth\\
in righteousness delight;\\
And with a pleasant countenance\\
beholdeth the upright.

%  % Separation Symbol
% \begin{center}
% \adforn{21}\quad\adforn{11}\quad\adforn{49}
% \end{center}

\subsection*{Psalm 12 \blfootnote{\emph{Psalm 12 Tune:} \sc {Farrant}}}
\addcontentsline{toc}{subsection}{Psalm 12}

To the chief Musician upon \emph{Sheminith},\\
A \emph{Psalm} of David.

1 Help, Lord, because the godly man\\
doth daily fade away;\\
And from among the sons of men\\
the faithful do decay.

2 Unto his neighbour ev’ry one\\
doth utter vanity:\\
They with a double heart do speak,\\
and lips of flattery.

3 God shall cut off all flatt’ring lips,\\
tongues that speak proudly thus,\\
4 We’ll with our tongue prevail, our lips\\
are ours: who’s lord o’er us?

5 For poor oppress’d, and for the sighs\\
of needy, rise will I,\\
Saith God, and him in safety set\\
from such as him defy.

6 The words of God are words most pure;\\
they be like silver try’d\\
In earthen furnace, seven times\\
that hath been purify’d.

7 Lord, thou shalt them preserve and keep\\
for ever from this race.\\
8 On each side walk the wicked, when\\
vile men are high in place.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 13 \blfootnote{\emph{Psalm 13 Tune:} \sc {Stornoway}}}
\addcontentsline{toc}{subsection}{Psalm 13}

To the chief Musician,\\
A \emph{Psalm} of David.

1 How long wilt thou forget me, Lord?\\
shall it for ever be?\\
O how long shall it be that thou\\
wilt hide thy face from me?

2 How long take counsel in my soul,\\
still sad in heart, shall I?\\
How long exalted over me\\
shall be mine enemy?

3 O Lord my God, consider well,\\
and answer to me make:\\
Mine eyes enlighten, lest the sleep\\
of death me overtake:

4 Lest that mine enemy should say,\\
Against him I prevail’d;\\
And those that trouble me rejoice,\\
when I am mov’d and fail’d.

5 But I have all my confidence\\
thy mercy set upon;\\
My heart within me shall rejoice\\
in thy salvation.

6 I will unto the Lord my God\\
sing praises cheerfully,\\
Because he hath his bounty shown\\
to me abundantly.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 14  \blfootnote{\emph{Psalm 14 Tune:} \sc {St. Thomas}}}
\addcontentsline{toc}{subsection}{Psalm 14}

To the chief Musician,\\
A \emph{Psalm} of David.

1 That there is not a God, the fool\\
doth in his heart conclude:\\
They are corrupt, their works are vile;\\
not one of them doth good.

2 Upon men’s sons the Lord from heav’n\\
did cast his eyes abroad,\\
To see if any understood,\\
and did seek after God.

3 They altogether filthy are,\\
they all aside are gone;\\
And there is none that doeth good,\\
yea, sure there is not one.

4 These workers of iniquity\\
do they not know at all,\\
That they my people eat as bread,\\
and on God do not call?

5 There fear’d they much; for God is with\\
the whole race of the just.\\
6 You shame the counsel of the poor,\\
because God is his trust.

7 Let Isr’el’s help from Sion come:\\
when back the Lord shall bring\\
His captives, Jacob shall rejoice,\\
and Israel shall sing.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 15 \blfootnote{\emph{Psalm 15 Tune:} \sc {Grafenberg}}}
\addcontentsline{toc}{subsection}{Psalm 15}

A \emph{Psalm} of David.

1 Within thy tabernacle, Lord,\\
who shall abide with thee?\\
And in thy high and holy hill\\
who shall a dweller be?

2 The man that walketh uprightly,\\
and worketh righteousness,\\
And as he thinketh in his heart,\\
so doth he truth express.

3 Who doth not slander with his tongue,\\
nor to his friend doth hurt;\\
Nor yet against his neighbour doth\\
take up an ill report.

4 In whose eyes vile men are despis’d;\\
but those that God do fear\\
He honoureth; and changeth not,\\
though to his hurt he swear.

5 His coin puts not to usury,\\
nor take reward will he\\
Against the guiltless. Who doth thus\\
shall never moved be.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 16}
\addcontentsline{toc}{subsection}{Psalm 16}

\emph{Michtam} of David.

1 Lord, keep me; for I trust in thee.\\
2 To God thus was my speech,\\
Thou art my Lord; and unto thee\\
my goodness doth not reach:

3 To saints on earth, to th’ excellent,\\
where my delight’s all plac’d.\\
4 Their sorrows shall be multiply’d\\
to other gods that haste:

Of their drink-offerings of blood\\
I will no off ‘ring make;\\
Yea, neither I their very names\\
up in my lips will take.

5 God is of mine inheritance\\
and cup the portion;\\
The lot that fallen is to me\\
thou dost maintain alone.

6 Unto me happily the lines\\
in pleasant places fell;\\
Yea, the inheritance I got\\
in beauty doth excel.

7 I bless the Lord, because he doth\\
by counsel me conduct;\\
And in the seasons of the night\\
my reins do me instruct.

8 Before me still the Lord I set:\\
sith it is so that he\\
Doth ever stand at my right hand,\\
I shall not moved be.

9 Because of this my heart is glad,\\
and joy shall be exprest\\
Ev’n by my glory; and my flesh\\
in confidence shall rest.

10 Because my soul in grave to dwell\\
shall not be left by thee;\\
Nor wilt thou give thine Holy One\\
corruption to see.

11 Thou wilt me shew the path of life:\\
of joys there is full store\\
Before thy face; at thy right hand\\
are pleasures evermore.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 17 \blfootnote{\emph{Psalm 17 Tune:} \sc{St. Bernard}}}
\addcontentsline{toc}{subsection}{Psalm 17}

A \emph{Prayer} of David.

1 Lord, hear the right, attend my cry,\\
unto my pray’r give heed,\\
That doth not in hypocrisy\\
from feigned lips proceed.

2 And from before thy presence forth\\
my sentence do thou send:\\
Toward these things that equal are\\
do thou thine eyes intend.

3 Thou prov’dst mine heart, thou visit’dst me\\
by night, thou didst me try,\\
Yet nothing found’st; for that my mouth\\
shall not sin, purpos’d I.

4 As for men’s works, I, by the word\\
that from thy lips doth flow,\\
Did me preserve out of the paths\\
wherein destroyers go.

5 Hold up my goings, Lord, me guide\\
in those thy paths divine,\\
So that my footsteps may not slide\\
out of those ways of thine.

6 I called have on thee, O God,\\
because thou wilt me hear:\\
That thou may’st hearken to my speech,\\
to me incline thine ear.

7 Thy wondrous loving-kindness show,\\
thou that, by thy right hand,\\
Sav’st them that trust in thee from those\\
that up against them stand.

8 As th’ apple of the eye me keep;\\
in thy wings shade me close\\
9 From lewd oppressors, compassing\\
me round, as deadly foes.

10 In their own fat they are inclos’d;\\
their mouth speaks loftily.\\
11 Our steps they compass’d; and to ground\\
down bowing set their eye.

12 He like unto a lion is\\
that’s greedy of his prey,\\
Or lion young, which lurking doth\\
in secret places stay.

13 Arise, and disappoint my foe,\\
and cast him down, O Lord:\\
My soul save from the wicked man,\\
the man which is thy sword.

14 From men, which are thy hand, O Lord,\\
from worldly men me save,\\
Which only in this present life\\
their part and portion have.

Whose belly with thy treasure hid\\
thou fill’st: they children have\\
In plenty; of their goods the rest\\
they to their children leave.

15 But as for me, I thine own face\\
in righteousness will see;\\
And with thy likeness, when I wake,\\
I satisfy’d shall be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 18 \blfootnote{\emph{Psalm 18 Tune:} \sc{Warwick}}}
\addcontentsline{toc}{subsection}{Psalm 18}


To the chief Musician, A \emph{Psalm} of David, the servant of the Lord, who spake unto the Lord the words of this song in the day that the Lord delivered him from the hand of all his enemies, and from the hand of Saul: And he said,

1 Thee will I love, O Lord, my strength.\\
2 My fortress is the Lord,\\
My rock, and he that doth to me\\
deliverance afford:

My God, my strength, whom I will trust,\\
a buckler unto me,\\
The horn of my salvation,\\
and my high tow’r, is he.

3 Upon the Lord, who worthy is\\
of praises, will I cry;\\
And then shall I preserved be\\
safe from mine enemy.

4 Floods of ill men affrighted me,\\
death’s pangs about me went;\\
5 Hell’s sorrows me environed;\\
death’s snares did me prevent.

6 In my distress I call’d on God,\\
cry to my God did I;\\
He from his temple heard my voice,\\
to his ears came my cry.

7 Th’ earth, as affrighted, then did shake,\\
trembling upon it seiz’d:\\
The hills’ foundations moved were,\\
because he was displeas’d.

8 Up from his nostrils came a smoke,\\
and from his mouth there came\\
Devouring fire, and coals by it\\
were turned into flame.

9 He also bowed down the heav’ns,\\
and thence he did descend;\\
And thickest clouds of darkness did\\
under his feet attend.

10 And he upon a cherub rode,\\
and thereon he did fly;\\
Yea, on the swift wings of the wind\\
his flight was from on high.

11 He darkness made his secret place:\\
about him, for his tent,\\
Dark waters were, and thickest clouds\\
of th’ airy firmament.

12 And at the brightness of that light,\\
which was before his eye,\\
His thick clouds pass’d away, hailstones\\
and coals of fire did fly.

13 The Lord God also in the heav’ns\\
did thunder in his ire;\\
And there the Highest gave his voice,\\
hailstones and coals of fire.

14 Yea, he his arrows sent abroad,\\
and them he scattered;\\
His lightnings also he shot out,\\
and them discomfited.

15 The waters’ channels then were seen,\\
the world’s foundations vast\\
At thy rebuke discover’d were,\\
and at thy nostrils’ blast.

16 And from above the Lord sent down,\\
and took me from below;\\
From many waters he me drew,\\
which would me overflow.

17 He me reliev’d from my strong foes,\\
and such as did me hate;\\
Because he saw that they for me\\
too strong were, and too great.

18 They me prevented in the day\\
of my calamity;\\
But even then the Lord himself\\
a stay was unto me.

19 He to a place where liberty\\
and room was hath me brought;\\
Because he took delight in me,\\
he my deliv’rance wrought.

20 According to my righteousness\\
he did me recompense,\\
He me repaid according to\\
my hands’ pure innocence.

21 For I God’s ways kept, from my God\\
did not turn wickedly.\\
His judgments were before me, I\\
his laws put not from me.

23 Sincere before him was my heart;\\
with him upright was I;\\
And watchfully I kept myself\\
from mine iniquity.

24 After my righteousness the Lord\\
hath recompensed me,\\
After the cleanness of my hands\\
appearing in his eye.

25 Thou gracious to the gracious art,\\
to upright men upright:\\
26 Pure to the pure, froward thou kyth’st\\
unto the froward wight.

27 For thou wilt the afflicted save\\
in grief that low do lie:\\
But wilt bring down the countenance\\
of them whose looks are high.

28 The Lord will light my candle so,\\
that it shall shine full bright:\\
The Lord my God will also make\\
my darkness to be light.

29 By thee through troops of men I break,\\
and them discomfit all;\\
And, by my God assisting me,\\
I overleap a wall.

30 As for God, perfect is his way:\\
the Lord his word is try’d;\\
He is a buckler to all those\\
who do in him confide.

31 Who but the Lord is God? but he\\
who is a rock and stay?\\
32 ‘Tis God that girdeth me with strength,\\
and perfect makes my way.

33 He made my feet swift as the hinds,\\
set me on my high places.\\
34 Mine hands to war he taught, mine arms\\
brake bows of steel in pieces.

35 The shield of thy salvation\\
thou didst on me bestow:\\
Thy right hand held me up, and great\\
thy kindness made me grow.

36 And in my way my steps thou hast\\
enlarged under me,\\
That I go safely, and my feet\\
are kept from sliding free.

37 Mine en’mies I pursued have,\\
and did them overtake;\\
Nor did I turn again till I\\
an end of them did make.

38 I wounded them, they could not rise;\\
they at my feet did fall.\\
39 Thou girdedst me with strength for war;\\
my foes thou brought’st down all:

40 And thou hast giv’n to me the necks\\
of all mine enemies;\\
That I might them destroy and slay,\\
who did against me rise.

41 They cried out, but there was none\\
that would or could them save;\\
Yea, they did cry unto the Lord,\\
but he no answer gave.

42 Then did I beat them small as dust\\
before the wind that flies;\\
And I did cast them out like dirt\\
upon the street that lies.

43 Thou mad’st me free from people’s strife,\\
and heathen’s head to be:\\
A people whom I have not known\\
shall service do to me.

44 At hearing they shall me obey,\\
to me they shall submit.\\
45 Strangers for fear shall fade away,\\
who in close places sit.

46 God lives, bless’d be my Rock; the God\\
of my health praised be.\\
47 God doth avenge me, and subdues\\
the people under me.

48 He saves me from mine enemies;\\
yea, thou hast lifted me\\
Above my foes; and from the man\\
of vi’lence set me free.

49 Therefore to thee will I give thanks\\
the heathen folk among;\\
And to thy name, O Lord, I will\\
sing praises in a song.

50 He great deliv’rance gives his king:\\
he mercy doth extend\\
To David, his anointed one,\\
and his seed without end.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 19 \blfootnote{\emph{Psalm 19 Tune:} \sc{Crediton / Moravia / St. Andrew}}}
\addcontentsline{toc}{subsection}{Psalm 19}

To the chief Musician,\\
A \emph{Psalm} of David.

1 The heav’ns God’s glory do declare,\\
the skies his hand-works preach:\\
2 Day utters speech to day, and night\\
to night doth knowledge teach.

3 There is no speech nor tongue to which\\
their voice doth not extend:\\
4 Their line is gone through all the earth,\\
their words to the world’s end.

In them he set the sun a tent;\\
5 Who, bridegroom-like, forth goes\\
From’s chamber, as a strong man doth\\
to run his race rejoice.

6 From heav’n’s end is his going forth,\\
circling to th’ end again;\\
And there is nothing from his heat\\
that hidden doth remain.

7 God’s law is perfect, and converts\\
the soul in sin that lies:\\
God’s testimony is most sure,\\
and makes the simple wise.

8 The statutes of the Lord are right,\\
and do rejoice the heart:\\
The Lord’s command is pure, and doth\\
light to the eyes impart.

9 Unspotted is the fear of God,\\
and doth endure for ever:\\
The judgments of the Lord are true\\
and righteous altogether.

10 They more than gold, yea, much fine gold,\\
to be desired are:\\
Than honey, honey from the comb\\
that droppeth, sweeter far.

11 Moreover, they thy servant warn\\
how he his life should frame:\\
A great reward provided is\\
for them that keep the same.

12 Who can his errors understand?\\
O cleanse thou me within\\
13 From secret faults. Thy servant keep\\
from all presumptuous sin:

And do not suffer them to have\\
dominion over me:\\
Then, righteous and innocent,\\
I from much sin shall be.

14 The words which from my mouth proceed,\\
the thoughts sent from my heart,\\
Accept, O Lord, for thou my strength\\
and my Redeemer art.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 20 \blfootnote{\emph{Psalm 20 Tune:} \sc{Orton}}}
\addcontentsline{toc}{subsection}{Psalm 20}

To the chief Musician,\\
A \emph{Psalm} of David.

1 Jehovah hear thee in the day\\
when trouble he doth send:\\
And let the name of Jacob’s God\\
thee from all ill defend.

2 O let him help send from above,\\
out of his sanctuary:\\
From Sion, his own holy hill,\\
let him give strength to thee.

3 Let him remember all thy gifts,\\
accept thy sacrifice:\\
4 Grant thee thine heart’s wish, and fulfil\\
thy thoughts and counsel wise.

5 In thy salvation we will joy;\\
in our God’s name we will\\
Display our banners: and the Lord\\
thy prayers all fulfil.

6 Now know I God his king doth save:\\
he from his holy heav’n\\
Will hear him, with the saving strength\\
by his own right hand giv’n.

7 In chariots some put confidence,\\
some horses trust upon:\\
But we remember will the name\\
of our Lord God alone.

8 We rise, and upright stand, when they\\
are bowed down, and fall.\\
9 Deliver, Lord; and let the King\\
us hear, when we do call.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 21 \blfootnote{\emph{Psalm 21 Tune:} \sc{Arnold}}}
\addcontentsline{toc}{subsection}{Psalm 21}

To the chief Musician,\\
A \emph{Psalm} of David.

1 The king in thy great strength, O Lord,\\
shall very joyful be:\\
In thy salvation rejoice\\
how veh’mently shall he!

2 Thou hast bestowed upon him\\
all that his heart would have;\\
And thou from him didst not withhold\\
whate’er his lips did crave.

3 For thou with blessings him prevent’st\\
of goodness manifold;\\
And thou hast set upon his head\\
a crown of purest gold.

4 When he desired life of thee,\\
thou life to him didst give;\\
Ev’n such a length of days, that he\\
for evermore should live.

5 In that salvation wrought by thee\\
his glory is made great;\\
Honour and comely majesty\\
thou hast upon him set.

6 Because that thou for evermore\\
most blessed hast him made;\\
And thou hast with thy countenance\\
made him exceeding glad.

7 Because the king upon the Lord\\
his confidence doth lay;\\
And through the grace of the most High\\
shall not be mov’d away.

8 Thine hand shall all those men find out\\
that en’mies are to thee;\\
Ev’n thy right hand shall find out those\\
of thee that haters be.

9 Like fiery ov’n thou shalt them make,\\
when kindled is thine ire;\\
God shall them swallow in his wrath,\\
devour them shall the fire.

10 Their fruit from earth thou shalt destroy,\\
their seed men from among:\\
11 For they beyond their might ‘gainst thee\\
did plot mischief and wrong.

12 Thou therefore shalt make them turn back,\\
when thou thy shafts shalt place\\
Upon thy strings, made ready all\\
to fly against their face.

13 In thy great pow’r and strength, O Lord,\\
be thou exalted high;\\
So shall we sing with joyful hearts,\\
thy power praise shall we.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 22 \blfootnote{\emph{Psalm 22 Tune:} \sc{Bangor / Old Twenty-Second}}}
\addcontentsline{toc}{subsection}{Psalm 22}

To the chief Musician upon \emph{Aijeleth Shahar},\\
A \emph{Psalm} of David.

1 My God, my God, why hast thou me\\
forsaken? why so far\\
Art thou from helping me, and from\\
my words that roaring are?

2 All day, my God, to thee I cry,\\
yet am not heard by thee;\\
And in the season of the night\\
I cannot silent be.

3 But thou art holy, thou that dost\\
inhabit Isr’el’s praise.\\
4 Our fathers hop’d in thee, they hop’d\\
and thou didst them release.

5 When unto thee they sent their cry,\\
to them deliv’rance came:\\
Because they put their trust in thee,\\
they were not put to shame.

6 But as for me, a worm I am,\\
and as no man am priz’d:\\
Reproach of men I am, and by\\
the people am despis’d.

7 All that me see laugh me to scorn;\\
shoot out the lip do they;\\
They nod and shake their heads at me,\\
and, mocking, thus do say,

8 This man did trust in God, that he\\
would free him by his might:\\
Let him deliver him, sith he\\
had in him such delight.

9 But thou art he out of the womb\\
that didst me safely take;\\
When I was on my mother’s breasts\\
thou me to hope didst make.

10 And I was cast upon thy care,\\
ev’n from the womb till now;\\
And from my mother’s belly, Lord,\\
my God and guide art thou.

11 Be not far off, for grief is near,\\
and none to help is found.\\
12 Bulls many compass me, strong bulls\\
of Bashan me surround.

13 Their mouths they open’d wide on me,\\
upon me gape did they,\\
Like to a lion ravening\\
and roaring for his prey.

14 Like water I’m pour’d out, my bones\\
all out of joint do part:\\
Amidst my bowels, as the wax,\\
so melted is my heart.

15 My strength is like a potsherd dry’d;\\
my tongue it cleaveth fast\\
Unto my jaws; and to the dust\\
of death thou brought me hast.

16 For dogs have compass’d me about:\\
the wicked, that did meet\\
In their assembly, me inclos’d;\\
they pierc’d my hands and feet.

17 I all my bones may tell; they do\\
upon me look and stare.\\
18 Upon my vesture lots they cast,\\
and clothes among them share.

19 But be not far, O Lord, my strength;\\
haste to give help to me.\\
20 From sword my soul, from pow’r of dogs\\
my darling set thou free.

21 Out of the roaring lion’s mouth\\
do thou me shield and save:\\
For from the horns of unicorns\\
an ear to me thou gave.

22 I will shew forth thy name unto\\
those that my brethren are;\\
Amidst the congregation\\
thy praise I will declare.

23 Praise ye the Lord, who do him fear;\\
him glorify all ye\\
The seed of Jacob: fear him all\\
that Isr’el’s children be.

24 For he despis’d not nor abhorr’d\\
th’ afflicted’s misery;\\
Nor from him hid his face, but heard\\
when he to him did cry.

25 Within the congregation great\\
my praise shall be of thee;\\
My vows before them that him fear\\
shall be perform’d by me.

26 The meek shall eat, and shall be fill’d;\\
they also praise shall give\\
Unto the Lord that do him seek:\\
your heart shall ever live.

27 All ends of th’ earth remember shall,\\
and turn the Lord unto;\\
All kindreds of the nations\\
to him shall homage do:

28 Because the kingdom to the Lord\\
doth appertain as his;\\
Likewise among the nations\\
the Governor he is.

29 Earth’s fat ones eat, and worship shall:\\
all who to dust descend\\
Shall bow to him; none of them can\\
his soul from death defend.

30 A seed shall service do to him;\\
unto the Lord it shall\\
Be for a generation\\
reckon’d in ages all.

31 They shall come, and they shall declare\\
his truth and righteousness\\
Unto a people yet unborn,\\
and that he hath done this.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 23 \blfootnote{\emph{Psalm 23 Tune:} \sc{Bays of Harris / Crimond}}}
\addcontentsline{toc}{subsection}{Psalm 23}

A \emph{Psalm} of David.

1 The Lord’s my shepherd, I’ll not want.\\
2 He makes me down to lie\\
In pastures green: he leadeth me\\
the quiet waters by.

3 My soul he doth restore again;\\
and me to walk doth make\\
Within the paths of righteousness,\\
ev’n for his own name’s sake.

4 Yea, though I walk in death’s dark vale,\\
yet will I fear none ill:\\
For thou art with me; and thy rod\\
and staff me comfort still.

5 My table thou hast furnished\\
in presence of my foes;\\
My head thou dost with oil anoint,\\
and my cup overflows.

6 Goodness and mercy all my life\\
shall surely follow me:\\
And in God’s house for evermore\\
my dwelling-place shall be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 24 \blfootnote{\emph{Psalm 24 Tune:} \sc{Rest / St. Georges Edinburgh}}}
\addcontentsline{toc}{subsection}{Psalm 24}

1 The earth belongs unto the Lord,\\
and all that it contains;\\
The world that is inhabited,\\
and all that there remains.

2 For the foundations thereof\\
he on the seas did lay,\\
And he hath it established\\
upon the floods to stay.

3 Who is the man that shall ascend\\
into the hill of God?\\
Or who within his holy place\\
shall have a firm abode?

4 Whose hands are clean, whose heart is pure,\\
and unto vanity\\
Who hath not lifted up his soul,\\
nor sworn deceitfully.

5 He from th’ Eternal shall receive\\
the blessing him upon,\\
And righteousness, ev’n from the God\\
of his salvation.

6 This is the generation\\
that after him enquire,\\
O Jacob, who do seek thy face\\
with their whole heart’s desire.

7 Ye gates, lift up your heads on high;\\
ye doors that last for aye,\\
Be lifted up, that so the King\\
of glory enter may.

8 But who of glory is the King?\\
The mighty Lord is this;\\
Ev’n that same Lord, that great in might\\
and strong in battle is.

9 Ye gates, lift up your heads; ye doors,\\
doors that do last for aye,\\
Be lifted up, that so the King\\
of glory enter may.

10 But who is he that is the King\\
of glory? who is this?\\
The Lord of hosts, and none but he,\\
the King of glory is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 25 \blfootnote{\emph{Psalm 25 S.M. Tune:} \sc{Dennis}}}
\addcontentsline{toc}{subsection}{Psalm 25}

\emph{First Version---Short Meter (S.M., 6.6.8.6)}

A \emph{Psalm} of David.

1 To thee I lift my soul:\\
2 O Lord, I trust in thee:\\
My God, let me not be asham’d,\\
nor foes triumph o’er me.

3 Let none that wait on thee\\
be put to shame at all;\\
But those that without cause transgress,\\
let shame upon them fall.

4 Shew me thy ways, O Lord;\\
thy paths, O teach thou me:\\
5 And do thou lead me in thy truth,\\
therein my teacher be:

For thou art God that dost\\
to me salvation send,\\
And I upon thee all the day\\
expecting do attend.

6 Thy tender mercies, Lord,\\
I pray thee to remember,\\
And loving-kindnesses; for they\\
have been of old for ever.

7 My sins and faults of youth\\
do thou, O Lord, forget:\\
After thy mercy think on me,\\
and for thy goodness great.

8 God good and upright is:\\
the way he’ll sinners show.\\
9 The meek in judgment he will guide,\\
and make his path to know.

10 The whole paths of the Lord\\
are truth and mercy sure,\\
To those that do his cov’nant keep,\\
and testimonies pure.

11 Now, for thine own name’s sake,\\
O Lord, I thee entreat\\
To pardon mine iniquity;\\
for it is very great.

12 What man is he that fears\\
the Lord, and doth him serve?\\
Him shall he teach the way that he\\
shall choose, and still observe.

13 His soul shall dwell at ease;\\
and his posterity\\
Shall flourish still, and of the earth\\
inheritors shall be.

14 With those that fear him is\\
the secret of the Lord;\\
The knowledge of his covenant\\
he will to them afford.

15 Mine eyes upon the Lord\\
continually are set:\\
For he it is that shall bring forth\\
my feet out of the net.

16 Turn unto me thy face,\\
and to me mercy show;\\
Because that I am desolate,\\
and am brought very low.

17 My heart’s griefs are increas’d:\\
me from distress relieve.\\
18 See mine affliction and my pain,\\
and all my sins forgive.

19 Consider thou my foes,\\
because they many are;\\
And it a cruel hatred is\\
which they against me bear.

20 O do thou keep my soul,\\
do thou deliver me:\\
And let me never be asham’d,\\
because I trust in thee.

21 Let uprightness and truth\\
keep me, who thee attend.\\
22 Redemption, Lord, to Israel\\
from all his troubles send.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (C.M., 8.6.8.6)}

1 To thee I lift my soul, O Lord:\\
2 My God, I trust in thee:\\
Let me not be asham’d; let not\\
my foes triumph o’er me.

3 Yea, let thou none ashamed be\\
that do on thee attend:\\
Ashamed let them be, O Lord,\\
who without cause offend.

4 Thy ways, Lord, shew; teach me thy paths:\\
5 Lead me in truth, teach me:\\
For of my safety thou art God;\\
all day I wait on thee.

6 Thy mercies, that most tender are,\\
do thou, O Lord, remember,\\
And loving-kindnesses; for they\\
have been of old for ever.

7 Let not the errors of my youth,\\
nor sins, remember’d be:\\
In mercy, for thy goodness’ sake,\\
O Lord, remember me.

8 The Lord is good and gracious,\\
he upright is also:\\
He therefore sinners will instruct\\
in ways that they should go.

9 The meek and lowly he will guide\\
in judgment just alway:\\
To meek and poor afflicted ones\\
he’ll clearly teach his way.

10 The whole paths of the Lord our God\\
are truth and mercy sure,\\
To such as keep his covenant,\\
and testimonies pure.

11 Now, for thine own name’s sake, O Lord,\\
I humbly thee entreat\\
To pardon mine iniquity;\\
for it is very great.

12 What man fears God? him shall he teach\\
the way that he shall chuse.\\
13 His soul shall dwell at ease; his seed\\
the earth, as heirs, shall use.

14 The secret of the Lord is with\\
such as do fear his name;\\
And he his holy covenant\\
will manifest to them.

15 Towards the Lord my waiting eyes\\
continually are set;\\
For he it is that shall bring forth\\
my feet out of the net.

16 O turn thee unto me, O God,\\
have mercy me upon;\\
Because I solitary am,\\
and in affliction.

17 Enlarg’d the griefs are of mine heart;\\
me from distress relieve.\\
18 See mine affliction and my pain,\\
and all my sins forgive.

19 Consider thou mine enemies,\\
because they many are;\\
And it a cruel hatred is\\
which they against me bear.

20 O do thou keep my soul; O God,\\
do thou deliver me:\\
Let me not be asham’d; for I\\
do put my trust in thee.

21 O let integrity and truth\\
keep me, who thee attend.\\
22 Redemption, Lord, to Israel\\
from all his troubles send.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 26 \blfootnote{\emph{Psalm 26 Tune:} \sc{St. Thomas}}}
\addcontentsline{toc}{subsection}{Psalm 26}

A \emph{Psalm} of David

1 Judge me, O Lord, for I have walk’d\\
in mine integrity:\\
I trusted also in the Lord;\\
slide therefore shall not I.

2 Examine me, and do me prove;\\
try heart and reins, O God:\\
3 For thy love is before mine eyes,\\
thy truth’s paths I have trode.

4 With persons vain I have not sat,\\
nor with dissemblers gone:\\
5 Th’ assembly of ill men I hate;\\
to sit with such I shun.

6 Mine hands in innocence, O Lord,\\
I’ll wash and purify;\\
So to thine holy altar go,\\
and compass it will I:

7 That I, with voice of thanksgiving,\\
may publish and declare,\\
And tell of all thy mighty works,\\
that great and wondrous are.

8 The habitation of thy house,\\
Lord, I have loved well;\\
Yea, in that place I do delight\\
where doth thine honour dwell.

9 With sinners gather not my soul,\\
and such as blood would spill:\\
10 Whose hands mischievous plots, right hand\\
corrupting bribes do fill.

11 But as for me, I will walk on\\
in mine integrity:\\
Do thou redeem me, and, O Lord,\\
be merciful to me.

12 My foot upon an even place\\
doth stand with stedfastness:\\
Within the congregations\\
th’ Eternal I will bless.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 27 \blfootnote{\emph{Psalm 27 Tune:} \sc{St. Botolph}}}
\addcontentsline{toc}{subsection}{Psalm 27}

A \emph{Psalm} of David.

1 The Lord’s my light and saving health,\\
who shall make me dismayed?\\
My life’s strength is the Lord, of whom\\
then shall I be afraid?

2 When as mine enemies and foes,\\
most wicked persons all,\\
To eat my flesh against me rose,\\
they stumbled and did fall.

3 Against me though an host encamp,\\
my heart yet fearless is:\\
Though war against me rise, I will\\
be confident in this.

4 One thing I of the Lord desired,\\
and will seek to obtain,\\
That all days of my life I may\\
within God’s house remain;

That I the beauty of the Lord\\
behold may and admire,\\
And that I in his holy place\\
may rev’rently enquire.

5 For he in his pavilion shall\\
me hide in evil days;\\
In secret of his tent me hide,\\
and on a rock me raise.

6 And now, ev’n at this present time,\\
mine head shall lifted be\\
Above all those that are my foes,\\
and round encompass me:

Therefore unto his tabernacle\\
I’ll sacrifices bring\\
Of joyfulness; I’ll sing, yea, I\\
to God will praises sing.

7 O Lord, give ear unto my voice,\\
when I do cry to thee;\\
Upon me also mercy have,\\
and do thou answer me.

8 When thou didst say, seek ye my face,\\
then unto thee reply\\
Thus did my heart, above all things\\
thy face, Lord, seek will I.

9 Far from me hide not thou thy face;\\
put not away from thee\\
Thy servant in thy wrath: thou hast\\
an helper been to me.

O God of my salvation,\\
leave me not, nor forsake:\\
10 Though me my parents both should leave,\\
the Lord will me up take.

11 O Lord, instruct me in thy way,\\
to me a leader be\\
In a plain path, because of those\\
that hatred bear to me.

12 Give me not to mine en’mies’ will;\\
for witnesses that lie\\
Against me risen are, and such\\
as breathe out cruelty.

13 I fainted had, unless that I\\
believed had to see\\
The Lord’s own goodness in the land\\
of them that living be.

14 Wait on the Lord, and be thou strong,\\
and he shall strength afford\\
Unto thine heart; yea, do thou wait,\\
I say, upon the Lord.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 28 \blfootnote{\emph{Psalm 28 Tune:} \sc{Caithness}}}
\addcontentsline{toc}{subsection}{Psalm 28}

A \emph{Psalm} of David.

1 To thee I’ll cry, O Lord, my rock;\\
hold not thy peace to me;\\
Lest like those that to pit descend\\
I by thy silence be.

2 The voice hear of my humble pray’rs,\\
when unto thee I cry;\\
When to thine holy oracle\\
I lift mine hands on high.

3 With ill men draw me not away\\
that work iniquity;\\
That speak peace to their friends, while in\\
their hearts doth mischief lie.

4 Give them according to their deeds\\
and ills endeavored:\\
And as their handy-works deserve,\\
to them be rendered.

5 God shall not build, but them destroy,\\
who would not understand\\
The Lord’s own works, nor did regard\\
the doing of his hand.

6 For ever blessed be the Lord,\\
for graciously he heard\\
The voice of my petitions,\\
and prayers did regard.

7 The Lord’s my strength and shield; my heart\\
upon him did rely;\\
And I am helped: hence my heart\\
doth joy exceedingly,

And with my song I will him praise.\\
8 My strength is God alone:\\
He also is the saving strength\\
of his anointed one.\\

9 O thine own people do thou save,\\
bless thine inheritance;\\
Them also do thou feed, and them\\
for evermore advance.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 29}
\addcontentsline{toc}{subsection}{Psalm 29}

A \emph{Psalm} of David.

1 Give ye unto the Lord, ye sons\\
that of the mighty be,\\
All strength and glory to the Lord\\
with cheerfulness give ye.

2 Unto the Lord the glory give\\
that to his name is due;\\
And in the beauty of holiness\\
unto Jehovah bow.

3 The Lord’s voice on the waters is;\\
the God of majesty\\
Doth thunder, and on multitudes\\
of waters sitteth he.

4 A pow’rful voice it is that comes\\
out from the Lord most high;\\
The voice of that great Lord is full\\
of glorious majesty.

5 The voice of the Eternal doth\\
asunder cedars tear;\\
Yea, God the Lord doth cedars break\\
that Lebanon doth bear.

6 He makes them like a calf to skip,\\
ev’n that great Lebanon,\\
And, like to a young unicorn,\\
the mountain Sirion.

7 God’s voice divides the flames of fire;\\
8 The desert it doth shake:\\
The Lord doth make the wilderness\\
of Kadesh all to quake.

9 God’s voice doth make the hinds to calve,\\
it makes the forest bare:\\
And in his temple ev’ry one\\
his glory doth declare.

10 The Lord sits on the floods; the Lord\\
sits King, and ever shall.\\
11 The Lord will give his people strength,\\
and with peace bless them all.\\

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 30}
\addcontentsline{toc}{subsection}{Psalm 30}

A \emph{Psalm} of David.

1 Lord, I will thee extol, for thou\\
hast lifted me on high,\\
And over me thou to rejoice\\
mad’st not mine enemy.

2 O thou who art the Lord my God,\\
I in distress to thee,\\
With loud cries lifted up my voice,\\
and thou hast healed me.

3 O Lord, my soul thou hast brought up,\\
and rescued from the grave;\\
That I to pit should not go down,\\
alive thou didst me save.

4 O ye that are his holy ones,\\
sing praise unto the Lord;\\
And give unto him thanks, when ye\\
his holiness record.

5 For but a moment lasts his wrath;\\
life in his favor lies:\\
Weeping may for a night endure,\\
at morn doth joy arise.

6 In my prosperity I said,\\
that nothing shall me move.\\
7 O Lord, thou hast my mountain made\\
to stand strong by thy love:

But when that thou, O gracious God,\\
didst hide thy face from me,\\
Then quickly was my prosp’rous state\\
turned into misery.

8 Wherefore unto the Lord my cry\\
I caused to ascend:\\
My humble supplication\\
I to the Lord did send.

9 What profit is there in my blood,\\
when I go down to pit?\\
Shall unto thee the dust give praise?\\
thy truth declare shall it?

10 Hear, Lord, have mercy; help me, Lord:\\
11 Thou turned hast my sadness\\
To dancing; yea, my sackcloth loosed,\\
and girded me with gladness;

12 That sing thy praise my glory may,\\
and never silent be.\\
O Lord my God, for evermore\\
I will give thanks to thee.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 31 \blfootnote{\emph{Psalm 31 Tune:} \sc{Consolation / St. Lawrence}}}
\addcontentsline{toc}{subsection}{Psalm 31}

To the chief Musician, A Psalm of David.

1 In thee, O Lord, I put my trust,\\
shamed let me never be;\\
According to thy righteousness\\
do thou deliver me.

2 Bow down thine ear to me, with speed\\
send me deliverance:\\
To save me, my strong rock be thou,\\
and my house of defense.

3 Because thou art my rock, and thee\\
I for my fortress take;\\
Therefore do thou me lead and guide,\\
ev’n for thine own name’s sake.

4 And since thou art my strength, therefore\\
pull me out of the net,\\
Which they in subtlety for me\\
so privily have set.

5 Into thine hands I do commit\\
my sp’rit: for thou art he,\\
O thou, Jehovah, God of truth,\\
that hast redeemed me.

6 Those that do lying vanities\\
regard, I have abhorred:\\
But as for me, my confidence\\
is fixed on the Lord.

7 I’ll in thy mercy gladly joy:\\
for thou my miseries\\
Considered hast; thou hast my soul\\
known in adversities:

8 And thou hast not enclosed me\\
within the en’my’s hand;\\
And by thee have my feet been made\\
in a large room to stand.

9 O Lord, upon me mercy have,\\
for trouble is on me:\\
Mine eye, my belly, and my soul,\\
with grief consumed be.

10 Because my life with grief is spent,\\
my years with sighs and groans:\\
My strength doth fail; and for my sin\\
consumed are my bones.

11 I was a scorn to all my foes,\\
and to my friends a fear;\\
And specially reproached of those\\
that were my neighbors near:

When they me saw they from me fled.\\
12 Ev’n so I am forgot,\\
As men are out of mind when dead:\\
I’m like a broken pot.

13 For slanders I of many heard;\\
fear compassed me, while they\\
Against me did consult, and plot\\
to take my life away.

14 But as for me, O Lord, my trust\\
upon thee I did lay;\\
And I to thee, Thou art my God,\\
did confidently say.

15 My times are wholly in thine hand:\\
do thou deliver me\\
From their hands that mine enemies\\
and persecutors be.

16 Thy countenance to shine do thou\\
upon thy servant make:\\
Unto me give salvation,\\
for thy great mercies’ sake.

17 Let me not be ashamed, O Lord,\\
for on thee called I have:\\
Let wicked men be shamed, let them\\
be silent in the grave.

18 To silence put the lying lips,\\
that grievous things do say,\\
And hard reports, in pride and scorn,\\
on righteous men do lay.

19 How great’s the goodness thou for them\\
that fear thee keep’st in store,\\
And wrought’st for them that trust in thee\\
the sons of men before!

20 In secret of thy presence thou\\
shalt hide them from man’s pride:\\
From strife of tongues thou closely shalt,\\
as in a tent, them hide.

21 All praise and thanks be to the Lord;\\
for he hath magnified\\
His wondrous love to me within\\
a city fortified.

22 For from thine eyes cut off I am,\\
I in my haste had said;\\
My voice yet heard’st thou, when to thee\\
with cries my moan I made.

23 O love the Lord, all ye his saints;\\
because the Lord doth guard\\
The faithful, and he plenteously\\
proud doers doth reward.

24 Be of good courage, and He strength\\
unto your heart shall send,\\
All ye whose hope and confidence\\
doth on the Lord depend.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 32 \blfootnote{\emph{Psalm 32 Tune:} \sc{Salzburg}}}
\addcontentsline{toc}{subsection}{Psalm 32}

A \emph{Psalm} of David, Maschil.

1 O blessed is the man to whom\\
is freely pardoned\\
All the transgression he hath done,\\
whose sin is covered.

2 Blessed is the man to whom the Lord\\
imputeth not his sin,\\
And in whose sp’rit there is no guile,\\
nor fraud is found therein.

3 When as I did refrain my speech,\\
and silent was my tongue,\\
My bones then waxed old, because\\
I roared all day long.

4 For upon me both day and night\\
thine hand did heavy lie,\\
So that my moisture turned is\\
in summer’s drought thereby.

5 I thereupon have unto thee\\
my sin acknowledged,\\
And likewise mine iniquity\\
I have not covered:

I will confess unto the Lord\\
my trespasses, said I;\\
And of my sin thou freely didst\\
forgive th’ iniquity.

6 For this shall ev’ry godly one\\
his prayer make to thee;\\
In such a time he shall thee seek,\\
as found thou mayest be.

Surely, when floods of waters great\\
do swell up to the brim,\\
They shall not overwhelm his soul,\\
nor once come near to him.

7 Thou art my hiding-place, thou shalt\\
from trouble keep me free:\\
Thou with songs of deliverance\\
about shalt compass me.

8 I will instruct thee, and thee teach\\
the way that thou shalt go;\\
And, with mine eye upon thee set,\\
I will direction show.

9 Then be not like the horse or mule,\\
which do not understand;\\
Whose mouth, lest they come near to thee,\\
a bridle must command.

10 Unto the man that wicked is\\
his sorrows shall abound;\\
But him that trusteth in the Lord\\
mercy shall compass round.

11 Ye righteous, in the Lord be glad,\\
in him do ye rejoice:\\
All ye that upright are in heart,\\
for joy lift up your voice.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 33 \blfootnote{\emph{Psalm 33 Tune:} \sc{Irish}}}
\addcontentsline{toc}{subsection}{Psalm 33}

1 Ye righteous, in the Lord rejoice;\\
it comely is and right,\\
That upright men, with thankful voice,\\
should praise the Lord of might.

2 Praise God with harp, and unto him\\
sing with the psaltery;\\
Upon a ten-stringed instrument\\
make ye sweet melody.

3 A new song to him sing, and play\\
with loud noise skilfully;\\
4 For right is God’s word, all his works\\
are done in verity.

5 To judgment and to righteousness\\
a love he beareth still;\\
The loving-kindness of the Lord\\
the earth throughout doth fill.

6 The heavens by the word of God\\
did their beginning take;\\
And by the breathing of his mouth\\
he all their hosts did make.

7 The waters of the seas he brings\\
together as an heap;\\
And in storehouses, as it were,\\
he layeth up the deep.

8 Let earth, and all that live therein,\\
with rev’rence fear the Lord;\\
Let all the world’s inhabitants\\
dread him with one accord.

9 For he did speak the word, and done\\
it was without delay;\\
Established it firmly stood,\\
whatever he did say.

10 God doth the counsel bring to naught\\
which heathen folk do take;\\
And what the people do devise\\
of none effect doth make.

11 O but the counsel of the Lord\\
doth stand for ever sure;\\
And of his heart the purposes\\
from age to age endure.

12 That nation blessed is, whose God\\
Jehovah is, and those\\
A blessed people are, whom for\\
his heritage he chose.

13 The Lord from heav’n sees and beholds\\
all sons of men full well:\\
14 He views all from his dwelling-place\\
that in the earth do dwell.

15 He forms their hearts alike, and all\\
their doings he observes.\\
16 Great hosts save not a king, much strength\\
no mighty man preserves.

17 An horse for preservation\\
is a deceitful thing;\\
And by the greatness of his strength\\
can no deliv’rance bring.

18 Behold on those that do him fear\\
the Lord doth set his eye;\\
Ev’n those who on his mercy do\\
with confidence rely.

19 From death to free their soul, in dearth\\
life unto them to yield.\\
20 Our soul doth wait upon the Lord;\\
he is our help and shield.

21 Since in his holy name we trust,\\
our heart shall joyful be.\\
22 Lord, let thy mercy be on us,\\
as we do hope in thee.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 34 \blfootnote{\emph{Psalm 34 Tune:} \sc{St. Matthew \ Wiltshire}}}
\addcontentsline{toc}{subsection}{Psalm 34}

\emph{A Psalm} of David, when he changed his behaviour before Abimelech; who drove him away, and he departed.

1 God will I bless all times; his praise\\
my mouth shall still express.\\
2 My soul shall boast in God: the meek\\
shall hear with joyfulness.

3 Extol the Lord with me, let us\\
exalt his name together.\\
4 I sought the Lord, he heard, and did\\
me from all fears deliver.

5 They looked to him, and lightened were:\\
not shamed were their faces.\\
6 This poor man cried, God heard, and saved\\
him from all his distresses.

7 The angel of the Lord encamps,\\
and round encompasseth\\
All those about that do him fear,\\
and them delivereth.

8 O taste and see that God is good:\\
who trusts in him is blessed.\\
9 Fear God his saints: none that him fear\\
shall be with want oppressed.

10 The lions young may hungry be,\\
and they may lack their food:\\
But they that truly seek the Lord\\
shall not lack any good.

11 O children, hither do ye come,\\
and unto me give ear;\\
I shall you teach to understand\\
how ye the Lord should fear.

12 What man is he that life desires,\\
to see good would live long?\\
13 Thy lips refrain from speaking guile,\\
and from ill words thy tongue.

14 Depart from ill, do good, seek peace,\\
pursue it earnestly.\\
15 God’s eyes are on the just; his ears\\
are open to their cry.

16 The face of God is set against\\
those that do wickedly,\\
That he may quite out from the earth\\
cut off their memory.

17 The righteous cry unto the Lord,\\
he unto them gives ear;\\
And they out of their troubles all\\
by him delivered are.

18 The Lord is ever nigh to them\\
that be of broken sp’rit;\\
To them he safety doth afford\\
that are in heart contrite.

19 The troubles that afflict the just\\
in number many be;\\
But yet at length out of them all\\
the Lord doth set him free.

20 He carefully his bones doth keep,\\
whatever can befall;\\
That not so much as one of them\\
can broken be at all.

21 Ill shall the wicked slay; laid waste\\
shall be who hate the just.\\
22 The Lord redeems his servants’ souls;\\
none perish that him trust.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 35}
\addcontentsline{toc}{subsection}{Psalm 35}

\emph{A Psalm} of David.

1 Plead, Lord, with those that plead; and fight\\
with those that fight with me.\\
2 Of shield and buckler take thou hold,\\
stand up mine help to be.

3 Draw also out the spear, and do\\
against them stop the way\\
That me pursue: unto my soul,\\
I’m thy salvation, say.

4 Let them confounded be and shamed\\
that for my soul have sought:\\
Who plot my hurt turned back be they,\\
and to confusion brought.

5 Let them be like unto the chaff\\
that flies before the wind;\\
And let the angel of the Lord\\
pursue them hard behind.

6 With darkness cover thou their way,\\
and let it slipp’ry prove;\\
And let the angel of the Lord\\
pursue them from above.

7 For without cause have they for me\\
their net hid in a pit,\\
They also have without a cause\\
for my soul digged it.

8 Let ruin seize him unawares;\\
his net he hid withal\\
Himself let catch; and in the same\\
destruction let him fall.

9 My soul in God shall joy; and glad\\
in his salvation be:\\
10 And all my bones shall say, O Lord,\\
who is like unto thee,

Which dost the poor set free from him\\
that is for him too strong;\\
The poor and needy from the man\\
that spoils and does him wrong?

11 False witnesses rose; to my charge\\
things I not knew they laid.\\
12 They, to the spoiling of my soul,\\
me ill for good repaid.

13 But as for me, when they were sick,\\
in sackcloth sad I mourned:\\
My humbled soul did fast, my pray’r\\
into my bosom turned.

14 Myself I did behave as he\\
had been my friend or brother;\\
I heavily bowed down, as one\\
that mourneth for his mother.

15 But in my trouble they rejoiced,\\
gath’ring themselves together;\\
Yea, objects vile together did\\
themselves against me gather:

I knew it not; they did me tear,\\
and quiet would not be.\\
16 With mocking hypocrites, at feasts\\
they gnashed their teeth at me.

17 How long, Lord, look’st thou on? from those\\
destructions they intend\\
Rescue my soul, from lions young\\
my darling do defend.

18 I will give thanks to thee, O Lord,\\
within th’ assembly great;\\
And where much people gathered are\\
thy praises forth will set.

19 Let not my wrongful enemies\\
proudly rejoice o’er me;\\
Nor who me hate without a cause,\\
let them wink with the eye.

20 For peace they do not speak at all;\\
but crafty plots prepare\\
Against all those within the land\\
that meek and quiet are.

21 With mouths set wide, they ‘gainst me said,\\
Ha, ha! our eye doth see.\\
22 Lord, thou hast seen, hold not thy peace;\\
Lord, be not far from me.

23 Stir up thyself; wake, that thou may’st\\
judgment to me afford,\\
Ev’n to my cause, O thou that art\\
my only God and Lord.

24 O Lord my God, do thou me judge\\
after thy righteousness;\\
And let them not their joy ‘gainst me\\
triumphantly express:

25 Nor let them say within their hearts,\\
Ah, we would have it thus;\\
Nor suffer them to say, that he\\
is swallowed up by us.

26 Shamed and confounded be they all\\
that at my hurt are glad;\\
Let those against me that do boast\\
with shame and scorn be clad.

27 Let them that love my righteous cause\\
be glad, shout, and not cease\\
To say, The Lord be magnified,\\
who loves his servant’s peace.

28 Thy righteousness shall also be\\
declared by my tongue;\\
The praises that belong to thee\\
speak shall it all day long.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 36 \blfootnote{\emph{Psalm 36 Tune:} \sc{Elijah}}}
\addcontentsline{toc}{subsection}{Psalm 36}

To the chief Musician,
\emph{A Psalm} of David, the servant of the Lord.

1 The wicked man’s transgression\\
within my heart thus says,\\
Undoubtedly the fear of God\\
is not before his eyes.

2 Because himself he flattereth\\
in his own blinded eye,\\
Until the hatefulness be found\\
of his iniquity.

3 Words from his mouth proceeding are,\\
fraud and iniquity:\\
He to be wise, and to do good,\\
hath left off utterly.

4 He mischief, lying on his bed,\\
most cunningly doth plot:\\
He sets himself in ways not good,\\
ill he abhorreth not.

5 Thy mercy, Lord, is in the heav’ns;\\
thy truth doth reach the clouds:\\
6 Thy justice is like mountains great;\\
thy judgments deep as floods:

Lord, thou preservest man and beast.\\
7 How precious is thy grace!\\
Therefore in shadow of thy wings\\
men’s sons their trust shall place.

8 They with the fatness of thy house\\
shall be well satisfied;\\
From rivers of thy pleasures thou\\
wilt drink to them provide.

9 Because of life the fountain pure\\
remains alone with thee;\\
And in that purest light of thine\\
we clearly light shall see.

10 Thy loving-kindness unto them\\
continue that thee know;\\
And still on men upright in heart\\
thy righteousness bestow.

11 Let not the foot of cruel pride\\
come, and against me stand;\\
And let me not removed be,\\
Lord, by the wicked’s hand.

12 There fallen are they, and ruined,\\
that work iniquities:\\
Cast down they are, and never shall\\
be able to arise.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 37 \blfootnote{\emph{Psalm 37 Tunes:} \sc{Azmon / Denfield}}}
\addcontentsline{toc}{subsection}{Psalm 37}

A Psalm of David.

1 For evil-doers fret thou not\\
thyself unquietly;\\
Nor do thou envy bear to those\\
that work iniquity.

2 For, even like unto the grass,\\
soon be cut down shall they;\\
And, like the green and tender herb,\\
they wither shall away.

3 Set thou thy trust upon the Lord,\\
and be thou doing good;\\
And so thou in the land shalt dwell,\\
and verily have food.

4 Delight thyself in God; he’ll give\\
thine heart’s desire to thee.\\
5 Thy way to God commit, him trust,\\
it bring to pass shall he.

6 And, like unto the light, he shall\\
thy righteousness display;\\
And he thy judgment shall bring forth\\
like noon-tide of the day.

7 Rest in the Lord, and patiently\\
wait for him: do not fret\\
For him who, prosp’ring in his way,\\
success in sin doth get.

8 Do thou from anger cease, and wrath\\
see thou forsake also:\\
Fret not thyself in any wise,\\
that evil thou should’st do.

9 For those that evil doers are\\
shall be cut off and fall:\\
But those that wait upon the Lord\\
the earth inherit shall.

10 For yet a little while, and then\\
the wicked shall not be;\\
His place thou shalt consider well,\\
but it thou shalt not see.

11 But by inheritance the earth\\
the meek ones shall possess:\\
They also shall delight themselves\\
in an abundant peace.

12 The wicked plots against the just,\\
and at him whets his teeth:\\
13 The Lord shall laugh at him, because\\
his day he coming seeth.

14 The wicked have drawn out the sword,\\
and bent their bow, to slay\\
The poor and needy, and to kill\\
men of an upright way.

15 But their own sword, which they have drawn,\\
shall enter their own heart:\\
Their bows which they have bent shall break,\\
and into pieces part.

16 A little that a just man hath\\
is more and better far\\
Than is the wealth of many such\\
as lewd and wicked are.

17 For sinners’ arms shall broken be;\\
but God the just sustains.\\
18 God knows the just man’s days, and still\\
their heritage remains.

19 They shall not be ashamed when they\\
the evil time do see;\\
And when the days of famine are,\\
they satisfied shall be.

20 But wicked men, and foes of God,\\
as fat of lambs, decay;\\
They shall consume, yea, into smoke\\
they shall consume away.

21 The wicked borrows, but the same\\
again he doth not pay;\\
Whereas the righteous mercy shows,\\
and gives his own away.

22 For such as blessed be of him\\
the earth inherit shall;\\
And they that cursed are of him\\
shall be destroyed all.

23 A good man’s footsteps by the Lord\\
are ordered aright;\\
And in the way wherein he walks\\
he greatly doth delight.

24 Although he fall, yet shall he not\\
be cast down utterly;\\
Because the Lord with his own hand\\
upholds him mightily.

25 I have been young, and now am old,\\
yet have I never seen\\
The just man left, nor that his seed\\
for bread have beggars been.

26 He’s ever merciful, and lends:\\
his seed is blessed therefore.\\
27 Depart from evil, and do good,\\
and dwell for evermore.

28 For God loves judgment, and his saints\\
leaves not in any case;\\
They are kept ever: but cut off\\
shall be the sinner’s race.

29 The just inherit shall the land,\\
and ever in it dwell:\\
30 The just man’s mouth doth wisdom speak;\\
his tongue doth judgment tell.

31 In’s heart the law is of his God,\\
his steps slide not away.\\
32 The wicked man doth watch the just,\\
and seeketh him to slay.

33 Yet him the Lord will not forsake,\\
nor leave him in his hands:\\
The righteous will he not condemn,\\
when he in judgment stands.

34 Wait on the Lord, and keep his way,\\
and thee exalt shall he\\
Th’ earth to inherit; when cut off\\
the wicked thou shalt see.

35 I saw the wicked great in pow’r,\\
spread like a green bay-tree:\\
36 He passed, yea, was not; him I sought,\\
but found he could not be.

37 Mark thou the perfect, and behold\\
the man of uprightness;\\
Because that surely of this man\\
the latter end is peace.

38 But those men that transgressors are\\
shall be destroyed together;\\
The latter end of wicked men\\
shall be cut off for ever.

39 But the salvation of the just\\
is from the Lord above;\\
He in the time of their distress\\
their stay and strength doth prove.

40 The Lord shall help, and them deliver:\\
he shall them free and save\\
From wicked men; because in him\\
their confidence they have.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 38}
\addcontentsline{toc}{subsection}{Psalm 38}

A Psalm of David, to bring to remembrance.

1 In thy great indignation,\\
O Lord, rebuke me not;\\
Nor on me lay thy chast’ning hand,\\
in thy displeasure hot.

2 For in me fast thine arrows stick,\\
thine hand doth press me sore:\\
3 And in my flesh there is no health,\\
nor soundness any more.

This grief I have, because thy wrath\\
is forth against me gone;\\
And in my bones there is no rest,\\
for sin that I have done.

4 Because gone up above mine head\\
my great transgressions be;\\
And, as a weighty burden, they\\
too heavy are for me.

5 My wounds do stink, and are corrupt;\\
my folly makes it so.\\
6 I troubled am, and much bowed down;\\
all day I mourning go.

7 For a disease that loathsome is\\
so fills my loins with pain,\\
That in my weak and weary flesh\\
no soundness doth remain.

8 So feeble and infirm am I,\\
and broken am so sore,\\
That, through disquiet of my heart,\\
I have been made to roar.

9 O Lord, all that I do desire\\
is still before thine eye;\\
And of my heart the secret groans\\
not hidden are from thee.

10 My heart doth pant incessantly,\\
my strength doth quite decay;\\
As for mine eyes, their wonted light\\
is from me gone away.

11 My lovers and my friends do stand\\
at distance from my sore;\\
And those do stand aloof that were\\
kinsmen and kind before.

12 Yea, they that seek my life lay snares:\\
who seek to do me wrong\\
Speak things mischievous, and deceits\\
imagine all day long.

13 But, as one deaf, that heareth not,\\
I suffered all to pass;\\
I as a dumb man did become,\\
whose mouth not opened was:

14 As one that hears not, in whose mouth\\
are no reproofs at all.\\
15 For, Lord, I hope in thee; my God,\\
thou’lt hear me when I call.

16 For I said, Hear me, lest they should\\
rejoice o’er me with pride;\\
And o’er me magnify themselves,\\
when as my foot doth slide.

17 For I am near to halt, my grief\\
is still before mine eye:\\
18 For I’ll declare my sin, and grieve\\
for mine iniquity.

19 But yet mine en’mies lively are,\\
and strong are they beside;\\
And they that hate me wrongfully\\
are greatly multiplied.

20 And they for good that render ill,\\
as en’mies me withstood;\\
Yea, ev’n for this, because that I\\
do follow what is good.

21 Forsake me not, O Lord; my God,\\
far from me never be.\\
22 O Lord, thou my salvation art,\\
haste to give help to me.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 39}
\addcontentsline{toc}{subsection}{Psalm 39}

To the chief Musician, even to Jeduthan,\\
A Psalm of David.

1 I said, I will look to my ways,\\
lest with my tongue I sin:\\
In sight of wicked men my mouth\\
with bridle I’ll keep in.

2 With silence I as dumb became,\\
I did myself restrain\\
From speaking good; but then the more\\
increased was my pain.

3 My heart within me waxed hot;\\
and, while I musing was,\\
The fire did burn; and from my tongue\\
these words I did let pass:

4 Mine end, and measure of my days,\\
O Lord, unto me show\\
What is the same; that I thereby\\
my frailty well may know.

5 Lo, thou my days an handbreadth mad’st;\\
mine age is in thine eye\\
As nothing: sure each man at best\\
is wholly vanity.

6 Sure each man walks in a vain show;\\
they vex themselves in vain:\\
He heaps up wealth, and doth not know\\
to whom it shall pertain.

7 And now, O Lord, what wait I for?\\
my hope is fix’d on thee.\\
8 Free me from all my trespasses,\\
the fool’s scorn make not me.

9 Dumb was I, op’ning not my mouth,\\
because this work was thine.\\
10 Thy stroke take from me; by the blow\\
of thine hand I do pine.

11 When with rebukes thou dost correct\\
man for iniquity,\\
Thou wastes his beauty like a moth:\\
sure each man’s vanity.

12 Attend my cry, Lord, at my tears\\
and pray’rs not silent be:\\
I sojourn as my fathers all,\\
and stranger am with thee.

13 O spare thou me, that I my strength\\
recover may again,\\
Before from hence I do depart,\\
and here no more remain.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 40}
\addcontentsline{toc}{subsection}{Psalm 40}

To the chief Musician,
A Psalm of David.

1 I waited for the Lord my God,\\
and patiently did bear;\\
At length to me he did incline\\
my voice and cry to hear.

2 He took me from a fearful pit,\\
and from the miry clay,\\
And on a rock he set my feet,\\
establishing my way.

3 He put a new song in my mouth,\\
our God to magnify:\\
Many shall see it, and shall fear,\\
and on the Lord rely.

4 O blessed is the man whose trust\\
upon the Lord relies;\\
Respecting not the proud, nor such\\
as turn aside to lies.

5 O Lord my God, full many are\\
the wonders thou hast done;\\
Thy gracious thoughts to us-ward far\\
above all thoughts are gone:

In order none can reckon them\\
to thee: if them declare,\\
And speak of them I would, they more\\
than can be numbered are.

6 No sacrifice nor offering\\
didst thou at all desire;\\
Mine ears thou bored: sin-off ‘ring thou\\
and burnt didst not require:

7 Then to the Lord these were my words,\\
I come, behold and see;\\
Within the volume of the book\\
it written is of me:

8 To do thy will I take delight,\\
O thou my God that art;\\
Yea, that most holy law of thine\\
I have within my heart.

9 Within the congregation great\\
I righteousness did preach:\\
Lo, thou dost know, O Lord, that I\\
refrained not my speech.

10 I never did within my heart\\
conceal thy righteousness;\\
I thy salvation have declared,\\
and shown thy faithfulness:

Thy kindness, which most loving is,\\
concealed have not I,\\
Nor from the congregation great\\
have hid thy verity.

11 Thy tender mercies, Lord, from me\\
O do thou not restrain;\\
Thy loving-kindness, and thy truth,\\
let them me still maintain.

12 For ills past reck’ning compass me,\\
and mine iniquities\\
Such hold upon me taken have,\\
I cannot lift mine eyes:

They more than hairs are on mine head,\\
thence is my heart dismayed.\\
13 Be pleased, Lord, to rescue me;\\
Lord, hasten to mine aid.

14 Shamed and confounded be they all\\
that seek my soul to kill;\\
Yea, let them backward driven be,\\
and shamed, that wish me ill.

15 For a reward of this their shame\\
confounded let them be.\\
That in this manner scoffing say,\\
Aha, aha! to me.

16 In thee let all be glad, and joy,\\
who seeking thee abide;\\
Who thy salvation love, say still,\\
The Lord be magnified.

17 I’m poor and needy, yet the Lord\\
of me a care doth take:\\
Thou art my help and savior,\\
my God, no tarrying make.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 41}
\addcontentsline{toc}{subsection}{Psalm 41}

To the chief Musician,
A Psalm of David.

1 Blessed is he that wisely doth\\
the poor man’s case consider;\\
For when the time of trouble is,\\
the Lord will him deliver.

2 God will him keep, yea, save alive;\\
on earth he blessed shall live;\\
And to his enemies’ desire\\
thou wilt him not up give.

3 God will give strength when he on bed\\
of languishing doth mourn;\\
And in his sickness sore, O Lord,\\
thou all his bed wilt turn.

4 I said, O Lord, do thou extend\\
thy mercy unto me;\\
O do thou heal my soul; for why?\\
I have offended thee.

5 Those that to me are enemies,\\
of me do evil say,\\
When shall he die, that so his name\\
may perish quite away?

6 To see me if he comes, he speaks\\
vain words: but then his heart\\
Heaps mischief to it, which he tells,\\
when forth he doth depart.

7 My haters jointly whispering,\\
‘gainst me my hurt devise.\\
8 Mischief, say they, cleaves fast to him;\\
he li’th, and shall not rise.

9 Yea, ev’n mine own familiar friend,\\
on whom I did rely,\\
Who ate my bread, ev’n he his heel\\
against me lifted high.

10 But, Lord, be merciful to me,\\
and up again me raise,\\
That I may justly them requite\\
according to their ways.

11 By this I know that certainly\\
I favored am by thee;\\
Because my hateful enemy\\
triumphs not over me.

12 But as for me, thou me uphold’st\\
in mine integrity;\\
And me before thy countenance\\
thou sett’st continually.

13 The Lord, the God of Israel,\\
be blessed for ever then,\\
From age to age eternally.\\
Amen, yea, and amen.
% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\end{multicols}

\begin{center}
\Large{\sc{Book II}}
\addcontentsline{toc}{section}{Book II}
\end{center}
\begin{multicols}{3}


\subsection*{Psalm 42}
\addcontentsline{toc}{subsection}{Psalm 42}

To the chief Musician, Maschil, for the sons of Korah.

1 Like as the hart for water-brooks\\
in thirst doth pant and bray;\\
So pants my longing soul, O God,\\
that come to thee I may.

2 My soul for God, the living God,\\
doth thirst: when shall I near\\
Unto thy countenance approach,\\
and in God’s sight appear?

3 My tears have unto me been meat,\\
both in the night and day,\\
While unto me continually,\\
Where is thy God? they say.

4 My soul is poured out in me,\\
when this I think upon;\\
Because that with the multitude\\
I heretofore had gone:

With them into God’s house I went,\\
with voice of joy and praise;\\
Yea, with the multitude that kept\\
the solemn holy days.

5 O why art thou cast down, my soul?\\
why in me so dismayed?\\
Trust God, for I shall praise him yet,\\
his count’nance is mine aid.

6 My God, my soul’s cast down in me;\\
thee therefore mind I will\\
From Jordan’s land, the Hermonites,\\
and ev’n from Mizar hill.

7 At the noise of thy water-spouts\\
deep unto deep doth call;\\
Thy breaking waves pass over me,\\
yea, and thy billows all.

8 His loving-kindness yet the Lord\\
command will in the day,\\
His song’s with me by night; to God,\\
by whom I live, I’ll pray:

9 And I will say to God my rock,\\
Why me forgett’st thou so?\\
Why, for my foes’ oppression,\\
thus mourning do I go?

10 ‘Tis as a sword within my bones,\\
when my foes me upbraid;\\
Ev’n when by them, Where is thy God?\\
’tis daily to me said.

11 O why art thou cast down, my soul?\\
why, thus with grief oppresed,\\
Art thou disquieted in me?\\
in God still hope and rest:

For yet I know I shall him praise,\\
who graciously to me\\
The health is of my countenance,\\
yea, mine own God is 

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 43}
\addcontentsline{toc}{subsection}{Psalm 43}

1 Judge me, O God, and plead my cause\\
against th’ ungodly nation;\\
From the unjust and crafty man,\\
O be thou my salvation.

2 For thou the God art of my strength;\\
why thrusts thou me thee fro’?\\
For th’ enemy’s oppression\\
why do I mourning go?

3 O send thy light forth and thy truth;\\
let them be guides to me,\\
And bring me to thine holy hill,\\
ev’n where thy dwellings be.

4 Then will I to God’s altar go,\\
to God my chiefest joy:\\
Yea, God, my God, thy name to praise\\
my harp I will employ.

5 Why art thou then cast down, my soul?\\
what should discourage thee?\\
And why with vexing thoughts art thou\\
disquieted in me?

Still trust in God; for him to praise\\
good cause I yet shall have:\\
He of my count’nance is the health,\\
my God that doth me save.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 44}
\addcontentsline{toc}{subsection}{Psalm 44}

To the chief Musician
for the sons of Korah, Maschil.

1 O God, we with our ears have heard,\\
our fathers have us told,\\
What works thou in their days hadst done,\\
ev’n in the days of old.

2 Thy hand did drive the heathen out,\\
and plant them in their place;\\
Thou didst afflict the nations,\\
but them thou didst increase.

3 For neither got their sword the land,\\
nor did their arm them save;\\
But thy right hand, arm, countenance;\\
for thou them favor gave.

4 Thou art my King: for Jacob, Lord,\\
deliv’rances command.\\
5 Through thee we shall push down our foes,\\
that do against us stand:

We, through thy name, shall tread down those\\
that ris’n against us have.\\
6 For in my bow I shall not trust,\\
nor shall my sword me save.

7 But from our foes thou hast us saved,\\
our haters put to shame.\\
8 In God we all the day do boast,\\
and ever praise thy name.

9 But now we are cast off by thee,\\
and us thou putt’st to shame;\\
And when our armies do go forth,\\
thou go’st not with the same.

10 Thou mak’st us from the enemy,\\
faint-hearted, to turn back;\\
And they who hate us for themselves\\
our spoils away do take.

11 Like sheep for meat thou gavest us;\\
‘mong heathen cast we be.\\
12 Thou didst for nought thy people sell;\\
their price enrich’d not thee.

13 Thou mak’st us a reproach to be\\
unto our neighbors near;\\
Derision and a scorn to them\\
that round about us are.

14 A by-word also thou dost us\\
among the heathen make;\\
The people, in contempt and spite,\\
at us their heads do shake.

15 Before me my confusion\\
continually abides;\\
And of my bashful countenance\\
the shame me ever hides:

16 For voice of him that doth reproach,\\
and speaketh blasphemy;\\
By reason of th’ avenging foe,\\
and cruel enemy.

17 All this is come on us, yet we\\
have not forgotten thee;\\
Nor falsely in thy covenant\\
behaved ourselves have we.

18 Back from thy way our heart not turned;\\
our steps no straying made;\\
19 Though us thou brak’st in dragons’ place,\\
and cover’dst with death’s shade.

20 If we God’s name forgot, or stretched\\
to a strange god our hands,\\
21 Shall not God search this out? for he\\
heart’s secrets understands.

22 Yea, for thy sake we’re killed all day,\\
counted as slaughter-sheep.\\
23 Rise, Lord, cast us not ever off;\\
awake, why dost thou sleep?

24 O wherefore hidest thou thy face?\\
forgett’st our cause distressed,\\
25 And our oppression? For our soul\\
is to the dust down pressed:

Our belly also on the earth\\
fast cleaving, hold doth take.\\
26 Rise for our help, and us redeem,\\
ev’n for thy mercies’ sake.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{First Version (C.M., 8.6.8.6)}

\subsection*{Psalm 45}
\addcontentsline{toc}{subsection}{Psalm 45}

To the chief Musician upon Shoshannim,
for the sons of Korah, Maschil, A Song of loves.

1 My heart brings forth a goodly thing;\\
my words that I indite\\
Concern the King: my tongue’s a pen\\
of one that swift doth write.

2 Thou fairer art than sons of men:\\
into thy lips is store\\
Of grace infused; God therefore thee\\
hath blessed for evermore.

3 O thou that art the mighty One,\\
thy sword gird on thy thigh;\\
Ev’n with thy glory excellent,\\
and with thy majesty.

4 For meekness, truth, and righteousness,\\
in state ride prosp’rously;\\
And thy right hand shall thee instruct\\
in things that fearful be.

5 Thine arrows sharply pierce the heart\\
of th’ en’mies of the King;\\
And under thy subjection\\
the people down do bring.

6 For ever and for ever is,\\
O God, thy throne of might;\\
The scepter of thy kingdom is\\
a scepter that is right.

7 Thou lovest right, and hatest ill;\\
for God, thy God, most high,\\
Above thy fellows hath with th’ oil\\
of joy anointed thee.

8 Of aloes, myrrh, and cassia,\\
a smell thy garments had,\\
Out of the iv’ry palaces,\\
whereby they made thee glad.

9 Among thy women honorable\\
kings’ daughters were at hand:\\
Upon thy right hand did the queen\\
in gold of Ophir stand.

10 O daughter, hearken and regard,\\
and do thine ear incline;\\
Likewise forget thy father’s house,\\
and people that are thine.

11 Then of the King desired shall be\\
thy beauty veh’mently:\\
Because he is thy Lord, do thou\\
him worship rev’rently.

12 The daughter there of Tyre shall be\\
with gifts and off ‘rings great:\\
Those of the people that are rich\\
thy favor shall entreat.

13 Behold, the daughter of the King\\
all glorious is within;\\
And with embroideries of gold\\
her garments wrought have been.

14 She shall be brought unto the King\\
in robes with needle wrought;\\
Her fellow-virgins following\\
shall unto thee be brought.

15 They shall be brought with gladness great,\\
and mirth on ev’ry side,\\
Into the palace of the King,\\
and there they shall abide.

16 Instead of those thy fathers dear,\\
thy children thou may’st take,\\
And in all places of the earth\\
them noble princes make.

17 Thy name remembered I will make\\
through ages all to be:\\
The people therefore evermore\\
shall praises give to thee.

 % Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (S.M, 6.6.8.6)}

To the chief Musician upon Shoshannim,
for the sons of Korah, Maschil, A Song of loves.

1 My heart inditing is\\
good matter in a song:\\
I speak the things that I have made,\\
which to the King belong:

My tongue shall be as quick,\\
his honour to indite,\\
As is the pen of any scribe\\
that useth fast to write.

2 Thou’rt fairest of all men;\\
grace in thy lips doth flow:\\
And therefore blessings evermore\\
on thee doth God bestow.

3 Thy sword gird on thy thigh,\\
thou that art most of might:\\
Appear in dreadful majesty,\\
and in thy glory bright.

4 For meekness, truth, and right,\\
ride prosp’rously in state;\\
And thy right hand shall teach to thee\\
things terrible and great.

5 Thy shafts shall pierce their hearts\\
that foes are to the King;\\
Whereby into subjection\\
the people thou shalt bring.

6 Thy royal seat, O Lord,\\
for ever shall remain:\\
The scepter of thy kingdom doth\\
all righteousness maintain.

7 Thou lov’st right, and hat’st ill;\\
for God, thy God, most high,\\
Above thy fellows hath with th’ oil\\
of joy anointed thee.

8 Of myrrh and spices sweet\\
a smell thy garments had,\\
Out of the iv’ry palaces,\\
whereby they made thee glad.

9 And in thy glorious train\\
kings’ daughters waiting stand;\\
And thy fair queen, in Ophir gold,\\
doth stand at thy right hand.

10 O daughter, take good heed,\\
incline, and give good ear;\\
Thou must forget thy kindred all,\\
and father’s house most dear.

11 Thy beauty to the King\\
shall then delightful be:\\
And do thou humbly worship him,\\
because thy Lord is he.

12 The daughter then of Tyre\\
there with a gift shall be,\\
And all the wealthy of the land\\
shall make their suit to thee.

13 The daughter of the King\\
all glorious is within;\\
And with embroideries of gold\\
her garments wrought have been.

14 She cometh to the King\\
in robes with needle wrought;\\
The virgins that do follow her\\
shall unto thee be brought.

15 They shall be brought with joy,\\
and mirth on ev’ry side,\\
Into the palace of the King,\\
and there they shall abide.

16 And in thy fathers’ stead,\\
thy children thou may’st take,\\
And in all places of the earth\\
them noble princes make.

17 I will show forth thy name\\
to generations all:\\
Therefore the people evermore\\
to thee give praises shal

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 46}
\addcontentsline{toc}{subsection}{Psalm 46}

To the chief Musician, for the sons of Korah,
A Song upon Alomoth.

1 God is our refuge and our strength,\\
in straits a present aid;\\
2 Therefore, although the earth remove,\\
we will not be afraid:

Though hills amidst the seas be cast;\\
3 Though waters roaring make,\\
And troubled be; yea, though the hills,\\
by swelling seas do shake.

4 A river is, whose streams do glad\\
the city of our God;\\
The holy place, wherein the Lord\\
most high hath his abode.

5 God in the midst of her doth dwell;\\
nothing shall her remove:\\
The Lord to her an helper will,\\
and that right early, prove.

6 The heathen raged tumultuously,\\
the kingdoms moved were:\\
The Lord God uttered his voice,\\
the earth did melt for fear.

7 The Lord of hosts upon our side\\
doth constantly remain:\\
The God of Jacob’s our refuge,\\
us safely to maintain.

8 Come, and behold what wondrous works\\
have by the Lord been wrought;\\
Come, see what desolations\\
he on the earth hath brought.

9 Unto the ends of all the earth\\
wars into peace he turns:\\
The bow he breaks, the spear he cuts,\\
in fire the chariot burns.

10 Be still, and know that I am God;\\
among the heathen I\\
Will be exalted; I on earth\\
will be exalted high.

11 Our God, who is the Lord of hosts,\\
is still upon our side;\\
The God of Jacob our refuge\\
for ever will abide.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 47}
\addcontentsline{toc}{subsection}{Psalm 47}

To the chief Musician,
A Psalm for the sons of Korah.

1 All people, clap your hands; to God\\
with voice of triumph shout:\\
2 For dreadful is the Lord most high,\\
great King the earth throughout.

3 The heathen people under us\\
he surely shall subdue;\\
And he shall make the nations\\
under our feet to bow.

4 The lot of our inheritance\\
choose out for us shall he,\\
Of Jacob, whom he loved well,\\
ev’n the excellency.

5 God is with shouts gone up, the Lord\\
with trumpets sounding high.\\
6 Sing praise to God, sing praise, sing praise,\\
praise to our King sing ye.

7 For God is King of all the earth;\\
with knowledge praise express.\\
8 God rules the nations: God sits on\\
his throne of holiness.

9 The princes of the people are\\
assembled willingly;\\
Ev’n of the God of Abraham\\
they who the people be.

For why? the shields that do defend\\
the earth are only his:\\
They to the Lord belong; yea, he\\
exalted greatly is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 48}
\addcontentsline{toc}{subsection}{Psalm 48}

A Song \emph{and} Psalm for the sons of Korah.

1 Great is the Lord, and greatly he\\
is to be praised still,\\
Within the city of our God,\\
upon his holy hill.

2 Mount Zion stands most beautiful,\\
the joy of all the land;\\
The city of the mighty King\\
on her north side doth stand.

3 The Lord within her palaces\\
is for a refuge known.\\
4 For, lo, the kings that gathered were\\
together, by have gone.

5 But when they did behold the same,\\
they, wond’ring, would not stay;\\
But, being troubled at the sight,\\
they thence did haste away.

6 Great terror there took hold on them;\\
they were possessed with fear;\\
Their grief came like a woman’s pain,\\
when she a child doth bear.

7 Thou Tarshish ships with east wind break’st:\\
8 As we have heard it told,\\
So, in the city of the Lord,\\
our eyes did it behold;

In our God’s city, which his hand\\
for ever stablish will.\\
9 We of thy loving-kindness thought,\\
Lord, in thy temple still.

10 O Lord, according to thy name,\\
through all the earth’s thy praise;\\
And thy right hand, O Lord, is full\\
of righteousness always.

11 Because thy judgments are made known,\\
let Zion mount rejoice;\\
Of Judah let the daughters all\\
send forth a cheerful voice.

12 Walk about Zion, and go round;\\
the high tow’rs thereof tell:\\
13 Consider ye her palaces,\\
and mark her bulwarks well;

That ye may tell posterity.\\
14 For this God doth abide\\
Our God for evermore; he will\\
ev’n unto death us guide.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 49}
\addcontentsline{toc}{subsection}{Psalm 49}

To the chief Musician,
A Psalm for the sons of Korah.

1 Hear this, all people, and give ear,\\
all in the world that dwell;\\
2 Both low and high, both rich and poor.\\
3 My mouth shall wisdom tell:

My heart shall knowledge meditate.\\
4 I will incline mine ear\\
To parables, and on the harp\\
my sayings dark declare.

5 Amidst those days that evil be,\\
why should I, fearing, doubt?\\
When of my heels th’ iniquity\\
shall compass me about.

6 Whoe’er they be that in their wealth\\
their confidence do pitch,\\
And boast themselves, because they are\\
become exceeding rich:

7 Yet none of these his brother can\\
redeem by any way;\\
Nor can he unto God for him\\
sufficient ransom pay,

8 (Their soul’s redemption precious is,\\
and it can never be,)\\
9 That still he should for ever live,\\
and not corruption see.

10 For why? he seeth that wise men die,\\
and brutish fools also\\
Do perish; and their wealth, when dead,\\
to others they let go.

11 Their inward thought is, that their house\\
and dwelling-places shall\\
Stand through all ages; they their lands\\
by their own names do call.

12 But yet in honor shall not man\\
abide continually;\\
But passing hence, may be compared\\
unto the beasts that die.

13 Thus brutish folly plainly is\\
their wisdom and their way;\\
Yet their posterity approve\\
what they do fondly say.

14 Like sheep they in the grave are laid,\\
and death shall them devour;\\
And in the morning upright men\\
shall over them have pow’r:

Their beauty from their dwelling shall\\
consume within the grave.\\
15 But from hell’s hand God will me free,\\
for he shall me receive.

16 Be thou not then afraid when one\\
enriched thou dost see,\\
Nor when the glory of his house\\
advanced is on high:

17 For he shall carry nothing hence\\
when death his days doth end;\\
Nor shall his glory after him\\
into the grave descend.

18 Although he his own soul did bless\\
whilst he on earth did live;\\
(And when thou to thyself dost well,\\
men will thee praises give;)

19 He to his fathers’ race shall go,\\
they never shall see light.\\
20 Man honored wanting knowledge is\\
like beasts that perish quite.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 50}
\addcontentsline{toc}{subsection}{Psalm 50}

\emph{First Version (S.M., 6.6.8.6)}

A Psalm of Asaph.

1 The mighty God, the Lord,\\
hath spoken, and did call\\
The earth, from rising of the sun,\\
to where he hath his fall.

2 From out of Zion hill,\\
which of excellency\\
And beauty the perfection is,\\
God shined gloriously.

3 Our God shall surely come,\\
keep silence shall not he:\\
Before him fire shall waste, great storms\\
shall round about him be.

4 Unto the heavens clear\\
he from above shall call,\\
And to the earth likewise, that he\\
may judge his people all.

5 Together let my saints\\
unto me gathered be,\\
Those that by sacrifice have made\\
a covenant with me.

6 And then the heavens shall\\
his righteousness declare:\\
Because the Lord himself is he\\
by whom men judged are.

7 My people Isr’el hear,\\
speak will I from on high,\\
Against thee I will testify;\\
God, ev’n thy God, am I.

8 I for thy sacrifice\\
no blame will on thee lay,\\
Nor for burnt-off’rings, which to me\\
thou offer’dst ev’ry day.

9 I’ll take no calf nor goats\\
from house or fold of thine:\\
10 For beasts of forests, cattle all\\
on thousand hills, are mine.

11 The fowls on mountains high\\
are all to me well known;\\
Wild beasts which in the fields do lie,\\
ev’n they are all mine own.

12 Then, if I hungry were,\\
I would not tell it thee;\\
Because the world, and fullness all\\
thereof, belongs to me.

13 Will I eat flesh of bulls?\\
or goats’ blood drink will I?\\
14 Thanks offer thou to God, and pay\\
thy vows to the most High.

15 And call upon me when\\
in trouble thou shalt be;\\
I will deliver thee, and thou\\
my name shalt glorify.

16 But to the wicked man\\
God saith, My laws and truth\\
Should’st thou declare? how dar’st thou take\\
my cov’nant in thy mouth?

17 Since thou instruction hat’st,\\
which should thy ways direct;\\
And since my words behind thy back\\
thou cast’st, and dost reject.

18 When thou a thief didst see,\\
with him thou didst consent;\\
And with the vile adulterers\\
partaker on thou went.

19 Thou giv’st thy mouth to ill,\\
thy tongue deceit doth frame;\\
20 Thou sitt’st, and ‘gainst thy brother speak’st,\\
thy mother’s son dost shame.

21 Because I silence kept,\\
while thou these things hast wrought;\\
That I was altogether like\\
thyself, hath been thy thought;

Yet I will thee reprove,\\
and set before thine eyes,\\
In order ranked, thy misdeeds\\
and thine iniquities.

22 Now, ye that God forget,\\
this carefully consider;\\
Lest I in pieces tear you all,\\
and none can you deliver.

23 Whoso doth offer praise\\
me glorifies; and I\\
Will show him God’s salvation,\\
that orders right his way.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\emph{Second Version (C.M., 8.6.8.6)}

A Psalm of Asaph.

1 The mighty God, the Lord, hath spoke,\\
and called the earth upon,\\
Ev’n from the rising of the sun\\
unto his going down.

2 From out of Zion, his own hill,\\
where the perfection high\\
Of beauty is, from thence the Lord\\
hath shined gloriously.

3 Our God shall come, and shall no more\\
be silent, but speak out:\\
Before him fire shall waste, great storms\\
shall compass him about.

4 He to the heavens from above,\\
and to the earth below,\\
Shall call, that he his judgments may\\
before his people show.

5 Let all my saints together be\\
unto me gathered;\\
Those that by sacrifice with me\\
a covenant have made.

6 And then the heavens shall declare\\
his righteousness abroad:\\
Because the Lord himself doth come;\\
none else is judge but God.

7 Hear, O my people, and I’ll speak;\\
O Israel by name,\\
Against thee I will testify;\\
God, ev’n thy God, I am.

8 I for thy sacrifices few\\
reprove thee never will,\\
Nor for burnt-off’rings to have been\\
before me offered still.

9 I’ll take no bullock nor he-goats\\
from house nor folds of thine:\\
10 For beasts of forests, cattle all\\
on thousand hills, are mine.

11 The fowls are all to me well known\\
that mountains high do yield;\\
And I do challenge as mine own\\
the wild beasts of the field.

12 If I were hungry, I would not\\
to thee for need complain;\\
For earth, and all its fullness, doth\\
to me of right pertain.

13 That I to eat the flesh of bulls\\
take pleasure dost thou think?\\
Or that I need, to quench my thirst,\\
the blood of goats to drink?

14 Nay, rather unto me, thy God,\\
thanksgiving offer thou:\\
To the most High perform thy word,\\
and fully pay thy vow:

15 And in the day of trouble great\\
see that thou call on me;\\
I will deliver thee, and thou\\
my name shalt glorify.

16 But God unto the wicked saith,\\
Why should’st thou mention make\\
Of my commands? how dar’st thou in\\
thy mouth my cov’nant take?

17 Since it is so that thou dost hate\\
all good instruction;\\
And since thou cast’st behind thy back,\\
and slight’st my words each one.

18 When thou a thief didst see, then straight\\
thou join’dst with him in sin,\\
And with the vile adulterers\\
thou hast partaker been.

19 Thy mouth to evil thou dost give,\\
thy tongue deceit doth frame.\\
20 Thou sitt’st, and ‘gainst thy brother speak’st,\\
thy mother’s son to shame.

21 These things thou wickedly hast done,\\
and I have silent been:\\
Thou thought’st that I was like thyself,\\
and did approve thy sin:

But I will sharply thee reprove,\\
and I will order right\\
Thy sins and thy transgressions\\
in presence of thy sight.

22 Consider this, and be afraid,\\
ye that forget the Lord,\\
Lest I in pieces tear you all,\\
when none can help afford.

23 Who off’reth praise me glorifies:\\
I will shew God’s salvation\\
To him that ordereth aright\\
his life and conversation.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 51}
\addcontentsline{toc}{subsection}{Psalm 51}

To the chief Musician, A Psalm of David,
when Nathan the prophet came unto him, after he had gone into Bathsheba.

1 After thy loving-kindness, Lord,\\
have mercy upon me:\\
For thy compassions great, blot out\\
all mine iniquity.

2 Me cleanse from sin, and thoroughly wash\\
from mine iniquity:\\
3 For my transgressions I confess;\\
my sin I ever see.

4 ‘Gainst thee, thee only, have I sinned,\\
in thy sight done this ill;\\
That when thou speak’st thou may’st be just,\\
and clear in judging still.

5 Behold, I in iniquity\\
was formed the womb within;\\
My mother also me conceived\\
in guiltiness and sin.

6 Behold, thou in the inward parts\\
with truth delighted art;\\
And wisdom thou shalt make me know\\
within the hidden part.

7 Do thou with hyssop sprinkle me,\\
I shall be cleansed so;\\
Yea, wash thou me, and then I shall\\
be whiter than the snow.

8 Of gladness and of joyfulness\\
make me to hear the voice;\\
That so these very bones which thou\\
hast broken may rejoice.

9 All mine iniquities blot out,\\
thy face hide from my sin.\\
10 Create a clean heart, Lord, renew\\
a right sp’rit me within.

11 Cast me not from thy sight, nor take\\
thy Holy Sp’rit away.\\
12 Restore me thy salvation’s joy;\\
with thy free Sp’rit me stay.

13 Then will I teach thy ways unto\\
those that transgressors be;\\
And those that sinners are shall then\\
be turned unto thee.

14 O God, of my salvation God,\\
me from blood-guiltiness\\
Set free; then shall my tongue aloud\\
sing of thy righteousness.

15 My closed lips, O Lord, by thee\\
let them be opened;\\
Then shall thy praises by my mouth\\
abroad be published.

16 For thou desir’st not sacrifice,\\
else would I give it thee;\\
Nor wilt thou with burnt-offering\\
at all delighted be.

17 A broken spirit is to God\\
a pleasing sacrifice:\\
A broken and a contrite heart,\\
Lord, thou wilt not despise.

18 Show kindness, and do good, O Lord,\\
to Zion, thine own hill:\\
The walls of thy Jerusalem\\
build up of thy good will.

19 Then righteous off’rings shall thee please,\\
and off’rings burnt, which they\\
With whole burnt-off’rings, and with calves,\\
shall on thine altar lay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 52}
\addcontentsline{toc}{subsection}{Psalm 52}

To the chief Musician, Maschil, \emph{A Psalm} of David,
when Doeg the Edomite came and told Saul, and said unto him, David is come to the house of Ahimelech.

1 Why dost thou boast, O mighty man,\\
of mischief and of ill?\\
The goodness of Almighty God\\
endureth ever still.

2 Thy tongue mischievous calumnies\\
deviseth subtilely,\\
Like to a razor sharp to cut,\\
working deceitfully.

3 Ill more than good, and more than truth\\
thou lovest to speak wrong:\hfill \emph{Selah}\\
4 Thou lovest all-devouring words,\\
O thou deceitful tongue.

5 So God shall thee destroy for aye,\\
remove thee, pluck thee out\\
Quite from thy house, out of the land\\
of life he shall thee root.\hfill\emph{Selah}

6 The righteous shall it see, and fear,\\
and laugh at him they shall:\\
7 Lo, this the man is that did not\\
make God his strength at all:

But he in his abundant wealth\\
his confidence did place;\\
And he took strength unto himself\\
from his own wickedness.

8 But I am in the house of God\\
like to an olive green:\\
My confidence for ever hath\\
upon God’s mercy been.

9 And I for ever will thee praise,\\
because thou hast done this:\\
I on thy name will wait; for good\\
before thy saints it is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 53}
\addcontentsline{toc}{subsection}{Psalm 53}

To the chief Musician upon Mahalath, Maschil,
\emph{A Psalm} of David.

1 That there is not a God, the fool\\
doth in his heart conclude:\\
They are corrupt, their works are vile,\\
not one of them doth good.

2 The Lord upon the sons of men\\
from heav’n did cast his eyes,\\
To see if any one there was\\
that sought God, and was wise.

3 They altogether filthy are,\\
they all are backward gone;\\
And there is none that doeth good,\\
no, not so much as one.

4 These workers of iniquity,\\
do they not know at all,\\
That they my people eat as bread,\\
and on God do not call?

5 Ev’n there they were afraid, and stood\\
with trembling, all dismayed,\\
Whereas there was no cause at all\\
why they should be afraid:

For God his bones that thee besieged\\
hath scattered all abroad;\\
Thou hast confounded them, for they\\
despised are of God.

6 Let Isr’el’s help from Zion come:\\
when back the Lord shall bring\\
His captives, Jacob shall rejoice,\\
and Israel shall sing.


% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 54}
\addcontentsline{toc}{subsection}{Psalm 54}

To the chief Musician on Neginoth, Maschil, \emph{A Psalm} of David, when the Ziphims came and said to Saul, Doth not David hide himself with us?

1 Save me, O God, by thy great name,\\
and judge me by thy strength:\\
2 My prayer hear, O God; give ear\\
unto my words at length.

3 For they that strangers are to me\\
do up against me rise;\\
Oppressors seek my soul, and God\\
set not before their eyes.

4 The Lord my God my helper is,\\
lo, therefore I am bold:\\
He taketh part with ev’ry one\\
that doth my soul uphold.

5 Unto mine enemies he shall\\
mischief and ill repay:\\
O for thy truth’s sake cut them off,\\
and sweep them clean away.

6 I will a sacrifice to thee\\
give with free willingness;\\
Thy name, O Lord, because ’tis good,\\
with praise I will confess.

7 For he hath me delivered\\
from all adversities;\\
And his desire mine eye hath seen\\
upon mine enemies.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 55}
\addcontentsline{toc}{subsection}{Psalm 55}

To the chief Musician on Neginoth, Maschil,
\emph{A Psalm} of David.

1 Lord, hear my pray’r, hide not thyself\\
from my entreating voice:\\
2 Attend and hear me; in my plaint\\
I mourn and make a noise.

3 Because of th’ en’my’s voice, and for\\
lewd men’s oppression great:\\
On me they cast iniquity,\\
and they in wrath me hate.

4 Sore pained within me is my heart:\\
death’s terrors on me fall.\\
5 On me comes trembling, fear and dread\\
o’erwhelmed me withal.

6 O that I, like a dove, had wings,\\
said I, then would I flee\\
Far hence, that I might find a place\\
where I in rest might be.

7 Lo, then far off I wander would,\\
and in the desert stay;\\
8 From windy storm and tempest I\\
would haste to ‘scape away.

9 O Lord, on them destruction bring,\\
and do their tongues divide;\\
For in the city violence\\
and strife I have espied.

10 They day and night upon the walls\\
do go about it round:\\
There mischief is, and sorrow there\\
in midst of it is found.

11 Abundant wickedness there is\\
within her inward part;\\
And from her streets deceitfulness\\
and guile do not depart.

12 He was no foe that me reproached,\\
then that endure I could;\\
Nor hater that did ‘gainst me boast,\\
from him me hide I would.

13 But thou, man, who mine equal, guide,\\
and mine acquaintance wast:\\
14 We joined sweet counsels, to God’s house\\
in company we past.

15 Let death upon them seize, and down\\
let them go quick to hell;\\
For wickedness doth much abound\\
among them where they dwell.

16 I’ll call on God: God will me save.\\
17 I’ll pray, and make a noise\\
At ev’ning, morning, and at noon;\\
and he shall hear my voice.

18 He hath my soul delivered,\\
that it in peace might be\\
From battle that against me was;\\
for many were with me.

19 The Lord shall hear, and them afflict,\\
of old who hath abode:\\
Because they never changes have,\\
therefore they fear not God.

20 ‘Gainst those that were at peace with him\\
he hath put forth his hand:\\
The covenant that he had made,\\
by breaking he profaned.

21 More smooth than butter were his words,\\
while in his heart was war;\\
His speeches were more soft than oil,\\
and yet drawn swords they are.

22 Cast thou thy burden on the Lord,\\
and he shall thee sustain;\\
Yea, he shall cause the righteous man\\
unmoved to remain.

23 But thou, O Lord my God, those men\\
in justice shalt o’erthrow,\\
And in destruction’s dungeon dark\\
at last shalt lay them low:

The bloody and deceitful men\\
shall not live half their days:\\
But upon thee with confidence\\
I will depend always.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 56}
\addcontentsline{toc}{subsection}{Psalm 56}


To the chief Musician upon Jonath-elem-rechokim, Michtam of David, when the Philistines took him in Gath.

1 Show mercy, Lord, to me, for man\\
would swallow me outright;\\
He me oppresseth, while he doth\\
against me daily fight.

2 They daily would me swallow up\\
that hate me spitefully;\\
For they be many that do fight\\
against me, O most High.

3 When I’m afraid I’ll trust in thee:\\
4 In God I’ll praise his word;\\
I will not fear what flesh can do,\\
my trust is in the Lord.

5 Each day they wrest my words; their thoughts\\
‘gainst me are all for ill.\\
6 They meet, they lurk, they mark my steps,\\
waiting my soul to kill.

7 But shall they by iniquity\\
escape thy judgments so?\\
O God, with indignation down\\
do thou the people throw.

8 My wand’rings all what they have been\\
thou know’st, their number took;\\
Into thy bottle put my tears:\\
are they not in thy book?

9 My foes shall, when I cry, turn back;\\
I know’t, God is for me.\\
10 In God his word I’ll praise; his word\\
in God shall praised be.

11 In God I trust; I will not fear\\
what man can do to me.\\
12 Thy vows upon me are, O God:\\
I’ll render praise to thee.

13 Wilt thou not, who from death me saved,\\
my feet from falls keep free,\\
To walk before God in the light\\
of those that living be?

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 57}
\addcontentsline{toc}{subsection}{Psalm 57}

To the chief Musician, Al-taschith, Michtam
of David, when he fled from Saul in the cave.

1 Be merciful to me, O God;\\
thy mercy unto me\\
Do thou extend; because my soul\\
doth put her trust in thee:

Yea, in the shadow of thy wings\\
my refuge I will place,\\
Until these sad calamities\\
do wholly overpass.

2 My cry I will cause to ascend\\
unto the Lord most high;\\
To God, who doth all things for me\\
perform most perfectly.

3 From heav’n he shall send down, and me\\
from his reproach defend\\
That would devour me: God his truth\\
and mercy forth shall send.

4 My soul among fierce lions is,\\
I firebrands live among,\\
Men’s sons, whose teeth are spears and darts,\\
a sharp sword is their tongue.

5 Be thou exalted very high\\
above the heav’ns, O God;\\
Let thou thy glory be advanced\\
o’er all the earth abroad.

6 My soul’s bowed down; for they a net\\
have laid, my steps to snare:\\
Into the pit which they have digged\\
for me, they fallen are.

7 My heart is fixed, my heart is fixed,\\
O God; I’ll sing and praise.\\
8 My glory wake; wake psalt’ry, harp;\\
myself I’ll early raise.

9 I’ll praise thee ‘mong the people, Lord;\\
‘mong nations sing will I:\\
10 For great to heav’n thy mercy is,\\
thy truth is to the sky.

11 O Lord, exalted be thy name\\
above the heav’ns to stand:\\
Do thou thy glory far advance\\
above both sea and land.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 58}
\addcontentsline{toc}{subsection}{Psalm 58}

To the chief Musician, Al-taschith, Michtam of David.

1 Do ye, O congregation,\\
indeed speak righteousness?\\
O ye that are the sons of men,\\
judge ye with uprightness?

2 Yea, ev’n within your very hearts\\
ye wickedness have done;\\
And ye the vi’lence of your hands\\
do weigh the earth upon.

3 The wicked men estranged are,\\
ev’n from the very womb;\\
They, speaking lies, do stray as soon\\
as to the world they come.

4 Unto a serpent’s poison like\\
their poison doth appear;\\
Yea, they are like the adder deaf,\\
that closely stops her ear;

5 That so she may not hear the voice\\
of one that charm her would,\\
No, not though he most cunning were,\\
and charm most wisely could.

6 Their teeth, O God, within their mouth\\
break thou in pieces small;\\
The great teeth break thou out, O Lord,\\
of these young lions all.

7 Let them like waters melt away,\\
which downward still do flow:\\
In pieces cut his arrows all,\\
when he shall bend his bow.

8 Like to a snail that melts away,\\
let each of them be gone;\\
Like woman’s birth untimely, that\\
they never see the sun.

9 He shall them take away before\\
your pots the thorns can find,\\
Both living, and in fury great,\\
as with a stormy wind.

10 The righteous, when he vengeance sees,\\
he shall be joyful then;\\
The righteous one shall wash his feet\\
in blood of wicked men.

11 So men shall say, The righteous man\\
reward shall never miss:\\
And verily upon the earth\\
a God to judge there is.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 59}
\addcontentsline{toc}{subsection}{Psalm 59}

To the chief Musician, Al-taschith, Michtam of David;
when Saul sent, and they watched the house to kill him.

1 My God, deliver me from those\\
that are mine enemies;\\
And do thou me defend from those\\
that up against me rise.

2 Do thou deliver me from them\\
that work iniquity;\\
And give me safety from the men\\
of bloody cruelty.

3 For, lo, they for my soul lay wait:\\
the mighty do combine\\
Against me, Lord; not for my fault,\\
nor any sin of mine.

4 They run, and, without fault in me,\\
themselves do ready make:\\
Awake to meet me with thy help;\\
and do thou notice take.

5 Awake therefore, Lord God of hosts,\\
thou God of Israel,\\
To visit heathen all: spare none\\
that wickedly rebel.

6 At ev’ning they go to and fro;\\
they make great noise and sound,\\
Like to a dog, and often walk\\
about the city round.

7 Behold, they belch out with their mouth,\\
and in their lips are swords:\\
For they do say thus, Who is he\\
that now doth hear our words?

8 But thou, O Lord, shalt laugh at them,\\
and all the heathen mock.\\
9 While he’s in pow’r I’ll wait on thee;\\
for God is my high rock.

10 He of my mercy that is God\\
betimes shall me prevent;\\
Upon mine en’mies God shall let\\
me see mine heart’s content.

11 Them slay not, lest my folk forget;\\
but scatter them abroad\\
By thy strong pow’r; and bring them down,\\
O thou our shield and God.

12 For their mouth’s sin, and for the words\\
that from their lips do fly,\\
Let them be taken in their pride;\\
because they curse and lie.

13 In wrath consume them, them consume,\\
that so they may not be:\\
And that in Jacob God doth rule\\
to th’ earth’s ends let them see.

14 At ev’ning let thou them return,\\
making great noise and sound,\\
Like to a dog, and often walk\\
about the city round.

15 And let them wander up and down,\\
in seeking food to eat;\\
And let them grudge when they shall not\\
be satisfied with meat.

16 But of thy pow’r I’ll sing aloud;\\
at morn thy mercy praise:\\
For thou to me my refuge wast,\\
and tow’r, in troublous days.

17 O God, thou art my strength, I will\\
sing praises unto thee;\\
For God is my defense, a God\\
of mercy unto me.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 60}
\addcontentsline{toc}{subsection}{Psalm 60}
To the chief Musician upon Shushan-eduth, Michtam of David, to teach; when he strove with Aram-naharaim, and with Aram-zobah, when Joab returned and smote of Edom, in the valley of Salt, twelve thousand.

1 O Lord, thou hast rejected us,\\
and scattered us abroad;\\
Thou justly hast displeased been;\\
return to us, O God.

2 The earth to tremble thou hast made;\\
therein didst breaches make:\\
Do thou thereof the breaches heal,\\
because the land doth shake.

3 Unto thy people thou hard things\\
hast showed, and on them sent;\\
And thou hast caused us to drink\\
wine of astonishment.

4 And yet a banner thou hast giv’n\\
to them who thee do fear;\\
That it by them, because of truth,\\
displayed may appear.

5 That thy beloved people may\\
delivered be from thrall,\\
Save with the pow’r of thy right hand,\\
and hear me when I call.

6 God in his holiness hath spoke;\\
herein I will take pleasure:\\
Shechem I will divide, and forth\\
will Succoth’s valley measure.

7 Gilead I claim as mine by right;\\
Manasseh mine shall be;\\
Ephraim is of mine head the strength;\\
Judah gives laws for me;

8 Moab’s my washing-pot; my shoe\\
I’ll over Edom throw;\\
And over Palestina’s land\\
I will in triumph go.

9 O who is he will bring me to\\
the city fortified?\\
O who is he that to the land\\
of Edom will me guide?

10 O God, which hadest us cast off,\\
this thing wilt thou not do?\\
Ev’n thou, O God, which didest not\\
forth with our armies go?

11 Help us from trouble; for the help\\
is vain which man supplies.\\
12 Through God we’ll do great acts; he shall\\
tread down our enemies.



% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 61}
\addcontentsline{toc}{subsection}{Psalm 61}

To the chief Musician upon Neginoth,
\emph{A Psalm} of David.

1 O God, give ear unto my cry;\\
unto my pray’r attend.\\
2 From th’ utmost corner of the land\\
my cry to thee I’ll send.

What time my heart is overwhelmed,\\
and in perplexity,\\
Do thou me lead unto the Rock\\
that higher is than I.

3 For thou hast for my refuge been\\
a shelter by thy pow’r;\\
And for defense against my foes\\
thou hast been a strong tow’r.

4 Within thy tabernacle I\\
for ever will abide;\\
And under covert of thy wings\\
with confidence me hide.

5 For thou the vows that I did make,\\
O Lord my God, didst hear:\\
Thou hast giv’n me the heritage\\
of those thy name that fear.

6 A life prolonged for many days\\
thou to the king shalt give;\\
Like many generations be\\
the years which he shall live.

7 He in God’s presence his abode\\
for evermore shall have:\\
O do thou truth and mercy both\\
prepare, that may him save.

8 And so will I perpetually\\
sing praise unto thy name;\\
That having made my vows, I may\\
each day perform the same.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 62}
\addcontentsline{toc}{subsection}{Psalm 62}

To the chief Musician, to Jeduthun, A Psalm of David.

1 My soul with expectation\\
depends on God indeed;\\
My strength and my salvation doth\\
from him alone proceed.

2 He only my salvation is,\\
and my strong rock is he:\\
He only is my sure defense;\\
much moved I shall not be.

3 How long will ye against a man\\
plot mischief? ye shall all\\
Be slain; ye as a tott’ring fence\\
shall be, and bowing wall.

4 They only plot to cast him down\\
from his excellency:\\
They joy in lies; with mouth they bless,\\
but they curse inwardly.

5 My soul, wait thou with patience\\
upon thy God alone;\\
On him dependeth all my hope\\
and expectation.

6 He only my salvation is,\\
and my strong rock is he;\\
He only is my sure defense:\\
I shall not moved be.

7 In God my glory placed is,\\
and my salvation sure;\\
In God the rock is of my strength,\\
my refuge most secure.

8 Ye people, place your confidence\\
in him continually;\\
Before him pour ye out your heart:\\
God is our refuge high.

9 Surely mean men are vanity,\\
and great men are a lie;\\
In balance laid, they wholly are\\
more light than vanity.

10 Trust ye not in oppression,\\
in robb’ry be not vain;\\
On wealth set not your hearts, when as\\
increased is your gain.

11 God hath it spoken once to me,\\
yea, this I heard again,\\
That power to Almighty God\\
alone doth appertain.

12 Yea, mercy also unto thee\\
belongs, O Lord, alone:\\
For thou according to his work\\
rewardest ev’ry one.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 63}
\addcontentsline{toc}{subsection}{Psalm 63}

A Psalm of David, when he was in the wilderness of Judah.

1 Lord, thee my God, I’ll early seek:\\
my soul doth thirst for thee;\\
My flesh longs in a dry parched land,\\
wherein no waters be:

2 That I thy power may behold,\\
and brightness of thy face,\\
As I have seen thee heretofore\\
within thy holy place.

3 Since better is thy love than life,\\
my lips thee praise shall give.\\
4 I in thy name will lift my hands,\\
and bless thee while I live.

5 Ev’n as with marrow and with fat\\
my soul shall filled be;\\
Then shall my mouth with joyful lips\\
sing praises unto thee:

6 When I do thee upon my bed\\
remember with delight,\\
And when on thee I meditate\\
in watches of the night.

7 In shadow of thy wings I’ll joy;\\
for thou mine help hast been.\\
8 My soul thee follows hard; and me\\
thy right hand doth sustain.

9 Who seek my soul to spill shall sink\\
down to earth’s lowest room.\\
10 They by the sword shall be cut off,\\
and foxes’ prey become.

11 Yet shall the king in God rejoice,\\
and each one glory shall\\
That swear by him: but stopped shall be\\
the mouth of liars all.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 64}
\addcontentsline{toc}{subsection}{Psalm 64}

To the chief Musician,
A Psalm of David.

1 When I to thee my prayer make,\\
Lord, to my voice give ear;\\
My life save from the enemy,\\
of whom I stand in fear.

2 Me from their secret counsel hide\\
who do live wickedly;\\
From insurrection of those men\\
that work iniquity:

3 Who do their tongues with malice whet,\\
and make them cut like swords;\\
In whose bent bows are arrows set,\\
ev’n sharp and bitter words:

4 That they may at the perfect man\\
in secret aim their shot;\\
Yea, suddenly they dare at him\\
to shoot, and fear it not.

5 In ill encourage they themselves,\\
and their snares close do lay:\\
Together conference they have;\\
Who shall them see? they say.

6 They have searched out iniquities,\\
a perfect search they keep:\\
Of each of them the inward thought,\\
and very heart, is deep.

7 God shall an arrow shoot at them,\\
and wound them suddenly:\\
8 So their own tongue shall them confound;\\
all who them see shall fly.

9 And on all men a fear shall fall,\\
God’s works they shall declare;\\
For they shall wisely notice take\\
what these his doings are.

10 In God the righteous shall rejoice,\\
and trust upon his might;\\
Yea, they shall greatly glory all\\
in heart that are upright.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 65}
\addcontentsline{toc}{subsection}{Psalm 65}

1 Praise waits for thee in Zion, Lord:\\
to thee vows paid shall be.\\
2 O thou that hearer art of pray’r,\\
all flesh shall come to thee.

3 Iniquities, I must confess,\\
prevail against me do:\\
But as for our transgressions,\\
them purge away shalt thou.

4 Blessed is the man whom thou dost choose,\\
and mak’st approach to thee,\\
That he within thy courts, O Lord,\\
may still a dweller be:

We surely shall be satisfied\\
with thy abundant grace,\\
And with the goodness of thy house,\\
ev’n of thy holy place.

5 O God of our salvation,\\
thou, in thy righteousness,\\
By fearful works unto our pray’rs\\
thine answer dost express:

Therefore the ends of all the earth,\\
and those afar that be\\
Upon the sea, their confidence,\\
O Lord, will place in thee.

6 Who, being girt with pow’r, sets fast\\
by his great strength the hills.\\
7 Who noise of seas, noise of their waves,\\
and people’s tumult, stills.

8 Those in the utmost parts that dwell\\
are at thy signs afraid:\\
Th’ outgoings of the morn and ev’n\\
by thee are joyful made.

9 The earth thou visit’st, wat’ring it;\\
thou mak’st it rich to grow\\
With God’s full flood; thou corn prepar’st,\\
when thou provid’st it so.

10 Her rigs thou wat’rest plenteously,\\
her furrows settelest:\\
With show’rs thou dost her mollify,\\
her spring by thee is blest.

11 So thou the year most lib’rally\\
dost with thy goodness crown;\\
And all thy paths abundantly\\
on us drop fatness down.

12 They drop upon the pastures wide,\\
that do in deserts lie;\\
The little hills on ev’ry side\\
rejoice right pleasantly.

13 With flocks the pastures clothed be,\\
the vales with corn are clad;\\
And now they shout and sing to thee,\\
for thou hast made them glad.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 66}
\addcontentsline{toc}{subsection}{Psalm 66}

To the chief Musician,
A Song \emph{or} Psalm.

1 All lands to God in joyful sounds,\\
aloft your voices raise.\\
2 Sing forth the honor of his name,\\
and glorious make his praise.

3 Say unto God, How terrible\\
in all thy works art thou!\\
Through thy great pow’r thy foes to thee\\
shall be constrained to bow.

4 All on the earth shall worship thee,\\
they shall thy praise proclaim\\
In songs: they shall sing cheerfully\\
unto thy holy name.

5 Come, and the works that God hath wrought\\
with admiration see:\\
In’s working to the sons of men\\
most terrible is he.

6 Into dry land the sea he turned,\\
and they a passage had;\\
Ev’n marching through the flood on foot,\\
there we in him were glad.

7 He ruleth ever by his pow’r;\\
his eyes the nations see:\\
O let not the rebellious ones\\
lift up themselves on high.

8 Ye people, bless our God; aloud\\
the voice speak of his praise:\\
9 Our soul in life who safe preserves,\\
our foot from sliding stays.

10 For thou didst prove and try us, Lord,\\
as men do silver try;\\
11 Brought’st us into the net, and mad’st\\
bands on our loins to lie.

12 Thou hast caused men ride o’er our heads;\\
and though that we did pass\\
Through fire and water, yet thou brought’st\\
us to a wealthy place.

13 I’ll bring burnt off ‘rings to thy house;\\
to thee my vows I’ll pay,\\
14 Which my lips uttered, my mouth spake,\\
when trouble on me lay.

15 Burnt-sacrifices of fat rams\\
with incense I will bring;\\
Of bullocks and of goats I will\\
present an offering.

16 All that fear God, come, hear, I’ll tell\\
what he did for my soul.\\
17 I with my mouth unto him cried,\\
my tongue did him extol.

18 If in my heart I sin regard,\\
the Lord me will not hear:\\
19 But surely God me heard, and to\\
my prayer’s voice gave ear.

20 O let the Lord, our gracious God,\\
for ever blessed be,\\
Who turned not my pray’r from him,\\
nor yet his grace from me.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 67}
\addcontentsline{toc}{subsection}{Psalm 67}

To the chief Musician on Neginoth,
A Psalm \emph{or} Song.

\emph{First Version---Short Meter (S.M. or 6.6.8.6)}

1 Lord, bless and pity us,\\
shine on us with thy face:\\
2 That th’ earth thy way, and nations all\\
may know thy saving grace.

3 Let people praise thee, Lord;\\
let people all thee praise.\\
4 O let the nations be glad,\\
in songs their voices raise:

Thou’lt justly people judge,\\
on earth rule nations all.\\
5 Let people praise thee, Lord; let them\\
praise thee, both great and small.

6 The earth her fruit shall yield,\\
our God shall blessing send.\\
7 God shall us bless; men shall him fear\\
unto earth’s utmost end.

\emph{Second Version---Common Meter (C.M. 8.6.8.6}

1 Lord, unto us be merciful,\\
do thou us also bless;\\
And graciously cause shine on us\\
the brightness of thy face:

2 That so thy way upon the earth\\
to all men may be known;\\
Also among the nations all\\
thy saving health be shown.

3 O let the people praise thee, Lord;\\
let people all thee praise.\\
4 O let the nations be glad,\\
and sing for joy always:

For rightly thou shalt people judge,\\
and nations rule on earth.\\
5 Let people praise thee, Lord; let all\\
the folk praise thee with mirth.

6 Then shall the earth yield her increase;\\
God, our God, bless us shall.\\
7 God shall us bless; and of the earth\\
the ends shall fear him all.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 68}
\addcontentsline{toc}{subsection}{Psalm 68}

To the chief Musician,
A Psalm \emph{or} Song of David.

1 Let God arise, and scattered\\
let all his en’mies be;\\
And let all those that do him hate\\
before his presence flee.

2 As smoke is driv’n, so drive thou them;\\
as fire melts wax away,\\
Before God’s face let wicked men\\
so perish and decay.

3 But let the righteous be glad:\\
let them before God’s sight\\
Be very joyful; yea, let them\\
rejoice with all their might.

4 To God sing, to his name sing praise;\\
extol him with your voice,\\
That rides on heav’n, by his name Jah,\\
before his face rejoice.

5 Because the Lord a father is\\
unto the fatherless;\\
God is the widow’s judge, within\\
his place of holiness.

6 God doth the solitary set\\
in fam’lies: and from bands\\
The chain’d doth free; but rebels do\\
inhabit parched lands.

7 O God, what time thou didst go forth\\
before thy people’s face;\\
And when through the great wilderness\\
thy glorious marching was;

8 Then at God’s presence shook the earth,\\
then drops from heaven fell;\\
This Sinai shook before the Lord,\\
the God of Israel.

9 O God, thou to thine heritage\\
didst send a plenteous rain,\\
Whereby thou, when it weary was,\\
didst it refresh again.

10 Thy congregation then did make\\
their habitation there:\\
Of thine own goodness for the poor,\\
O God, thou didst prepare.

11 The Lord himself did give the word,\\
the word abroad did spread;\\
Great was the company of them\\
the same who published.

12 Kings of great armies foiled were,\\
and forced to flee away;\\
And women, who remained at home,\\
did distribute the prey.

13 Though ye have lien among the pots,\\
like doves ye shall appear,\\
Whose wings with silver, and with gold\\
whose feathers covered are.

14 When there th’ Almighty scatter’d kings,\\
like Salmon’s snow ’twas white.\\
15 God’s hill is like to Bashan hill,\\
like Bashan hill for height.

16 Why do ye leap, ye mountains high?\\
this is the hill where God\\
Desires to dwell; yea, God in it\\
for aye will make abode.

17 God’s chariots twenty thousand are,\\
thousands of angels strong;\\
In’s holy place God is, as in\\
mount Sinai, them among.

18 Thou hast, O Lord, most glorious,\\
ascended up on high;\\
And in triumph victorious led\\
captive captivity:

Thou hast received gifts for men,\\
for such as did rebel;\\
Yea, ev’n for them, that God the Lord\\
in midst of them might dwell.

19 Blessed be the Lord, who is to us\\
of our salvation God;\\
Who daily with his benefits\\
us plenteously doth load.

20 He of salvation is the God,\\
who is our God most strong;\\
And unto God the Lord from death\\
the issues do belong.

21 But surely God shall wound the head\\
of those that are his foes;\\
The hairy scalp of him that still\\
on in his trespass goes.

22 God said, My people I will bring\\
again from Bashan hill;\\
Yea, from the sea’s devouring depths\\
them bring again I will;

23 That in the blood of enemies\\
thy foot imbrued may be,\\
And of thy dogs dipped in the same\\
the tongues thou mayest see.

24 Thy goings they have seen, O God;\\
the steps of majesty\\
Of my God, and my mighty King,\\
within the sanctuary.

25 Before went singers, players next\\
on instruments took way;\\
And them among the damsels were\\
that did on timbrels play.

26 Within the congregations\\
bless God with one accord:\\
From Isr’el’s fountain do ye bless\\
and praise the mighty Lord.

27 With their prince, little Benjamin,\\
princes and council there\\
Of Judah were, there Zabulon’s\\
and Napht’li’s princes were.

28 Thy God commands thy strength; make strong\\
what thou wrought’st for us, Lord.\\
29 For thy house at Jerusalem\\
kings shall thee gifts afford.

30 The spearmen’s host, the multitude\\
of bulls, which fiercely look,\\
Those calves which people have forth sent,\\
O Lord our God, rebuke,

Till ev’ry one submit himself,\\
and silver pieces bring:\\
The people that delight in war\\
disperse, O God and King.

31 Those that be princes great shall then\\
come out of Egypt lands;\\
And Ethiopia to God\\
shall soon stretch out her hands.

32 O all ye kingdoms of the earth,\\
sing praises to this King;\\
For he is Lord that ruleth all,\\
unto him praises sing.

33 To him that rides on heav’ns of heav’ns,\\
which he of old did found;\\
Lo, he sends out his voice, a voice\\
in might that doth abound.

34 Strength unto God do ye ascribe;\\
for his excellency\\
Is over Israel, his strength\\
is in the clouds most high.

35 Thou’rt from thy temple dreadful, Lord;\\
Isr’el’s own God is he,\\
Who gives his people strength and pow’r:\\
O let God blessed be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}
\subsection*{Psalm 69}
\addcontentsline{toc}{subsection}{Psalm 69}

To the chief Musician upon Shoshannim,
\emph{A Psalm} of David.

1 Save me, O God, because the floods\\
do so environ me,\\
That ev’n unto my very soul\\
come in the waters be.

2 I downward in deep mire do sink,\\
where standing there is none:\\
I am into deep waters come,\\
where floods have o’er me gone.

3 I weary with my crying am,\\
my throat is also dried;\\
Mine eyes do fail, while for my God\\
I waiting do abide.

4 Those men that do without a cause\\
bear hatred unto me,\\
Than are the hairs upon my head\\
in number more they be:

They that would me destroy, and are\\
mine en’mies wrongfully,\\
Are mighty: so what I took not,\\
to render forced was I.

5 Lord, thou my folly know’st, my sins\\
not covered are from thee.\\
6 Let none that wait on thee be shamed,\\
Lord God of hosts, for me.

O Lord, the God of Israel,\\
let none, who search do make,\\
And seek thee, be at any time\\
confounded for my sake.

7 For I have borne reproach for thee,\\
my face is hid with shame.\\
8 To brethren strange, to mother’s sons\\
an alien I became.

9 Because the zeal did eat me up,\\
which to thine house I bear;\\
And the reproaches cast at thee,\\
upon me fallen are.

10 My tears and fasts, t’ afflict my soul,\\
were turned to my shame.\\
11 When sackcloth I did wear, to them\\
a proverb I became.

12 The men that in the gate do sit\\
against me evil spake;\\
They also that vile drunkards were\\
of me their song did make.

13 But, in an acceptable time,\\
my pray’r, Lord, is to thee:\\
In truth of thy salvation, Lord,\\
and mercy great, hear me.

14 Deliver me out of the mire,\\
from sinking do me keep;\\
Free me from those that do me hate,\\
and from the waters deep.

15 Let not the flood on me prevail,\\
whose water overflows;\\
Nor deep me swallow, nor the pit\\
her mouth upon me close.

16 Hear me, O Lord, because thy love\\
and kindness is most good;\\
Turn unto me, according to\\
thy mercies’ multitude.

17 Nor from thy servant hide thy face:\\
I’m troubled, soon attend.\\
18 Draw near my soul, and it redeem;\\
me from my foes defend.

19 To thee is my reproach well known,\\
my shame, and my disgrace:\\
Those that mine adversaries be\\
are all before thy face.

20 Reproach hath broke my heart; I’m full\\
of grief: I looked for one\\
To pity me, but none I found;\\
comforters found I none.

21 They also bitter gall did give\\
unto me for my meat:\\
They gave me vinegar to drink,\\
when as my thirst was great.

22 Before them let their table prove\\
a snare; and do thou make\\
Their welfare and prosperity\\
a trap themselves to take.

23 Let thou their eyes so darkened be,\\
that sight may them forsake;\\
And let their loins be made by thee\\
continually to shake.

24 Thy fury pour thou out on them,\\
and indignation;\\
And let thy wrathful anger, Lord,\\
fast hold take them upon.

25 All waste and desolate let be\\
their habitation;\\
And in their tabernacles all\\
inhabitants be none.

26 Because him they do persecute,\\
whom thou didst smite before;\\
They talk unto the grief of those\\
whom thou hast wounded sore.

27 Add thou iniquity unto\\
their former wickedness;\\
And do not let them come at all\\
into thy righteousness.

28 Out of the book of life let them\\
be razed and blotted quite;\\
Among the just and righteous\\
let not their names be writ.

29 But now become exceeding poor\\
and sorrowful am I:\\
By thy salvation, O my God,\\
let me be set on high.

30 The name of God I with a song\\
most cheerfully will praise;\\
And I, in giving thanks to him,\\
his name shall highly raise.

31 This to the Lord a sacrifice\\
more gracious shall prove\\
Than bullock, ox, or any beast\\
that hath both horn and hoof.

32 When this the humble men shall see,\\
it joy to them shall give:\\
O all ye that do seek the Lord,\\
your hearts shall ever live.

33 For God the poor hears, and will not\\
his prisoners contemn.\\
34 Let heav’n, and earth, and seas, him praise,\\
and all that move in them.

35 For God will Judah’s cities build,\\
and he will Zion save,\\
That they may dwell therein, and it\\
in sure possession have.

36 And they that are his servants’ seed\\
inherit shall the same;\\
So shall they have their dwelling there\\
that love his blessed name.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 70}
\addcontentsline{toc}{subsection}{Psalm 70}

To the chief Musician,
\emph{A Psalm} of David, to bring to remembrance.

\emph{First Version---Short Meter (S.M. 6.6.8.6)}

1 Lord, haste me to deliver;\\
with speed, Lord, succor me.\\
2 Let them that for my soul do seek\\
shamed and confounded be:

Turned back be they, and shamed,\\
that in my hurt delight.\\
3 Turned back be they, Ha, ha! that say,\\
their shaming to requite.

4 In thee let all be glad,\\
and joy that seek for thee:\\
Let them who thy salvation love\\
say still, God praised be.

5 I poor and needy am;\\
come, Lord, and make no stay:\\
My help thou and deliv’rer art;\\
O Lord, make no delay.

\emph{Second Version---Common Meter (C.M. 8.6.8.6.)}

To the chief Musician,
\emph{A Psalm} of David, to bring to remembrance.

1 Make haste, O God, me to preserve;\\
with speed, Lord, succor me.\\
2 Let them that for my soul do seek\\
shamed and confounded be:

Let them be turned back, and shamed,\\
that in my hurt delight.\\
3 Turned back be they, Ha, ha! that say,\\
their shaming to requite.

4 O Lord, in thee let all be glad,\\
and joy that seek for thee:\\
Let them who thy salvation love\\
say still, God praised be.

5 But I both poor and needy am;\\
come, Lord, and make no stay:\\
My help thou and deliv’rer art;\\
O Lord, make no delay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}
\subsection*{Psalm 71}
\addcontentsline{toc}{subsection}{Psalm 71}

1 O Lord, my hope and confidence\\
is placed in thee alone;\\
Then let thy servant never be\\
put to confusion.

2 And let me, in thy righteousness,\\
from thee deliv’rance have;\\
Cause me escape, incline thine ear\\
unto me, and me save.

3 Be thou my dwelling-rock, to which\\
I ever may resort:\\
Thou gav’st commandment me to save,\\
for thou’rt my rock and fort.

4 Free me, my God, from wicked hands,\\
hands cruel and unjust:\\
5 For thou, O Lord God, art my hope,\\
and from my youth my trust.

6 Thou from the womb didst hold me up;\\
thou art the same that me\\
Out of my mother’s bowels took;\\
I ever will praise thee.

7 To many I a wonder am;\\
but thou’rt my refuge strong.\\
8 Filled let my mouth be with thy praise\\
and honor all day long.

9 O do not cast me off, when as\\
old age doth overtake me;\\
And when my strength decayed is,\\
then do not thou forsake me.

10 For those that are mine enemies\\
against me speak with hate;\\
And they together counsel take\\
that for my soul lay wait.

11 They said, God leaves him; him pursue\\
and take: none will him save.\\
12 Be thou not far from me, my God:\\
thy speedy help I crave.

13 Confound, consume them, that unto\\
my soul are enemies:\\
Clothed be they with reproach and shame\\
that do my hurt devise.

14 But I with expectation\\
will hope continually;\\
And yet with praises more and more\\
I will thee magnify.

15 Thy justice and salvation\\
my mouth abroad shall show,\\
Ev’n all the day; for I thereof\\
the numbers do not know.

16 And I will constantly go on\\
in strength of God the Lord;\\
And thine own righteousness, ev’n thine\\
alone, I will record.

17 For even from my youth, O God,\\
by thee I have been taught;\\
And hitherto I have declared\\
the wonders thou hast wrought.

18 And now, Lord, leave me not, when I\\
old and gray-headed grow:\\
Till to this age thy strength and pow’r\\
to all to come I show.

19 And thy most perfect righteousness\\
O Lord, is very high,\\
Who hast so great things done: O God,\\
who is like unto thee?

20 Thou, Lord, who great adversities,\\
and sore, to me didst show,\\
Shalt quicken, and bring me again\\
from depths of earth below.

21 My greatness and my pow’r thou wilt\\
increase, and far extend:\\
On ev’ry side against all grief\\
thou wilt me comfort send.

22 Thee, ev’n thy truth, I’ll also praise,\\
my God, with psaltery:\\
Thou Holy One of Israel,\\
with harp I’ll sing to thee.

23 My lips shall much rejoice in thee,\\
when I thy praises sound;\\
My soul, which thou redeemed hast,\\
in joy shall much abound.

24 My tongue thy justice shall proclaim,\\
continuing all day long;\\
For they confounded are, and shamed,\\
that seek to do me wrong.

1 O Lord, thy judgments give the king,\\
his son thy righteousness.\\
2 With right he shall thy people judge,\\
thy poor with uprightness.

3 The lofty mountains shall bring forth\\
unto the people peace;\\
Likewise the little hills the same\\
shall do by righteousness.

4 The people’s poor ones he shall judge,\\
the needy’s children save;\\
And those shall he in pieces break\\
who them oppressed have.

5 They shall thee fear, while sun and moon\\
do last, through ages all.\\
6 Like rain on mown grass he shall drop,\\
or show’rs on earth that fall.

7 The just shall flourish in his days,\\
and prosper in his reign:\\
He shall, while doth the moon endure,\\
abundant peace maintain.

8 His large and great dominion shall\\
from sea to sea extend:\\
It from the river shall reach forth\\
unto earth’s utmost end.

9 They in the wilderness that dwell\\
bow down before him must;\\
And they that are his enemies\\
shall lick the very dust.

10 The kings of Tarshish, and the isles,\\
to him shall presents bring;\\
And unto him shall offer gifts\\
Sheba’s and Seba’s king.

11 Yea, all the mighty kings on earth\\
before him down shall fall;\\
And all the nations of the world\\
do service to him shall.

12 For he the needy shall preserve,\\
when he to him doth call;\\
The poor also, and him that hath\\
no help of man at all.

13 The poor man and the indigent\\
in mercy he shall spare;\\
He shall preserve alive the souls\\
of those that needy are.

14 Both from deceit and violence\\
their soul he shall set free;\\
And in his sight right precious\\
and dear their blood shall be.

15 Yea, he shall live, and giv’n to him\\
shall be of Sheba’s gold:\\
For him still shall they pray, and he\\
shall daily be extolled.

16 Of corn an handful in the earth\\
on tops of mountains high,\\
With prosp’rous fruit shall shake, like trees\\
on Lebanon that be.

The city shall be flourishing,\\
her citizens abound\\
In number shall, like to the grass\\
that grows upon the ground.

17 His name for ever shall endure;\\
last like the sun it shall:\\
Men shall be blessed in him, and blessed\\
all nations shall him call.

18 Now blessed be the Lord our God,\\
the God of Israel,\\
For he alone doth wondrous works,\\
in glory that excel.

19 And blessed be his glorious name\\
to all eternity:\\
The whole earth let his glory fill.\\
Amen, so let it be.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 72}
\addcontentsline{toc}{subsection}{Psalm 72}


\emph{A Psalm} for Solomon.



% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 70}
\addcontentsline{toc}{subsection}{Psalm 66}

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 70}
\addcontentsline{toc}{subsection}{Psalm 66}

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 70}
\addcontentsline{toc}{subsection}{Psalm 66}

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\end{multicols}

\begin{center}
\Large{\sc{Book III}}
\addcontentsline{toc}{section}{Book III}
\end{center}
\begin{multicols}{3}

\end{multicols}

\begin{center}
\Large{\sc{Book IV}}
\addcontentsline{toc}{section}{Book IV}
\end{center}
\begin{multicols}{3}


\end{multicols}

\begin{center}
\Large{\sc{Book V}}
\addcontentsline{toc}{section}{Book V}
\end{center}
\begin{multicols}{3}


\subsection*{Psalm 119}
\addcontentsline{toc}{subsection}{Psalm 119}

\subsubsection*{Aleph}

1 Blessed are they that undefiled,\\
and straight are in the way;\\
Who in the Lord’s most holy law\\
do walk, and do not stray.

2 Blessed are they who to observe\\
his statutes are inclined;\\
And who do seek the living God\\
with their whole heart and mind.

3 Such in his ways do walk, and they\\
do no iniquity.\\
4 Thou hast commanded us to keep\\
thy precepts carefully.

5 O that thy statutes to observe\\
thou would’st my ways direct!\\
6 Then shall I not be shamed, when I\\
thy precepts all respect.

7 Then with integrity of heart\\
thee will I praise and bless,\\
When I the judgments all have learned\\
of thy pure righteousness.

8 That I will keep thy statutes all\\
firmly resolved have I:\\
O do not then, most gracious God,\\
forsake me utterly.

\subsubsection*{Beth}

9 By what means shall a young man learn\\
his way to purify?\\
If he according to thy word\\
thereto attentive be.

10 Unfeignedly thee have I sought\\
with all my soul and heart:\\
O let me not from the right path\\
of thy commands depart.

11 Thy word I in my heart have hid,\\
that I offend not thee.\\
12 O Lord, thou ever blessed art,\\
thy statutes teach thou me.

13 The judgments of thy mouth each one\\
my lips declared have:\\
14 More joy thy testimonies' way\\
than riches all me gave.

15 I will thy holy precepts make\\
my meditation;\\
And carefully I’ll have respect\\
unto thy ways each one.

16 Upon thy statutes my delight\\
shall constantly be set:\\
And, by thy grace, I never will\\
thy holy word forget.

\subsubsection*{Gimel}

17 With me thy servant, in thy grace,\\
deal bountifully, Lord;\\
That by thy favor I may live,\\
and duly keep thy word.

18 Open mine eyes, that of thy law\\
the wonders I may see.\\
19 I am a stranger on this earth,\\
hide not thy laws from me.

20 My soul within me breaks, and doth\\
much fainting still endure,\\
Through longing that it hath all times\\
unto thy judgments pure.

21 Thou hast rebuked the cursed proud,\\
who from thy precepts swerve.\\
22 Reproach and shame remove from me,\\
for I thy laws observe.

23 Against me princes spake with spite,\\
while they in council sat:\\
But I thy servant did upon\\
thy statutes meditate.

24 My comfort, and my heart’s delight,\\
thy testimonies be;\\
And they, in all my doubts and fears,\\
are counselors to me

\subsubsection*{Daleth}

25 My soul to dust cleaves: quicken me,\\
according to thy word.\\
26 My ways I showed, and me thou heard’st:\\
teach me thy statutes, Lord.

27 The way of thy commandements\\
make me aright to know;\\
So all thy works that wondrous are\\
I shall to others show.

28 My soul doth melt, and drop away,\\
for heaviness and grief:\\
To me, according to thy word,\\
give strength, and send relief.

29 From me the wicked way of lies\\
let far removed be;\\
And graciously thy holy law\\
do thou grant unto me.

30 I chosen have the perfect way\\
of truth and verity:\\
Thy judgments that most righteous are\\
before me laid have I.

31 I to thy testimonies cleave;\\
shame do not on me cast.\\
32 I’ll run thy precepts’ way, when thou\\
my heart enlarged hast.

\subsubsection*{He}

33 Teach me, O Lord, the perfect way\\
of thy precepts divine,\\
And to observe it to the end\\
I shall my heart incline.

34 Give understanding unto me,\\
so keep thy law shall I;\\
Yea, ev’n with my whole heart I shall\\
observe it carefully.

35 In thy law’s path make me to go;\\
for I delight therein.\\
36 My heart unto thy testimonies,\\
and not to greed, incline.

37 Turn thou away my sight and eyes\\
from viewing vanity;\\
And in thy good and holy way\\
be pleased to quicken me.

38 Confirm to me thy gracious word,\\
which I did gladly hear,\\
Ev’n to thy servant, Lord, who is\\
devoted to thy fear.

39 Turn thou away my fear’d reproach;\\
for good thy judgments be.\\
40 Lo, for thy precepts I have long’d;\\
in thy truth quicken me.

\subsubsection*{Vau}

41 Let thy sweet mercies also come\\
and visit me, O Lord;\\
Ev’n thy benign salvation,\\
according to thy word.

42 So shall I have wherewith I may\\
give him an answer just,\\
Who spitefully reproacheth me;\\
for in thy word I trust.

43 The word of truth out of my mouth\\
take thou not utterly;\\
For on thy judgments righteous\\
my hope doth still rely.

44 So shall I keep for evermore\\
thy law continually.\\
45 And, sith that I thy precepts seek,\\
I’ll walk at liberty.

46 I’ll speak thy word to kings, and I\\
with shame shall not be moved;\\
47 And will delight myself always\\
in thy laws, which I loved.

48 To thy commandements, which I loved,\\
my hands lift up I will;\\
And I will also meditate\\
upon thy statutes still

\subsubsection*{Zain}

49 Remember, Lord, thy gracious word\\
thou to thy servant spake,\\
Which, for a ground of my sure hope,\\
thou causedst me to take.

50 This word of thine my comfort is\\
in mine affliction:\\
For in my straits I am revived\\
by this thy word alone.

51 The men whose hearts with pride are stuffed\\
did greatly me deride;\\
Yet from thy straight commandements\\
I have not turned aside.

52 Thy judgments righteous, O Lord,\\
which thou of old forth gave,\\
I did remember, and myself\\
by them comforted have.

53 Horror took hold on me, because\\
ill men thy law forsake.\\
54 I in my house of pilgrimage\\
thy laws my songs do make.

55 Thy name by night, Lord, I did mind,\\
and I have kept thy law.\\
56 And this I had, because thy word\\
I kept, and stood in awe.

\subsubsection*{Cheth}

57 Thou my sure portion art alone,\\
which I did choose, O Lord:\\
I have resolved, and said, that I\\
would keep thy holy word.

58 With my whole heart I did entreat\\
thy face and favor free:\\
According to thy gracious word\\
be merciful to me.

59 I thought upon my former ways,\\
and did my life well try;\\
And to thy testimonies pure\\
my feet then turned I.

60 I did not stay, nor linger long,\\
as those that slothful are;\\
But hastily thy laws to keep\\
myself I did prepare.

61 Bands of ill men me robbed; yet I\\
thy precepts did not slight.\\
62 I’ll rise at midnight thee to praise,\\
ev’n for thy judgments right.

63 I am companion to all those\\
who fear, and thee obey.\\
64 O Lord, thy mercy fills the earth:\\
teach me thy laws, I pray.

\subsubsection*{Teth}

65 Well hast thou with thy servant dealt,\\
as thou didst promise give.\\
66 Good judgment me, and knowledge teach,\\
for I thy word believe.

67 Ere I afflicted was I strayed;\\
but now I keep thy word.\\
68 Both good thou art, and good thou do’st:\\
teach me thy statutes, Lord.

69 The men that are puffed up with pride\\
against me forged a lie;\\
Yet thy commandements observe\\
with my whole heart will I.

70 Their hearts, through worldly ease and wealth,\\
as fat as grease they be:\\
But in thy holy law I take\\
delight continually.

71 It hath been very good for me\\
that I afflicted was,\\
That I might well instructed be,\\
and learn thy holy laws.

72 The word that cometh from thy mouth\\
is better unto me\\
Than many thousands and great sums\\
of gold and silver be.

\subsubsection*{Jod}

73 Thou mad’st and fashion’dst me: thy laws\\
to know give wisdom, Lord.\\
74 So who thee fear shall joy to see\\
me trusting in thy word.

75 That very right thy judgments are\\
I know, and do confess;\\
And that thou hast afflicted me\\
in truth and faithfulness.

76 O let thy kindness merciful,\\
I pray thee, comfort me,\\
As to thy servant faithfully\\
was promised by thee.

77 And let thy tender mercies come\\
to me, that I may live;\\
Because thy holy laws to me\\
sweet delectation give.

78 Lord, let the proud ashamed be;\\
for they, without a cause,\\
With me perversely dealt: but I\\
will muse upon thy laws.

79 Let such as fear thee, and have known\\
thy statutes, turn to me.\\
80 My heart let in thy laws be sound,\\
that shamed I never be.


Caph

81 My soul for thy salvation faints;\\
yet I thy word believe.\\
82 Mine eyes fail for thy word: I say,\\
When wilt thou comfort give?

83 For like a bottle I’m become,\\
that in the smoke is set:\\
I’m black, and parched with grief; yet I\\
thy statutes not forget.

84 How many are thy servant’s days?\\
when wilt thou execute\\
Just judgment on these wicked men\\
that do me persecute?

85 The proud have digged pits for me,\\
which is against thy laws.\\
86 Thy words all faithful are: help me,\\
pursued without a cause.

87 They so consumed me, that on earth\\
my life they scarce did leave:\\
Thy precepts yet forsook I not,\\
but close to them did cleave.

88 After thy loving-kindness, Lord,\\
me quicken, and preserve:\\
The testimony of thy mouth\\
so shall I still observe.

\subsubsection*{Lamed}

89 Thy word for ever is, O Lord,\\
in heaven settled fast;\\
90 Unto all generations\\
thy faithfulness doth last:

The earth thou hast established,\\
and it abides by thee.\\
91 This day they stand as thou ordain’dst;\\
for all thy servants be.

92 Unless in thy most perfect law\\
my soul delights had found,\\
I should have perished, when as\\
my troubles did abound.

93 Thy precepts I will ne’er forget;\\
they quick’ning to me brought.\\
94 Lord, I am thine; O save thou me:\\
thy precepts I have sought.

95 For me the wicked have laid wait,\\
me seeking to destroy:\\
But I thy testimonies true\\
consider will with joy.

96 An end of all perfection\\
here have I seen, O God:\\
But as for thy commandment,\\
it is exceeding broad.

\subsubsection*{Mem}

97 O how love I thy law! it is\\
my study all the day:\\
98 It makes me wiser than my foes;\\
for it doth with me stay.

99 Than all my teachers now I have\\
more understanding far;\\
Because my meditation\\
thy testimonies are.

100 In understanding I excel\\
those that are ancients;\\
For I endeavored to keep\\
all thy commandements.

101 My feet from each ill way I stayed,\\
that I may keep thy word.\\
102 I from thy judgments have not swerved;\\
for thou hast taught me, Lord.

103 How sweet unto my taste, O Lord,\\
are all thy words of truth!\\
Yea, I do find them sweeter far\\
than honey to my mouth.

104 I through thy precepts, that are pure,\\
do understanding get;\\
I therefore ev’ry way that’s false\\
with all my heart do hate

\subsubsection*{Nun}

105 Thy word is to my feet a lamp,\\
and to my path a light.\\
106 I sworn have, and I will perform,\\
to keep thy judgments right.

107 I am with sore affliction\\
ev’n overwhelmed, O Lord:\\
In mercy raise and quicken me,\\
according to thy word.

108 The free-will-off ‘rings of my mouth\\
accept, I thee beseech:\\
And unto me thy servant, Lord,\\
thy judgments clearly teach.

109 Though still my soul be in my hand,\\
thy laws I’ll not forget.\\
110 I erred not from them, though for me\\
the wicked snares did set.

111 I of thy testimonies have\\
above all things made choice,\\
To be my heritage for aye;\\
for they my heart rejoice.

112 I carefully inclined have\\
my heart still to attend;\\
That I thy statutes may perform\\
alway unto the end.

\subsubsection*{Samech}

113 I hate the thoughts of vanity,\\
but love thy law do I.\\
114 My shield and hiding-place thou art:\\
I on thy word rely.

115 All ye that evil-doers are\\
from me depart away;\\
For the commandements of my God\\
I purpose to obey.

116 According to thy faithful word\\
uphold and stablish me,\\
That I may live, and of my hope\\
ashamed never be.

117 Hold thou me up, so shall I be\\
in peace and safety still;\\
And to thy statutes have respect\\
continually I will.

118 Thou tread’st down all that love to stray;\\
false their deceit doth prove.\\
119 Lewd men, like dross, away thou putt’st;\\
therefore thy law I love.

120 For fear of thee my very flesh\\
doth tremble, all dismayed;\\
And of thy righteous judgments, Lord,\\
my soul is much afraid.

\subsubsection*{Ain}

121 To all men I have judgment done,\\
performing justice right;\\
Then let me not be left unto\\
my fierce oppressors’ might.

122 For good unto thy servant, Lord,\\
thy servant’s surety be:\\
From the oppression of the proud\\
do thou deliver me.

123 Mine eyes do fail with looking long\\
for thy salvation,\\
The word of thy pure righteousness\\
while I do wait upon.

124 In mercy with thy servant deal,\\
thy laws me teach and show.\\
125 I am thy servant, wisdom give,\\
that I thy laws may know.

126 ‘Tis time thou work, Lord; for they have\\
made void thy law divine.\\
127 Therefore thy precepts more I love\\
than gold, yea, gold most fine.

128 Concerning all things thy commands\\
all right I judge therefore;\\
And ev’ry false and wicked way\\
I perfectly abhor.


\subsubsection*{Pe}

129 Thy statutes, Lord, are wonderful,\\
my soul them keeps with care.\\
130 The entrance of thy words gives light,\\
makes wise who simple are.

131 My mouth I have wide opened,\\
and panted earnestly,\\
While after thy commandements\\
I longed exceedingly.

132 Look on me, Lord, and merciful\\
do thou unto me prove,\\
As thou art wont to do to those\\
thy name who truly love.

133 O let my footsteps in thy word\\
aright still ordered be:\\
Let no iniquity obtain\\
dominion over me.

134 From man’s oppression save thou me;\\
so keep thy laws I will.\\
135 Thy face make on thy servant shine;\\
teach me thy statutes still.

136 Rivers of waters from mine eyes\\
did run down, when I saw\\
How wicked men run on in sin,\\
and do not keep thy law.

\subsubsection*{Tzaddi}

137 O Lord, thou art most righteous;\\
thy judgments are upright.\\
138 Thy testimonies thou command’st\\
most faithful are and right.

139 My zeal hath ev’n consumed me,\\
because mine enemies\\
Thy holy words forgotten have,\\
and do thy laws despise.

140 Thy word’s most pure, therefore on it\\
thy servant’s love is set.\\
141 Small, and despised I am, yet I\\
thy precepts not forget.

142 Thy righteousness is righteousness\\
which ever doth endure:\\
Thy holy law, Lord, also is\\
the very truth most pure.

143 Trouble and anguish have me found,\\
and taken hold on me:\\
Yet in my trouble my delight\\
thy just commandements be.

144 Eternal righteousness is in\\
thy testimonies all:\\
Lord, to me understanding give,\\
and ever live I shall.

\subsubsection*{Koph}

145 With my whole heart I cried, Lord, hear;\\
I will thy word obey.\\
146 I cried to thee; save me, and I\\
will keep thy laws alway.

147 I of the morning did prevent\\
the dawning, and did cry:\\
For all mine expectation\\
did on thy word rely.

148 Mine eyes did timeously prevent\\
the watches of the night,\\
That in thy word with careful mind\\
then meditate I might.

149 After thy loving-kindness hear\\
my voice, that calls on thee:\\
According to thy judgment, Lord,\\
revive and quicken me.

150 Who follow mischief they draw nigh;\\
they from thy law are far:\\
151 But thou art near, Lord; most firm truth\\
all thy commandements are.

152 As for thy testimonies all,\\
of old this have I tried,\\
That thou hast surely founded them\\
for ever to abide.

\subsubsection*{Resh}

153 Consider mine affliction,\\
in safety do me set:\\
Deliver me, O Lord, for I\\
thy law do not forget.

154 After thy word revive thou me:\\
save me, and plead my cause.\\
155 Salvation is from sinners far;\\
for they seek not thy laws.

156 O Lord, both great and manifold\\
thy tender mercies be:\\
According to thy judgments just,\\
revive and quicken me.

157 My persecutors many are,\\
and foes that do combine;\\
Yet from thy testimonies pure\\
my heart doth not decline.

158 I saw transgressors, and was grieved;\\
for they keep not thy word.\\
159 See how I love thy law! as thou\\
art kind, me quicken, Lord.

160 From the beginning all thy word\\
hath been most true and sure:\\
Thy righteous judgments ev’ry one\\
for evermore endure.

\subsubsection*{Schin}

161 Princes have persecuted me,\\
although no cause they saw:\\
But still of thy most holy word\\
my heart doth stand in awe.

162 I at thy word rejoice, as one\\
of spoil that finds great store.\\
163 Thy law I love; but lying all\\
I hate and do abhor.

164 Sev’n times a-day it is my care\\
to give due praise to thee;\\
Because of all thy judgments, Lord,\\
which righteous ever be.

165 Great peace have they who love thy law;\\
offence they shall have none.\\
166 I hoped for thy salvation, Lord,\\
and thy commands have done.

167 My soul thy testimonies pure\\
observed carefully;\\
On them my heart is set, and them\\
I love exceedingly.

168 Thy testimonies and thy laws\\
I kept with special care;\\
For all my works and ways each one\\
before thee open are.

\subsubsection*{Tau}

169 O let my earnest pray’r and cry\\
come near before thee, Lord:\\
Give understanding unto me,\\
according to thy word.

170 Let my request before thee come:\\
after thy word me free.\\
171 My lips shall utter praise, when thou\\
hast taught thy laws to me.

172 My tongue of thy most blessed word\\
shall speak, and it confess;\\
Because all thy commandements\\
are perfect righteousness.

173 Let thy strong hand make help to me:\\
thy precepts are my choice.\\
174 I longed for thy salvation, Lord,\\
and in thy law rejoice.

175 O let my soul live, and it shall\\
give praises unto thee;\\
And let thy judgments gracious\\
be helpful unto me.

176 I, like a lost sheep, went astray;\\
thy servant seek, and find:\\
For thy commands I suffered not\\
to slip out of my mind.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 120}
\addcontentsline{toc}{subsection}{Psalm 120}

A Song of degrees.

1 In my distress to God I cried,\\
and he gave ear to me.\\
2 From lying lips, and guileful tongue,\\
O Lord, my soul set free.

3 What shall be giv’n thee? or what shall\\
be done to thee, false tongue?\\
4 Ev’n burning coals of juniper,\\
sharp arrows of the strong.

5 Woe’s me that I in Mesech am\\
a sojourner so long;\\
That I in tabernacles dwell\\
to Kedar that belong.

6 My soul with him that hateth peace\\
hath long a dweller been.\\
7 I am for peace; but when I speak,\\
for battle they are keen.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 121}
\addcontentsline{toc}{subsection}{Psalm 121}

A Song of degrees.

1 I to the hills will lift mine eyes,\\
from whence doth come mine aid.\\
2 My safety cometh from the Lord,\\
who heav’n and earth hath made.

3 Thy foot he’ll not let slide, nor will\\
he slumber that thee keeps.\\
4 Behold, he that keeps Israel,\\
he slumbers not, nor sleeps.

5 The Lord thee keeps, the Lord thy shade\\
on thy right hand doth stay:\\
6 The moon by night thee shall not smite,\\
nor yet the sun by day.

7 The Lord shall keep thy soul; he shall\\
preserve thee from all ill.\\
8 Henceforth thy going out and in\\
God keep for ever will.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 122}
\addcontentsline{toc}{subsection}{Psalm 122}

A Song of degrees of David.

1 I joyed when to the house of God,\\
Go up, they said to me.\\
2 Jerusalem, within thy gates\\
our feet shall standing be.

3 Jerus’lem, as a city, is\\
compactly built together:\\
4 Unto that place the tribes go up,\\
the tribes of God go thither:

To Isr’el’s testimony, there\\
to God’s name thanks to pay.\\
5 For thrones of judgment, ev’n the thrones\\
of David’s house, there stay.

6 Pray that Jerusalem may have\\
peace and felicity:\\
Let them that love thee and thy peace\\
have still prosperity.

7 Therefore I wish that peace may still\\
within thy walls remain,\\
And ever may thy palaces\\
prosperity retain.

8 Now, for my friends’ and brethren’s sakes,\\
Peace be in thee, I’ll say.\\
9 And for the house of God our Lord,\\
I’ll seek thy good alway.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 123}
\addcontentsline{toc}{subsection}{Psalm 123}

A Song of degrees.

1 O thou that dwellest in the heav’ns,\\
I lift mine eyes to thee.\\
2 Behold, as servants’ eyes do look\\
their masters’ hand to see,

As handmaid’s eyes her mistress’ hand;\\
so do our eyes attend\\
Upon the Lord our God, until\\
to us he mercy send.

3 O Lord, be gracious unto us,\\
unto us gracious be;\\
Because replenished with contempt\\
exceedingly are we.

4 Our soul is filled with scorn of those\\
that at their ease abide,\\
And with the insolent contempt\\
of those that swell in pride.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 124}
\addcontentsline{toc}{subsection}{Psalm 124}

A Song of degrees of David.

1 Had not the Lord been on our side,\\
may Israel now say;\\
2 Had not the Lord been on our side,\\
when men rose us to slay;

3 They had us swallow’d quick, when as\\
their wrath ’gainst us did flame:\\
4 Waters had cover’d us, our soul\\
had sunk beneath the stream.

5 Then had the waters, swelling high,\\
over our soul made way.\\
6 Bless’d be the Lord, who to their teeth\\
us gave not for a prey.

7 Our soul’s escaped, as a bird\\
out of the fowler’s snare;\\
The snare asunder broken is,\\
and we escaped are.

8 Our sure and all-sufficient help\\
is in JEHOVAH’s name;\\
His name who did the heav’n create,\\
and who the earth did frame.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 125}
\addcontentsline{toc}{subsection}{Psalm 125}

A Song of degrees.

1 They in the Lord that firmly trust\\
shall be like Zion hill,\\
Which at no time can be removed,\\
but standeth ever still.

2 As round about Jerusalem\\
the mountains stand alway,\\
The Lord his folk doth compass so,\\
from henceforth and for aye.

3 For ill men’s rod upon the lot\\
of just men shall not lie;\\
Lest righteous men stretch forth their hands\\
unto iniquity.

4 Do thou to all those that be good\\
thy goodness, Lord, impart;\\
And do thou good to those that are\\
upright within their heart.

5 But as for such as turn aside\\
after their crooked way,\\
God shall lead forth with wicked men:\\
on Isr’el peace shall stay.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 126}
\addcontentsline{toc}{subsection}{Psalm 126}

A Song of degrees.

1 When Zion’s bondage God turned back,\\
as men that dreamed were we.\\
2 Then filled with laughter was our mouth,\\
our tongue with melody:

They ‘mong the heathen said, The Lord\\
great things for them hath wrought.\\
3 The Lord hath done great things for us,\\
whence joy to us is brought.

4 As streams of water in the south,\\
our bondage, Lord, recall.\\
5 Who sow in tears, a reaping time\\
of joy enjoy they shall.

6 That man who, bearing precious seed,\\
in going forth doth mourn,\\
He doubtless, bringing back his sheaves,\\
rejoicing shall return.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 127}
\addcontentsline{toc}{subsection}{Psalm 127}

A Song of degrees for Solomon.

1 Except the Lord do build the house,\\
the builders lose their pain:\\
Except the Lord the city keep,\\
the watchmen watch in vain.

2 ‘Tis vain for you to rise betimes,\\
or late from rest to keep,\\
To feed on sorrows’ bread; so gives\\
he his beloved sleep.

3 Lo, children are God’s heritage,\\
the womb’s fruit his reward.\\
4 The sons of youth as arrows are,\\
for strong men’s hands prepared.

5 O happy is the man that hath\\
his quiver filled with those;\\
They unashamed in the gate\\
shall speak unto their foes.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 128}
\addcontentsline{toc}{subsection}{Psalm 128}

A Song of degrees.

1 Blessed is each one that fears the Lord,\\
and walketh in his ways;\\
2 For of thy labour thou shalt eat,\\
and happy be always.

3 Thy wife shall as a fruitful vine\\
by thy house’ sides be found:\\
Thy children like to olive-plants\\
about thy table round.

4 Behold, the man that fears the Lord,\\
thus blessed shall he be.\\
5 The Lord shall out of Zion give\\
his blessing unto thee:

Thou shalt Jerus’lem’s good behold\\
whilst thou on earth dost dwell.\\
6 Thou shalt thy children’s children see,\\
and peace on Israel.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}

\subsection*{Psalm 129}
\addcontentsline{toc}{subsection}{Psalm 129}


A Song of degrees.

1 Oft did they vex me from my youth,\\
may Isr’el now declare;\\
2 Oft did they vex me from my youth,\\
yet not victorious were.

3 The plowers plowed upon my back;\\
they long their furrows drew.\\
4 The righteous Lord did cut the cords\\
of the ungodly crew.

5 Let Zion’s haters all be turned\\
back with confusion.\\
6 As grass on houses’ tops be they,\\
which fades ere it be grown:

7 Whereof enough to fill his hand\\
the mower cannot find;\\
Nor can the man his bosom fill,\\
whose work is sheaves to bind.

8 Neither say they who do go by,\\
God’s blessing on you rest:\\
We in the name of God the Lord\\
do wish you to be blest.

% Separation Symbol
\begin{center}
\adforn{21}\quad\adforn{11}\quad\adforn{49}
\end{center}


\subsection*{Psalm 130}
\addcontentsline{toc}{subsection}{Psalm 130}

A Song of degrees.

1 Lord, from the depths to thee I cried.\\
2 My voice, Lord, do thou hear:\\
Unto my supplication’s voice\\
give an attentive ear.

3 Lord, who shall stand, if thou, O Lord,\\
should’st mark iniquity?\\
4 But yet with thee forgiveness is,\\
that feared thou mayest be.

5 I wait for God, my soul doth wait,\\
my hope is in his word.\\
6 More than they that for morning watch,\\
my soul waits for the Lord;

I say, more than they that do watch\\
the morning light to see.\\
7 Let Israel hope in the Lord,\\
for with him mercies be;

And plenteous redemption\\
is ever found with him.\\
8 And from all his iniquities\\
he Isr’el shall redeem.

\end{multicols}



